### 2023-02-05 Paper/paper.tex
add paper.tex

\documentclass[12pt]{report}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ulem} % underline
\usepackage{hyperref} % hyperlinks
\usepackage{hyphenat}
\usepackage{bookmark}
\usepackage{systeme} % systems of equation
\usepackage{parcolumns} % multiple column parallel typesetting
\usepackage{float} % H float specifier
\usepackage[english]{babel} % comment this for Angelsächsisch
\usepackage{graphicx} % including graphics
\usepackage{sourcecodepro} % best font
\usepackage{listings} % source code listings
\usepackage[utf8]{inputenc} % necessary for proper unicode support
% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry

% umlauts in lstlistings env
\lstset{
    literate={ö}{{\"o}}1
    {ä}{{\"a}}1
    {ü}{{\"u}}1
    {Ö}{{\"O}}1
    {Ä}{{\"A}}1
    {Ü}{{\"U}}1
}

\lstset{
    basicstyle=\footnotesize\ttfamily,
    breaklines=true
}

\lstset{
    numbers=left,
    stepnumber=1,
    firstnumber=1,
    numberfirstline=true
}

\providecommand{\lxor}{\veebar}
\renewcommand{\proofname}{Beweis.}
\newcommand{\sref}[1]{\textsuperscript{\ref{#1}}}

\title{
\textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\
{\footnotesize Bachelorarbeit}
}
\author{Bastian Engel}
\date{\today}

\begin{document}
\maketitle

\chapter*{Abstract}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\chapter*{Declaration}

I hereby declare that ...

\tableofcontents

\chapter{Introduction}
\section{Block ciphers}

Modern-day communication relies on ...

\subsection{GIFT}

GIFT is a lightweight block cipher based on ...

\subsubsection{Substitution layer}

\subsubsection{Permutation layer}

\subsubsection{Round key addition}

\subsubsection{Round key extraction}

\subsection{Camellia}

\section{The ARMv8 platform}

With small devices, embedded processors and ASICs becoming ever more ubiquitous
and essential in areas like medicine or automotive design, the need for ...

\chapter{A simple implementation}

\chapter{Optimizations through bitslicing}

\chapter{Leveraging NEON advanced SIMD instructions}

\chapter*{Acknowledgements}

I want to thank ...

\end{document}

### 2023-02-09 Paper/paper.tex
add refs.bib, start work on introduction

@@
 \usepackage{ulem} % underline
+\usepackage{float} % H float specifier
 \usepackage{hyperref} % hyperlinks
@@
 \usepackage{parcolumns} % multiple column parallel typesetting
-\usepackage{float} % H float specifier
 \usepackage[english]{babel} % comment this for Angelsächsisch
@@
 \begin{document}
+
 \maketitle
@@
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
+Lorem ipsum dolor \cite{gift:2017} sit amet, consectetur adipisicing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
@@
 
-Modern-day communication relies on ...
+Securing communication channels between different parties has been a long-term
+subject of study for cryptographers and engineers which is essential to our
+modern world to cope with ever-increasing amounts of devices producing and
+sharing data. The main way to facilitate high-throughput, confidential
+communications nowadays is through the use of symmetric cryptography in which
+two parties share a common secret, called a key, which allows them to encrypt,
+share and subsequently decrypt messages to achieve confidentiality against
+third parties. Ciphers can be divided into two categories; block ciphers, which
+always encrypt fixed-sized messages called blocks, and stream ciphers, which
+continuously provide encryption for an arbitrarily long, constant stream of
+data.
+
+A block cipher can be defined as a bijection between the input block (the
+message) and the output block (the ciphertext). For any block cipher with block
+size $n$, we denote the key-dependent encryption and decryption functions as
+$E_K,D_K:\mathbb{F}_2^n\rightarrow \mathbb{F}_2^n$. The simplest way to
+characterize this bijection is through a lookup table which yields the highest
+possible performance as each block can be encrypted by one simple lookup
+depending on the key and the message. This is not practical though due to most
+ciphers working with block and key sizes $n,|K|\geq 64$. For a block cipher
+with $n=64,|K|=128$, a space of $2^{64}2^{128}=2^{192}$ is necessary.
+Considering modern consumer hard disks being able to store data in the order of
+$2^{40}$, it is easy to see that a lookup table is wholly impractical. We
+therefore describe block ciphers algorithmically which opens up possibilities
+for different tradeoffs and security concerns.
 
@@
 
-GIFT is a lightweight block cipher based on ...
+\texttt{GIFT}\cite{gift:2017}, first presented in the \textit{CHES 2017}
+cryptographic hardware and embedded systems conference, is a lightweight block
+cipher based on a previous design called \texttt|PRESENT|, developed in 2007. Its
+goal is to offer maximum security while being extremely light on resources.
+Modern battery-powered devices like RFID tags or low-latency operations like
+on-the-fly disc encryption present strong hardware and power constraints. GIFT
+aims to be a simple, low-energy cipher suited for these kinds of applications.
+
+\texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
+and \texttt{GIFT-128} working with 128-bit blocks. In both cases, the key is 128
+bits long. The design is a very simple, round-based substitution-permutation
+network (SPN). One round consists in a sequential application of the confusion
+layer by means of 4-bit S-boxes and subsequent diffusion through bit
+permutation. After the bit permutation, a round key is added to the cipher
+state and the single round is complete. \texttt{GIFT-64} uses 28 rounds while
+\texttt{GIFT-128} uses 40 rounds.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/GIFT-64.pdf}
+    \caption{Two rounds of GIFT-64}
+    \label{fig:gift64}
+\end{figure}
 
@@
 
+The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
+16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \verb|GIFT-128|. The S-box
+$S:\mathbb{F}_2^4\rightarrow \mathbb{F}_2^4$ is defined as follows:
+
+\[
+    \begin{array}{l|cccccccccccccccc}
+        x & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & a & b & c & d & e & f \\
+        \hline
+        S(x) & 1 & a & 4 & c & 6 & f & 3 & 9 & 2 & d & b & 7 & 5 & 0 & 8 & e
+    \end{array}
+\]
+
 \subsubsection{Permutation layer}
+
+The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
+\forall i\in\{0,1,\dots,n-1\}$. The different permutations for \verb|GIFT-64|
+and \verb|GIFT-128| can be expressed by:
+
+\begin{align*}
+    P_{64}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+16\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
+    P_{128}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+32\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
+\end{align*}
 
@@
 
-\subsubsection{Round key extraction}
+\subsubsection{Round key extraction and key schedule}
 
@@
 
+\bibliography{refs}
+\bibliographystyle{ieeetr}
+
 \end{document}

### 2023-02-15 Paper/paper.tex
add implementation strategy chapter

@@
 \usepackage{systeme} % systems of equation
+\usepackage[utf8]{inputenc} % necessary for proper unicode support
 \usepackage{parcolumns} % multiple column parallel typesetting
@@
 \usepackage{listings} % source code listings
-\usepackage[utf8]{inputenc} % necessary for proper unicode support
+\usepackage{biblatex} % better citation
 % \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry
@@
 
-\lstset{
+\lstset {
     numbers=left,
@@
 \newcommand{\sref}[1]{\textsuperscript{\ref{#1}}}
+\newcommand{\F}[1]{\mathbb{F}_2^{#1}}
+
+\addbibresource{refs.bib}
 
@@
 size $n$, we denote the key-dependent encryption and decryption functions as
-$E_K,D_K:\mathbb{F}_2^n\rightarrow \mathbb{F}_2^n$. The simplest way to
+$E_K,D_K:\F{n}\rightarrow \F{n}$. The simplest way to
 characterize this bijection is through a lookup table which yields the highest
@@
 ciphers working with block and key sizes $n,|K|\geq 64$. For a block cipher
-with $n=64,|K|=128$, a space of $2^{64}2^{128}=2^{192}$ is necessary.
+with $n=64,|K|=128$, a space of $2^{64}2^{128}64=2^{198}$ is necessary.
 Considering modern consumer hard disks being able to store data in the order of
@@
 The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
-16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \verb|GIFT-128|. The S-box
-$S:\mathbb{F}_2^4\rightarrow \mathbb{F}_2^4$ is defined as follows:
+16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \texttt{GIFT-128}. The S-box
+$S:\F{4}\rightarrow \F{4}$ is defined as follows:
 
@@
 The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
-\forall i\in\{0,1,\dots,n-1\}$. The different permutations for \verb|GIFT-64|
-and \verb|GIFT-128| can be expressed by:
+\forall i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64}
+and \texttt{GIFT-128} can be expressed by:
 
@@
 
+The last step of each round consists in XORing a round key $R_i$ to the cipher
+state. The new cipher state $s_{i+1}$ after each full round is therefore given
+by
+
+\[
+    s_{i+1}=P(S(s_i))\oplus R_i
+\]
+
 \subsubsection{Round key extraction and key schedule}
+
+Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
+$K=k7||k6||\dots||k0$ denote the $128$-bit key state.
+
+\paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
+state. $u_i$ and $v_i$ are XORed to $r_{4i+1}$ and $r_{4i}$ of the round key
+$R$ respectively.
+
+\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k4||k1||k_0$ from
+the key state. $u_i$ and $v_i$ are XORed to $r_{4i+2}$ and $b_{4i+1}$ of the
+round key $R$ respectively.
+
+In both cases, we additionally XOR a round constant $C=c_5c_4c_3c_2c_1c_0$ to
+bit positions $n-1,23,19,15,11,7,3$. The round constants are generated using a
+6-bit affine linear-feedback shift register and have the following values:\\
+
+\begin{tabular}{r|l}
+    \textbf{Rounds} & \textbf{Constants} \\
+    \hline
+    \textbf{1 - 16} &  \small\texttt{01,03,07,0F,1F,3E,3D,3B,37,2F,1E,3C,39,33,27,0E} \\
+    \textbf{17 - 32} & \small\texttt{1D,3A,35,2B,16,2C,18,30,21,02,05,0B,17,2E,1C,38} \\
+    \textbf{33 - 48} & \small\texttt{31,23,06,0D,1B,36,2D,1A,34,29,12,24,08,11,22,04}
+\end{tabular}\\
+
+The key state is then updated by setting $k_1\leftarrow k_1\ggg 2$,
+$k_0\leftarrow k_0\ggg 12$ and rotating the new state $32$ bits to the right:
+
+\[
+    k_7||k_6||\dots||k_1||k_0\leftarrow k_1\ggg 2||k_0\ggg 12||k_7||k_6||\dots||k_3||k_2
+\]
 
@@
 
-\chapter{A simple implementation}
+\chapter{Implementation strategies}
 
-\chapter{Optimizations through bitslicing}
+Due to the structural differences of SPN- and Feistel network-based ciphers, we
+shall analyze these two separately.
 
-\chapter{Leveraging NEON advanced SIMD instructions}
+\section{Strategies for SPN}
+
+Three implementation strategies for substitution-permutation networks are
+introduced by \cite{implx86:2014}:
+
+\begin{itemize}
+    \item Table-based implementations
+    \item \texttt{vperm} implementations
+    \item Bitslice implementations
+\end{itemize}
+
+\subsection{Table-based}
+
+Table-driven programming is a simple way to increase performance of operations
+by tabulating the results, therefore requiring only a single memory access to
+acquire the result. This approach is obviously limited to manageable table
+sizes, so while tabulating a function like the AES S-box
+$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space,
+tabulating the \texttt{GIFT} permutation layer
+$P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
+$2^{70}$ space, which is totally unfeasible.
+
+A common approach is to tabulate the output of each S-box, including the
+diffusion layer, and then XORing the results together. Let $n$ denote the
+internal cipher state size and $s$ the size of a single S-box in bits. For each
+S-box $S_i,i\in\{0,\dots,\frac{n}{s}\}$, we can construct a mapping
+$T_i:\F{s}\rightarrow \F{n}$ representing substitution with subsequent
+permutation of that single S-box. The cipher state before round key addition is
+then given by $\bigoplus_{i=0}^{\frac{n}{s}-1}{T_i(m_i)}$ for each $s$-bit
+message chunk $m_i$. This approach requires space of
+$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for \texttt{GIFT-64},
+results in a manageable size of $\frac{64^2 2^4}{4}=2^{14}$ bits which equals
+$16$ KiB.
+
+\subsubsection{Constructing the tables}
+
+For \texttt{GIFT-64}, table construction is relatively straightforward and can
+be done as follows:
+
+\begin{lstlisting}[frame=single, caption={Table construction algorithm}, escapechar=|]
+    tables <- [][]
+    for sbox_index from 0 to 15 do
+        for sbox_input from 0 to 15 do|\label{lst:tablesbox}|
+            output <- sbox(sbox_input)
+            output <- permute(output << (4 * sbox_index))
+            tables[sbox_index][sbox_input] <- output
+\end{lstlisting}
+
+Implementing this algorithm gives us the following table representing the first
+and second S-box.
+
+\[
+    \begin{array}{l|l|l|c}\label{table:gift64subperm}
+        x & T_0(x) & T_1(x) & \dots \\
+        \hline
+        0x0 & 0x1               & 0x1000000000000   & \dots \\
+        0x1 & 0x8000000020000   & 0x800000002       & \dots \\
+        0x2 & 0x400000000       & 0x40000           & \dots \\
+        0x3 & 0x8000400000000   & 0x800040000       & \dots \\
+        0x4 & 0x400020000       & 0x40002           & \dots \\
+        0x5 & 0x8000400020001   & 0x1000800040002   & \dots \\
+        0x6 & 0x20001           & 0x1000000000002   & \dots \\
+        0x7 & 0x8000000000001   & 0x1000800000000   & \dots \\
+        0x8 & 0x20000           & 0x2               & \dots \\
+        0x9 & 0x8000400000001   & 0x1000800040000   & \dots \\
+        0xa & 0x8000000020001   & 0x1000800000002   & \dots \\
+        0xb & 0x400020001       & 0x1000000040002   & \dots \\
+        0xc & 0x400000001       & 0x1000000040000   & \dots \\
+        0xd & 0x0               & 0x0               & \dots \\
+        0xe & 0x8000000000000   & 0x800000000       & \dots \\
+        0xf & 0x8000400020000   & 0x800040002       & \dots
+    \end{array}
+\]
+
+The tables for \texttt{GIFT-128} can be generated in a similar way by looping
+through all 32 S-boxes instead of 16 on line \ref{lst:tablesbox}.
+
+\chapter{Benchmarks}
 
@@
 
-\bibliography{refs}
-\bibliographystyle{ieeetr}
+\printbibliography
 

### 2023-02-15 scripts/table_64.py

@@ -1,0 +1,33 @@
+#!/usr/bin/env python3
+
+sbox = (
+    0x1, 0xa, 0x4, 0xc, 0x6, 0xf, 0x3, 0x9,
+    0x2, 0xd, 0xb, 0x7, 0x5, 0x0, 0x8, 0xe
+)
+
+permute = (
+    0, 17, 34, 51, 48, 1, 18, 35, 32, 49, 2, 19, 16, 33, 50, 3,
+    4, 21, 38, 55, 52, 5, 22, 39, 36, 53, 6, 23, 20, 37, 54, 7,
+    8, 25, 42, 59, 56, 9, 26, 43, 40, 57, 10, 27, 24, 41, 58, 11,
+    12, 29, 46, 63, 60, 13, 30, 47, 44, 61, 14, 31, 28, 45, 62, 15
+)
+
+def perm(x):
+    result = 0
+    for i in range(64):
+        bit = (x >> i) & 0x1
+        result |= bit << permute[i]
+    return result
+
+tables = []
+for sbox_index in range(16):
+    table = []
+    for sbox_input in range(16):
+        output = sbox[sbox_input]
+        output = perm(output << (4 * sbox_index))
+        table.append(output)
+    tables.append(table)
+
+for x in tables[1]:
+    print(hex(x), end=' ')
+print('')

### 2023-02-27 Paper/paper.tex
start work on vperm

@@
 \usepackage{hyperref} % hyperlinks
-\usepackage{hyphenat}
+% \usepackage{hyphenat}
 \usepackage{bookmark}
@@
 
-\chapter{Benchmarks}
+\subsection{Using \texttt{vperm}}
+
+Nowadays, most instructions set architectures support single-instruction,
+multiple-data processing. The idea of such an SIMD system is to work on
+multiple data stored in vectors at once to speed up calculations. For A64, two
+types of vector processing are available:
+
+\begin{enumerate}
+    \item Advanced SIMD, known as NEON
+    \item Scalable Vector Extension (SVE)
+\end{enumerate}
+
+We will take a look at NEON as this is the type of vector processing supported
+by the Cortex A-73 processor.
+
+\subsubsection{ARM Neon}
+
+The register file of the NEON unit is made up of 32 quad-word (128-bit)
+registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
+registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
+sized lanes on which the vector instructions operate. Valid ways to interpret
+for example the register \texttt{V0} are:
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/V_register.pdf}
+    \caption{Divisions of the V register}
+    \label{fig:gift64}
+\end{figure}
+
+NEON instructions interpret their operands' layouts (i.e. lane count and width)
+through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For instance,
+adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
+together and storing the result in \texttt{V0} can be done as follows:
+
+\begin{center}
+    \texttt{ADD V0.8H, V1.8H, V2.8H}
+\end{center}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/vector_add.pdf}
+    \caption{Addition of two vector registers}
+    \label{fig:gift64}
+\end{figure}
+
+The plenitude of different processing instructions allow flexible ways to
+further speed up algorithms having reached their optimizational limit on
+non-SIMD platforms. \texttt{vperm}, a general term standing for \textit{vector
+permute}, is a common instruction on SIMD machines. Called \texttt{TBL} on
+NEON, it is used for parallel table lookups and arbitrary permutations. It takes
+two inputs and performs a lanewise lookup:
+
+\begin{enumerate}
+    \item A register with lookup values
+    \item Two or more registers containing data
+\end{enumerate}
+
+\subsubsection{S-box lookup}
+
+This instruction can be used to implement S-box lookup of all 16 S-boxes in a
+single instruction. We do this by packing our 64-bit cipher state
+$s=s_{15}||s_{14}||\dots||s_0$ into a vector register $V_0$. Because we can
+only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane. We
+then put the S-box itself into register $V_1$ which will be used as the data
+register for the table lookup.
+
+\subsection{Bitslicing}
+
+\chapter{Implementation}
+
+\chapter{Evaluation}
+
+\section{Limitations}
+
+\section{Benchmarks}
 

### 2023-03-03 Paper/paper.tex
start work on bitslicing

@@
     \caption{Two rounds of GIFT-64}
-    \label{fig:gift64}
 \end{figure}
@@
 \[
-    \begin{array}{l|l|l|c}\label{table:gift64subperm}
+    \begin{array}{l|l|l|c}
         x & T_0(x) & T_1(x) & \dots \\
@@
     \includegraphics[width=\textwidth]{Figures/V_register.pdf}
-    \caption{Divisions of the V register}
-    \label{fig:gift64}
+    \caption{Divisions of the V0 register}
 \end{figure}
@@
 NEON instructions interpret their operands' layouts (i.e. lane count and width)
-through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For instance,
+through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
 adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
@@
     \caption{Addition of two vector registers}
-    \label{fig:gift64}
 \end{figure}
@@
 NEON, it is used for parallel table lookups and arbitrary permutations. It takes
-two inputs and performs a lanewise lookup:
+two inputs to perform a lanewise lookup:
 
@@
 $s=s_{15}||s_{14}||\dots||s_0$ into a vector register $V_0$. Because we can
-only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane. We
-then put the S-box itself into register $V_1$ which will be used as the data
-register for the table lookup.
+only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane which
+neatly fits into the 128-bit registers. We then put the S-box itself into
+register $V_1$ which will be used as the data register for the table lookup.
+
+The confusion layer can now be performed through one \texttt{TBL} instruction:
+
+\begin{center}
+    \texttt{TBL V0.16B, V1.16B, V0.16B}
+\end{center}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/tbl_example.pdf}
+    \caption{Performing the S-Box lookup in parallel}
+\end{figure}
 
 \subsection{Bitslicing}
+
+Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
+achieve a more efficient representation to operate on. The structure of \texttt{GIFT}
+naturally offers possibilities for bitslicing. We split the cipher state bits $b_{63}b_{62}\dots b_0$
+into four slices $S_i,0\leq i\leq 3$:
+
+\[
+    S=\begin{bmatrix}
+        S_0\\
+        S_1\\
+        S_2\\
+        S_3
+    \end{bmatrix}
+    =\begin{bmatrix}
+        b_{60}b_{56}b_{52}\dots b_0\\
+        b_{61}b_{57}b_{53}\dots b_1\\
+        b_{62}b_{58}b_{54}\dots b_2\\
+        b_{63}b_{59}b_{55}\dots b_3
+    \end{bmatrix}
+\]
 

### 2023-03-05 Paper/paper.tex
bitslicing perm and sub

@@
 The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
-\forall i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64}
-and \texttt{GIFT-128} can be expressed by:
+i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64} and
+\texttt{GIFT-128} can be expressed by:
 
@@
 We will take a look at NEON as this is the type of vector processing supported
-by the Cortex A-73 processor.
+by the Cortex-A73 processor.
 
@@
 Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
-achieve a more efficient representation to operate on. The structure of \texttt{GIFT}
-naturally offers possibilities for bitslicing. We split the cipher state bits $b_{63}b_{62}\dots b_0$
-into four slices $S_i,0\leq i\leq 3$:
+achieve a more efficient representation to operate on. The structure of
+\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
+cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
+i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
+individual S-boxes.
 
@@
 
+This representation offers multiple advantages. We first note that computation
+of the S-box can be executed in parallel, similar to the \texttt{vperm}
+technique above. This can be done by finding an algorithmic way to apply the
+S-box which has already been proposed by the original \texttt{GIFT} authors:
+
+\begin{align*}
+    S_1&\leftarrow S_1\oplus (S_0\land S_2) \\
+    t&\leftarrow S_0\oplus (S_1\land S_3) \\
+    S_2&\leftarrow S_2\oplus (t\lor S_1) \\
+    S_0&\leftarrow S_3\oplus S_2 \\
+    S_1&\leftarrow S_1\oplus S_0 \\
+    S_0&\leftarrow \lnot S_0 \\
+    S_2&\leftarrow S_2\oplus (t\land S_1) \\
+    S_3&\leftarrow t
+\end{align*}
+
+This is very efficient as it only requires six XOR-, three AND and one OR
+operation.
+
+An important property of the permutation is the fact that bits always stay in
+their slice. This means we can decompose the permutation $P$ into four
+permutations $P_i,i\in\{0,1,2,3\}$ and apply these permutations
+separately to each slice. One possible way to implement a permutation $P_i$ in
+software is to mask off all bits individually, shift them to their correct
+position and OR them together:
+
+\[
+    P_i(S_i)=\bigvee_{k=0}^{15}{(S_i\land m_i) \ll s_i}
+\]
+
+This approach requires $47$ operations, meaning all four permutations require
+over $150$ operations which would present a major bottleneck to the round
+function. We can improve on this by working on multiple message blocks at once
+and using the aforementioned \texttt{vperm} instruction to implement the bit
+shuffling. We then need only four \texttt{TBL} instructions for the complete
+diffusion layer.
+
+\subsubsection{Using \texttt{vperm} for slice permutation}
+
+We cannot use the \texttt{TBL} instruction directly as we need to shuffle
+individual bits, but the smallest data we can operate on are bytes. We
+therefore encrypt $8n$ messages at once which allows us to create bytewise
+groupings. These messages are put into $4m$ registers with register $R_{4i}$
+containing $S_0$, register $R_{4i+1}$ containing $S_1$ and so forth. With block
+size $BS$ and register size $RS$, the following must hold:
+
+\[
+    8n\cdot BS=4m\cdot RS
+\]
+
+In the case of \texttt{GIFT-64} with $BS=64$ and ARM NEON with $RS=128$, we get
+
+\[
+    8n\cdot 64=4m\cdot 128\Leftrightarrow n=m
+\]
+
+We choose $n=m=1$, so for each encryption, we divide eight messages into four
+registers.
+
+\subsubsection{Packing the data into bitslice format}
+
+Let $a,b,c,d,e,f,g,h$ be eight messages with subscripts denoting individual
+bits. We first put these messages into four SIMD registers $V_0,V_1,V_2,V_3$:
+
+\begin{align*}
+    V_0&=b||a \\
+    V_1&=d||c \\
+    V_2&=f||e \\
+    V_3&=h||g
+\end{align*}
+
+We then use the \texttt{SWAPMOVE} technique to bring the data into bitslice
+format. This operation operates on two registers $A,B$ using mask $M$ and shift
+value $N$. It swaps bits in $A$ masked by $M$ with bits masked by $M\ll N$ in
+B using only three XOR-, one AND- and two shift operations.
+
+\begin{align*}
+    &\text{SWAPMOVE}(A,B,M,N): \\
+    &\qquad T=((A\gg N)\oplus B)\land M \\
+    &\qquad B=B\oplus T \\
+    &\qquad A=A\oplus (T\ll N) \\
+\end{align*}
+
+The following operations put all of the first bits of each S-Box into $V_0$,
+each second bit into $V_1$ and so forth and regroups these bits into bytes.
+
+\begin{align*}
+    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_0,V_0,0xf0f0\dots 0f,4) \\
+    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_1,V_1,0xf0f0\dots 0f,4) \\
+    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_2,V_2,0xf0f0\dots 0f,4) \\
+    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_3,V_3,0xf0f0\dots 0f,4)
+\end{align*}
+
+With numbers denoting byte groups (i.e. $3$ corresponding to
+$h_3g_3f_3e_3d_3c_3b_3a_3$), our data now has the following
+permutation-friendly format:
+
+\begin{alignat*}{7}
+    V_0&=&&124&&||\dots&&||12&&||8 &&||4&&||0 \\
+    V_1&=&&125&&||\dots&&||13&&||9 &&||5&&||1 \\
+    V_2&=&&126&&||\dots&&||14&&||10&&||6&&||2 \\
+    V_3&=&&127&&||\dots&&||15&&||11&&||7&&||3 \\
+\end{alignat*}
+
+We can now create permutation tables using the specification of the individual
+slice permutations $P_i$:
+
+\[
+    \begin{array}{c|llllllllllllllll}
+        j & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\
+        \hline
+        P_0(j) & 0 & 12 & 8 & 4 & 1 & 13 & 9 & 5 & 2 & 14 & 10 & 6 & 3 & 15 & 11 & 7 \\
+        P_1(j) & 4 & 0 & 12 & 8 & 5 & 1 & 13 & 9 & 6 & 2 & 14 & 10 & 7 & 3 & 15 & 11 \\
+        P_2(j) & 8 & 4 & 0 & 12 & 9 & 5 & 1 & 13 & 10 & 6 & 2 & 14 & 11 & 7 & 3 & 15 \\
+        P_3(j) & 12 & 8 & 4 & 0 & 13 & 9 & 5 & 1 & 14 & 10 & 6 & 2 & 15 & 11 & 7 & 3
+    \end{array}
+\]
+
+One thing to take note of is the original permutation values only showing where
+a given byte should land, not which byte belongs to a certain position - i.e.
+for $P_0$, byte 1 should land in position 12, but the byte belonging to
+position 1 is byte 4. Because \texttt{vperm} works in the latter way, we have
+to do some trivial rearrangements. Putting these values into registers
+$V_5,V_6,V_7,V_8$ now allows us to compute the permutation layer in only four
+instructions:
+
+\begin{align*}
+    &\texttt{TBL V0, V0, V5} &\texttt{TBL V1, V1, V6} \\
+    &\texttt{TBL V2, V2, V7} &\texttt{TBL V3, V3, V8}
+\end{align*}
+
 \chapter{Implementation}

### 2023-03-05 impl/benchmark.c

@@ -99,6 +99,7 @@ void benchmark_gift_64_vec_sbox(void)
 {
         uint64_t key[2];
         key_rand(key);
+        gift_64_vec_sbox_init();
 
         uint8x16_t m;
         m = vsetq_lane_u32(rand(), m, 0);

### 2023-03-05 impl/test.c

@@ -152,6 +152,8 @@ void test_gift_64_table(void)
 
 void test_gift_64_vec_sbox(void)
 {
+        gift_64_vec_sbox_init();
+
         // test encrypt to known value
         printf("testing GIFT_64_VEC_SBOX encrytion to known value...\n");
         uint64_t key[2] = { 0x5085772fe6916616UL, 0x3c9d8c18fdd20608UL };

### 2023-03-05 impl/vector/gift_vec_sbox.c

@@ -7,6 +7,28 @@
 static uint8x16_t sbox_vec;
 static uint8x16_t sbox_vec_inv;
 
+#define U64_TO_V128(V,M)\
+        V = vsetq_lane_u64(\
+        (uint64_t)((M >> 4 * 0) & 0xf) << 8 * 0 |\
+        (uint64_t)((M >> 4 * 1) & 0xf) << 8 * 1 |\
+        (uint64_t)((M >> 4 * 2) & 0xf) << 8 * 2 |\
+        (uint64_t)((M >> 4 * 3) & 0xf) << 8 * 3 |\
+        (uint64_t)((M >> 4 * 4) & 0xf) << 8 * 4 |\
+        (uint64_t)((M >> 4 * 5) & 0xf) << 8 * 5 |\
+        (uint64_t)((M >> 4 * 6) & 0xf) << 8 * 6 |\
+        (uint64_t)((M >> 4 * 7) & 0xf) << 8 * 7,\
+        V, 0);\
+        V = vsetq_lane_u64(\
+        (uint64_t)((M >> 4 * 8)  & 0xf) << 8 * 0 |\
+        (uint64_t)((M >> 4 * 9)  & 0xf) << 8 * 1 |\
+        (uint64_t)((M >> 4 * 10) & 0xf) << 8 * 2 |\
+        (uint64_t)((M >> 4 * 11) & 0xf) << 8 * 3 |\
+        (uint64_t)((M >> 4 * 12) & 0xf) << 8 * 4 |\
+        (uint64_t)((M >> 4 * 13) & 0xf) << 8 * 5 |\
+        (uint64_t)((M >> 4 * 14) & 0xf) << 8 * 6 |\
+        (uint64_t)((M >> 4 * 15) & 0xf) << 8 * 7,\
+        V, 1);
+
 static const size_t perm_64[] = {
         0, 17, 34, 51, 48, 1, 18, 35, 32, 49, 2, 19, 16, 33, 50, 3,
         4, 21, 38, 55, 52, 5, 22, 39, 36, 53, 6, 23, 20, 37, 54, 7,
@@ -63,24 +85,8 @@ uint8x16_t gift_64_vec_sbox_permute(const uint8x16_t cipher_state)
                 }
         }
 
-        // load into vector register (yes, this is slow. too bad!)
         uint8x16_t ret;
[…]

### 2023-03-05 impl/vector/gift_vec_sbox.h

@@ -9,11 +9,13 @@
 // expose for benchmarking
 uint8x16_t gift_64_vec_sbox_subcells(const uint8x16_t cipher_state);
 uint8x16_t gift_64_vec_sbox_subcells_inv(const uint8x16_t cipher_state);
-// can't really use vector registers here as
 uint8x16_t gift_64_vec_sbox_permute(const uint8x16_t cipher_state);
 uint8x16_t gift_64_vec_sbox_permute_inv(const uint8x16_t cipher_state);
 void       gift_64_vec_sbox_generate_round_keys(uint8x16_t round_keys[ROUNDS_GIFT_64],
                                                 const uint64_t key[2]);
+
+// construct tables
+void gift_64_vec_sbox_init(void);
 
 uint64_t gift_64_vec_sbox_encrypt(const uint64_t m, const uint64_t key[2]);
 uint64_t gift_64_vec_sbox_decrypt(const uint64_t c, const uint64_t key[2]);

### 2023-03-05 impl/vector/gift_vec_sliced.c

@@ -1,0 +1,133 @@
+#include "gift_vec_sliced.h"
+
+static uint8x16_t perm_s0;
+static uint8x16_t perm_s1;
+static uint8x16_t perm_s2;
+static uint8x16_t perm_s3;
+
+static uint8x16_t perm_inv_s0;
+static uint8x16_t perm_inv_s1;
+static uint8x16_t perm_inv_s2;
+static uint8x16_t perm_inv_s3;
+
+static uint8x16_t pack_mask_0;
+static uint8x16_t pack_mask_1;
+static uint8x16_t pack_mask_2;
+
+void swapmove(uint8x16_t *a, uint8x16_t *b, uint8x16_t m, int n)
+{
+
+}
+
+void bits_pack(uint8x16x4_t m)
+{
+        // take care not to shift mask bits out of the register
+        swapmove(m.val[0], m.val[1], 0x5555555555555555UL, 1);
+        swapmove(m.val[2], m.val[3], 0x5555555555555555UL, 1);
+        swapmove(m.val[4], m.val[5], 0x5555555555555555UL, 1);
+        swapmove(m.val[6], m.val[7], 0x5555555555555555UL, 1);
+
+        swapmove(m.val[0], m.val[2], 0x3333333333333333UL, 2);
+        swapmove(m.val[1], m.val[3], 0x3333333333333333UL, 2);
+        swapmove(m.val[4], m.val[6], 0x3333333333333333UL, 2);
+        swapmove(m.val[5], m.val[7], 0x3333333333333333UL, 2);
+
+        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
+        swapmove(m.val[0], m.val[4], 0x0f0f0f0f0f0f0f0fUL, 4);
+        swapmove(m.val[2], m.val[6], 0x0f0f0f0f0f0f0f0fUL, 4);
+        swapmove(m.val[1], m.val[5], 0x0f0f0f0f0f0f0f0fUL, 4);
+        swapmove(m.val[3], m.val[7], 0x0f0f0f0f0f0f0f0fUL, 4);
+}
+
+void gift_64_vec_sliced_subcells(uint8x16x4_t cs)
+{
+        cs.val[1]       = veorq_u8(cs.val[1], vandq_u8(cs.val[0], cs.val[2]));
[…]

### 2023-03-05 impl/vector/gift_vec_sliced.h

@@ -1,0 +1,25 @@
+#ifndef GIFT_VEC_SLICED_H
+#define GIFT_VEC_SLICED_H
+
+#include <stdint.h>
+#include <arm_neon.h>
+
+#define ROUNDS_GIFT_64 28
+
+// expose for benchmarking
+void swapmove(uint8x16_t *a, uint8x16_t *b, uint8x16_t m, int n);
+void bits_pack(uint8x16x4_t m);
+
+void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state);
+void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state);
+void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state);
+void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state);
+void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[ROUNDS_GIFT_64],
+                                            const uint64_t key[2]);
+
+void gift_64_vec_sliced_init(void);
+
+void gift_64_vec_sbox_encrypt(uint64_t c[8], const uint64_t m[8], const uint64_t key[2]);
+void gift_64_vec_sbox_decrypt(uint64_t m[8], const uint64_t c[8], const uint64_t key[2]);
+
+#endif

### 2023-03-05 scripts/perm_64_sliced_better.py

@@ -1,0 +1,54 @@
+#!/usr/bin/env python3
+
+p0 = (0, 12, 8, 4, 1, 13, 9, 5, 2, 14, 10, 6, 3, 15, 11, 7)
+p1 = (4, 0, 12, 8, 5, 1, 13, 9, 6, 2, 14, 10, 7, 3, 15, 11)
+p2 = (8, 4, 0, 12, 9, 5, 1, 13, 10, 6, 2, 14, 11, 7, 3, 15)
+p3 = (12, 8, 4, 0, 13, 9, 5, 1, 14, 10, 6, 2, 15, 11, 7, 3)
+
+for i in range(16):
+    for j, p0j in enumerate(p0):
+        if p0j == i:
+            print(j, end=', ')
+
+print('')
+
+for i in range(16):
+    for j, p1j in enumerate(p1):
+        if p1j == i:
+            print(j, end=', ')
+
+print('')
+
+for i in range(16):
+    for j, p2j in enumerate(p2):
+        if p2j == i:
+            print(j, end=', ')
+
+print('')
+
+for i in range(16):
+    for j, p3j in enumerate(p3):
+        if p3j == i:
+            print(j, end=', ')
+
+print('\n0x', end='')
+
+for i in range(16):
+    print(f"{p0[15 - i]:02x}", end='')
+
+print('\n0x', end='')
+
+for i in range(16):
+    print(f"{p1[15 - i]:02x}", end='')
+
+print('\n0x', end='')
+
+for i in range(16):
+    print(f"{p2[15 - i]:02x}", end='')
+
+print('\n0x', end='')
+
+for i in range(16):
+    print(f"{p3[15 - i]:02x}", end='')
+
+print('')

### 2023-03-06 Paper/paper.tex
shl/shr for vec registers

@@
 i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
-individual S-boxes.
+individual S-boxes. This is equivalent to transposing the bit matrix.
 
@@
 
-Let $a,b,c,d,e,f,g,h$ be eight messages with subscripts denoting individual
-bits. We first put these messages into four SIMD registers $V_0,V_1,V_2,V_3$:
+Let $a,b,c,d,e,f,g,h$ be eight messages of length $64$ with subscripts denoting
+individual bits. We first put these messages into four SIMD registers
+$V_0,V_1,V_2,V_3$:
 
@@
 
+We will provide implementations for the presented strategies in the C
+programming language. Although directly writing Assembler code could result in
+a small performance benefit, this generally increases the work necessary by an
+order of magnitude for only limited results.
+
 \chapter{Evaluation}
-
-\section{Limitations}
-
-\section{Benchmarks}
 

### 2023-03-06 impl/vector/gift_vec_sliced.c

@@ -14,29 +14,52 @@ static uint8x16_t pack_mask_0;
 static uint8x16_t pack_mask_1;
 static uint8x16_t pack_mask_2;
 
-void swapmove(uint8x16_t *a, uint8x16_t *b, uint8x16_t m, int n)
+uint8x16_t shl(uint8x16_t v, int n)
 {
-
+        uint64_t l0 = vgetq_lane_u64(v, 0);
+        uint64_t l1 = vgetq_lane_u64(v, 1);
+        l1 = l1 << n | (l0 >> (64 - n));
+        l0 <<= n;
+        uint8x16_t ret;
+        ret = vsetq_lane_u64(l0, ret, 0);
+        ret = vsetq_lane_u64(l1, ret, 1);
+        return ret;
 }
 
-void bits_pack(uint8x16x4_t m)
+uint8x16_t shr(uint8x16_t v, int n)
+{
+        uint64_t l0 = vgetq_lane_u64(v, 0);
+        uint64_t l1 = vgetq_lane_u64(v, 1);
+        l0 = l0 >> n | (((l1 << (64 - n)) >> (64 - n)) << (64 - n));
+        l1 >>= n;
+        uint8x16_t ret;
+        ret = vsetq_lane_u64(l0, ret, 0);
+        ret = vsetq_lane_u64(l1, ret, 1);
+        return ret;
+}
+
+void gift_64_vec_sliced_swapmove(uint8x16_t *a, uint8x16_t *b, uint8x16_t m, int n)
+{
+
+        uint8x16_t t = vandq_u8(veorq_u8(shl(*a, n), *b), m);
+        *b = veorq_u8(*b, t);
+        *a = veorq_u8(*a, shl(t, n));
+}
+
+void gift_64_vec_sliced_bits_pack(uint8x16x4_t m)
 {
         // take care not to shift mask bits out of the register
-        swapmove(m.val[0], m.val[1], 0x5555555555555555UL, 1);
-        swapmove(m.val[2], m.val[3], 0x5555555555555555UL, 1);
-        swapmove(m.val[4], m.val[5], 0x5555555555555555UL, 1);
-        swapmove(m.val[6], m.val[7], 0x5555555555555555UL, 1);
+        gift_64_vec_sliced_swapmove(&m.val[0], &m.val[1], pack_mask_0, 1);
[…]

### 2023-03-06 impl/vector/gift_vec_sliced.h

@@ -7,8 +7,10 @@
 #define ROUNDS_GIFT_64 28
 
 // expose for benchmarking
-void swapmove(uint8x16_t *a, uint8x16_t *b, uint8x16_t m, int n);
-void bits_pack(uint8x16x4_t m);
+uint8x16_t shl(uint8x16_t v, int n);
+uint8x16_t shr(uint8x16_t v, int n);
+void gift_64_vec_sliced_swapmove(uint8x16_t *a, uint8x16_t *b, uint8x16_t m, int n);
+void gift_64_vec_sliced_bits_pack(uint8x16x4_t m);
 
 void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state);
 void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state);
@@ -19,7 +21,7 @@ void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[ROUNDS_GIFT_
 
 void gift_64_vec_sliced_init(void);
 
-void gift_64_vec_sbox_encrypt(uint64_t c[8], const uint64_t m[8], const uint64_t key[2]);
-void gift_64_vec_sbox_decrypt(uint64_t m[8], const uint64_t c[8], const uint64_t key[2]);
+void gift_64_vec_sliced_encrypt(uint64_t c[8], const uint64_t m[8], const uint64_t key[2]);
+void gift_64_vec_sliced_decrypt(uint64_t m[8], const uint64_t c[8], const uint64_t key[2]);
 
 #endif

### 2023-03-13 Paper/paper.tex
encryption working

@@
 
-We choose $n=m=1$, so for each encryption, we divide eight messages into four
-registers.
+$n=m=1$ would be a valid choice which yields eight messages divided into four
+registers. We choose $n=m=2$ so we can directly utilize the algorithm for bit
+packing presented by the original GIFT authors, although it is simple to adapt
+this algorithm to only four registers and eight messages by dividing each
+128-bit NEON register into two separate 64-bit registers and adjusting the
+\texttt{SWAPMOVE} shift and mask values.
 
@@
 
-Let $a,b,c,d,e,f,g,h$ be eight messages of length $64$ with subscripts denoting
-individual bits. We first put these messages into four SIMD registers
-$V_0,V_1,V_2,V_3$:
+Let $a,b,\dots,p$ be sixteen messages of length $64$ with subscripts denoting
+individual bits. We first put these messages into eight SIMD registers
+$V_0,V_1,\dots,V_7$:
 
-\begin{align*}
-    V_0&=b||a \\
-    V_1&=d||c \\
-    V_2&=f||e \\
-    V_3&=h||g
-\end{align*}
+\begin{alignat*}{3}
+    V_0&=b||a\qquad V_4&&=j||&&i \\
+    V_1&=d||c\qquad V_5&&=l||&&k \\
+    V_2&=f||e\qquad V_6&&=n||&&m \\
+    V_3&=h||g\qquad V_7&&=p||&&o
+\end{alignat*}
 
@@
 format. This operation operates on two registers $A,B$ using mask $M$ and shift
-value $N$. It swaps bits in $A$ masked by $M$ with bits masked by $M\ll N$ in
-B using only three XOR-, one AND- and two shift operations.
+value $N$. It swaps bits in $A$ masked by $(M\ll N)$ with bits in $B$ masked by
+$M$ in using only three XOR-, one AND- and two shift operations.
 
@@
 
-The following operations put all of the first bits of each S-Box into $V_0$,
-each second bit into $V_1$ and so forth and regroups these bits into bytes.
+The following operations group all $i$th bits of the messages $a,c,\dots,o$
+into bytes and puts these into the lower half of the registers $V_{i\bmod 8}$.
+The same is done for messages $b,d,\dots,p$, only differing in that the bytes
+are put into the upper half of the registers.
 
 \begin{align*}
-    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_0,V_0,0xf0f0\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_1,V_1,0xf0f0\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_2,V_2,0xf0f0\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_3,V_3,0xf0f0\dots 0f,4)
+    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_4,V_5,0x5555\dots 55,1) \\
+    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_6,V_7,0x5555\dots 55,1) \\
+    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_4,V_6,0x3333\dots 33,2) \\
+    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_5,V_7,0x3333\dots 33,2) \\
+    &\text{SWAPMOVE}(V_0,V_4,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_1,V_5,0x0f0f\dots 0f,4) \\
+    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4) \\
 \end{align*}
 
-With numbers denoting byte groups (i.e. $3$ corresponding to
-$h_3g_3f_3e_3d_3c_3b_3a_3$), our data now has the following
-permutation-friendly format:
+With $Ax=o_xm_xk_xj_xg_xe_xc_xa_x$ and $Bx=p_xn_xl_xi_xh_xf_xd_xb_x$ denoting
+byte groups, our data now has the following permutation-friendly format:
 
-\begin{alignat*}{7}
-    V_0&=&&124&&||\dots&&||12&&||8 &&||4&&||0 \\
-    V_1&=&&125&&||\dots&&||13&&||9 &&||5&&||1 \\
-    V_2&=&&126&&||\dots&&||14&&||10&&||6&&||2 \\
-    V_3&=&&127&&||\dots&&||15&&||11&&||7&&||3 \\
-\end{alignat*}
+\[
+    \begin{array}{c|llllllll}
+        n & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
+        \hline
+        V_0 & A56 & A48 & A40 & A32 & A24 & A16 & A8 & A0 \\
+        V_1 & A57 & A49 & A41 & A33 & A25 & A17 & A9 & A1 \\
+        V_2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
+        V_3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
+        V_4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
+        V_5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
+        V_6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
+        V_7 & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
+    \end{array}
+\]
+
+\[
+    \begin{array}{c|llllllll}
+        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 \\
+        \hline
+        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8 & B0 \\
+        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9 & B1 \\
+        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 \\
+        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 \\
+        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 \\
+        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 \\
+        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 \\
+        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7
+    \end{array}
+\]
 
@@
 position 1 is byte 4. Because \texttt{vperm} works in the latter way, we have
-to do some trivial rearrangements. Putting these values into registers
-$V_5,V_6,V_7,V_8$ now allows us to compute the permutation layer in only four
-instructions:
+to do some trivial rearrangements. In addition, because data for each slice is
+split between two registers $(V_0,V_4),(V_1,V_5),\dots$, the \texttt{vperm}
+operation needs to use these two registers as a data source. Using two data
+sources is supported by the \texttt{TBL} instruction by means of an additional
+parameter.
 
-\begin{align*}
-    &\texttt{TBL V0, V0, V5} &\texttt{TBL V1, V1, V6} \\
-    &\texttt{TBL V2, V2, V7} &\texttt{TBL V3, V3, V8}
-\end{align*}
+Assuming the correct permutation values are put into registers
+$V_9,V_{10},\dots,V_{16}$, this now allows us to compute the permutation layer for
+all 16 blocks in only eight permutation instructions plus four additional vector
+copy instructions used for saving the original values of $V_0,V_1,V_2,V_3$
+before the permutation is applied.
+
+\begin{alignat*}{2}
+    &\texttt{MOV V17, V0}\qquad &&\texttt{MOV V18, V1} \\
+    &\texttt{MOV V19, V2}\qquad &&\texttt{MOV V20, V3} \\
+    &\texttt{TBL V0, \{V0, V4\}, V9}\qquad &&\texttt{TBL V1, \{V1, V5\}, V10} \\
+    &\texttt{TBL V2, \{V2, V6\}, V9}\qquad &&\texttt{TBL V3, \{V3, V7\}, V10} \\
+    &\texttt{TBL V4, \{V17, V4\}, V9}\qquad &&\texttt{TBL V5, \{V18, V5\}, V10} \\
+    &\texttt{TBL V6, \{V19, V6\}, V9}\qquad &&\texttt{TBL V7, \{V20, V7\}, V10}
+\end{alignat*}
 

### 2023-03-13 impl/test.c

@@ -184,11 +184,19 @@ void test_gift_64_vec_sliced(void)
         // test encrypt to known value
         printf("testing GIFT_64_VEC_SLICED encrytion to known value...\n");
         uint64_t key[2] = { 0x5085772fe6916616UL, 0x3c9d8c18fdd20608UL };
-        uint64_t m[8] = { 0xffffffffffffffffUL };
+        uint64_t m[16] = {
+                0x4dcfd3bdd61810f0UL, 0x4dcfd3bdd61810f0UL,
+                0x4dcfd3bdd61810f0UL, 0x4dcfd3bdd61810f0UL,
+                0x4dcfd3bdd61810f0UL, 0x4dcfd3bdd61810f0UL,
+                0x4dcfd3bdd61810f0UL, 0x4dcfd3bdd61810f0UL,
+                0x4dcfd3bdd61810f0UL, 0x4dcfd3bdd61810f0UL,
+                0x4dcfd3bdd61810f0UL, 0x4dcfd3bdd61810f0UL,
+                0x4dcfd3bdd61810f0UL, 0x4dcfd3bdd61810f0UL,
+                0x4dcfd3bdd61810f0UL, 0x4dcfd3bdd61810f0UL,
+        };
         uint64_t c_expected = 0xb11d30b8d39763e1UL;
-        uint64_t c[8];
+        uint64_t c[16];
         gift_64_vec_sliced_encrypt(c, m, key);
-        printf("%lx\n", c[0]);
         ASSERT_EQUALS(c[0], c_expected);
         ASSERT_EQUALS(c[1], c_expected);
 
@@ -196,13 +204,13 @@ void test_gift_64_vec_sliced(void)
         printf("testing GIFT_64_VEC_SLICED encrypt-decrypt...\n");
         for (int i = 0; i < 100; i++) {
                 key_rand(key);
-                m_rand(m, 8 * 8);
+                m_rand((uint8_t*)m, sizeof(m));
 
                 gift_64_vec_sliced_encrypt(c, m, key);
                 // only encryption for table approach
-                uint64_t m_actual[8];
+                uint64_t m_actual[16];
                 gift_64_vec_sliced_decrypt(m_actual, c, key);
-                ASSERT_TRUE(memcmp(m, m_actual, 8 * 8) == 0);
+                ASSERT_TRUE(memcmp(m, m_actual, sizeof(m_actual)) == 0);
         }
 }
 

### 2023-03-13 impl/vector/gift_vec_sliced.c

@@ -1,16 +1,33 @@
 #include "gift_vec_sliced.h"
 
 #include <stddef.h>
+#include <stdio.h>
+#include <string.h>
 
-static uint8x16_t perm_s0;
-static uint8x16_t perm_s1;
-static uint8x16_t perm_s2;
-static uint8x16_t perm_s3;
+static uint64_t perm_u64[16] = {
+        0x1410050115110400UL, 0x1410050115110400UL,
+        0x0501151104001410UL, 0x0501151104001410UL,
+        0x1511040014100501UL, 0x1511040014100501UL,
+        0x0400141005011511UL, 0x0400141005011511UL,
+        0x1612070317130602UL, 0x1612070317130602UL,
+        0x0703171306021612UL, 0x0703171306021612UL,
+        0x1713060216120703UL, 0x1713060216120703UL,
+        0x0602161207031713UL, 0x0602161207031713UL
+};
 
-static uint8x16_t perm_inv_s0;
-static uint8x16_t perm_inv_s1;
-static uint8x16_t perm_inv_s2;
-static uint8x16_t perm_inv_s3;
+static uint64_t perm_inv_u64[16] = {
+        0x1511050114100400UL, 0x1511050114100400UL,
+        0x1713070316120602UL, 0x1713070316120602UL,
+        0x1115010510140004UL, 0x1115010510140004UL,
+        0x1317030712160206UL, 0x1317030712160206UL,
+        0x1317030712160206UL, 0x1317030712160206UL,
+        0x1511050114100400UL, 0x1511050114100400UL,
+        0x1713070316120602UL, 0x1713070316120602UL,
+        0x1115010510140004UL, 0x1115010510140004UL,
+};
+
+static uint8x16x4_t perm[2];
+static uint8x16x4_t perm_inv[2];
 
 static uint8x16_t pack_mask_0;
 static uint8x16_t pack_mask_1;
@@ -50,73 +67,118 @@ uint8x16_t shr(uint8x16_t v, int n)
         return ret;
 }
 
-void gift_64_vec_sliced_swapmove(uint8x16_t *a, uint8x16_t *b, uint8x16_t m, int n)
[…]

### 2023-03-13 impl/vector/gift_vec_sliced.h

@@ -9,19 +9,23 @@
 // expose for benchmarking
 uint8x16_t shl(uint8x16_t v, int n);
 uint8x16_t shr(uint8x16_t v, int n);
-void gift_64_vec_sliced_swapmove(uint8x16_t *a, uint8x16_t *b, uint8x16_t m, int n);
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t *m);
+void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n);
+void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
 
-void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state);
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state);
-void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state);
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state);
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[ROUNDS_GIFT_64],
-                                            const uint64_t key[2]);
+void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
+void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
+void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
+void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
+void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
+                                            const uint64_t key[restrict 2]);
 
 void gift_64_vec_sliced_init(void);
 
-void gift_64_vec_sliced_encrypt(uint64_t c[8], const uint64_t m[8], const uint64_t key[2]);
-void gift_64_vec_sliced_decrypt(uint64_t m[8], const uint64_t c[8], const uint64_t key[2]);
+void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
+                                const uint64_t m[restrict 16],
+                                const uint64_t key[restrict 2]);
+void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
+                                const uint64_t c[restrict 16],
+                                const uint64_t key[restrict 2]);
 
 #endif

### 2023-03-13 scripts/perm_128_sliced.py

@@ -1,0 +1,47 @@
+#!/usr/bin/env python3
+
+p = ((0, 4, 8, 12, 3, 7, 11, 15, 2, 6, 10, 14, 1, 5, 9, 13),
+     (1, 5, 9, 13, 0, 4, 8, 12, 3, 7, 11, 15, 2, 6, 10, 14),
+     (2, 6, 10, 14, 1, 5, 9, 13, 0, 4, 8, 12, 3, 7, 11, 15),
+     (3, 7, 11, 15, 2, 6, 10, 14, 1, 5, 9, 13, 0, 4, 8, 12))
+
+p_inv = ((0, 12, 8, 4, 1, 13, 9, 5, 2, 14, 10, 6, 3, 15, 11, 7),
+         (4, 0, 12, 8, 5, 1, 13, 9, 6, 2, 14, 10, 7, 3, 15, 11),
+         (8, 4, 0, 12, 9, 5, 1, 13, 10, 6, 2, 14, 11, 7, 3, 15),
+         (12, 8, 4, 0, 13, 9, 5, 1, 14, 10, 6, 2, 15, 11, 7, 3))
+
+regs = [list() for _ in range(8)]
+for r in range(4):
+    # r + 0, low
+    for byte in range(14, -1, -2):
+        byte_perm = p[r][byte]
+        regs[r].append(byte_perm // 2 if byte_perm % 2 == 0 else byte_perm // 2 + 16)
+
+    # r + 4, low
+    for byte in range(15, -1, -2):
+        byte_perm = p[r][byte]
+        regs[r + 4].append(byte_perm // 2 if byte_perm % 2 == 0 else byte_perm // 2 + 16)
+
+for r in range(8):
+    for v in range(8):
+        print(f"{regs[r][v]:02x}", end='')
+    print('')
+
+# inverse
+
+regs = [list() for _ in range(8)]
+for r in range(4):
+    # r + 0, low
+    for byte in range(14, -1, -2):
+        byte_perm = p_inv[r][byte]
+        regs[r].append(byte_perm // 2 if byte_perm % 2 == 0 else byte_perm // 2 + 16)
+
+    # r + 4, low
+    for byte in range(15, -1, -2):
+        byte_perm = p_inv[r][byte]
+        regs[r + 4].append(byte_perm // 2 if byte_perm % 2 == 0 else byte_perm // 2 + 16)
+
+for r in range(8):
+    for v in range(8):
+        print(f"{regs[r][v]:02x}", end='')
+    print('')

### 2023-03-18 Paper/paper.tex
add some code

@@
 \lstset{
-    basicstyle=\footnotesize\ttfamily,
+    basicstyle=\scriptsize\ttfamily,
     breaklines=true
@@
 \lstset {
+    frame=single,
     numbers=left,
@@
     firstnumber=1,
-    numberfirstline=true
+    numberfirstline=true,
+    texcl=false
 }
@@
 \chapter{Introduction}
+\section{Notation}
+
 \section{Block ciphers}
@@
 
-\begin{lstlisting}[frame=single, caption={Table construction algorithm}, escapechar=|]
+\begin{lstlisting}[caption={Table construction algorithm}, escapechar=|]
     tables <- [][]
@@
 
+\subsubsection{Parallel S-Boxes}
+
 This representation offers multiple advantages. We first note that computation
@@
 packing presented by the original GIFT authors, although it is simple to adapt
-this algorithm to only four registers and eight messages by dividing each
-128-bit NEON register into two separate 64-bit registers and adjusting the
+this algorithm to only four registers and eight messages by adjusting the
 \texttt{SWAPMOVE} shift and mask values.
@@
 \end{align*}
+
+One caveat of this approach is the fact that NEON registers cannot be shifted
+in their entirety due to the fact bits are not able to cross lanes. This leads
+to the problem of being able to shift at most two lanes of 64 bits at once. We
+thus need to implement the \texttt{shr(V,n)} and \texttt{shl(V,n)} operations
+on our own. This can be done by first extracting the 64-bit lanes $a,b$ out of
+$V=b||a$, shifting the lanes individually and finally shifting and ORing the
+crossing bits back into the other lane.
+
+\begin{alignat*}{2}
+    &\text{shl}(V,n): \\
+    &\qquad a,b&&=V[0],V[1] \\
+    &\qquad c&&=(a\gg (64-n)) \\
+    &\qquad a&&=(a\ll n) \\
+    &\qquad b&&=(b\ll n)\lor c \\
+    &\qquad V[0],V[1]&&=a,b
+\end{alignat*}
 
@@
 
+\subsubsection{Round key function}
+
+In contrast to packing and unpacking of data which is only done once in the
+beginning and end, a round key is derived for every round, so the round key
+derivation function needs to be as fast as possible. A simple but naive
+approach for one round would be to generate a single round key, copy it 15
+times and pack the resulting registers similar to how we proceed with the
+messages. Due to the cost of packing the messages, this is prohibitivly
+expensive. Because we know where each byte group ends up after packing, we can
+directly set the round key bits at the correct position. Extending these bits
+to bytes can then be done simply by repeatedly shifting and ORing the registers
+together.
+
+\subsection{Strategies for Camellia}
+
 \chapter{Implementation}
 
-We will provide implementations for the presented strategies in the C
-programming language. Although directly writing Assembler code could result in
-a small performance benefit, this generally increases the work necessary by an
-order of magnitude for only limited results.
+We will provide and discuss implementations for the presented strategies in the
+C programming language. Although directly writing Assembler code could result
+in a small performance benefit, this generally increases the work necessary by
+an order of magnitude for only limited results. This is why we use NEON
+intrinsics.
+
+\subsection{NEON intrinsics}
+
+The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
+definitions including vector data types and C functions for working with these
+vectors. These functions are known as NEON intrinsics and give the programmer a
+high-level interface to most NEON instructions. Major advantages of this
+approach include the ease of development as the compiler takes over register
+allocation and load/store operations as well as performance benefits through
+compiler optimizations.
+
+Standard vector data types have the format \texttt{uintnxm\_t} with lane width
+$n$ in bits and and lane count $m$. Array types of the format
+\texttt{uintnxmxc\_t}, $c\in\{1,2,3,4\}$ are also defined which are used in
+operations requiring multiple parameters like \texttt{TBL} or pairwise
+load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
+after the operation name to indicate operation on a 128-bit register and end
+with the lane data format. Multiplying eight pairs of 16-bit numbers \texttt{a,b} for
+example can be done via the following:
+
+\begin{center}
+    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
+\end{center}
+
+The compiler allocates vector registers for \texttt{a}, \texttt{b} and
+\texttt{result} and assembles the intrinsic to \texttt{MUL Vr.8H, Va.8H,
+Vb.8H}. Necessary load and stores for the result and parameters are also
+handled automatically.
 
 \chapter{Evaluation}
+
+\chapter{Appendix}
+
+\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.h}]
+#ifndef GIFT_VEC_SLICED_H
+#define GIFT_VEC_SLICED_H
+
+#include <stdint.h>
+#include <arm_neon.h>
+
+#define ROUNDS_GIFT_64 28
+
+// expose for benchmarking
+uint8x16_t shl(uint8x16_t v, int n);
+uint8x16_t shr(uint8x16_t v, int n);
+void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
+                                 uint8x16_t m, int n);
+void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
+
+void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
+void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
+void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
+void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
+void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
+                                            const uint64_t key[restrict 2]);
+
+void gift_64_vec_sliced_init(void);
+
+void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
+                                const uint64_t m[restrict 16],
+                                const uint64_t key[restrict 2]);
+void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
+                                const uint64_t c[restrict 16],
+                                const uint64_t key[restrict 2]);
+
+#endif
+\end{lstlisting}
+
+\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.c}]
+#include "gift_vec_sliced.h"
+
+#include <stddef.h>
+#include <stdio.h>
+#include <string.h>
+
+static uint64_t perm_u64[16] = {
+        0x1410050115110400UL, 0x1410050115110400UL,
+        0x0501151104001410UL, 0x0501151104001410UL,
+        0x1511040014100501UL, 0x1511040014100501UL,
+        0x0400141005011511UL, 0x0400141005011511UL,
+        0x1612070317130602UL, 0x1612070317130602UL,
+        0x0703171306021612UL, 0x0703171306021612UL,
+        0x1713060216120703UL, 0x1713060216120703UL,
+        0x0602161207031713UL, 0x0602161207031713UL
+};
+
+static uint64_t perm_inv_u64[16] = {
+        0x1511050114100400UL, 0x1511050114100400UL,
+        0x1713070316120602UL, 0x1713070316120602UL,
+        0x1115010510140004UL, 0x1115010510140004UL,
+        0x1317030712160206UL, 0x1317030712160206UL,
+        0x1317030712160206UL, 0x1317030712160206UL,
+        0x1511050114100400UL, 0x1511050114100400UL,
+        0x1713070316120602UL, 0x1713070316120602UL,
+        0x1115010510140004UL, 0x1115010510140004UL
+};
+
+static uint8x16x4_t perm[2];
+static uint8x16x4_t perm_inv[2];
+
+static uint8x16_t pack_mask_0;
+static uint8x16_t pack_mask_1;
+static uint8x16_t pack_mask_2;
+
+static const int round_constant[] = {
+        // rounds 0-15
+        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
+        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
+        // rounds 16-31
+        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
+        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
+        // rounds 32-47
+        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
+        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
+};
+
+uint8x16_t shl(uint8x16_t v, int n)
+{
+        // TODO check out vst
+        uint64_t l0 = vgetq_lane_u64(v, 0);
+        uint64_t l1 = vgetq_lane_u64(v, 1);
+        l1 = l1 << n | (l0 >> (64 - n));
+        l0 <<= n;
+        uint8x16_t ret;
+        ret = vsetq_lane_u64(l0, ret, 0);
+        ret = vsetq_lane_u64(l1, ret, 1);
+        return ret;
+}
+
+uint8x16_t shr(uint8x16_t v, int n)
+{
+        uint64_t l0 = vgetq_lane_u64(v, 0);
+        uint64_t l1 = vgetq_lane_u64(v, 1);
+        l0 = l0 >> n | (((l1 << (64 - n)) >> (64 - n)) << (64 - n));
+        l1 >>= n;
+        uint8x16_t ret;
+        ret = vsetq_lane_u64(l0, ret, 0);
+        ret = vsetq_lane_u64(l1, ret, 1);
+        return ret;
+}
+
+void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
+{
+
+        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
+        *b = veorq_u8(*b, t);
+        *a = veorq_u8(*a, shl(t, n));
+}
+
+void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
+{
+        // take care not to shift mask bits out of the register
+        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[1], pack_mask_0, 1);
+        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[0].val[3], pack_mask_0, 1);
+        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[1], pack_mask_0, 1);
+        gift_64_vec_sliced_swapmove(&m[1].val[2], &m[1].val[3], pack_mask_0, 1);
+
+        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[2], pack_mask_1, 2);
+        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[0].val[3], pack_mask_1, 2);
+        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[2], pack_mask_1, 2);
+        gift_64_vec_sliced_swapmove(&m[1].val[1], &m[1].val[3], pack_mask_1, 2);
+
+        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
+        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[1].val[0], pack_mask_2, 4);
+        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[1].val[2], pack_mask_2, 4);
+        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[1].val[1], pack_mask_2, 4);
+        gift_64_vec_sliced_swapmove(&m[0].val[3], &m[1].val[3], pack_mask_2, 4);
+}
+
+void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
+{
+        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
+        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
+        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
+        cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
+        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
+        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
+        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
+        cs[0].val[3] = t;
+
+        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
+        t            = veorq_u8(cs[1].val[0], vandq_u8(cs[1].val[1], cs[1].val[3]));
+        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
+        cs[1].val[0] = veorq_u8(cs[1].val[3], cs[1].val[2]);
+        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
+        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
+        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
+        cs[1].val[3] = t;
+}
+
+void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cs[restrict 2])
+{
+        uint8x16_t t = cs[0].val[3];
+        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
+        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
+        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
+        cs[0].val[3] = veorq_u8(cs[0].val[0], cs[0].val[2]);
+        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
+        cs[0].val[0] = veorq_u8(t, vandq_u8(cs[0].val[1], cs[0].val[3]));
+        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
+
+        t            = cs[1].val[3];
+        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
+        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
+        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
+        cs[1].val[3] = veorq_u8(cs[1].val[0], cs[1].val[2]);
+        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
+        cs[1].val[0] = veorq_u8(t, vandq_u8(cs[1].val[1], cs[1].val[3]));
+        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
+}
+
+void gift_64_vec_sliced_permute(uint8x16x4_t cs[restrict 2])
+{
+        uint8x16x2_t pairs[4] = {
+                { .val = { cs[0].val[0], cs[1].val[0] }},
+                { .val = { cs[0].val[1], cs[1].val[1] }},
+                { .val = { cs[0].val[2], cs[1].val[2] }},
+                { .val = { cs[0].val[3], cs[1].val[3] }},
+        };
+
+        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm[0].val[0]);
+        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm[0].val[1]);
+        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm[0].val[2]);
+        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm[0].val[3]);
+
+        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm[1].val[0]);
+        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm[1].val[1]);
+        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm[1].val[2]);
+        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm[1].val[3]);
+}
+
+void gift_64_vec_sliced_permute_inv(uint8x16x4_t cs[restrict 2])
+{
+        uint8x16x2_t pairs[4] = {
+                { .val = { cs[0].val[0], cs[1].val[0] }},
+                { .val = { cs[0].val[1], cs[1].val[1] }},
+                { .val = { cs[0].val[2], cs[1].val[2] }},
+                { .val = { cs[0].val[3], cs[1].val[3] }},
+        };
+
+        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm_inv[0].val[0]);
+        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm_inv[0].val[1]);
+        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm_inv[0].val[2]);
+        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm_inv[0].val[3]);
+
+        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm_inv[1].val[0]);
+        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm_inv[1].val[1]);
+        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm_inv[1].val[2]);
+        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm_inv[1].val[3]);
+}
+
+void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
+                                            const uint64_t key[restrict 2])
+{
+        // TODO shift-and-or
+
+        uint64_t key_state[] = {key[0], key[1]};
+        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
+                int v = (key_state[0] >> 0 ) & 0xffff;
+                int u = (key_state[0] >> 16) & 0xffff;
+
+                // add round key (RK=U||V)
+                uint64_t round_key[8] = { 0x0UL };
+                for (size_t i = 0; i < 8; i++) {
+                        int key_bit_v   = (v >> (2 * i + 0)) & 0x1;
+                        int key_bit_u   = (u >> (2 * i + 0)) & 0x1;
+                        round_key[0] ^= (uint64_t)key_bit_v << (i * 8);
+                        round_key[1] ^= (uint64_t)key_bit_u << (i * 8);
+
+                        key_bit_v   = (v >> (2 * i + 1)) & 0x1;
+                        key_bit_u   = (u >> (2 * i + 1)) & 0x1;
+                        round_key[4] ^= (uint64_t)key_bit_v << (i * 8);
+                        round_key[5] ^= (uint64_t)key_bit_u << (i * 8);
+                }
+
+                // add single bit
+                round_key[7] ^= 1UL << (7 * 8);
+
+                // add round constants
+                round_key[3] ^= ((round_constant[round] >> 0) & 0x1) << (0 * 8);
+                round_key[7] ^= ((round_constant[round] >> 1) & 0x1) << (0 * 8);
+                round_key[3] ^= ((round_constant[round] >> 2) & 0x1) << (1 * 8);
+                round_key[7] ^= ((round_constant[round] >> 3) & 0x1) << (1 * 8);
+                round_key[3] ^= ((round_constant[round] >> 4) & 0x1) << (2 * 8);
+                round_key[7] ^= ((round_constant[round] >> 5) & 0x1) << (2 * 8);
+
+                // extend bits to bytes
+                for (size_t i = 0; i < 8; i++) {
+                        round_key[i] |= round_key[i] << 1;
+                        round_key[i] |= round_key[i] << 2;
+                        round_key[i] |= round_key[i] << 4;
+                }
+
+                // load into vector registers (upper and lower part are same)
+                round_keys[round][0].val[0] = vdupq_n_u64(round_key[0]);
+                round_keys[round][0].val[1] = vdupq_n_u64(round_key[1]);
+                round_keys[round][0].val[2] = vdupq_n_u64(round_key[2]);
+                round_keys[round][0].val[3] = vdupq_n_u64(round_key[3]);
+                round_keys[round][1].val[0] = vdupq_n_u64(round_key[4]);
+                round_keys[round][1].val[1] = vdupq_n_u64(round_key[5]);
+                round_keys[round][1].val[2] = vdupq_n_u64(round_key[6]);
+                round_keys[round][1].val[3] = vdupq_n_u64(round_key[7]);
+
+                // update key state
+                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
+                int k1 = (key_state[0] >> 16) & 0xffffUL;
+                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
+                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
+                key_state[0] >>= 32;
+                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
+                key_state[1] >>= 32;
+                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
+        }
+}
+
+void gift_64_vec_sliced_init(void)
+{
+        // permutations
+        perm[0] = vld1q_u8_x4((uint8_t*)&perm_u64[0]);
+        perm[1] = vld1q_u8_x4((uint8_t*)&perm_u64[8]);
+
+        // inverse permutations
+        perm_inv[0] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[0]);
+        perm_inv[1] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[8]);
+
+        // packing masks
+        pack_mask_0 = vdupq_n_u8(0x55);
+        pack_mask_1 = vdupq_n_u8(0x33);
+        pack_mask_2 = vdupq_n_u8(0x0f);
+}
+
+void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
+                                const uint64_t m[restrict 16],
+                                const uint64_t key[restrict 2])
+{
+        uint8x16x4_t s[2];
+        s[0] = vld1q_u8_x4((uint8_t*)&m[0]);
+        s[1] = vld1q_u8_x4((uint8_t*)&m[8]);
+        gift_64_vec_sliced_bits_pack(s);
+
+        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
+        gift_64_vec_sliced_generate_round_keys(round_keys, key);
+
+        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
+                gift_64_vec_sliced_subcells(s);
+                gift_64_vec_sliced_permute(s);
+
+                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
+                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
+                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
+                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
+                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
+                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
+                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
+                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
+        }
+
+        gift_64_vec_sliced_bits_pack(s);
+        vst1q_u8_x4((uint8_t*)&c[0], s[0]);
+        vst1q_u8_x4((uint8_t*)&c[8], s[1]);
+}
+
+void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
+                                const uint64_t c[restrict 16],
+                                const uint64_t key[restrict 2])
+{
+        uint8x16x4_t s[2];
+        s[0] = vld1q_u8_x4((uint8_t*)&c[0]);
+        s[1] = vld1q_u8_x4((uint8_t*)&c[8]);
+        gift_64_vec_sliced_bits_pack(s);
+
+        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
+        gift_64_vec_sliced_generate_round_keys(round_keys, key);
+
+        for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
+                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
+                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
+                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
+                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
+                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
+                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
+                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
+                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
+
+                gift_64_vec_sliced_permute_inv(s);
+                gift_64_vec_sliced_subcells_inv(s);
+        }
+
+        gift_64_vec_sliced_bits_pack(s);
+        vst1q_u8_x4((uint8_t*)&m[0], s[0]);
+        vst1q_u8_x4((uint8_t*)&m[8], s[1]);
+}
+\end{lstlisting}
 

### 2023-03-20 Paper/paper.tex
intrinsics table

@@
 \usepackage{hyperref} % hyperlinks
+\usepackage{tabularx} % tabular with column specifier X to break lines
 % \usepackage{hyphenat}
@@
 $n$ in bits and and lane count $m$. Array types of the format
-\texttt{uintnxmxc\_t}, $c\in\{1,2,3,4\}$ are also defined which are used in
+\texttt{uintnxmxc\_t}, $c\in\{2,3,4\}$ are also defined which are used in
 operations requiring multiple parameters like \texttt{TBL} or pairwise
 load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
-after the operation name to indicate operation on a 128-bit register and end
-with the lane data format. Multiplying eight pairs of 16-bit numbers \texttt{a,b} for
-example can be done via the following:
+after the operation name to indicate operation on a 128-bit register, and end
+with the lane data format. Multiplying eight pairs of 16-bit numbers
+\texttt{a,b} for example can be done via the following:
 
@@
 
-The compiler allocates vector registers for \texttt{a}, \texttt{b} and
-\texttt{result} and assembles the intrinsic to \texttt{MUL Vr.8H, Va.8H,
-Vb.8H}. Necessary load and stores for the result and parameters are also
-handled automatically.
+In this case, the compiler allocates vector registers for \texttt{a},
+\texttt{b} and \texttt{result} and assembles the intrinsic to \texttt{MUL
+Vr.8H, Va.8H, Vb.8H}. Necessary load and stores for the result and parameters
+are also handled automatically. Of special interest to us are the following
+intrinsics, each existing in different variants with different lane widths and
+also array types: \\
+
+\begin{table}[h!]
+    \centering
+    \footnotesize
+    \begin{tabularx}{\textwidth}{ll|X}
+        Intrinsic && Description \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vreinterpretq\_u8\_u64(uint64x2\_t)} & Explicit casting \\
+        \hline
+        \texttt{uint64\_t} & \texttt{vgetq\_lane\_u64(void)} & Extract a single lane \\
+        \hline
+        \texttt{void} & \texttt{vsetq\_lane\_u64(uint64\_t)} & Insert a single lane \\
+        \hline
+        \texttt{uint64x2\_t} & \texttt{vdupq\_n\_u64(uint64\_t)} & Initialize all lanes to same value \\
+        \hline
+        \texttt{void} & \texttt{vst1q\_u64(uint64\_t*, uint64x2\_t)} & Store from register to memory \\
+        \hline
+        \texttt{uint64x2\_t} & \texttt{vld1q\_u64(uint64\_t*, uint64x2\_t)} & Load from memory to register \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{veorq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise XOR \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vandq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise AND \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vorrq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise OR \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vmvnq\_u8(uint8x16\_t)} & bitwise NOT \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
+    \end{tabularx}
+\end{table}
+
+The implementations closely follow the aforementioned strategies;
+instruction-level optimization is left to the compiler.
 
@@
 {
-        // TODO check out vst
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l1 = l1 << n | (l0 >> (64 - n));
-        l0 <<= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
+        uint64_t l[2];
+        vst1q_u64(l, v);
+        l[1] = l[1] << n | (l[0] >> (64 - n));
+        l[1] <<= n;
+        return vreinterpretq_u8_u64(vld1q_u64(l));
 }
@@
 {
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l0 = l0 >> n | (((l1 << (64 - n)) >> (64 - n)) << (64 - n));
-        l1 >>= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
+        uint64_t l[2];
+        vst1q_u64(l, v);
+        l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
+        l[1] >>= n;
+        return vreinterpretq_u8_u64(vld1q_u64(l));
 }

### 2023-03-20 impl/benchmark.c

@@ -19,13 +19,7 @@
         asm volatile("mrs %[c], PMCCNTR_EL0" : [c] "=r"(t1));\
         printf(#code ": took %ld cycles\n", t1 - t0);
 
-void key_rand(uint64_t k[])
-{
-        k[0] = rand();
-        k[1] = rand();
-}
-
-void m_rand(uint8_t m[], size_t n)
+void rand_bytes(uint8_t m[], size_t n)
 {
         for (size_t i = 0; i < n; i++) {
                 m[i] = rand();
@@ -35,10 +29,10 @@ void m_rand(uint8_t m[], size_t n)
 void benchmark_gift_64(void)
 {
         uint64_t key[2];
-        key_rand(key);
+        rand_bytes((uint8_t*)key, sizeof(key));
 
         uint64_t m;
-        m_rand((uint8_t*)&m, sizeof(m));
+        rand_bytes((uint8_t*)&m, sizeof(m));
 
         uint64_t round_keys[ROUNDS_GIFT_64];
 
@@ -57,10 +51,10 @@ void benchmark_gift_64(void)
 void benchmark_gift_128(void)
 {
         uint64_t key[2];
-        key_rand(key);
+        rand_bytes((uint8_t*)key, sizeof(key));
 
         uint8_t m[16];
-        m_rand(m, sizeof(m));
+        rand_bytes(m, sizeof(m));
 
         uint8_t c[16];
         uint8_t round_keys[ROUNDS_GIFT_128][32];
@@ -80,10 +74,10 @@ void benchmark_gift_128(void)
 void benchmark_gift_64_sliced(void)
 {
         uint64_t key[2];
-        key_rand(key);
+        rand_bytes((uint8_t*)key, sizeof(key));
 
         uint8_t m[16];
-        m_rand(m, sizeof(m));
+        rand_bytes(m, sizeof(m));
 
         uint8_t c[16];
         uint8_t round_keys[ROUNDS_GIFT_128][32];
@@ -99,7 +93,7 @@ void benchmark_gift_64_sliced(void)
 void benchmark_gift_64_vec_sbox(void)
 {
         uint64_t key[2];
-        key_rand(key);
+        rand_bytes((uint8_t*)key, sizeof(key));
         gift_64_vec_sbox_init();
 
         uint8x16_t m;
@@ -121,17 +115,17 @@ void benchmark_gift_64_vec_sbox(void)
         }
 
         uint64_t m_;
[…]

### 2023-03-20 impl/vector/gift_vec_sliced.c

@@ -44,27 +44,20 @@ static const int round_constant[] = {
 
 uint8x16_t shl(uint8x16_t v, int n)
 {
-        // TODO check out vst
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l1 = l1 << n | (l0 >> (64 - n));
-        l0 <<= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
+        uint64_t l[2];
+        vst1q_u64(l, v);
+        l[1] = l[1] << n | (l[0] >> (64 - n));
+        l[1] <<= n;
+        return vreinterpretq_u8_u64(vld1q_u64(l));
 }
 
 uint8x16_t shr(uint8x16_t v, int n)
 {
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l0 = l0 >> n | (((l1 << (64 - n)) >> (64 - n)) << (64 - n));
-        l1 >>= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
+        uint64_t l[2];
+        vst1q_u64(l, v);
+        l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
+        l[1] >>= n;
+        return vreinterpretq_u8_u64(vld1q_u64(l));
 }
 
 void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)

### 2023-03-20 Paper/paper.tex
change in representation yet again

@@
 cryptographic hardware and embedded systems conference, is a lightweight block
-cipher based on a previous design called \texttt|PRESENT|, developed in 2007. Its
+cipher based on a previous design called \texttt{PRESENT}, developed in 2007. Its
 goal is to offer maximum security while being extremely light on resources.
@@
 and using the aforementioned \texttt{vperm} instruction to implement the bit
-shuffling. We then need only four \texttt{TBL} instructions for the complete
-diffusion layer.
+shuffling. We then need only four instructions for the complete diffusion
+layer.
 
@@
 \[
-    \begin{array}{c|llllllll}
-        n & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
-        \hline
-        V_0 & A56 & A48 & A40 & A32 & A24 & A16 & A8 & A0 \\
-        V_1 & A57 & A49 & A41 & A33 & A25 & A17 & A9 & A1 \\
-        V_2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
-        V_3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
-        V_4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
-        V_5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
-        V_6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
-        V_7 & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
+    \scriptsize
+    \begin{array}{c|llllllll|llllllll}
+        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
+        \hline
+        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8 & B0 & A56 & A48 & A40 & A32 & A24 & A16 & A8 & A0 \\
+        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9 & B1 & A57 & A49 & A41 & A33 & A25 & A17 & A9 & A1 \\
+        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
+        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
+        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
+        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
+        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
+        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7  & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
     \end{array}
@@
 
-\[
-    \begin{array}{c|llllllll}
-        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 \\
-        \hline
-        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8 & B0 \\
-        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9 & B1 \\
-        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 \\
-        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 \\
-        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 \\
-        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 \\
-        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 \\
-        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7
-    \end{array}
-\]
+Although this would already work, we prefer to have only bits of the same
+messages in each register - otherwise the permutation would need to operate on
+two source registers with the added requirement of storing the pre-permutation
+values for the first four registers, slowing down the round function through
+superfluous load/stores. This transformation is trivial by use of
+\texttt{TBL} with two data source operands.
 
@@
 
-One thing to take note of is the original permutation values only showing where
-a given byte should land, not which byte belongs to a certain position - i.e.
-for $P_0$, byte 1 should land in position 12, but the byte belonging to
-position 1 is byte 4. Because \texttt{vperm} works in the latter way, we have
-to do some trivial rearrangements. In addition, because data for each slice is
-split between two registers $(V_0,V_4),(V_1,V_5),\dots$, the \texttt{vperm}
-operation needs to use these two registers as a data source. Using two data
-sources is supported by the \texttt{TBL} instruction by means of an additional
-parameter.
+Another thing to take note of is the original permutation values only showing
+where a given byte should land, not which byte belongs to a certain position -
+i.e. for $P_0$, byte 1 should land in position 12, but the byte belonging to
+position 1 is byte 4. Because \texttt{TBL} works in the latter way, we have
+to do some trivial rearrangements.
 
 Assuming the correct permutation values are put into registers
-$V_9,V_{10},\dots,V_{16}$, this now allows us to compute the permutation layer for
-all 16 blocks in only eight permutation instructions plus four additional vector
-copy instructions used for saving the original values of $V_0,V_1,V_2,V_3$
-before the permutation is applied.
+$V_8,V_{10},\dots,V_{15}$, this now allows us to compute the permutation layer
+for all 16 blocks in only eight permutation instructions.
 
 \begin{alignat*}{2}
-    &\texttt{MOV V17, V0}\qquad &&\texttt{MOV V18, V1} \\
-    &\texttt{MOV V19, V2}\qquad &&\texttt{MOV V20, V3} \\
-    &\texttt{TBL V0, \{V0, V4\}, V9}\qquad &&\texttt{TBL V1, \{V1, V5\}, V10} \\
-    &\texttt{TBL V2, \{V2, V6\}, V9}\qquad &&\texttt{TBL V3, \{V3, V7\}, V10} \\
-    &\texttt{TBL V4, \{V17, V4\}, V9}\qquad &&\texttt{TBL V5, \{V18, V5\}, V10} \\
-    &\texttt{TBL V6, \{V19, V6\}, V9}\qquad &&\texttt{TBL V7, \{V20, V7\}, V10}
+    &\texttt{TBL V0, V0, V8}\qquad &&\texttt{TBL V1, V1, V9} \\
+    &\texttt{TBL V2, V2, V10}\qquad &&\texttt{TBL V3, V3, V11} \\
+    &\texttt{TBL V4, V4, V12}\qquad &&\texttt{TBL V5, V5, V13} \\
+    &\texttt{TBL V6, V6, V14}\qquad &&\texttt{TBL V7, V7, V15}
 \end{alignat*}

### 2023-03-20 impl/test.c

@@ -184,6 +184,14 @@ void test_gift_64_vec_sliced(void)
 {
         gift_64_vec_sliced_init();
 
+        printf("testing GIFT_64_VEC_SLICED substitution...\n");
+        uint8x16x4_t s[2];
+        s[0].val[0] = vdupq_n_u64(0x0123456789abcdefUL);
+        gift_64_vec_sliced_bits_pack(s);
+        gift_64_vec_sliced_subcells(s);
+        gift_64_vec_sliced_bits_unpack(s);
+        ASSERT_EQUALS(vgetq_lane_u64(s[0].val[0], 0), 0x1a4c6f392db7508e);
+
         // test encrypt to known value
         printf("testing GIFT_64_VEC_SLICED encrytion to known value...\n");
         uint64_t key[2] = { 0x5085772fe6916616UL, 0x3c9d8c18fdd20608UL };

### 2023-03-20 impl/vector/gift_vec_sliced.c

@@ -4,28 +4,42 @@
 #include <stdio.h>
 #include <string.h>
 
+static uint64_t pack_shf_u64[16] = {
+        0x1707160615051404UL, 0x1303120211011000UL, // S0
+        0x1707160615051404UL, 0x1303120211011000UL, // S1
+        0x1707160615051404UL, 0x1303120211011000UL, // S2
+        0x1707160615051404UL, 0x1303120211011000UL, // S3
+        0x1f0f1e0e1d0d1c0cUL, 0x1b0b1a0a19091808UL, // S4
+        0x1f0f1e0e1d0d1c0cUL, 0x1b0b1a0a19091808UL, // S5
+        0x1f0f1e0e1d0d1c0cUL, 0x1b0b1a0a19091808UL, // S6
+        0x1f0f1e0e1d0d1c0cUL, 0x1b0b1a0a19091808UL  // S7
+};
+
+// TODO find reverse
+
 static uint64_t perm_u64[16] = {
-        0x1410050115110400UL, 0x1410050115110400UL,
-        0x0501151104001410UL, 0x0501151104001410UL,
-        0x1511040014100501UL, 0x1511040014100501UL,
-        0x0400141005011511UL, 0x0400141005011511UL,
-        0x1612070317130602UL, 0x1612070317130602UL,
-        0x0703171306021612UL, 0x0703171306021612UL,
-        0x1713060216120703UL, 0x1713060216120703UL,
-        0x0602161207031713UL, 0x0602161207031713UL
+        0x0d0905010e0a0602UL, 0x0f0b07030c080400UL, // S0
+        0x0105090d0004080cUL, 0x03070b0f02060a0eUL, // S1
+        0x02060a0e0105090dUL, 0x0004080c03070b0fUL, // S2
+        0x03070b0f02060a0eUL, 0x0105090d0004080cUL, // S3
+        0x0d0905010e0a0602UL, 0x0f0b07030c080400UL, // S4
+        0x0105090d0004080cUL, 0x03070b0f02060a0eUL, // S5
+        0x02060a0e0105090dUL, 0x0004080c03070b0fUL, // S6
+        0x03070b0f02060a0eUL, 0x0105090d0004080cUL  // S7
 };
 
 static uint64_t perm_inv_u64[16] = {
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL,
[…]

### 2023-03-20 impl/vector/gift_vec_sliced.h

@@ -11,6 +11,7 @@ uint8x16_t shl(uint8x16_t v, int n);
 uint8x16_t shr(uint8x16_t v, int n);
 void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n);
 void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
+void gift_64_vec_sliced_bits_unpack(uint8x16x4_t m[restrict 2]);
 
 void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
 void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);

### 2023-03-20 scripts/64_vec_sliced_pack.py

@@ -1,0 +1,8 @@
+#!/usr/bin/env python3
+
+print('0x', end='')
+for i in range(15, -1, -1):
+    src = (i % 2) * 16 + i // 2 + 8
+    print(f'{src:02x}', end='')
+
+print('')

### 2023-03-20 scripts/perm_64_sliced_better.py

@@ -5,31 +5,37 @@ p1 = (4, 0, 12, 8, 5, 1, 13, 9, 6, 2, 14, 10, 7, 3, 15, 11)
 p2 = (8, 4, 0, 12, 9, 5, 1, 13, 10, 6, 2, 14, 11, 7, 3, 15)
 p3 = (12, 8, 4, 0, 13, 9, 5, 1, 14, 10, 6, 2, 15, 11, 7, 3)
 
-for i in range(16):
+print('0x', end='')
+for i in range(15, -1, -1):
     for j, p0j in enumerate(p0):
         if p0j == i:
-            print(j, end=', ')
+            print(f"{j:02x}", end='')
 
 print('')
 
+print('0x', end='')
 for i in range(16):
     for j, p1j in enumerate(p1):
         if p1j == i:
-            print(j, end=', ')
+            print(f"{j:02x}", end='')
 
 print('')
 
+print('0x', end='')
 for i in range(16):
     for j, p2j in enumerate(p2):
         if p2j == i:
-            print(j, end=', ')
+            print(f"{j:02x}", end='')
+
 
 print('')
 
+print('0x', end='')
 for i in range(16):
     for j, p3j in enumerate(p3):
         if p3j == i:
-            print(j, end=', ')
+            print(f"{j:02x}", end='')
+
 
 print('\n0x', end='')
 

### 2023-03-23 Paper/chapters/abstract.tex
move chapters into own files

--- 2023-03-18 Paper/paper.tex
@@
-\documentclass[12pt]{report}
-
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{ulem} % underline
-\usepackage{float} % H float specifier
-\usepackage{hyperref} % hyperlinks
-% \usepackage{hyphenat}
-\usepackage{bookmark}
-\usepackage{systeme} % systems of equation
-\usepackage[utf8]{inputenc} % necessary for proper unicode support
-\usepackage{parcolumns} % multiple column parallel typesetting
-\usepackage[english]{babel} % comment this for Angelsächsisch
-\usepackage{graphicx} % including graphics
-\usepackage{sourcecodepro} % best font
-\usepackage{listings} % source code listings
-\usepackage{biblatex} % better citation
-% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry
-
-% umlauts in lstlistings env
-\lstset{
-    literate={ö}{{\"o}}1
-    {ä}{{\"a}}1
-    {ü}{{\"u}}1
-    {Ö}{{\"O}}1
-    {Ä}{{\"A}}1
-    {Ü}{{\"U}}1
-}
-
-\lstset{
-    basicstyle=\scriptsize\ttfamily,
-    breaklines=true
-}
-
-\lstset {
-    frame=single,
-    numbers=left,
-    stepnumber=1,
-    firstnumber=1,
-    numberfirstline=true,
-    texcl=false
-}
-
-\providecommand{\lxor}{\veebar}
-\renewcommand{\proofname}{Beweis.}
-\newcommand{\sref}[1]{\textsuperscript{\ref{#1}}}
-\newcommand{\F}[1]{\mathbb{F}_2^{#1}}
-
-\addbibresource{refs.bib}
-
-\title{
-    \textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\
-    {\footnotesize Bachelorarbeit}
-}
-\author{Bastian Engel}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
 \chapter*{Abstract}
@@
 culpa qui officia deserunt mollit anim id est laborum.
-
-\chapter*{Declaration}
-
-I hereby declare that ...
-
-\tableofcontents
-
-\chapter{Introduction}
-\section{Notation}
-
-\section{Block ciphers}
-
-Securing communication channels between different parties has been a long-term
-subject of study for cryptographers and engineers which is essential to our
-modern world to cope with ever-increasing amounts of devices producing and
-sharing data. The main way to facilitate high-throughput, confidential
-communications nowadays is through the use of symmetric cryptography in which
-two parties share a common secret, called a key, which allows them to encrypt,
-share and subsequently decrypt messages to achieve confidentiality against
-third parties. Ciphers can be divided into two categories; block ciphers, which
-always encrypt fixed-sized messages called blocks, and stream ciphers, which
-continuously provide encryption for an arbitrarily long, constant stream of
-data.
-
-A block cipher can be defined as a bijection between the input block (the
-message) and the output block (the ciphertext). For any block cipher with block
-size $n$, we denote the key-dependent encryption and decryption functions as
-$E_K,D_K:\F{n}\rightarrow \F{n}$. The simplest way to
-characterize this bijection is through a lookup table which yields the highest
-possible performance as each block can be encrypted by one simple lookup
-depending on the key and the message. This is not practical though due to most
-ciphers working with block and key sizes $n,|K|\geq 64$. For a block cipher
-with $n=64,|K|=128$, a space of $2^{64}2^{128}64=2^{198}$ is necessary.
-Considering modern consumer hard disks being able to store data in the order of
-$2^{40}$, it is easy to see that a lookup table is wholly impractical. We
-therefore describe block ciphers algorithmically which opens up possibilities
-for different tradeoffs and security concerns.
-
-\subsection{GIFT}
-
-\texttt{GIFT}\cite{gift:2017}, first presented in the \textit{CHES 2017}
-cryptographic hardware and embedded systems conference, is a lightweight block
-cipher based on a previous design called \texttt|PRESENT|, developed in 2007. Its
-goal is to offer maximum security while being extremely light on resources.
-Modern battery-powered devices like RFID tags or low-latency operations like
-on-the-fly disc encryption present strong hardware and power constraints. GIFT
-aims to be a simple, low-energy cipher suited for these kinds of applications.
-
-\texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
-and \texttt{GIFT-128} working with 128-bit blocks. In both cases, the key is 128
-bits long. The design is a very simple, round-based substitution-permutation
-network (SPN). One round consists in a sequential application of the confusion
-layer by means of 4-bit S-boxes and subsequent diffusion through bit
-permutation. After the bit permutation, a round key is added to the cipher
-state and the single round is complete. \texttt{GIFT-64} uses 28 rounds while
-\texttt{GIFT-128} uses 40 rounds.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/GIFT-64.pdf}
-    \caption{Two rounds of GIFT-64}
-\end{figure}
-
-\subsubsection{Substitution layer}
-
-The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
-16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \texttt{GIFT-128}. The S-box
-$S:\F{4}\rightarrow \F{4}$ is defined as follows:
-
-\[
-    \begin{array}{l|cccccccccccccccc}
-        x & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & a & b & c & d & e & f \\
-        \hline
-        S(x) & 1 & a & 4 & c & 6 & f & 3 & 9 & 2 & d & b & 7 & 5 & 0 & 8 & e
-    \end{array}
-\]
-
-\subsubsection{Permutation layer}
-
-The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
-i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64} and
-\texttt{GIFT-128} can be expressed by:
-
-\begin{align*}
-    P_{64}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+16\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-    P_{128}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+32\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-\end{align*}
-
-\subsubsection{Round key addition}
-
-The last step of each round consists in XORing a round key $R_i$ to the cipher
-state. The new cipher state $s_{i+1}$ after each full round is therefore given
-by
-
-\[
-    s_{i+1}=P(S(s_i))\oplus R_i
-\]
-
-\subsubsection{Round key extraction and key schedule}
-
-Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
-$K=k7||k6||\dots||k0$ denote the $128$-bit key state.
-
-\paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
-state. $u_i$ and $v_i$ are XORed to $r_{4i+1}$ and $r_{4i}$ of the round key
-$R$ respectively.
-
-\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k4||k1||k_0$ from
-the key state. $u_i$ and $v_i$ are XORed to $r_{4i+2}$ and $b_{4i+1}$ of the
-round key $R$ respectively.
-
-In both cases, we additionally XOR a round constant $C=c_5c_4c_3c_2c_1c_0$ to
-bit positions $n-1,23,19,15,11,7,3$. The round constants are generated using a
-6-bit affine linear-feedback shift register and have the following values:\\
-
-\begin{tabular}{r|l}
-    \textbf{Rounds} & \textbf{Constants} \\
-    \hline
-    \textbf{1 - 16} &  \small\texttt{01,03,07,0F,1F,3E,3D,3B,37,2F,1E,3C,39,33,27,0E} \\
-    \textbf{17 - 32} & \small\texttt{1D,3A,35,2B,16,2C,18,30,21,02,05,0B,17,2E,1C,38} \\
-    \textbf{33 - 48} & \small\texttt{31,23,06,0D,1B,36,2D,1A,34,29,12,24,08,11,22,04}
-\end{tabular}\\
-
-The key state is then updated by setting $k_1\leftarrow k_1\ggg 2$,
-$k_0\leftarrow k_0\ggg 12$ and rotating the new state $32$ bits to the right:
-
-\[
-    k_7||k_6||\dots||k_1||k_0\leftarrow k_1\ggg 2||k_0\ggg 12||k_7||k_6||\dots||k_3||k_2
-\]
-
-\subsection{Camellia}
-
-\section{The ARMv8 platform}
-
-With small devices, embedded processors and ASICs becoming ever more ubiquitous
-and essential in areas like medicine or automotive design, the need for ...
-
-\chapter{Implementation strategies}
-
-Due to the structural differences of SPN- and Feistel network-based ciphers, we
-shall analyze these two separately.
-
-\section{Strategies for SPN}
-
-Three implementation strategies for substitution-permutation networks are
-introduced by \cite{implx86:2014}:
-
-\begin{itemize}
-    \item Table-based implementations
-    \item \texttt{vperm} implementations
-    \item Bitslice implementations
-\end{itemize}
-
-\subsection{Table-based}
-
-Table-driven programming is a simple way to increase performance of operations
-by tabulating the results, therefore requiring only a single memory access to
-acquire the result. This approach is obviously limited to manageable table
-sizes, so while tabulating a function like the AES S-box
-$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space,
-tabulating the \texttt{GIFT} permutation layer
-$P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
-$2^{70}$ space, which is totally unfeasible.
-
-A common approach is to tabulate the output of each S-box, including the
-diffusion layer, and then XORing the results together. Let $n$ denote the
-internal cipher state size and $s$ the size of a single S-box in bits. For each
-S-box $S_i,i\in\{0,\dots,\frac{n}{s}\}$, we can construct a mapping
-$T_i:\F{s}\rightarrow \F{n}$ representing substitution with subsequent
-permutation of that single S-box. The cipher state before round key addition is
-then given by $\bigoplus_{i=0}^{\frac{n}{s}-1}{T_i(m_i)}$ for each $s$-bit
-message chunk $m_i$. This approach requires space of
-$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for \texttt{GIFT-64},
-results in a manageable size of $\frac{64^2 2^4}{4}=2^{14}$ bits which equals
-$16$ KiB.
-
-\subsubsection{Constructing the tables}
-
-For \texttt{GIFT-64}, table construction is relatively straightforward and can
-be done as follows:
-
-\begin{lstlisting}[caption={Table construction algorithm}, escapechar=|]
-    tables <- [][]
-    for sbox_index from 0 to 15 do
-        for sbox_input from 0 to 15 do|\label{lst:tablesbox}|
-            output <- sbox(sbox_input)
-            output <- permute(output << (4 * sbox_index))
-            tables[sbox_index][sbox_input] <- output
-\end{lstlisting}
-
-Implementing this algorithm gives us the following table representing the first
-and second S-box.
-
-\[
-    \begin{array}{l|l|l|c}
-        x & T_0(x) & T_1(x) & \dots \\
-        \hline
-        0x0 & 0x1               & 0x1000000000000   & \dots \\
-        0x1 & 0x8000000020000   & 0x800000002       & \dots \\
-        0x2 & 0x400000000       & 0x40000           & \dots \\
-        0x3 & 0x8000400000000   & 0x800040000       & \dots \\
-        0x4 & 0x400020000       & 0x40002           & \dots \\
-        0x5 & 0x8000400020001   & 0x1000800040002   & \dots \\
-        0x6 & 0x20001           & 0x1000000000002   & \dots \\
-        0x7 & 0x8000000000001   & 0x1000800000000   & \dots \\
-        0x8 & 0x20000           & 0x2               & \dots \\
-        0x9 & 0x8000400000001   & 0x1000800040000   & \dots \\
-        0xa & 0x8000000020001   & 0x1000800000002   & \dots \\
-        0xb & 0x400020001       & 0x1000000040002   & \dots \\
-        0xc & 0x400000001       & 0x1000000040000   & \dots \\
-        0xd & 0x0               & 0x0               & \dots \\
-        0xe & 0x8000000000000   & 0x800000000       & \dots \\
-        0xf & 0x8000400020000   & 0x800040002       & \dots
-    \end{array}
-\]
-
-The tables for \texttt{GIFT-128} can be generated in a similar way by looping
-through all 32 S-boxes instead of 16 on line \ref{lst:tablesbox}.
-
-\subsection{Using \texttt{vperm}}
-
-Nowadays, most instructions set architectures support single-instruction,
-multiple-data processing. The idea of such an SIMD system is to work on
-multiple data stored in vectors at once to speed up calculations. For A64, two
-types of vector processing are available:
-
-\begin{enumerate}
-    \item Advanced SIMD, known as NEON
-    \item Scalable Vector Extension (SVE)
-\end{enumerate}
-
-We will take a look at NEON as this is the type of vector processing supported
-by the Cortex-A73 processor.
-
-\subsubsection{ARM Neon}
-
-The register file of the NEON unit is made up of 32 quad-word (128-bit)
-registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
-registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
-sized lanes on which the vector instructions operate. Valid ways to interpret
-for example the register \texttt{V0} are:
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/V_register.pdf}
-    \caption{Divisions of the V0 register}
-\end{figure}
-
-NEON instructions interpret their operands' layouts (i.e. lane count and width)
-through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
-adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
-together and storing the result in \texttt{V0} can be done as follows:
-
-\begin{center}
-    \texttt{ADD V0.8H, V1.8H, V2.8H}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/vector_add.pdf}
-    \caption{Addition of two vector registers}
-\end{figure}
-
-The plenitude of different processing instructions allow flexible ways to
-further speed up algorithms having reached their optimizational limit on
-non-SIMD platforms. \texttt{vperm}, a general term standing for \textit{vector
-permute}, is a common instruction on SIMD machines. Called \texttt{TBL} on
-NEON, it is used for parallel table lookups and arbitrary permutations. It takes
-two inputs to perform a lanewise lookup:
-
-\begin{enumerate}
-    \item A register with lookup values
-    \item Two or more registers containing data
-\end{enumerate}
-
-\subsubsection{S-box lookup}
-
-This instruction can be used to implement S-box lookup of all 16 S-boxes in a
-single instruction. We do this by packing our 64-bit cipher state
-$s=s_{15}||s_{14}||\dots||s_0$ into a vector register $V_0$. Because we can
-only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane which
-neatly fits into the 128-bit registers. We then put the S-box itself into
-register $V_1$ which will be used as the data register for the table lookup.
-
-The confusion layer can now be performed through one \texttt{TBL} instruction:
-
-\begin{center}
-    \texttt{TBL V0.16B, V1.16B, V0.16B}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/tbl_example.pdf}
-    \caption{Performing the S-Box lookup in parallel}
-\end{figure}
-
-\subsection{Bitslicing}
-
-Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
-achieve a more efficient representation to operate on. The structure of
-\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
-cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
-i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
-individual S-boxes. This is equivalent to transposing the bit matrix.
-
-\[
-    S=\begin{bmatrix}
-        S_0\\
-        S_1\\
-        S_2\\
-        S_3
-    \end{bmatrix}
-    =\begin{bmatrix}
-        b_{60}b_{56}b_{52}\dots b_0\\
-        b_{61}b_{57}b_{53}\dots b_1\\
-        b_{62}b_{58}b_{54}\dots b_2\\
-        b_{63}b_{59}b_{55}\dots b_3
-    \end{bmatrix}
-\]
-
-\subsubsection{Parallel S-Boxes}
-
-This representation offers multiple advantages. We first note that computation
-of the S-box can be executed in parallel, similar to the \texttt{vperm}
-technique above. This can be done by finding an algorithmic way to apply the
-S-box which has already been proposed by the original \texttt{GIFT} authors:
-
-\begin{align*}
-    S_1&\leftarrow S_1\oplus (S_0\land S_2) \\
-    t&\leftarrow S_0\oplus (S_1\land S_3) \\
-    S_2&\leftarrow S_2\oplus (t\lor S_1) \\
-    S_0&\leftarrow S_3\oplus S_2 \\
-    S_1&\leftarrow S_1\oplus S_0 \\
-    S_0&\leftarrow \lnot S_0 \\
-    S_2&\leftarrow S_2\oplus (t\land S_1) \\
-    S_3&\leftarrow t
-\end{align*}
-
-This is very efficient as it only requires six XOR-, three AND and one OR
-operation.
-
-An important property of the permutation is the fact that bits always stay in
-their slice. This means we can decompose the permutation $P$ into four
-permutations $P_i,i\in\{0,1,2,3\}$ and apply these permutations
-separately to each slice. One possible way to implement a permutation $P_i$ in
-software is to mask off all bits individually, shift them to their correct
-position and OR them together:
-
-\[
-    P_i(S_i)=\bigvee_{k=0}^{15}{(S_i\land m_i) \ll s_i}
-\]
-
-This approach requires $47$ operations, meaning all four permutations require
-over $150$ operations which would present a major bottleneck to the round
-function. We can improve on this by working on multiple message blocks at once
-and using the aforementioned \texttt{vperm} instruction to implement the bit
-shuffling. We then need only four \texttt{TBL} instructions for the complete
-diffusion layer.
-
-\subsubsection{Using \texttt{vperm} for slice permutation}
-
-We cannot use the \texttt{TBL} instruction directly as we need to shuffle
-individual bits, but the smallest data we can operate on are bytes. We
-therefore encrypt $8n$ messages at once which allows us to create bytewise
-groupings. These messages are put into $4m$ registers with register $R_{4i}$
-containing $S_0$, register $R_{4i+1}$ containing $S_1$ and so forth. With block
-size $BS$ and register size $RS$, the following must hold:
-
-\[
-    8n\cdot BS=4m\cdot RS
-\]
-
-In the case of \texttt{GIFT-64} with $BS=64$ and ARM NEON with $RS=128$, we get
-
-\[
-    8n\cdot 64=4m\cdot 128\Leftrightarrow n=m
-\]
-
-$n=m=1$ would be a valid choice which yields eight messages divided into four
-registers. We choose $n=m=2$ so we can directly utilize the algorithm for bit
-packing presented by the original GIFT authors, although it is simple to adapt
-this algorithm to only four registers and eight messages by adjusting the
-\texttt{SWAPMOVE} shift and mask values.
-
-\subsubsection{Packing the data into bitslice format}
-
-Let $a,b,\dots,p$ be sixteen messages of length $64$ with subscripts denoting
-individual bits. We first put these messages into eight SIMD registers
-$V_0,V_1,\dots,V_7$:
-
-\begin{alignat*}{3}
-    V_0&=b||a\qquad V_4&&=j||&&i \\
-    V_1&=d||c\qquad V_5&&=l||&&k \\
-    V_2&=f||e\qquad V_6&&=n||&&m \\
-    V_3&=h||g\qquad V_7&&=p||&&o
-\end{alignat*}
-
-We then use the \texttt{SWAPMOVE} technique to bring the data into bitslice
-format. This operation operates on two registers $A,B$ using mask $M$ and shift
-value $N$. It swaps bits in $A$ masked by $(M\ll N)$ with bits in $B$ masked by
-$M$ in using only three XOR-, one AND- and two shift operations.
-
-\begin{align*}
-    &\text{SWAPMOVE}(A,B,M,N): \\
-    &\qquad T=((A\gg N)\oplus B)\land M \\
-    &\qquad B=B\oplus T \\
-    &\qquad A=A\oplus (T\ll N) \\
-\end{align*}
-
-One caveat of this approach is the fact that NEON registers cannot be shifted
-in their entirety due to the fact bits are not able to cross lanes. This leads
-to the problem of being able to shift at most two lanes of 64 bits at once. We
-thus need to implement the \texttt{shr(V,n)} and \texttt{shl(V,n)} operations
-on our own. This can be done by first extracting the 64-bit lanes $a,b$ out of
-$V=b||a$, shifting the lanes individually and finally shifting and ORing the
-crossing bits back into the other lane.
-
-\begin{alignat*}{2}
-    &\text{shl}(V,n): \\
-    &\qquad a,b&&=V[0],V[1] \\
-    &\qquad c&&=(a\gg (64-n)) \\
-    &\qquad a&&=(a\ll n) \\
-    &\qquad b&&=(b\ll n)\lor c \\
-    &\qquad V[0],V[1]&&=a,b
-\end{alignat*}
-
-The following operations group all $i$th bits of the messages $a,c,\dots,o$
-into bytes and puts these into the lower half of the registers $V_{i\bmod 8}$.
-The same is done for messages $b,d,\dots,p$, only differing in that the bytes
-are put into the upper half of the registers.
-
-\begin{align*}
-    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_4,V_5,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_6,V_7,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_4,V_6,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_5,V_7,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_0,V_4,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_1,V_5,0x0f0f\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4) \\
-\end{align*}
-
-With $Ax=o_xm_xk_xj_xg_xe_xc_xa_x$ and $Bx=p_xn_xl_xi_xh_xf_xd_xb_x$ denoting
-byte groups, our data now has the following permutation-friendly format:
-
-\[
-    \begin{array}{c|llllllll}
-        n & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
-        \hline
-        V_0 & A56 & A48 & A40 & A32 & A24 & A16 & A8 & A0 \\
-        V_1 & A57 & A49 & A41 & A33 & A25 & A17 & A9 & A1 \\
-        V_2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
-        V_3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
-        V_4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
-        V_5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
-        V_6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
-        V_7 & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
-    \end{array}
-\]
-
-\[
-    \begin{array}{c|llllllll}
-        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 \\
-        \hline
-        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8 & B0 \\
-        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9 & B1 \\
-        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 \\
-        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 \\
-        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 \\
-        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 \\
-        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 \\
-        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7
-    \end{array}
-\]
-
-We can now create permutation tables using the specification of the individual
-slice permutations $P_i$:
-
-\[
-    \begin{array}{c|llllllllllllllll}
-        j & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\
-        \hline
-        P_0(j) & 0 & 12 & 8 & 4 & 1 & 13 & 9 & 5 & 2 & 14 & 10 & 6 & 3 & 15 & 11 & 7 \\
-        P_1(j) & 4 & 0 & 12 & 8 & 5 & 1 & 13 & 9 & 6 & 2 & 14 & 10 & 7 & 3 & 15 & 11 \\
-        P_2(j) & 8 & 4 & 0 & 12 & 9 & 5 & 1 & 13 & 10 & 6 & 2 & 14 & 11 & 7 & 3 & 15 \\
-        P_3(j) & 12 & 8 & 4 & 0 & 13 & 9 & 5 & 1 & 14 & 10 & 6 & 2 & 15 & 11 & 7 & 3
-    \end{array}
-\]
-
-One thing to take note of is the original permutation values only showing where
-a given byte should land, not which byte belongs to a certain position - i.e.
-for $P_0$, byte 1 should land in position 12, but the byte belonging to
-position 1 is byte 4. Because \texttt{vperm} works in the latter way, we have
-to do some trivial rearrangements. In addition, because data for each slice is
-split between two registers $(V_0,V_4),(V_1,V_5),\dots$, the \texttt{vperm}
-operation needs to use these two registers as a data source. Using two data
-sources is supported by the \texttt{TBL} instruction by means of an additional
-parameter.
-
-Assuming the correct permutation values are put into registers
-$V_9,V_{10},\dots,V_{16}$, this now allows us to compute the permutation layer for
-all 16 blocks in only eight permutation instructions plus four additional vector
-copy instructions used for saving the original values of $V_0,V_1,V_2,V_3$
-before the permutation is applied.
-
-\begin{alignat*}{2}
-    &\texttt{MOV V17, V0}\qquad &&\texttt{MOV V18, V1} \\
-    &\texttt{MOV V19, V2}\qquad &&\texttt{MOV V20, V3} \\
-    &\texttt{TBL V0, \{V0, V4\}, V9}\qquad &&\texttt{TBL V1, \{V1, V5\}, V10} \\
-    &\texttt{TBL V2, \{V2, V6\}, V9}\qquad &&\texttt{TBL V3, \{V3, V7\}, V10} \\
-    &\texttt{TBL V4, \{V17, V4\}, V9}\qquad &&\texttt{TBL V5, \{V18, V5\}, V10} \\
-    &\texttt{TBL V6, \{V19, V6\}, V9}\qquad &&\texttt{TBL V7, \{V20, V7\}, V10}
-\end{alignat*}
-
-\subsubsection{Round key function}
-
-In contrast to packing and unpacking of data which is only done once in the
-beginning and end, a round key is derived for every round, so the round key
-derivation function needs to be as fast as possible. A simple but naive
-approach for one round would be to generate a single round key, copy it 15
-times and pack the resulting registers similar to how we proceed with the
-messages. Due to the cost of packing the messages, this is prohibitivly
-expensive. Because we know where each byte group ends up after packing, we can
-directly set the round key bits at the correct position. Extending these bits
-to bytes can then be done simply by repeatedly shifting and ORing the registers
-together.
-
-\subsection{Strategies for Camellia}
-
-\chapter{Implementation}
-
-We will provide and discuss implementations for the presented strategies in the
-C programming language. Although directly writing Assembler code could result
-in a small performance benefit, this generally increases the work necessary by
-an order of magnitude for only limited results. This is why we use NEON
-intrinsics.
-
-\subsection{NEON intrinsics}
-
-The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
-definitions including vector data types and C functions for working with these
-vectors. These functions are known as NEON intrinsics and give the programmer a
-high-level interface to most NEON instructions. Major advantages of this
-approach include the ease of development as the compiler takes over register
-allocation and load/store operations as well as performance benefits through
-compiler optimizations.
-
-Standard vector data types have the format \texttt{uintnxm\_t} with lane width
-$n$ in bits and and lane count $m$. Array types of the format
-\texttt{uintnxmxc\_t}, $c\in\{1,2,3,4\}$ are also defined which are used in
-operations requiring multiple parameters like \texttt{TBL} or pairwise
-load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
-after the operation name to indicate operation on a 128-bit register and end
-with the lane data format. Multiplying eight pairs of 16-bit numbers \texttt{a,b} for
-example can be done via the following:
-
-\begin{center}
-    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
-\end{center}
-
-The compiler allocates vector registers for \texttt{a}, \texttt{b} and
-\texttt{result} and assembles the intrinsic to \texttt{MUL Vr.8H, Va.8H,
-Vb.8H}. Necessary load and stores for the result and parameters are also
-handled automatically.
-
-\chapter{Evaluation}
-
-\chapter{Appendix}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.h}]
-#ifndef GIFT_VEC_SLICED_H
-#define GIFT_VEC_SLICED_H
-
-#include <stdint.h>
-#include <arm_neon.h>
-
-#define ROUNDS_GIFT_64 28
-
-// expose for benchmarking
-uint8x16_t shl(uint8x16_t v, int n);
-uint8x16_t shr(uint8x16_t v, int n);
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
-                                 uint8x16_t m, int n);
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2]);
-
-void gift_64_vec_sliced_init(void);
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2]);
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2]);
-
-#endif
-\end{lstlisting}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.c}]
-#include "gift_vec_sliced.h"
-
-#include <stddef.h>
-#include <stdio.h>
-#include <string.h>
-
-static uint64_t perm_u64[16] = {
-        0x1410050115110400UL, 0x1410050115110400UL,
-        0x0501151104001410UL, 0x0501151104001410UL,
-        0x1511040014100501UL, 0x1511040014100501UL,
-        0x0400141005011511UL, 0x0400141005011511UL,
-        0x1612070317130602UL, 0x1612070317130602UL,
-        0x0703171306021612UL, 0x0703171306021612UL,
-        0x1713060216120703UL, 0x1713060216120703UL,
-        0x0602161207031713UL, 0x0602161207031713UL
-};
-
-static uint64_t perm_inv_u64[16] = {
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL
-};
-
-static uint8x16x4_t perm[2];
-static uint8x16x4_t perm_inv[2];
-
-static uint8x16_t pack_mask_0;
-static uint8x16_t pack_mask_1;
-static uint8x16_t pack_mask_2;
-
-static const int round_constant[] = {
-        // rounds 0-15
-        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
-        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
-        // rounds 16-31
-        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
-        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
-        // rounds 32-47
-        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
-        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
-};
-
-uint8x16_t shl(uint8x16_t v, int n)
-{
-        // TODO check out vst
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l1 = l1 << n | (l0 >> (64 - n));
-        l0 <<= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
-}
-
-uint8x16_t shr(uint8x16_t v, int n)
-{
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l0 = l0 >> n | (((l1 << (64 - n)) >> (64 - n)) << (64 - n));
-        l1 >>= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
-}
-
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
-{
-
-        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
-        *b = veorq_u8(*b, t);
-        *a = veorq_u8(*a, shl(t, n));
-}
-
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
-{
-        // take care not to shift mask bits out of the register
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[0].val[3], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[2], &m[1].val[3], pack_mask_0, 1);
-
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[0].val[3], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[1], &m[1].val[3], pack_mask_1, 2);
-
-        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[1].val[0], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[1].val[2], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[1].val[1], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[3], &m[1].val[3], pack_mask_2, 4);
-}
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
-{
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[3] = t;
-
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-        t            = veorq_u8(cs[1].val[0], vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(cs[1].val[3], cs[1].val[2]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[3] = t;
-}
-
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16_t t = cs[0].val[3];
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[3] = veorq_u8(cs[0].val[0], cs[0].val[2]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(t, vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-
-        t            = cs[1].val[3];
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[3] = veorq_u8(cs[1].val[0], cs[1].val[2]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(t, vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-}
-
-void gift_64_vec_sliced_permute(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm[1].val[3]);
-}
-
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm_inv[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm_inv[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm_inv[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm_inv[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm_inv[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm_inv[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm_inv[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm_inv[1].val[3]);
-}
-
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2])
-{
-        // TODO shift-and-or
-
-        uint64_t key_state[] = {key[0], key[1]};
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                int v = (key_state[0] >> 0 ) & 0xffff;
-                int u = (key_state[0] >> 16) & 0xffff;
-
-                // add round key (RK=U||V)
-                uint64_t round_key[8] = { 0x0UL };
-                for (size_t i = 0; i < 8; i++) {
-                        int key_bit_v   = (v >> (2 * i + 0)) & 0x1;
-                        int key_bit_u   = (u >> (2 * i + 0)) & 0x1;
-                        round_key[0] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[1] ^= (uint64_t)key_bit_u << (i * 8);
-
-                        key_bit_v   = (v >> (2 * i + 1)) & 0x1;
-                        key_bit_u   = (u >> (2 * i + 1)) & 0x1;
-                        round_key[4] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[5] ^= (uint64_t)key_bit_u << (i * 8);
-                }
-
-                // add single bit
-                round_key[7] ^= 1UL << (7 * 8);
-
-                // add round constants
-                round_key[3] ^= ((round_constant[round] >> 0) & 0x1) << (0 * 8);
-                round_key[7] ^= ((round_constant[round] >> 1) & 0x1) << (0 * 8);
-                round_key[3] ^= ((round_constant[round] >> 2) & 0x1) << (1 * 8);
-                round_key[7] ^= ((round_constant[round] >> 3) & 0x1) << (1 * 8);
-                round_key[3] ^= ((round_constant[round] >> 4) & 0x1) << (2 * 8);
-                round_key[7] ^= ((round_constant[round] >> 5) & 0x1) << (2 * 8);
-
-                // extend bits to bytes
-                for (size_t i = 0; i < 8; i++) {
-                        round_key[i] |= round_key[i] << 1;
-                        round_key[i] |= round_key[i] << 2;
-                        round_key[i] |= round_key[i] << 4;
-                }
-
-                // load into vector registers (upper and lower part are same)
-                round_keys[round][0].val[0] = vdupq_n_u64(round_key[0]);
-                round_keys[round][0].val[1] = vdupq_n_u64(round_key[1]);
-                round_keys[round][0].val[2] = vdupq_n_u64(round_key[2]);
-                round_keys[round][0].val[3] = vdupq_n_u64(round_key[3]);
-                round_keys[round][1].val[0] = vdupq_n_u64(round_key[4]);
-                round_keys[round][1].val[1] = vdupq_n_u64(round_key[5]);
-                round_keys[round][1].val[2] = vdupq_n_u64(round_key[6]);
-                round_keys[round][1].val[3] = vdupq_n_u64(round_key[7]);
-
-                // update key state
-                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
-                int k1 = (key_state[0] >> 16) & 0xffffUL;
-                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
-                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
-                key_state[0] >>= 32;
-                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
-                key_state[1] >>= 32;
-                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
-        }
-}
-
-void gift_64_vec_sliced_init(void)
-{
-        // permutations
-        perm[0] = vld1q_u8_x4((uint8_t*)&perm_u64[0]);
-        perm[1] = vld1q_u8_x4((uint8_t*)&perm_u64[8]);
-
-        // inverse permutations
-        perm_inv[0] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[0]);
-        perm_inv[1] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[8]);
-
-        // packing masks
-        pack_mask_0 = vdupq_n_u8(0x55);
-        pack_mask_1 = vdupq_n_u8(0x33);
-        pack_mask_2 = vdupq_n_u8(0x0f);
-}
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&m[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&m[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                gift_64_vec_sliced_subcells(s);
-                gift_64_vec_sliced_permute(s);
-
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&c[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&c[8], s[1]);
-}
-
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&c[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&c[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-
-                gift_64_vec_sliced_permute_inv(s);
-                gift_64_vec_sliced_subcells_inv(s);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&m[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&m[8], s[1]);
-}
-\end{lstlisting}
-
-\chapter*{Acknowledgements}
-
-I want to thank ...
-
-\printbibliography
-
-\end{document}

### 2023-03-23 Paper/chapters/appendix.tex
move chapters into own files

--- 2023-03-18 Paper/paper.tex
@@
-\documentclass[12pt]{report}
-
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{ulem} % underline
-\usepackage{float} % H float specifier
-\usepackage{hyperref} % hyperlinks
-% \usepackage{hyphenat}
-\usepackage{bookmark}
-\usepackage{systeme} % systems of equation
-\usepackage[utf8]{inputenc} % necessary for proper unicode support
-\usepackage{parcolumns} % multiple column parallel typesetting
-\usepackage[english]{babel} % comment this for Angelsächsisch
-\usepackage{graphicx} % including graphics
-\usepackage{sourcecodepro} % best font
-\usepackage{listings} % source code listings
-\usepackage{biblatex} % better citation
-% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry
-
-% umlauts in lstlistings env
-\lstset{
-    literate={ö}{{\"o}}1
-    {ä}{{\"a}}1
-    {ü}{{\"u}}1
-    {Ö}{{\"O}}1
-    {Ä}{{\"A}}1
-    {Ü}{{\"U}}1
-}
-
-\lstset{
-    basicstyle=\scriptsize\ttfamily,
-    breaklines=true
-}
-
-\lstset {
-    frame=single,
-    numbers=left,
-    stepnumber=1,
-    firstnumber=1,
-    numberfirstline=true,
-    texcl=false
-}
-
-\providecommand{\lxor}{\veebar}
-\renewcommand{\proofname}{Beweis.}
-\newcommand{\sref}[1]{\textsuperscript{\ref{#1}}}
-\newcommand{\F}[1]{\mathbb{F}_2^{#1}}
-
-\addbibresource{refs.bib}
-
-\title{
-    \textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\
-    {\footnotesize Bachelorarbeit}
-}
-\author{Bastian Engel}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\chapter*{Abstract}
-
-Lorem ipsum dolor \cite{gift:2017} sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
-nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
-fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-\chapter*{Declaration}
-
-I hereby declare that ...
-
-\tableofcontents
-
-\chapter{Introduction}
-\section{Notation}
-
-\section{Block ciphers}
-
-Securing communication channels between different parties has been a long-term
-subject of study for cryptographers and engineers which is essential to our
-modern world to cope with ever-increasing amounts of devices producing and
-sharing data. The main way to facilitate high-throughput, confidential
-communications nowadays is through the use of symmetric cryptography in which
-two parties share a common secret, called a key, which allows them to encrypt,
-share and subsequently decrypt messages to achieve confidentiality against
-third parties. Ciphers can be divided into two categories; block ciphers, which
-always encrypt fixed-sized messages called blocks, and stream ciphers, which
-continuously provide encryption for an arbitrarily long, constant stream of
-data.
-
-A block cipher can be defined as a bijection between the input block (the
-message) and the output block (the ciphertext). For any block cipher with block
-size $n$, we denote the key-dependent encryption and decryption functions as
-$E_K,D_K:\F{n}\rightarrow \F{n}$. The simplest way to
-characterize this bijection is through a lookup table which yields the highest
-possible performance as each block can be encrypted by one simple lookup
-depending on the key and the message. This is not practical though due to most
-ciphers working with block and key sizes $n,|K|\geq 64$. For a block cipher
-with $n=64,|K|=128$, a space of $2^{64}2^{128}64=2^{198}$ is necessary.
-Considering modern consumer hard disks being able to store data in the order of
-$2^{40}$, it is easy to see that a lookup table is wholly impractical. We
-therefore describe block ciphers algorithmically which opens up possibilities
-for different tradeoffs and security concerns.
-
-\subsection{GIFT}
-
-\texttt{GIFT}\cite{gift:2017}, first presented in the \textit{CHES 2017}
-cryptographic hardware and embedded systems conference, is a lightweight block
-cipher based on a previous design called \texttt|PRESENT|, developed in 2007. Its
-goal is to offer maximum security while being extremely light on resources.
-Modern battery-powered devices like RFID tags or low-latency operations like
-on-the-fly disc encryption present strong hardware and power constraints. GIFT
-aims to be a simple, low-energy cipher suited for these kinds of applications.
-
-\texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
-and \texttt{GIFT-128} working with 128-bit blocks. In both cases, the key is 128
-bits long. The design is a very simple, round-based substitution-permutation
-network (SPN). One round consists in a sequential application of the confusion
-layer by means of 4-bit S-boxes and subsequent diffusion through bit
-permutation. After the bit permutation, a round key is added to the cipher
-state and the single round is complete. \texttt{GIFT-64} uses 28 rounds while
-\texttt{GIFT-128} uses 40 rounds.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/GIFT-64.pdf}
-    \caption{Two rounds of GIFT-64}
-\end{figure}
-
-\subsubsection{Substitution layer}
-
-The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
-16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \texttt{GIFT-128}. The S-box
-$S:\F{4}\rightarrow \F{4}$ is defined as follows:
-
-\[
-    \begin{array}{l|cccccccccccccccc}
-        x & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & a & b & c & d & e & f \\
-        \hline
-        S(x) & 1 & a & 4 & c & 6 & f & 3 & 9 & 2 & d & b & 7 & 5 & 0 & 8 & e
-    \end{array}
-\]
-
-\subsubsection{Permutation layer}
-
-The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
-i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64} and
-\texttt{GIFT-128} can be expressed by:
-
-\begin{align*}
-    P_{64}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+16\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-    P_{128}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+32\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-\end{align*}
-
-\subsubsection{Round key addition}
-
-The last step of each round consists in XORing a round key $R_i$ to the cipher
-state. The new cipher state $s_{i+1}$ after each full round is therefore given
-by
-
-\[
-    s_{i+1}=P(S(s_i))\oplus R_i
-\]
-
-\subsubsection{Round key extraction and key schedule}
-
-Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
-$K=k7||k6||\dots||k0$ denote the $128$-bit key state.
-
-\paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
-state. $u_i$ and $v_i$ are XORed to $r_{4i+1}$ and $r_{4i}$ of the round key
-$R$ respectively.
-
-\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k4||k1||k_0$ from
-the key state. $u_i$ and $v_i$ are XORed to $r_{4i+2}$ and $b_{4i+1}$ of the
-round key $R$ respectively.
-
-In both cases, we additionally XOR a round constant $C=c_5c_4c_3c_2c_1c_0$ to
-bit positions $n-1,23,19,15,11,7,3$. The round constants are generated using a
-6-bit affine linear-feedback shift register and have the following values:\\
-
-\begin{tabular}{r|l}
-    \textbf{Rounds} & \textbf{Constants} \\
-    \hline
-    \textbf{1 - 16} &  \small\texttt{01,03,07,0F,1F,3E,3D,3B,37,2F,1E,3C,39,33,27,0E} \\
-    \textbf{17 - 32} & \small\texttt{1D,3A,35,2B,16,2C,18,30,21,02,05,0B,17,2E,1C,38} \\
-    \textbf{33 - 48} & \small\texttt{31,23,06,0D,1B,36,2D,1A,34,29,12,24,08,11,22,04}
-\end{tabular}\\
-
-The key state is then updated by setting $k_1\leftarrow k_1\ggg 2$,
-$k_0\leftarrow k_0\ggg 12$ and rotating the new state $32$ bits to the right:
-
-\[
-    k_7||k_6||\dots||k_1||k_0\leftarrow k_1\ggg 2||k_0\ggg 12||k_7||k_6||\dots||k_3||k_2
-\]
-
-\subsection{Camellia}
-
-\section{The ARMv8 platform}
-
-With small devices, embedded processors and ASICs becoming ever more ubiquitous
-and essential in areas like medicine or automotive design, the need for ...
-
-\chapter{Implementation strategies}
-
-Due to the structural differences of SPN- and Feistel network-based ciphers, we
-shall analyze these two separately.
-
-\section{Strategies for SPN}
-
-Three implementation strategies for substitution-permutation networks are
-introduced by \cite{implx86:2014}:
-
-\begin{itemize}
-    \item Table-based implementations
-    \item \texttt{vperm} implementations
-    \item Bitslice implementations
-\end{itemize}
-
-\subsection{Table-based}
-
-Table-driven programming is a simple way to increase performance of operations
-by tabulating the results, therefore requiring only a single memory access to
-acquire the result. This approach is obviously limited to manageable table
-sizes, so while tabulating a function like the AES S-box
-$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space,
-tabulating the \texttt{GIFT} permutation layer
-$P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
-$2^{70}$ space, which is totally unfeasible.
-
-A common approach is to tabulate the output of each S-box, including the
-diffusion layer, and then XORing the results together. Let $n$ denote the
-internal cipher state size and $s$ the size of a single S-box in bits. For each
-S-box $S_i,i\in\{0,\dots,\frac{n}{s}\}$, we can construct a mapping
-$T_i:\F{s}\rightarrow \F{n}$ representing substitution with subsequent
-permutation of that single S-box. The cipher state before round key addition is
-then given by $\bigoplus_{i=0}^{\frac{n}{s}-1}{T_i(m_i)}$ for each $s$-bit
-message chunk $m_i$. This approach requires space of
-$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for \texttt{GIFT-64},
-results in a manageable size of $\frac{64^2 2^4}{4}=2^{14}$ bits which equals
-$16$ KiB.
-
-\subsubsection{Constructing the tables}
-
-For \texttt{GIFT-64}, table construction is relatively straightforward and can
-be done as follows:
-
-\begin{lstlisting}[caption={Table construction algorithm}, escapechar=|]
-    tables <- [][]
-    for sbox_index from 0 to 15 do
-        for sbox_input from 0 to 15 do|\label{lst:tablesbox}|
-            output <- sbox(sbox_input)
-            output <- permute(output << (4 * sbox_index))
-            tables[sbox_index][sbox_input] <- output
-\end{lstlisting}
-
-Implementing this algorithm gives us the following table representing the first
-and second S-box.
-
-\[
-    \begin{array}{l|l|l|c}
-        x & T_0(x) & T_1(x) & \dots \\
-        \hline
-        0x0 & 0x1               & 0x1000000000000   & \dots \\
-        0x1 & 0x8000000020000   & 0x800000002       & \dots \\
-        0x2 & 0x400000000       & 0x40000           & \dots \\
-        0x3 & 0x8000400000000   & 0x800040000       & \dots \\
-        0x4 & 0x400020000       & 0x40002           & \dots \\
-        0x5 & 0x8000400020001   & 0x1000800040002   & \dots \\
-        0x6 & 0x20001           & 0x1000000000002   & \dots \\
-        0x7 & 0x8000000000001   & 0x1000800000000   & \dots \\
-        0x8 & 0x20000           & 0x2               & \dots \\
-        0x9 & 0x8000400000001   & 0x1000800040000   & \dots \\
-        0xa & 0x8000000020001   & 0x1000800000002   & \dots \\
-        0xb & 0x400020001       & 0x1000000040002   & \dots \\
-        0xc & 0x400000001       & 0x1000000040000   & \dots \\
-        0xd & 0x0               & 0x0               & \dots \\
-        0xe & 0x8000000000000   & 0x800000000       & \dots \\
-        0xf & 0x8000400020000   & 0x800040002       & \dots
-    \end{array}
-\]
-
-The tables for \texttt{GIFT-128} can be generated in a similar way by looping
-through all 32 S-boxes instead of 16 on line \ref{lst:tablesbox}.
-
-\subsection{Using \texttt{vperm}}
-
-Nowadays, most instructions set architectures support single-instruction,
-multiple-data processing. The idea of such an SIMD system is to work on
-multiple data stored in vectors at once to speed up calculations. For A64, two
-types of vector processing are available:
-
-\begin{enumerate}
-    \item Advanced SIMD, known as NEON
-    \item Scalable Vector Extension (SVE)
-\end{enumerate}
-
-We will take a look at NEON as this is the type of vector processing supported
-by the Cortex-A73 processor.
-
-\subsubsection{ARM Neon}
-
-The register file of the NEON unit is made up of 32 quad-word (128-bit)
-registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
-registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
-sized lanes on which the vector instructions operate. Valid ways to interpret
-for example the register \texttt{V0} are:
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/V_register.pdf}
-    \caption{Divisions of the V0 register}
-\end{figure}
-
-NEON instructions interpret their operands' layouts (i.e. lane count and width)
-through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
-adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
-together and storing the result in \texttt{V0} can be done as follows:
-
-\begin{center}
-    \texttt{ADD V0.8H, V1.8H, V2.8H}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/vector_add.pdf}
-    \caption{Addition of two vector registers}
-\end{figure}
-
-The plenitude of different processing instructions allow flexible ways to
-further speed up algorithms having reached their optimizational limit on
-non-SIMD platforms. \texttt{vperm}, a general term standing for \textit{vector
-permute}, is a common instruction on SIMD machines. Called \texttt{TBL} on
-NEON, it is used for parallel table lookups and arbitrary permutations. It takes
-two inputs to perform a lanewise lookup:
-
-\begin{enumerate}
-    \item A register with lookup values
-    \item Two or more registers containing data
-\end{enumerate}
-
-\subsubsection{S-box lookup}
-
-This instruction can be used to implement S-box lookup of all 16 S-boxes in a
-single instruction. We do this by packing our 64-bit cipher state
-$s=s_{15}||s_{14}||\dots||s_0$ into a vector register $V_0$. Because we can
-only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane which
-neatly fits into the 128-bit registers. We then put the S-box itself into
-register $V_1$ which will be used as the data register for the table lookup.
-
-The confusion layer can now be performed through one \texttt{TBL} instruction:
-
-\begin{center}
-    \texttt{TBL V0.16B, V1.16B, V0.16B}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/tbl_example.pdf}
-    \caption{Performing the S-Box lookup in parallel}
-\end{figure}
-
-\subsection{Bitslicing}
-
-Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
-achieve a more efficient representation to operate on. The structure of
-\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
-cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
-i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
-individual S-boxes. This is equivalent to transposing the bit matrix.
-
-\[
-    S=\begin{bmatrix}
-        S_0\\
-        S_1\\
-        S_2\\
-        S_3
-    \end{bmatrix}
-    =\begin{bmatrix}
-        b_{60}b_{56}b_{52}\dots b_0\\
-        b_{61}b_{57}b_{53}\dots b_1\\
-        b_{62}b_{58}b_{54}\dots b_2\\
-        b_{63}b_{59}b_{55}\dots b_3
-    \end{bmatrix}
-\]
-
-\subsubsection{Parallel S-Boxes}
-
-This representation offers multiple advantages. We first note that computation
-of the S-box can be executed in parallel, similar to the \texttt{vperm}
-technique above. This can be done by finding an algorithmic way to apply the
-S-box which has already been proposed by the original \texttt{GIFT} authors:
-
-\begin{align*}
-    S_1&\leftarrow S_1\oplus (S_0\land S_2) \\
-    t&\leftarrow S_0\oplus (S_1\land S_3) \\
-    S_2&\leftarrow S_2\oplus (t\lor S_1) \\
-    S_0&\leftarrow S_3\oplus S_2 \\
-    S_1&\leftarrow S_1\oplus S_0 \\
-    S_0&\leftarrow \lnot S_0 \\
-    S_2&\leftarrow S_2\oplus (t\land S_1) \\
-    S_3&\leftarrow t
-\end{align*}
-
-This is very efficient as it only requires six XOR-, three AND and one OR
-operation.
-
-An important property of the permutation is the fact that bits always stay in
-their slice. This means we can decompose the permutation $P$ into four
-permutations $P_i,i\in\{0,1,2,3\}$ and apply these permutations
-separately to each slice. One possible way to implement a permutation $P_i$ in
-software is to mask off all bits individually, shift them to their correct
-position and OR them together:
-
-\[
-    P_i(S_i)=\bigvee_{k=0}^{15}{(S_i\land m_i) \ll s_i}
-\]
-
-This approach requires $47$ operations, meaning all four permutations require
-over $150$ operations which would present a major bottleneck to the round
-function. We can improve on this by working on multiple message blocks at once
-and using the aforementioned \texttt{vperm} instruction to implement the bit
-shuffling. We then need only four \texttt{TBL} instructions for the complete
-diffusion layer.
-
-\subsubsection{Using \texttt{vperm} for slice permutation}
-
-We cannot use the \texttt{TBL} instruction directly as we need to shuffle
-individual bits, but the smallest data we can operate on are bytes. We
-therefore encrypt $8n$ messages at once which allows us to create bytewise
-groupings. These messages are put into $4m$ registers with register $R_{4i}$
-containing $S_0$, register $R_{4i+1}$ containing $S_1$ and so forth. With block
-size $BS$ and register size $RS$, the following must hold:
-
-\[
-    8n\cdot BS=4m\cdot RS
-\]
-
-In the case of \texttt{GIFT-64} with $BS=64$ and ARM NEON with $RS=128$, we get
-
-\[
-    8n\cdot 64=4m\cdot 128\Leftrightarrow n=m
-\]
-
-$n=m=1$ would be a valid choice which yields eight messages divided into four
-registers. We choose $n=m=2$ so we can directly utilize the algorithm for bit
-packing presented by the original GIFT authors, although it is simple to adapt
-this algorithm to only four registers and eight messages by adjusting the
-\texttt{SWAPMOVE} shift and mask values.
-
-\subsubsection{Packing the data into bitslice format}
-
-Let $a,b,\dots,p$ be sixteen messages of length $64$ with subscripts denoting
-individual bits. We first put these messages into eight SIMD registers
-$V_0,V_1,\dots,V_7$:
-
-\begin{alignat*}{3}
-    V_0&=b||a\qquad V_4&&=j||&&i \\
-    V_1&=d||c\qquad V_5&&=l||&&k \\
-    V_2&=f||e\qquad V_6&&=n||&&m \\
-    V_3&=h||g\qquad V_7&&=p||&&o
-\end{alignat*}
-
-We then use the \texttt{SWAPMOVE} technique to bring the data into bitslice
-format. This operation operates on two registers $A,B$ using mask $M$ and shift
-value $N$. It swaps bits in $A$ masked by $(M\ll N)$ with bits in $B$ masked by
-$M$ in using only three XOR-, one AND- and two shift operations.
-
-\begin{align*}
-    &\text{SWAPMOVE}(A,B,M,N): \\
-    &\qquad T=((A\gg N)\oplus B)\land M \\
-    &\qquad B=B\oplus T \\
-    &\qquad A=A\oplus (T\ll N) \\
-\end{align*}
-
-One caveat of this approach is the fact that NEON registers cannot be shifted
-in their entirety due to the fact bits are not able to cross lanes. This leads
-to the problem of being able to shift at most two lanes of 64 bits at once. We
-thus need to implement the \texttt{shr(V,n)} and \texttt{shl(V,n)} operations
-on our own. This can be done by first extracting the 64-bit lanes $a,b$ out of
-$V=b||a$, shifting the lanes individually and finally shifting and ORing the
-crossing bits back into the other lane.
-
-\begin{alignat*}{2}
-    &\text{shl}(V,n): \\
-    &\qquad a,b&&=V[0],V[1] \\
-    &\qquad c&&=(a\gg (64-n)) \\
-    &\qquad a&&=(a\ll n) \\
-    &\qquad b&&=(b\ll n)\lor c \\
-    &\qquad V[0],V[1]&&=a,b
-\end{alignat*}
-
-The following operations group all $i$th bits of the messages $a,c,\dots,o$
-into bytes and puts these into the lower half of the registers $V_{i\bmod 8}$.
-The same is done for messages $b,d,\dots,p$, only differing in that the bytes
-are put into the upper half of the registers.
-
-\begin{align*}
-    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_4,V_5,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_6,V_7,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_4,V_6,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_5,V_7,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_0,V_4,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_1,V_5,0x0f0f\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4) \\
-\end{align*}
-
-With $Ax=o_xm_xk_xj_xg_xe_xc_xa_x$ and $Bx=p_xn_xl_xi_xh_xf_xd_xb_x$ denoting
-byte groups, our data now has the following permutation-friendly format:
-
-\[
-    \begin{array}{c|llllllll}
-        n & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
-        \hline
-        V_0 & A56 & A48 & A40 & A32 & A24 & A16 & A8 & A0 \\
-        V_1 & A57 & A49 & A41 & A33 & A25 & A17 & A9 & A1 \\
-        V_2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
-        V_3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
-        V_4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
-        V_5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
-        V_6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
-        V_7 & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
-    \end{array}
-\]
-
-\[
-    \begin{array}{c|llllllll}
-        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 \\
-        \hline
-        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8 & B0 \\
-        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9 & B1 \\
-        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 \\
-        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 \\
-        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 \\
-        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 \\
-        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 \\
-        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7
-    \end{array}
-\]
-
-We can now create permutation tables using the specification of the individual
-slice permutations $P_i$:
-
-\[
-    \begin{array}{c|llllllllllllllll}
-        j & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\
-        \hline
-        P_0(j) & 0 & 12 & 8 & 4 & 1 & 13 & 9 & 5 & 2 & 14 & 10 & 6 & 3 & 15 & 11 & 7 \\
-        P_1(j) & 4 & 0 & 12 & 8 & 5 & 1 & 13 & 9 & 6 & 2 & 14 & 10 & 7 & 3 & 15 & 11 \\
-        P_2(j) & 8 & 4 & 0 & 12 & 9 & 5 & 1 & 13 & 10 & 6 & 2 & 14 & 11 & 7 & 3 & 15 \\
-        P_3(j) & 12 & 8 & 4 & 0 & 13 & 9 & 5 & 1 & 14 & 10 & 6 & 2 & 15 & 11 & 7 & 3
-    \end{array}
-\]
-
-One thing to take note of is the original permutation values only showing where
-a given byte should land, not which byte belongs to a certain position - i.e.
-for $P_0$, byte 1 should land in position 12, but the byte belonging to
-position 1 is byte 4. Because \texttt{vperm} works in the latter way, we have
-to do some trivial rearrangements. In addition, because data for each slice is
-split between two registers $(V_0,V_4),(V_1,V_5),\dots$, the \texttt{vperm}
-operation needs to use these two registers as a data source. Using two data
-sources is supported by the \texttt{TBL} instruction by means of an additional
-parameter.
-
-Assuming the correct permutation values are put into registers
-$V_9,V_{10},\dots,V_{16}$, this now allows us to compute the permutation layer for
-all 16 blocks in only eight permutation instructions plus four additional vector
-copy instructions used for saving the original values of $V_0,V_1,V_2,V_3$
-before the permutation is applied.
-
-\begin{alignat*}{2}
-    &\texttt{MOV V17, V0}\qquad &&\texttt{MOV V18, V1} \\
-    &\texttt{MOV V19, V2}\qquad &&\texttt{MOV V20, V3} \\
-    &\texttt{TBL V0, \{V0, V4\}, V9}\qquad &&\texttt{TBL V1, \{V1, V5\}, V10} \\
-    &\texttt{TBL V2, \{V2, V6\}, V9}\qquad &&\texttt{TBL V3, \{V3, V7\}, V10} \\
-    &\texttt{TBL V4, \{V17, V4\}, V9}\qquad &&\texttt{TBL V5, \{V18, V5\}, V10} \\
-    &\texttt{TBL V6, \{V19, V6\}, V9}\qquad &&\texttt{TBL V7, \{V20, V7\}, V10}
-\end{alignat*}
-
-\subsubsection{Round key function}
-
-In contrast to packing and unpacking of data which is only done once in the
-beginning and end, a round key is derived for every round, so the round key
-derivation function needs to be as fast as possible. A simple but naive
-approach for one round would be to generate a single round key, copy it 15
-times and pack the resulting registers similar to how we proceed with the
-messages. Due to the cost of packing the messages, this is prohibitivly
-expensive. Because we know where each byte group ends up after packing, we can
-directly set the round key bits at the correct position. Extending these bits
-to bytes can then be done simply by repeatedly shifting and ORing the registers
-together.
-
-\subsection{Strategies for Camellia}
-
-\chapter{Implementation}
-
-We will provide and discuss implementations for the presented strategies in the
-C programming language. Although directly writing Assembler code could result
-in a small performance benefit, this generally increases the work necessary by
-an order of magnitude for only limited results. This is why we use NEON
-intrinsics.
-
-\subsection{NEON intrinsics}
-
-The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
-definitions including vector data types and C functions for working with these
-vectors. These functions are known as NEON intrinsics and give the programmer a
-high-level interface to most NEON instructions. Major advantages of this
-approach include the ease of development as the compiler takes over register
-allocation and load/store operations as well as performance benefits through
-compiler optimizations.
-
-Standard vector data types have the format \texttt{uintnxm\_t} with lane width
-$n$ in bits and and lane count $m$. Array types of the format
-\texttt{uintnxmxc\_t}, $c\in\{1,2,3,4\}$ are also defined which are used in
-operations requiring multiple parameters like \texttt{TBL} or pairwise
-load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
-after the operation name to indicate operation on a 128-bit register and end
-with the lane data format. Multiplying eight pairs of 16-bit numbers \texttt{a,b} for
-example can be done via the following:
-
-\begin{center}
-    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
-\end{center}
-
-The compiler allocates vector registers for \texttt{a}, \texttt{b} and
-\texttt{result} and assembles the intrinsic to \texttt{MUL Vr.8H, Va.8H,
-Vb.8H}. Necessary load and stores for the result and parameters are also
-handled automatically.
-
-\chapter{Evaluation}
-
 \chapter{Appendix}
@@
 \end{lstlisting}
-
-\chapter*{Acknowledgements}
-
-I want to thank ...
-
-\printbibliography
-
-\end{document}

### 2023-03-23 Paper/chapters/implementation-strategies.tex
move chapters into own files

--- 2023-03-20 Paper/paper.tex
@@
-\documentclass[12pt]{report}
-
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{ulem} % underline
-\usepackage{float} % H float specifier
-\usepackage{hyperref} % hyperlinks
-\usepackage{tabularx} % tabular with column specifier X to break lines
-% \usepackage{hyphenat}
-\usepackage{bookmark}
-\usepackage{systeme} % systems of equation
-\usepackage[utf8]{inputenc} % necessary for proper unicode support
-\usepackage{parcolumns} % multiple column parallel typesetting
-\usepackage[english]{babel} % comment this for Angelsächsisch
-\usepackage{graphicx} % including graphics
-\usepackage{sourcecodepro} % best font
-\usepackage{listings} % source code listings
-\usepackage{biblatex} % better citation
-% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry
-
-% umlauts in lstlistings env
-\lstset{
-    literate={ö}{{\"o}}1
-    {ä}{{\"a}}1
-    {ü}{{\"u}}1
-    {Ö}{{\"O}}1
-    {Ä}{{\"A}}1
-    {Ü}{{\"U}}1
-}
-
-\lstset{
-    basicstyle=\scriptsize\ttfamily,
-    breaklines=true
-}
-
-\lstset {
-    frame=single,
-    numbers=left,
-    stepnumber=1,
-    firstnumber=1,
-    numberfirstline=true,
-    texcl=false
-}
-
-\providecommand{\lxor}{\veebar}
-\renewcommand{\proofname}{Beweis.}
-\newcommand{\sref}[1]{\textsuperscript{\ref{#1}}}
-\newcommand{\F}[1]{\mathbb{F}_2^{#1}}
-
-\addbibresource{refs.bib}
-
-\title{
-    \textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\
-    {\footnotesize Bachelorarbeit}
-}
-\author{Bastian Engel}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\chapter*{Abstract}
-
-Lorem ipsum dolor \cite{gift:2017} sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
-nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
-fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-\chapter*{Declaration}
-
-I hereby declare that ...
-
-\tableofcontents
-
-\chapter{Introduction}
-\section{Notation}
-
-\section{Block ciphers}
-
-Securing communication channels between different parties has been a long-term
-subject of study for cryptographers and engineers which is essential to our
-modern world to cope with ever-increasing amounts of devices producing and
-sharing data. The main way to facilitate high-throughput, confidential
-communications nowadays is through the use of symmetric cryptography in which
-two parties share a common secret, called a key, which allows them to encrypt,
-share and subsequently decrypt messages to achieve confidentiality against
-third parties. Ciphers can be divided into two categories; block ciphers, which
-always encrypt fixed-sized messages called blocks, and stream ciphers, which
-continuously provide encryption for an arbitrarily long, constant stream of
-data.
-
-A block cipher can be defined as a bijection between the input block (the
-message) and the output block (the ciphertext). For any block cipher with block
-size $n$, we denote the key-dependent encryption and decryption functions as
-$E_K,D_K:\F{n}\rightarrow \F{n}$. The simplest way to
-characterize this bijection is through a lookup table which yields the highest
-possible performance as each block can be encrypted by one simple lookup
-depending on the key and the message. This is not practical though due to most
-ciphers working with block and key sizes $n,|K|\geq 64$. For a block cipher
-with $n=64,|K|=128$, a space of $2^{64}2^{128}64=2^{198}$ is necessary.
-Considering modern consumer hard disks being able to store data in the order of
-$2^{40}$, it is easy to see that a lookup table is wholly impractical. We
-therefore describe block ciphers algorithmically which opens up possibilities
-for different tradeoffs and security concerns.
-
-\subsection{GIFT}
-
-\texttt{GIFT}\cite{gift:2017}, first presented in the \textit{CHES 2017}
-cryptographic hardware and embedded systems conference, is a lightweight block
-cipher based on a previous design called \texttt{PRESENT}, developed in 2007. Its
-goal is to offer maximum security while being extremely light on resources.
-Modern battery-powered devices like RFID tags or low-latency operations like
-on-the-fly disc encryption present strong hardware and power constraints. GIFT
-aims to be a simple, low-energy cipher suited for these kinds of applications.
-
-\texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
-and \texttt{GIFT-128} working with 128-bit blocks. In both cases, the key is 128
-bits long. The design is a very simple, round-based substitution-permutation
-network (SPN). One round consists in a sequential application of the confusion
-layer by means of 4-bit S-boxes and subsequent diffusion through bit
-permutation. After the bit permutation, a round key is added to the cipher
-state and the single round is complete. \texttt{GIFT-64} uses 28 rounds while
-\texttt{GIFT-128} uses 40 rounds.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/GIFT-64.pdf}
-    \caption{Two rounds of GIFT-64}
-\end{figure}
-
-\subsubsection{Substitution layer}
-
-The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
-16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \texttt{GIFT-128}. The S-box
-$S:\F{4}\rightarrow \F{4}$ is defined as follows:
-
-\[
-    \begin{array}{l|cccccccccccccccc}
-        x & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & a & b & c & d & e & f \\
-        \hline
-        S(x) & 1 & a & 4 & c & 6 & f & 3 & 9 & 2 & d & b & 7 & 5 & 0 & 8 & e
-    \end{array}
-\]
-
-\subsubsection{Permutation layer}
-
-The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
-i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64} and
-\texttt{GIFT-128} can be expressed by:
-
-\begin{align*}
-    P_{64}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+16\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-    P_{128}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+32\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-\end{align*}
-
-\subsubsection{Round key addition}
-
-The last step of each round consists in XORing a round key $R_i$ to the cipher
-state. The new cipher state $s_{i+1}$ after each full round is therefore given
-by
-
-\[
-    s_{i+1}=P(S(s_i))\oplus R_i
-\]
-
-\subsubsection{Round key extraction and key schedule}
-
-Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
-$K=k7||k6||\dots||k0$ denote the $128$-bit key state.
-
-\paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
-state. $u_i$ and $v_i$ are XORed to $r_{4i+1}$ and $r_{4i}$ of the round key
-$R$ respectively.
-
-\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k4||k1||k_0$ from
-the key state. $u_i$ and $v_i$ are XORed to $r_{4i+2}$ and $b_{4i+1}$ of the
-round key $R$ respectively.
-
-In both cases, we additionally XOR a round constant $C=c_5c_4c_3c_2c_1c_0$ to
-bit positions $n-1,23,19,15,11,7,3$. The round constants are generated using a
-6-bit affine linear-feedback shift register and have the following values:\\
-
-\begin{tabular}{r|l}
-    \textbf{Rounds} & \textbf{Constants} \\
-    \hline
-    \textbf{1 - 16} &  \small\texttt{01,03,07,0F,1F,3E,3D,3B,37,2F,1E,3C,39,33,27,0E} \\
-    \textbf{17 - 32} & \small\texttt{1D,3A,35,2B,16,2C,18,30,21,02,05,0B,17,2E,1C,38} \\
-    \textbf{33 - 48} & \small\texttt{31,23,06,0D,1B,36,2D,1A,34,29,12,24,08,11,22,04}
-\end{tabular}\\
-
-The key state is then updated by setting $k_1\leftarrow k_1\ggg 2$,
-$k_0\leftarrow k_0\ggg 12$ and rotating the new state $32$ bits to the right:
-
-\[
-    k_7||k_6||\dots||k_1||k_0\leftarrow k_1\ggg 2||k_0\ggg 12||k_7||k_6||\dots||k_3||k_2
-\]
-
-\subsection{Camellia}
-
-\section{The ARMv8 platform}
-
-With small devices, embedded processors and ASICs becoming ever more ubiquitous
-and essential in areas like medicine or automotive design, the need for ...
-
 \chapter{Implementation strategies}
@@
 superfluous load/stores. This transformation is trivial by use of
-\texttt{TBL} with two data source operands.
+\texttt{TBL} with two data source operands. The final data format we operate on
+is as follows:
+
+\[
+    \scriptsize
+    \begin{array}{c|llllllll|llllllll}
+        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
+        \hline
+        V_0 & A60 & A56 & A52 & A48 & A44 & A40 & A36 & A32 & A28 & A24 & A20 & A16 & A12 & A8  & A4 & A0 \\
+        V_1 & A61 & A57 & A53 & A49 & A45 & A41 & A37 & A33 & A29 & A25 & A21 & A17 & A13 & A9  & A5 & A1 \\
+        V_2 & A62 & A58 & A54 & A50 & A46 & A42 & A38 & A34 & A30 & A26 & A22 & A18 & A14 & A10 & A6 & A2 \\
+        V_3 & A63 & A59 & A55 & A51 & A47 & A43 & A39 & A35 & A31 & A27 & A23 & A19 & A15 & A11 & A7 & A3 \\
+        V_4 & B60 & B56 & B52 & B48 & B44 & B40 & B36 & B32 & B28 & B24 & B20 & B16 & B12 & B8  & B4 & B0 \\
+        V_5 & B61 & B57 & B53 & B49 & B45 & B41 & B37 & B33 & B29 & B25 & B21 & B17 & B13 & B9  & B5 & B1 \\
+        V_6 & B62 & B58 & B54 & B50 & B46 & B42 & B38 & B34 & B30 & B26 & B22 & B18 & B14 & B10 & B6 & B2 \\
+        V_7 & B63 & B59 & B55 & B51 & B47 & B43 & B39 & B35 & B31 & B27 & B23 & B19 & B15 & B11 & B7 & B3
+    \end{array}
+\]
 
 We can now create permutation tables using the specification of the individual
-slice permutations $P_i$:
+slice permutations $P_i$ which are then applied to $V_i$ and $V_{i+4}$:
 
@@
 
-Another thing to take note of is the original permutation values only showing
-where a given byte should land, not which byte belongs to a certain position -
-i.e. for $P_0$, byte 1 should land in position 12, but the byte belonging to
-position 1 is byte 4. Because \texttt{TBL} works in the latter way, we have
-to do some trivial rearrangements.
+One thing to take note of is the original permutation values only show where a
+given byte should land, not which byte belongs to a certain position - i.e. for
+$P_0$, byte 1 should land in position 12, but the byte belonging to position 1
+is byte 4. Because \texttt{TBL} works in the latter way, we have to do some
+trivial rearrangements.
 
 Assuming the correct permutation values are put into registers
-$V_8,V_{10},\dots,V_{15}$, this now allows us to compute the permutation layer
+$V_8,V_9,V_{10},V_{11}$, this now allows us to compute the permutation layer
 for all 16 blocks in only eight permutation instructions.
@@
     &\texttt{TBL V0, V0, V8}\qquad &&\texttt{TBL V1, V1, V9} \\
+    &\texttt{TBL V4, V4, V8}\qquad &&\texttt{TBL V5, V5, V9} \\
     &\texttt{TBL V2, V2, V10}\qquad &&\texttt{TBL V3, V3, V11} \\
-    &\texttt{TBL V4, V4, V12}\qquad &&\texttt{TBL V5, V5, V13} \\
-    &\texttt{TBL V6, V6, V14}\qquad &&\texttt{TBL V7, V7, V15}
+    &\texttt{TBL V6, V6, V10}\qquad &&\texttt{TBL V7, V7, V11}
 \end{alignat*}
@@
 expensive. Because we know where each byte group ends up after packing, we can
-directly set the round key bits at the correct position. Extending these bits
+directly XOR the round key bits to the correct position. Extending these bits
 to bytes can then be done simply by repeatedly shifting and ORing the registers
@@
 \subsection{Strategies for Camellia}
-
-\chapter{Implementation}
-
-We will provide and discuss implementations for the presented strategies in the
-C programming language. Although directly writing Assembler code could result
-in a small performance benefit, this generally increases the work necessary by
-an order of magnitude for only limited results. This is why we use NEON
-intrinsics.
-
-\subsection{NEON intrinsics}
-
-The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
-definitions including vector data types and C functions for working with these
-vectors. These functions are known as NEON intrinsics and give the programmer a
-high-level interface to most NEON instructions. Major advantages of this
-approach include the ease of development as the compiler takes over register
-allocation and load/store operations as well as performance benefits through
-compiler optimizations.
-
-Standard vector data types have the format \texttt{uintnxm\_t} with lane width
-$n$ in bits and and lane count $m$. Array types of the format
-\texttt{uintnxmxc\_t}, $c\in\{2,3,4\}$ are also defined which are used in
-operations requiring multiple parameters like \texttt{TBL} or pairwise
-load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
-after the operation name to indicate operation on a 128-bit register, and end
-with the lane data format. Multiplying eight pairs of 16-bit numbers
-\texttt{a,b} for example can be done via the following:
-
-\begin{center}
-    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
-\end{center}
-
-In this case, the compiler allocates vector registers for \texttt{a},
-\texttt{b} and \texttt{result} and assembles the intrinsic to \texttt{MUL
-Vr.8H, Va.8H, Vb.8H}. Necessary load and stores for the result and parameters
-are also handled automatically. Of special interest to us are the following
-intrinsics, each existing in different variants with different lane widths and
-also array types: \\
-
-\begin{table}[h!]
-    \centering
-    \footnotesize
-    \begin{tabularx}{\textwidth}{ll|X}
-        Intrinsic && Description \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vreinterpretq\_u8\_u64(uint64x2\_t)} & Explicit casting \\
-        \hline
-        \texttt{uint64\_t} & \texttt{vgetq\_lane\_u64(void)} & Extract a single lane \\
-        \hline
-        \texttt{void} & \texttt{vsetq\_lane\_u64(uint64\_t)} & Insert a single lane \\
-        \hline
-        \texttt{uint64x2\_t} & \texttt{vdupq\_n\_u64(uint64\_t)} & Initialize all lanes to same value \\
-        \hline
-        \texttt{void} & \texttt{vst1q\_u64(uint64\_t*, uint64x2\_t)} & Store from register to memory \\
-        \hline
-        \texttt{uint64x2\_t} & \texttt{vld1q\_u64(uint64\_t*, uint64x2\_t)} & Load from memory to register \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{veorq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise XOR \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vandq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise AND \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vorrq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise OR \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vmvnq\_u8(uint8x16\_t)} & bitwise NOT \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
-    \end{tabularx}
-\end{table}
-
-The implementations closely follow the aforementioned strategies;
-instruction-level optimization is left to the compiler.
-
-\chapter{Evaluation}
-
-\chapter{Appendix}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.h}]
-#ifndef GIFT_VEC_SLICED_H
-#define GIFT_VEC_SLICED_H
-
-#include <stdint.h>
-#include <arm_neon.h>
-
-#define ROUNDS_GIFT_64 28
-
-// expose for benchmarking
-uint8x16_t shl(uint8x16_t v, int n);
-uint8x16_t shr(uint8x16_t v, int n);
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
-                                 uint8x16_t m, int n);
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2]);
-
-void gift_64_vec_sliced_init(void);
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2]);
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2]);
-
-#endif
-\end{lstlisting}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.c}]
-#include "gift_vec_sliced.h"
-
-#include <stddef.h>
-#include <stdio.h>
-#include <string.h>
-
-static uint64_t perm_u64[16] = {
-        0x1410050115110400UL, 0x1410050115110400UL,
-        0x0501151104001410UL, 0x0501151104001410UL,
-        0x1511040014100501UL, 0x1511040014100501UL,
-        0x0400141005011511UL, 0x0400141005011511UL,
-        0x1612070317130602UL, 0x1612070317130602UL,
-        0x0703171306021612UL, 0x0703171306021612UL,
-        0x1713060216120703UL, 0x1713060216120703UL,
-        0x0602161207031713UL, 0x0602161207031713UL
-};
-
-static uint64_t perm_inv_u64[16] = {
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL
-};
-
-static uint8x16x4_t perm[2];
-static uint8x16x4_t perm_inv[2];
-
-static uint8x16_t pack_mask_0;
-static uint8x16_t pack_mask_1;
-static uint8x16_t pack_mask_2;
-
-static const int round_constant[] = {
-        // rounds 0-15
-        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
-        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
-        // rounds 16-31
-        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
-        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
-        // rounds 32-47
-        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
-        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
-};
-
-uint8x16_t shl(uint8x16_t v, int n)
-{
-        uint64_t l[2];
-        vst1q_u64(l, v);
-        l[1] = l[1] << n | (l[0] >> (64 - n));
-        l[1] <<= n;
-        return vreinterpretq_u8_u64(vld1q_u64(l));
-}
-
-uint8x16_t shr(uint8x16_t v, int n)
-{
-        uint64_t l[2];
-        vst1q_u64(l, v);
-        l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
-        l[1] >>= n;
-        return vreinterpretq_u8_u64(vld1q_u64(l));
-}
-
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
-{
-
-        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
-        *b = veorq_u8(*b, t);
-        *a = veorq_u8(*a, shl(t, n));
-}
-
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
-{
-        // take care not to shift mask bits out of the register
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[0].val[3], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[2], &m[1].val[3], pack_mask_0, 1);
-
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[0].val[3], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[1], &m[1].val[3], pack_mask_1, 2);
-
-        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[1].val[0], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[1].val[2], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[1].val[1], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[3], &m[1].val[3], pack_mask_2, 4);
-}
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
-{
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[3] = t;
-
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-        t            = veorq_u8(cs[1].val[0], vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(cs[1].val[3], cs[1].val[2]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[3] = t;
-}
-
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16_t t = cs[0].val[3];
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[3] = veorq_u8(cs[0].val[0], cs[0].val[2]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(t, vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-
-        t            = cs[1].val[3];
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[3] = veorq_u8(cs[1].val[0], cs[1].val[2]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(t, vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-}
-
-void gift_64_vec_sliced_permute(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm[1].val[3]);
-}
-
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm_inv[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm_inv[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm_inv[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm_inv[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm_inv[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm_inv[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm_inv[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm_inv[1].val[3]);
-}
-
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2])
-{
-        // TODO shift-and-or
-
-        uint64_t key_state[] = {key[0], key[1]};
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                int v = (key_state[0] >> 0 ) & 0xffff;
-                int u = (key_state[0] >> 16) & 0xffff;
-
-                // add round key (RK=U||V)
-                uint64_t round_key[8] = { 0x0UL };
-                for (size_t i = 0; i < 8; i++) {
-                        int key_bit_v   = (v >> (2 * i + 0)) & 0x1;
-                        int key_bit_u   = (u >> (2 * i + 0)) & 0x1;
-                        round_key[0] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[1] ^= (uint64_t)key_bit_u << (i * 8);
-
-                        key_bit_v   = (v >> (2 * i + 1)) & 0x1;
-                        key_bit_u   = (u >> (2 * i + 1)) & 0x1;
-                        round_key[4] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[5] ^= (uint64_t)key_bit_u << (i * 8);
-                }
-
-                // add single bit
-                round_key[7] ^= 1UL << (7 * 8);
-
-                // add round constants
-                round_key[3] ^= ((round_constant[round] >> 0) & 0x1) << (0 * 8);
-                round_key[7] ^= ((round_constant[round] >> 1) & 0x1) << (0 * 8);
-                round_key[3] ^= ((round_constant[round] >> 2) & 0x1) << (1 * 8);
-                round_key[7] ^= ((round_constant[round] >> 3) & 0x1) << (1 * 8);
-                round_key[3] ^= ((round_constant[round] >> 4) & 0x1) << (2 * 8);
-                round_key[7] ^= ((round_constant[round] >> 5) & 0x1) << (2 * 8);
-
-                // extend bits to bytes
-                for (size_t i = 0; i < 8; i++) {
-                        round_key[i] |= round_key[i] << 1;
-                        round_key[i] |= round_key[i] << 2;
-                        round_key[i] |= round_key[i] << 4;
-                }
-
-                // load into vector registers (upper and lower part are same)
-                round_keys[round][0].val[0] = vdupq_n_u64(round_key[0]);
-                round_keys[round][0].val[1] = vdupq_n_u64(round_key[1]);
-                round_keys[round][0].val[2] = vdupq_n_u64(round_key[2]);
-                round_keys[round][0].val[3] = vdupq_n_u64(round_key[3]);
-                round_keys[round][1].val[0] = vdupq_n_u64(round_key[4]);
-                round_keys[round][1].val[1] = vdupq_n_u64(round_key[5]);
-                round_keys[round][1].val[2] = vdupq_n_u64(round_key[6]);
-                round_keys[round][1].val[3] = vdupq_n_u64(round_key[7]);
-
-                // update key state
-                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
-                int k1 = (key_state[0] >> 16) & 0xffffUL;
-                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
-                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
-                key_state[0] >>= 32;
-                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
-                key_state[1] >>= 32;
-                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
-        }
-}
-
-void gift_64_vec_sliced_init(void)
-{
-        // permutations
-        perm[0] = vld1q_u8_x4((uint8_t*)&perm_u64[0]);
-        perm[1] = vld1q_u8_x4((uint8_t*)&perm_u64[8]);
-
-        // inverse permutations
-        perm_inv[0] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[0]);
-        perm_inv[1] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[8]);
-
-        // packing masks
-        pack_mask_0 = vdupq_n_u8(0x55);
-        pack_mask_1 = vdupq_n_u8(0x33);
-        pack_mask_2 = vdupq_n_u8(0x0f);
-}
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&m[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&m[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                gift_64_vec_sliced_subcells(s);
-                gift_64_vec_sliced_permute(s);
-
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&c[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&c[8], s[1]);
-}
-
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&c[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&c[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-
-                gift_64_vec_sliced_permute_inv(s);
-                gift_64_vec_sliced_subcells_inv(s);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&m[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&m[8], s[1]);
-}
-\end{lstlisting}
-
-\chapter*{Acknowledgements}
-
-I want to thank ...
-
-\printbibliography
-
-\end{document}

### 2023-03-23 Paper/chapters/implementation.tex
move chapters into own files

--- 2023-03-20 Paper/paper.tex
@@
-\documentclass[12pt]{report}
-
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{ulem} % underline
-\usepackage{float} % H float specifier
-\usepackage{hyperref} % hyperlinks
-\usepackage{tabularx} % tabular with column specifier X to break lines
-% \usepackage{hyphenat}
-\usepackage{bookmark}
-\usepackage{systeme} % systems of equation
-\usepackage[utf8]{inputenc} % necessary for proper unicode support
-\usepackage{parcolumns} % multiple column parallel typesetting
-\usepackage[english]{babel} % comment this for Angelsächsisch
-\usepackage{graphicx} % including graphics
-\usepackage{sourcecodepro} % best font
-\usepackage{listings} % source code listings
-\usepackage{biblatex} % better citation
-% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry
-
-% umlauts in lstlistings env
-\lstset{
-    literate={ö}{{\"o}}1
-    {ä}{{\"a}}1
-    {ü}{{\"u}}1
-    {Ö}{{\"O}}1
-    {Ä}{{\"A}}1
-    {Ü}{{\"U}}1
-}
-
-\lstset{
-    basicstyle=\scriptsize\ttfamily,
-    breaklines=true
-}
-
-\lstset {
-    frame=single,
-    numbers=left,
-    stepnumber=1,
-    firstnumber=1,
-    numberfirstline=true,
-    texcl=false
-}
-
-\providecommand{\lxor}{\veebar}
-\renewcommand{\proofname}{Beweis.}
-\newcommand{\sref}[1]{\textsuperscript{\ref{#1}}}
-\newcommand{\F}[1]{\mathbb{F}_2^{#1}}
-
-\addbibresource{refs.bib}
-
-\title{
-    \textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\
-    {\footnotesize Bachelorarbeit}
-}
-\author{Bastian Engel}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\chapter*{Abstract}
-
-Lorem ipsum dolor \cite{gift:2017} sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
-nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
-fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-\chapter*{Declaration}
-
-I hereby declare that ...
-
-\tableofcontents
-
-\chapter{Introduction}
-\section{Notation}
-
-\section{Block ciphers}
-
-Securing communication channels between different parties has been a long-term
-subject of study for cryptographers and engineers which is essential to our
-modern world to cope with ever-increasing amounts of devices producing and
-sharing data. The main way to facilitate high-throughput, confidential
-communications nowadays is through the use of symmetric cryptography in which
-two parties share a common secret, called a key, which allows them to encrypt,
-share and subsequently decrypt messages to achieve confidentiality against
-third parties. Ciphers can be divided into two categories; block ciphers, which
-always encrypt fixed-sized messages called blocks, and stream ciphers, which
-continuously provide encryption for an arbitrarily long, constant stream of
-data.
-
-A block cipher can be defined as a bijection between the input block (the
-message) and the output block (the ciphertext). For any block cipher with block
-size $n$, we denote the key-dependent encryption and decryption functions as
-$E_K,D_K:\F{n}\rightarrow \F{n}$. The simplest way to
-characterize this bijection is through a lookup table which yields the highest
-possible performance as each block can be encrypted by one simple lookup
-depending on the key and the message. This is not practical though due to most
-ciphers working with block and key sizes $n,|K|\geq 64$. For a block cipher
-with $n=64,|K|=128$, a space of $2^{64}2^{128}64=2^{198}$ is necessary.
-Considering modern consumer hard disks being able to store data in the order of
-$2^{40}$, it is easy to see that a lookup table is wholly impractical. We
-therefore describe block ciphers algorithmically which opens up possibilities
-for different tradeoffs and security concerns.
-
-\subsection{GIFT}
-
-\texttt{GIFT}\cite{gift:2017}, first presented in the \textit{CHES 2017}
-cryptographic hardware and embedded systems conference, is a lightweight block
-cipher based on a previous design called \texttt{PRESENT}, developed in 2007. Its
-goal is to offer maximum security while being extremely light on resources.
-Modern battery-powered devices like RFID tags or low-latency operations like
-on-the-fly disc encryption present strong hardware and power constraints. GIFT
-aims to be a simple, low-energy cipher suited for these kinds of applications.
-
-\texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
-and \texttt{GIFT-128} working with 128-bit blocks. In both cases, the key is 128
-bits long. The design is a very simple, round-based substitution-permutation
-network (SPN). One round consists in a sequential application of the confusion
-layer by means of 4-bit S-boxes and subsequent diffusion through bit
-permutation. After the bit permutation, a round key is added to the cipher
-state and the single round is complete. \texttt{GIFT-64} uses 28 rounds while
-\texttt{GIFT-128} uses 40 rounds.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/GIFT-64.pdf}
-    \caption{Two rounds of GIFT-64}
-\end{figure}
-
-\subsubsection{Substitution layer}
-
-The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
-16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \texttt{GIFT-128}. The S-box
-$S:\F{4}\rightarrow \F{4}$ is defined as follows:
-
-\[
-    \begin{array}{l|cccccccccccccccc}
-        x & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & a & b & c & d & e & f \\
-        \hline
-        S(x) & 1 & a & 4 & c & 6 & f & 3 & 9 & 2 & d & b & 7 & 5 & 0 & 8 & e
-    \end{array}
-\]
-
-\subsubsection{Permutation layer}
-
-The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
-i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64} and
-\texttt{GIFT-128} can be expressed by:
-
-\begin{align*}
-    P_{64}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+16\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-    P_{128}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+32\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-\end{align*}
-
-\subsubsection{Round key addition}
-
-The last step of each round consists in XORing a round key $R_i$ to the cipher
-state. The new cipher state $s_{i+1}$ after each full round is therefore given
-by
-
-\[
-    s_{i+1}=P(S(s_i))\oplus R_i
-\]
-
-\subsubsection{Round key extraction and key schedule}
-
-Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
-$K=k7||k6||\dots||k0$ denote the $128$-bit key state.
-
-\paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
-state. $u_i$ and $v_i$ are XORed to $r_{4i+1}$ and $r_{4i}$ of the round key
-$R$ respectively.
-
-\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k4||k1||k_0$ from
-the key state. $u_i$ and $v_i$ are XORed to $r_{4i+2}$ and $b_{4i+1}$ of the
-round key $R$ respectively.
-
-In both cases, we additionally XOR a round constant $C=c_5c_4c_3c_2c_1c_0$ to
-bit positions $n-1,23,19,15,11,7,3$. The round constants are generated using a
-6-bit affine linear-feedback shift register and have the following values:\\
-
-\begin{tabular}{r|l}
-    \textbf{Rounds} & \textbf{Constants} \\
-    \hline
-    \textbf{1 - 16} &  \small\texttt{01,03,07,0F,1F,3E,3D,3B,37,2F,1E,3C,39,33,27,0E} \\
-    \textbf{17 - 32} & \small\texttt{1D,3A,35,2B,16,2C,18,30,21,02,05,0B,17,2E,1C,38} \\
-    \textbf{33 - 48} & \small\texttt{31,23,06,0D,1B,36,2D,1A,34,29,12,24,08,11,22,04}
-\end{tabular}\\
-
-The key state is then updated by setting $k_1\leftarrow k_1\ggg 2$,
-$k_0\leftarrow k_0\ggg 12$ and rotating the new state $32$ bits to the right:
-
-\[
-    k_7||k_6||\dots||k_1||k_0\leftarrow k_1\ggg 2||k_0\ggg 12||k_7||k_6||\dots||k_3||k_2
-\]
-
-\subsection{Camellia}
-
-\section{The ARMv8 platform}
-
-With small devices, embedded processors and ASICs becoming ever more ubiquitous
-and essential in areas like medicine or automotive design, the need for ...
-
-\chapter{Implementation strategies}
-
-Due to the structural differences of SPN- and Feistel network-based ciphers, we
-shall analyze these two separately.
-
-\section{Strategies for SPN}
-
-Three implementation strategies for substitution-permutation networks are
-introduced by \cite{implx86:2014}:
-
-\begin{itemize}
-    \item Table-based implementations
-    \item \texttt{vperm} implementations
-    \item Bitslice implementations
-\end{itemize}
-
-\subsection{Table-based}
-
-Table-driven programming is a simple way to increase performance of operations
-by tabulating the results, therefore requiring only a single memory access to
-acquire the result. This approach is obviously limited to manageable table
-sizes, so while tabulating a function like the AES S-box
-$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space,
-tabulating the \texttt{GIFT} permutation layer
-$P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
-$2^{70}$ space, which is totally unfeasible.
-
-A common approach is to tabulate the output of each S-box, including the
-diffusion layer, and then XORing the results together. Let $n$ denote the
-internal cipher state size and $s$ the size of a single S-box in bits. For each
-S-box $S_i,i\in\{0,\dots,\frac{n}{s}\}$, we can construct a mapping
-$T_i:\F{s}\rightarrow \F{n}$ representing substitution with subsequent
-permutation of that single S-box. The cipher state before round key addition is
-then given by $\bigoplus_{i=0}^{\frac{n}{s}-1}{T_i(m_i)}$ for each $s$-bit
-message chunk $m_i$. This approach requires space of
-$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for \texttt{GIFT-64},
-results in a manageable size of $\frac{64^2 2^4}{4}=2^{14}$ bits which equals
-$16$ KiB.
-
-\subsubsection{Constructing the tables}
-
-For \texttt{GIFT-64}, table construction is relatively straightforward and can
-be done as follows:
-
-\begin{lstlisting}[caption={Table construction algorithm}, escapechar=|]
-    tables <- [][]
-    for sbox_index from 0 to 15 do
-        for sbox_input from 0 to 15 do|\label{lst:tablesbox}|
-            output <- sbox(sbox_input)
-            output <- permute(output << (4 * sbox_index))
-            tables[sbox_index][sbox_input] <- output
-\end{lstlisting}
-
-Implementing this algorithm gives us the following table representing the first
-and second S-box.
-
-\[
-    \begin{array}{l|l|l|c}
-        x & T_0(x) & T_1(x) & \dots \\
-        \hline
-        0x0 & 0x1               & 0x1000000000000   & \dots \\
-        0x1 & 0x8000000020000   & 0x800000002       & \dots \\
-        0x2 & 0x400000000       & 0x40000           & \dots \\
-        0x3 & 0x8000400000000   & 0x800040000       & \dots \\
-        0x4 & 0x400020000       & 0x40002           & \dots \\
-        0x5 & 0x8000400020001   & 0x1000800040002   & \dots \\
-        0x6 & 0x20001           & 0x1000000000002   & \dots \\
-        0x7 & 0x8000000000001   & 0x1000800000000   & \dots \\
-        0x8 & 0x20000           & 0x2               & \dots \\
-        0x9 & 0x8000400000001   & 0x1000800040000   & \dots \\
-        0xa & 0x8000000020001   & 0x1000800000002   & \dots \\
-        0xb & 0x400020001       & 0x1000000040002   & \dots \\
-        0xc & 0x400000001       & 0x1000000040000   & \dots \\
-        0xd & 0x0               & 0x0               & \dots \\
-        0xe & 0x8000000000000   & 0x800000000       & \dots \\
-        0xf & 0x8000400020000   & 0x800040002       & \dots
-    \end{array}
-\]
-
-The tables for \texttt{GIFT-128} can be generated in a similar way by looping
-through all 32 S-boxes instead of 16 on line \ref{lst:tablesbox}.
-
-\subsection{Using \texttt{vperm}}
-
-Nowadays, most instructions set architectures support single-instruction,
-multiple-data processing. The idea of such an SIMD system is to work on
-multiple data stored in vectors at once to speed up calculations. For A64, two
-types of vector processing are available:
-
-\begin{enumerate}
-    \item Advanced SIMD, known as NEON
-    \item Scalable Vector Extension (SVE)
-\end{enumerate}
-
-We will take a look at NEON as this is the type of vector processing supported
-by the Cortex-A73 processor.
-
-\subsubsection{ARM Neon}
-
-The register file of the NEON unit is made up of 32 quad-word (128-bit)
-registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
-registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
-sized lanes on which the vector instructions operate. Valid ways to interpret
-for example the register \texttt{V0} are:
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/V_register.pdf}
-    \caption{Divisions of the V0 register}
-\end{figure}
-
-NEON instructions interpret their operands' layouts (i.e. lane count and width)
-through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
-adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
-together and storing the result in \texttt{V0} can be done as follows:
-
-\begin{center}
-    \texttt{ADD V0.8H, V1.8H, V2.8H}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/vector_add.pdf}
-    \caption{Addition of two vector registers}
-\end{figure}
-
-The plenitude of different processing instructions allow flexible ways to
-further speed up algorithms having reached their optimizational limit on
-non-SIMD platforms. \texttt{vperm}, a general term standing for \textit{vector
-permute}, is a common instruction on SIMD machines. Called \texttt{TBL} on
-NEON, it is used for parallel table lookups and arbitrary permutations. It takes
-two inputs to perform a lanewise lookup:
-
-\begin{enumerate}
-    \item A register with lookup values
-    \item Two or more registers containing data
-\end{enumerate}
-
-\subsubsection{S-box lookup}
-
-This instruction can be used to implement S-box lookup of all 16 S-boxes in a
-single instruction. We do this by packing our 64-bit cipher state
-$s=s_{15}||s_{14}||\dots||s_0$ into a vector register $V_0$. Because we can
-only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane which
-neatly fits into the 128-bit registers. We then put the S-box itself into
-register $V_1$ which will be used as the data register for the table lookup.
-
-The confusion layer can now be performed through one \texttt{TBL} instruction:
-
-\begin{center}
-    \texttt{TBL V0.16B, V1.16B, V0.16B}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/tbl_example.pdf}
-    \caption{Performing the S-Box lookup in parallel}
-\end{figure}
-
-\subsection{Bitslicing}
-
-Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
-achieve a more efficient representation to operate on. The structure of
-\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
-cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
-i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
-individual S-boxes. This is equivalent to transposing the bit matrix.
-
-\[
-    S=\begin{bmatrix}
-        S_0\\
-        S_1\\
-        S_2\\
-        S_3
-    \end{bmatrix}
-    =\begin{bmatrix}
-        b_{60}b_{56}b_{52}\dots b_0\\
-        b_{61}b_{57}b_{53}\dots b_1\\
-        b_{62}b_{58}b_{54}\dots b_2\\
-        b_{63}b_{59}b_{55}\dots b_3
-    \end{bmatrix}
-\]
-
-\subsubsection{Parallel S-Boxes}
-
-This representation offers multiple advantages. We first note that computation
-of the S-box can be executed in parallel, similar to the \texttt{vperm}
-technique above. This can be done by finding an algorithmic way to apply the
-S-box which has already been proposed by the original \texttt{GIFT} authors:
-
-\begin{align*}
-    S_1&\leftarrow S_1\oplus (S_0\land S_2) \\
-    t&\leftarrow S_0\oplus (S_1\land S_3) \\
-    S_2&\leftarrow S_2\oplus (t\lor S_1) \\
-    S_0&\leftarrow S_3\oplus S_2 \\
-    S_1&\leftarrow S_1\oplus S_0 \\
-    S_0&\leftarrow \lnot S_0 \\
-    S_2&\leftarrow S_2\oplus (t\land S_1) \\
-    S_3&\leftarrow t
-\end{align*}
-
-This is very efficient as it only requires six XOR-, three AND and one OR
-operation.
-
-An important property of the permutation is the fact that bits always stay in
-their slice. This means we can decompose the permutation $P$ into four
-permutations $P_i,i\in\{0,1,2,3\}$ and apply these permutations
-separately to each slice. One possible way to implement a permutation $P_i$ in
-software is to mask off all bits individually, shift them to their correct
-position and OR them together:
-
-\[
-    P_i(S_i)=\bigvee_{k=0}^{15}{(S_i\land m_i) \ll s_i}
-\]
-
-This approach requires $47$ operations, meaning all four permutations require
-over $150$ operations which would present a major bottleneck to the round
-function. We can improve on this by working on multiple message blocks at once
-and using the aforementioned \texttt{vperm} instruction to implement the bit
-shuffling. We then need only four instructions for the complete diffusion
-layer.
-
-\subsubsection{Using \texttt{vperm} for slice permutation}
-
-We cannot use the \texttt{TBL} instruction directly as we need to shuffle
-individual bits, but the smallest data we can operate on are bytes. We
-therefore encrypt $8n$ messages at once which allows us to create bytewise
-groupings. These messages are put into $4m$ registers with register $R_{4i}$
-containing $S_0$, register $R_{4i+1}$ containing $S_1$ and so forth. With block
-size $BS$ and register size $RS$, the following must hold:
-
-\[
-    8n\cdot BS=4m\cdot RS
-\]
-
-In the case of \texttt{GIFT-64} with $BS=64$ and ARM NEON with $RS=128$, we get
-
-\[
-    8n\cdot 64=4m\cdot 128\Leftrightarrow n=m
-\]
-
-$n=m=1$ would be a valid choice which yields eight messages divided into four
-registers. We choose $n=m=2$ so we can directly utilize the algorithm for bit
-packing presented by the original GIFT authors, although it is simple to adapt
-this algorithm to only four registers and eight messages by adjusting the
-\texttt{SWAPMOVE} shift and mask values.
-
-\subsubsection{Packing the data into bitslice format}
-
-Let $a,b,\dots,p$ be sixteen messages of length $64$ with subscripts denoting
-individual bits. We first put these messages into eight SIMD registers
-$V_0,V_1,\dots,V_7$:
-
-\begin{alignat*}{3}
-    V_0&=b||a\qquad V_4&&=j||&&i \\
-    V_1&=d||c\qquad V_5&&=l||&&k \\
-    V_2&=f||e\qquad V_6&&=n||&&m \\
-    V_3&=h||g\qquad V_7&&=p||&&o
-\end{alignat*}
-
-We then use the \texttt{SWAPMOVE} technique to bring the data into bitslice
-format. This operation operates on two registers $A,B$ using mask $M$ and shift
-value $N$. It swaps bits in $A$ masked by $(M\ll N)$ with bits in $B$ masked by
-$M$ in using only three XOR-, one AND- and two shift operations.
-
-\begin{align*}
-    &\text{SWAPMOVE}(A,B,M,N): \\
-    &\qquad T=((A\gg N)\oplus B)\land M \\
-    &\qquad B=B\oplus T \\
-    &\qquad A=A\oplus (T\ll N) \\
-\end{align*}
-
-One caveat of this approach is the fact that NEON registers cannot be shifted
-in their entirety due to the fact bits are not able to cross lanes. This leads
-to the problem of being able to shift at most two lanes of 64 bits at once. We
-thus need to implement the \texttt{shr(V,n)} and \texttt{shl(V,n)} operations
-on our own. This can be done by first extracting the 64-bit lanes $a,b$ out of
-$V=b||a$, shifting the lanes individually and finally shifting and ORing the
-crossing bits back into the other lane.
-
-\begin{alignat*}{2}
-    &\text{shl}(V,n): \\
-    &\qquad a,b&&=V[0],V[1] \\
-    &\qquad c&&=(a\gg (64-n)) \\
-    &\qquad a&&=(a\ll n) \\
-    &\qquad b&&=(b\ll n)\lor c \\
-    &\qquad V[0],V[1]&&=a,b
-\end{alignat*}
-
-The following operations group all $i$th bits of the messages $a,c,\dots,o$
-into bytes and puts these into the lower half of the registers $V_{i\bmod 8}$.
-The same is done for messages $b,d,\dots,p$, only differing in that the bytes
-are put into the upper half of the registers.
-
-\begin{align*}
-    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_4,V_5,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_6,V_7,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_4,V_6,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_5,V_7,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_0,V_4,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_1,V_5,0x0f0f\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4) \\
-\end{align*}
-
-With $Ax=o_xm_xk_xj_xg_xe_xc_xa_x$ and $Bx=p_xn_xl_xi_xh_xf_xd_xb_x$ denoting
-byte groups, our data now has the following permutation-friendly format:
-
-\[
-    \scriptsize
-    \begin{array}{c|llllllll|llllllll}
-        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
-        \hline
-        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8 & B0 & A56 & A48 & A40 & A32 & A24 & A16 & A8 & A0 \\
-        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9 & B1 & A57 & A49 & A41 & A33 & A25 & A17 & A9 & A1 \\
-        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
-        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
-        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
-        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
-        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
-        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7  & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
-    \end{array}
-\]
-
-Although this would already work, we prefer to have only bits of the same
-messages in each register - otherwise the permutation would need to operate on
-two source registers with the added requirement of storing the pre-permutation
-values for the first four registers, slowing down the round function through
-superfluous load/stores. This transformation is trivial by use of
-\texttt{TBL} with two data source operands.
-
-We can now create permutation tables using the specification of the individual
-slice permutations $P_i$:
-
-\[
-    \begin{array}{c|llllllllllllllll}
-        j & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\
-        \hline
-        P_0(j) & 0 & 12 & 8 & 4 & 1 & 13 & 9 & 5 & 2 & 14 & 10 & 6 & 3 & 15 & 11 & 7 \\
-        P_1(j) & 4 & 0 & 12 & 8 & 5 & 1 & 13 & 9 & 6 & 2 & 14 & 10 & 7 & 3 & 15 & 11 \\
-        P_2(j) & 8 & 4 & 0 & 12 & 9 & 5 & 1 & 13 & 10 & 6 & 2 & 14 & 11 & 7 & 3 & 15 \\
-        P_3(j) & 12 & 8 & 4 & 0 & 13 & 9 & 5 & 1 & 14 & 10 & 6 & 2 & 15 & 11 & 7 & 3
-    \end{array}
-\]
-
-Another thing to take note of is the original permutation values only showing
-where a given byte should land, not which byte belongs to a certain position -
-i.e. for $P_0$, byte 1 should land in position 12, but the byte belonging to
-position 1 is byte 4. Because \texttt{TBL} works in the latter way, we have
-to do some trivial rearrangements.
-
-Assuming the correct permutation values are put into registers
-$V_8,V_{10},\dots,V_{15}$, this now allows us to compute the permutation layer
-for all 16 blocks in only eight permutation instructions.
-
-\begin{alignat*}{2}
-    &\texttt{TBL V0, V0, V8}\qquad &&\texttt{TBL V1, V1, V9} \\
-    &\texttt{TBL V2, V2, V10}\qquad &&\texttt{TBL V3, V3, V11} \\
-    &\texttt{TBL V4, V4, V12}\qquad &&\texttt{TBL V5, V5, V13} \\
-    &\texttt{TBL V6, V6, V14}\qquad &&\texttt{TBL V7, V7, V15}
-\end{alignat*}
-
-\subsubsection{Round key function}
-
-In contrast to packing and unpacking of data which is only done once in the
-beginning and end, a round key is derived for every round, so the round key
-derivation function needs to be as fast as possible. A simple but naive
-approach for one round would be to generate a single round key, copy it 15
-times and pack the resulting registers similar to how we proceed with the
-messages. Due to the cost of packing the messages, this is prohibitivly
-expensive. Because we know where each byte group ends up after packing, we can
-directly set the round key bits at the correct position. Extending these bits
-to bytes can then be done simply by repeatedly shifting and ORing the registers
-together.
-
-\subsection{Strategies for Camellia}
-
 \chapter{Implementation}
@@
 instruction-level optimization is left to the compiler.
-
-\chapter{Evaluation}
-
-\chapter{Appendix}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.h}]
-#ifndef GIFT_VEC_SLICED_H
-#define GIFT_VEC_SLICED_H
-
-#include <stdint.h>
-#include <arm_neon.h>
-
-#define ROUNDS_GIFT_64 28
-
-// expose for benchmarking
-uint8x16_t shl(uint8x16_t v, int n);
-uint8x16_t shr(uint8x16_t v, int n);
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
-                                 uint8x16_t m, int n);
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2]);
-
-void gift_64_vec_sliced_init(void);
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2]);
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2]);
-
-#endif
-\end{lstlisting}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.c}]
-#include "gift_vec_sliced.h"
-
-#include <stddef.h>
-#include <stdio.h>
-#include <string.h>
-
-static uint64_t perm_u64[16] = {
-        0x1410050115110400UL, 0x1410050115110400UL,
-        0x0501151104001410UL, 0x0501151104001410UL,
-        0x1511040014100501UL, 0x1511040014100501UL,
-        0x0400141005011511UL, 0x0400141005011511UL,
-        0x1612070317130602UL, 0x1612070317130602UL,
-        0x0703171306021612UL, 0x0703171306021612UL,
-        0x1713060216120703UL, 0x1713060216120703UL,
-        0x0602161207031713UL, 0x0602161207031713UL
-};
-
-static uint64_t perm_inv_u64[16] = {
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL
-};
-
-static uint8x16x4_t perm[2];
-static uint8x16x4_t perm_inv[2];
-
-static uint8x16_t pack_mask_0;
-static uint8x16_t pack_mask_1;
-static uint8x16_t pack_mask_2;
-
-static const int round_constant[] = {
-        // rounds 0-15
-        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
-        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
-        // rounds 16-31
-        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
-        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
-        // rounds 32-47
-        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
-        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
-};
-
-uint8x16_t shl(uint8x16_t v, int n)
-{
-        uint64_t l[2];
-        vst1q_u64(l, v);
-        l[1] = l[1] << n | (l[0] >> (64 - n));
-        l[1] <<= n;
-        return vreinterpretq_u8_u64(vld1q_u64(l));
-}
-
-uint8x16_t shr(uint8x16_t v, int n)
-{
-        uint64_t l[2];
-        vst1q_u64(l, v);
-        l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
-        l[1] >>= n;
-        return vreinterpretq_u8_u64(vld1q_u64(l));
-}
-
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
-{
-
-        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
-        *b = veorq_u8(*b, t);
-        *a = veorq_u8(*a, shl(t, n));
-}
-
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
-{
-        // take care not to shift mask bits out of the register
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[0].val[3], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[2], &m[1].val[3], pack_mask_0, 1);
-
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[0].val[3], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[1], &m[1].val[3], pack_mask_1, 2);
-
-        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[1].val[0], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[1].val[2], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[1].val[1], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[3], &m[1].val[3], pack_mask_2, 4);
-}
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
-{
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[3] = t;
-
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-        t            = veorq_u8(cs[1].val[0], vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(cs[1].val[3], cs[1].val[2]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[3] = t;
-}
-
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16_t t = cs[0].val[3];
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[3] = veorq_u8(cs[0].val[0], cs[0].val[2]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(t, vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-
-        t            = cs[1].val[3];
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[3] = veorq_u8(cs[1].val[0], cs[1].val[2]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(t, vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-}
-
-void gift_64_vec_sliced_permute(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm[1].val[3]);
-}
-
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm_inv[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm_inv[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm_inv[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm_inv[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm_inv[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm_inv[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm_inv[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm_inv[1].val[3]);
-}
-
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2])
-{
-        // TODO shift-and-or
-
-        uint64_t key_state[] = {key[0], key[1]};
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                int v = (key_state[0] >> 0 ) & 0xffff;
-                int u = (key_state[0] >> 16) & 0xffff;
-
-                // add round key (RK=U||V)
-                uint64_t round_key[8] = { 0x0UL };
-                for (size_t i = 0; i < 8; i++) {
-                        int key_bit_v   = (v >> (2 * i + 0)) & 0x1;
-                        int key_bit_u   = (u >> (2 * i + 0)) & 0x1;
-                        round_key[0] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[1] ^= (uint64_t)key_bit_u << (i * 8);
-
-                        key_bit_v   = (v >> (2 * i + 1)) & 0x1;
-                        key_bit_u   = (u >> (2 * i + 1)) & 0x1;
-                        round_key[4] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[5] ^= (uint64_t)key_bit_u << (i * 8);
-                }
-
-                // add single bit
-                round_key[7] ^= 1UL << (7 * 8);
-
-                // add round constants
-                round_key[3] ^= ((round_constant[round] >> 0) & 0x1) << (0 * 8);
-                round_key[7] ^= ((round_constant[round] >> 1) & 0x1) << (0 * 8);
-                round_key[3] ^= ((round_constant[round] >> 2) & 0x1) << (1 * 8);
-                round_key[7] ^= ((round_constant[round] >> 3) & 0x1) << (1 * 8);
-                round_key[3] ^= ((round_constant[round] >> 4) & 0x1) << (2 * 8);
-                round_key[7] ^= ((round_constant[round] >> 5) & 0x1) << (2 * 8);
-
-                // extend bits to bytes
-                for (size_t i = 0; i < 8; i++) {
-                        round_key[i] |= round_key[i] << 1;
-                        round_key[i] |= round_key[i] << 2;
-                        round_key[i] |= round_key[i] << 4;
-                }
-
-                // load into vector registers (upper and lower part are same)
-                round_keys[round][0].val[0] = vdupq_n_u64(round_key[0]);
-                round_keys[round][0].val[1] = vdupq_n_u64(round_key[1]);
-                round_keys[round][0].val[2] = vdupq_n_u64(round_key[2]);
-                round_keys[round][0].val[3] = vdupq_n_u64(round_key[3]);
-                round_keys[round][1].val[0] = vdupq_n_u64(round_key[4]);
-                round_keys[round][1].val[1] = vdupq_n_u64(round_key[5]);
-                round_keys[round][1].val[2] = vdupq_n_u64(round_key[6]);
-                round_keys[round][1].val[3] = vdupq_n_u64(round_key[7]);
-
-                // update key state
-                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
-                int k1 = (key_state[0] >> 16) & 0xffffUL;
-                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
-                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
-                key_state[0] >>= 32;
-                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
-                key_state[1] >>= 32;
-                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
-        }
-}
-
-void gift_64_vec_sliced_init(void)
-{
-        // permutations
-        perm[0] = vld1q_u8_x4((uint8_t*)&perm_u64[0]);
-        perm[1] = vld1q_u8_x4((uint8_t*)&perm_u64[8]);
-
-        // inverse permutations
-        perm_inv[0] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[0]);
-        perm_inv[1] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[8]);
-
-        // packing masks
-        pack_mask_0 = vdupq_n_u8(0x55);
-        pack_mask_1 = vdupq_n_u8(0x33);
-        pack_mask_2 = vdupq_n_u8(0x0f);
-}
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&m[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&m[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                gift_64_vec_sliced_subcells(s);
-                gift_64_vec_sliced_permute(s);
-
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&c[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&c[8], s[1]);
-}
-
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&c[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&c[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-
-                gift_64_vec_sliced_permute_inv(s);
-                gift_64_vec_sliced_subcells_inv(s);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&m[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&m[8], s[1]);
-}
-\end{lstlisting}
-
-\chapter*{Acknowledgements}
-
-I want to thank ...
-
-\printbibliography
-
-\end{document}

### 2023-03-23 Paper/chapters/introduction.tex
move chapters into own files

--- 2023-03-18 Paper/paper.tex
@@
-\documentclass[12pt]{report}
-
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{ulem} % underline
-\usepackage{float} % H float specifier
-\usepackage{hyperref} % hyperlinks
-% \usepackage{hyphenat}
-\usepackage{bookmark}
-\usepackage{systeme} % systems of equation
-\usepackage[utf8]{inputenc} % necessary for proper unicode support
-\usepackage{parcolumns} % multiple column parallel typesetting
-\usepackage[english]{babel} % comment this for Angelsächsisch
-\usepackage{graphicx} % including graphics
-\usepackage{sourcecodepro} % best font
-\usepackage{listings} % source code listings
-\usepackage{biblatex} % better citation
-% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry
-
-% umlauts in lstlistings env
-\lstset{
-    literate={ö}{{\"o}}1
-    {ä}{{\"a}}1
-    {ü}{{\"u}}1
-    {Ö}{{\"O}}1
-    {Ä}{{\"A}}1
-    {Ü}{{\"U}}1
-}
-
-\lstset{
-    basicstyle=\scriptsize\ttfamily,
-    breaklines=true
-}
-
-\lstset {
-    frame=single,
-    numbers=left,
-    stepnumber=1,
-    firstnumber=1,
-    numberfirstline=true,
-    texcl=false
-}
-
-\providecommand{\lxor}{\veebar}
-\renewcommand{\proofname}{Beweis.}
-\newcommand{\sref}[1]{\textsuperscript{\ref{#1}}}
-\newcommand{\F}[1]{\mathbb{F}_2^{#1}}
-
-\addbibresource{refs.bib}
-
-\title{
-    \textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\
-    {\footnotesize Bachelorarbeit}
-}
-\author{Bastian Engel}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\chapter*{Abstract}
-
-Lorem ipsum dolor \cite{gift:2017} sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
-nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
-fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-\chapter*{Declaration}
-
-I hereby declare that ...
-
-\tableofcontents
-
 \chapter{Introduction}
@@
 and essential in areas like medicine or automotive design, the need for ...
-
-\chapter{Implementation strategies}
-
-Due to the structural differences of SPN- and Feistel network-based ciphers, we
-shall analyze these two separately.
-
-\section{Strategies for SPN}
-
-Three implementation strategies for substitution-permutation networks are
-introduced by \cite{implx86:2014}:
-
-\begin{itemize}
-    \item Table-based implementations
-    \item \texttt{vperm} implementations
-    \item Bitslice implementations
-\end{itemize}
-
-\subsection{Table-based}
-
-Table-driven programming is a simple way to increase performance of operations
-by tabulating the results, therefore requiring only a single memory access to
-acquire the result. This approach is obviously limited to manageable table
-sizes, so while tabulating a function like the AES S-box
-$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space,
-tabulating the \texttt{GIFT} permutation layer
-$P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
-$2^{70}$ space, which is totally unfeasible.
-
-A common approach is to tabulate the output of each S-box, including the
-diffusion layer, and then XORing the results together. Let $n$ denote the
-internal cipher state size and $s$ the size of a single S-box in bits. For each
-S-box $S_i,i\in\{0,\dots,\frac{n}{s}\}$, we can construct a mapping
-$T_i:\F{s}\rightarrow \F{n}$ representing substitution with subsequent
-permutation of that single S-box. The cipher state before round key addition is
-then given by $\bigoplus_{i=0}^{\frac{n}{s}-1}{T_i(m_i)}$ for each $s$-bit
-message chunk $m_i$. This approach requires space of
-$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for \texttt{GIFT-64},
-results in a manageable size of $\frac{64^2 2^4}{4}=2^{14}$ bits which equals
-$16$ KiB.
-
-\subsubsection{Constructing the tables}
-
-For \texttt{GIFT-64}, table construction is relatively straightforward and can
-be done as follows:
-
-\begin{lstlisting}[caption={Table construction algorithm}, escapechar=|]
-    tables <- [][]
-    for sbox_index from 0 to 15 do
-        for sbox_input from 0 to 15 do|\label{lst:tablesbox}|
-            output <- sbox(sbox_input)
-            output <- permute(output << (4 * sbox_index))
-            tables[sbox_index][sbox_input] <- output
-\end{lstlisting}
-
-Implementing this algorithm gives us the following table representing the first
-and second S-box.
-
-\[
-    \begin{array}{l|l|l|c}
-        x & T_0(x) & T_1(x) & \dots \\
-        \hline
-        0x0 & 0x1               & 0x1000000000000   & \dots \\
-        0x1 & 0x8000000020000   & 0x800000002       & \dots \\
-        0x2 & 0x400000000       & 0x40000           & \dots \\
-        0x3 & 0x8000400000000   & 0x800040000       & \dots \\
-        0x4 & 0x400020000       & 0x40002           & \dots \\
-        0x5 & 0x8000400020001   & 0x1000800040002   & \dots \\
-        0x6 & 0x20001           & 0x1000000000002   & \dots \\
-        0x7 & 0x8000000000001   & 0x1000800000000   & \dots \\
-        0x8 & 0x20000           & 0x2               & \dots \\
-        0x9 & 0x8000400000001   & 0x1000800040000   & \dots \\
-        0xa & 0x8000000020001   & 0x1000800000002   & \dots \\
-        0xb & 0x400020001       & 0x1000000040002   & \dots \\
-        0xc & 0x400000001       & 0x1000000040000   & \dots \\
-        0xd & 0x0               & 0x0               & \dots \\
-        0xe & 0x8000000000000   & 0x800000000       & \dots \\
-        0xf & 0x8000400020000   & 0x800040002       & \dots
-    \end{array}
-\]
-
-The tables for \texttt{GIFT-128} can be generated in a similar way by looping
-through all 32 S-boxes instead of 16 on line \ref{lst:tablesbox}.
-
-\subsection{Using \texttt{vperm}}
-
-Nowadays, most instructions set architectures support single-instruction,
-multiple-data processing. The idea of such an SIMD system is to work on
-multiple data stored in vectors at once to speed up calculations. For A64, two
-types of vector processing are available:
-
-\begin{enumerate}
-    \item Advanced SIMD, known as NEON
-    \item Scalable Vector Extension (SVE)
-\end{enumerate}
-
-We will take a look at NEON as this is the type of vector processing supported
-by the Cortex-A73 processor.
-
-\subsubsection{ARM Neon}
-
-The register file of the NEON unit is made up of 32 quad-word (128-bit)
-registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
-registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
-sized lanes on which the vector instructions operate. Valid ways to interpret
-for example the register \texttt{V0} are:
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/V_register.pdf}
-    \caption{Divisions of the V0 register}
-\end{figure}
-
-NEON instructions interpret their operands' layouts (i.e. lane count and width)
-through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
-adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
-together and storing the result in \texttt{V0} can be done as follows:
-
-\begin{center}
-    \texttt{ADD V0.8H, V1.8H, V2.8H}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/vector_add.pdf}
-    \caption{Addition of two vector registers}
-\end{figure}
-
-The plenitude of different processing instructions allow flexible ways to
-further speed up algorithms having reached their optimizational limit on
-non-SIMD platforms. \texttt{vperm}, a general term standing for \textit{vector
-permute}, is a common instruction on SIMD machines. Called \texttt{TBL} on
-NEON, it is used for parallel table lookups and arbitrary permutations. It takes
-two inputs to perform a lanewise lookup:
-
-\begin{enumerate}
-    \item A register with lookup values
-    \item Two or more registers containing data
-\end{enumerate}
-
-\subsubsection{S-box lookup}
-
-This instruction can be used to implement S-box lookup of all 16 S-boxes in a
-single instruction. We do this by packing our 64-bit cipher state
-$s=s_{15}||s_{14}||\dots||s_0$ into a vector register $V_0$. Because we can
-only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane which
-neatly fits into the 128-bit registers. We then put the S-box itself into
-register $V_1$ which will be used as the data register for the table lookup.
-
-The confusion layer can now be performed through one \texttt{TBL} instruction:
-
-\begin{center}
-    \texttt{TBL V0.16B, V1.16B, V0.16B}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/tbl_example.pdf}
-    \caption{Performing the S-Box lookup in parallel}
-\end{figure}
-
-\subsection{Bitslicing}
-
-Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
-achieve a more efficient representation to operate on. The structure of
-\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
-cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
-i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
-individual S-boxes. This is equivalent to transposing the bit matrix.
-
-\[
-    S=\begin{bmatrix}
-        S_0\\
-        S_1\\
-        S_2\\
-        S_3
-    \end{bmatrix}
-    =\begin{bmatrix}
-        b_{60}b_{56}b_{52}\dots b_0\\
-        b_{61}b_{57}b_{53}\dots b_1\\
-        b_{62}b_{58}b_{54}\dots b_2\\
-        b_{63}b_{59}b_{55}\dots b_3
-    \end{bmatrix}
-\]
-
-\subsubsection{Parallel S-Boxes}
-
-This representation offers multiple advantages. We first note that computation
-of the S-box can be executed in parallel, similar to the \texttt{vperm}
-technique above. This can be done by finding an algorithmic way to apply the
-S-box which has already been proposed by the original \texttt{GIFT} authors:
-
-\begin{align*}
-    S_1&\leftarrow S_1\oplus (S_0\land S_2) \\
-    t&\leftarrow S_0\oplus (S_1\land S_3) \\
-    S_2&\leftarrow S_2\oplus (t\lor S_1) \\
-    S_0&\leftarrow S_3\oplus S_2 \\
-    S_1&\leftarrow S_1\oplus S_0 \\
-    S_0&\leftarrow \lnot S_0 \\
-    S_2&\leftarrow S_2\oplus (t\land S_1) \\
-    S_3&\leftarrow t
-\end{align*}
-
-This is very efficient as it only requires six XOR-, three AND and one OR
-operation.
-
-An important property of the permutation is the fact that bits always stay in
-their slice. This means we can decompose the permutation $P$ into four
-permutations $P_i,i\in\{0,1,2,3\}$ and apply these permutations
-separately to each slice. One possible way to implement a permutation $P_i$ in
-software is to mask off all bits individually, shift them to their correct
-position and OR them together:
-
-\[
-    P_i(S_i)=\bigvee_{k=0}^{15}{(S_i\land m_i) \ll s_i}
-\]
-
-This approach requires $47$ operations, meaning all four permutations require
-over $150$ operations which would present a major bottleneck to the round
-function. We can improve on this by working on multiple message blocks at once
-and using the aforementioned \texttt{vperm} instruction to implement the bit
-shuffling. We then need only four \texttt{TBL} instructions for the complete
-diffusion layer.
-
-\subsubsection{Using \texttt{vperm} for slice permutation}
-
-We cannot use the \texttt{TBL} instruction directly as we need to shuffle
-individual bits, but the smallest data we can operate on are bytes. We
-therefore encrypt $8n$ messages at once which allows us to create bytewise
-groupings. These messages are put into $4m$ registers with register $R_{4i}$
-containing $S_0$, register $R_{4i+1}$ containing $S_1$ and so forth. With block
-size $BS$ and register size $RS$, the following must hold:
-
-\[
-    8n\cdot BS=4m\cdot RS
-\]
-
-In the case of \texttt{GIFT-64} with $BS=64$ and ARM NEON with $RS=128$, we get
-
-\[
-    8n\cdot 64=4m\cdot 128\Leftrightarrow n=m
-\]
-
-$n=m=1$ would be a valid choice which yields eight messages divided into four
-registers. We choose $n=m=2$ so we can directly utilize the algorithm for bit
-packing presented by the original GIFT authors, although it is simple to adapt
-this algorithm to only four registers and eight messages by adjusting the
-\texttt{SWAPMOVE} shift and mask values.
-
-\subsubsection{Packing the data into bitslice format}
-
-Let $a,b,\dots,p$ be sixteen messages of length $64$ with subscripts denoting
-individual bits. We first put these messages into eight SIMD registers
-$V_0,V_1,\dots,V_7$:
-
-\begin{alignat*}{3}
-    V_0&=b||a\qquad V_4&&=j||&&i \\
-    V_1&=d||c\qquad V_5&&=l||&&k \\
-    V_2&=f||e\qquad V_6&&=n||&&m \\
-    V_3&=h||g\qquad V_7&&=p||&&o
-\end{alignat*}
-
-We then use the \texttt{SWAPMOVE} technique to bring the data into bitslice
-format. This operation operates on two registers $A,B$ using mask $M$ and shift
-value $N$. It swaps bits in $A$ masked by $(M\ll N)$ with bits in $B$ masked by
-$M$ in using only three XOR-, one AND- and two shift operations.
-
-\begin{align*}
-    &\text{SWAPMOVE}(A,B,M,N): \\
-    &\qquad T=((A\gg N)\oplus B)\land M \\
-    &\qquad B=B\oplus T \\
-    &\qquad A=A\oplus (T\ll N) \\
-\end{align*}
-
-One caveat of this approach is the fact that NEON registers cannot be shifted
-in their entirety due to the fact bits are not able to cross lanes. This leads
-to the problem of being able to shift at most two lanes of 64 bits at once. We
-thus need to implement the \texttt{shr(V,n)} and \texttt{shl(V,n)} operations
-on our own. This can be done by first extracting the 64-bit lanes $a,b$ out of
-$V=b||a$, shifting the lanes individually and finally shifting and ORing the
-crossing bits back into the other lane.
-
-\begin{alignat*}{2}
-    &\text{shl}(V,n): \\
-    &\qquad a,b&&=V[0],V[1] \\
-    &\qquad c&&=(a\gg (64-n)) \\
-    &\qquad a&&=(a\ll n) \\
-    &\qquad b&&=(b\ll n)\lor c \\
-    &\qquad V[0],V[1]&&=a,b
-\end{alignat*}
-
-The following operations group all $i$th bits of the messages $a,c,\dots,o$
-into bytes and puts these into the lower half of the registers $V_{i\bmod 8}$.
-The same is done for messages $b,d,\dots,p$, only differing in that the bytes
-are put into the upper half of the registers.
-
-\begin{align*}
-    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_4,V_5,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_6,V_7,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_4,V_6,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_5,V_7,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_0,V_4,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_1,V_5,0x0f0f\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4) \\
-\end{align*}
-
-With $Ax=o_xm_xk_xj_xg_xe_xc_xa_x$ and $Bx=p_xn_xl_xi_xh_xf_xd_xb_x$ denoting
-byte groups, our data now has the following permutation-friendly format:
-
-\[
-    \begin{array}{c|llllllll}
-        n & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
-        \hline
-        V_0 & A56 & A48 & A40 & A32 & A24 & A16 & A8 & A0 \\
-        V_1 & A57 & A49 & A41 & A33 & A25 & A17 & A9 & A1 \\
-        V_2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
-        V_3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
-        V_4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
-        V_5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
-        V_6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
-        V_7 & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
-    \end{array}
-\]
-
-\[
-    \begin{array}{c|llllllll}
-        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 \\
-        \hline
-        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8 & B0 \\
-        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9 & B1 \\
-        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 \\
-        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 \\
-        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 \\
-        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 \\
-        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 \\
-        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7
-    \end{array}
-\]
-
-We can now create permutation tables using the specification of the individual
-slice permutations $P_i$:
-
-\[
-    \begin{array}{c|llllllllllllllll}
-        j & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\
-        \hline
-        P_0(j) & 0 & 12 & 8 & 4 & 1 & 13 & 9 & 5 & 2 & 14 & 10 & 6 & 3 & 15 & 11 & 7 \\
-        P_1(j) & 4 & 0 & 12 & 8 & 5 & 1 & 13 & 9 & 6 & 2 & 14 & 10 & 7 & 3 & 15 & 11 \\
-        P_2(j) & 8 & 4 & 0 & 12 & 9 & 5 & 1 & 13 & 10 & 6 & 2 & 14 & 11 & 7 & 3 & 15 \\
-        P_3(j) & 12 & 8 & 4 & 0 & 13 & 9 & 5 & 1 & 14 & 10 & 6 & 2 & 15 & 11 & 7 & 3
-    \end{array}
-\]
-
-One thing to take note of is the original permutation values only showing where
-a given byte should land, not which byte belongs to a certain position - i.e.
-for $P_0$, byte 1 should land in position 12, but the byte belonging to
-position 1 is byte 4. Because \texttt{vperm} works in the latter way, we have
-to do some trivial rearrangements. In addition, because data for each slice is
-split between two registers $(V_0,V_4),(V_1,V_5),\dots$, the \texttt{vperm}
-operation needs to use these two registers as a data source. Using two data
-sources is supported by the \texttt{TBL} instruction by means of an additional
-parameter.
-
-Assuming the correct permutation values are put into registers
-$V_9,V_{10},\dots,V_{16}$, this now allows us to compute the permutation layer for
-all 16 blocks in only eight permutation instructions plus four additional vector
-copy instructions used for saving the original values of $V_0,V_1,V_2,V_3$
-before the permutation is applied.
-
-\begin{alignat*}{2}
-    &\texttt{MOV V17, V0}\qquad &&\texttt{MOV V18, V1} \\
-    &\texttt{MOV V19, V2}\qquad &&\texttt{MOV V20, V3} \\
-    &\texttt{TBL V0, \{V0, V4\}, V9}\qquad &&\texttt{TBL V1, \{V1, V5\}, V10} \\
-    &\texttt{TBL V2, \{V2, V6\}, V9}\qquad &&\texttt{TBL V3, \{V3, V7\}, V10} \\
-    &\texttt{TBL V4, \{V17, V4\}, V9}\qquad &&\texttt{TBL V5, \{V18, V5\}, V10} \\
-    &\texttt{TBL V6, \{V19, V6\}, V9}\qquad &&\texttt{TBL V7, \{V20, V7\}, V10}
-\end{alignat*}
-
-\subsubsection{Round key function}
-
-In contrast to packing and unpacking of data which is only done once in the
-beginning and end, a round key is derived for every round, so the round key
-derivation function needs to be as fast as possible. A simple but naive
-approach for one round would be to generate a single round key, copy it 15
-times and pack the resulting registers similar to how we proceed with the
-messages. Due to the cost of packing the messages, this is prohibitivly
-expensive. Because we know where each byte group ends up after packing, we can
-directly set the round key bits at the correct position. Extending these bits
-to bytes can then be done simply by repeatedly shifting and ORing the registers
-together.
-
-\subsection{Strategies for Camellia}
-
-\chapter{Implementation}
-
-We will provide and discuss implementations for the presented strategies in the
-C programming language. Although directly writing Assembler code could result
-in a small performance benefit, this generally increases the work necessary by
-an order of magnitude for only limited results. This is why we use NEON
-intrinsics.
-
-\subsection{NEON intrinsics}
-
-The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
-definitions including vector data types and C functions for working with these
-vectors. These functions are known as NEON intrinsics and give the programmer a
-high-level interface to most NEON instructions. Major advantages of this
-approach include the ease of development as the compiler takes over register
-allocation and load/store operations as well as performance benefits through
-compiler optimizations.
-
-Standard vector data types have the format \texttt{uintnxm\_t} with lane width
-$n$ in bits and and lane count $m$. Array types of the format
-\texttt{uintnxmxc\_t}, $c\in\{1,2,3,4\}$ are also defined which are used in
-operations requiring multiple parameters like \texttt{TBL} or pairwise
-load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
-after the operation name to indicate operation on a 128-bit register and end
-with the lane data format. Multiplying eight pairs of 16-bit numbers \texttt{a,b} for
-example can be done via the following:
-
-\begin{center}
-    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
-\end{center}
-
-The compiler allocates vector registers for \texttt{a}, \texttt{b} and
-\texttt{result} and assembles the intrinsic to \texttt{MUL Vr.8H, Va.8H,
-Vb.8H}. Necessary load and stores for the result and parameters are also
-handled automatically.
-
-\chapter{Evaluation}
-
-\chapter{Appendix}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.h}]
-#ifndef GIFT_VEC_SLICED_H
-#define GIFT_VEC_SLICED_H
-
-#include <stdint.h>
-#include <arm_neon.h>
-
-#define ROUNDS_GIFT_64 28
-
-// expose for benchmarking
-uint8x16_t shl(uint8x16_t v, int n);
-uint8x16_t shr(uint8x16_t v, int n);
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
-                                 uint8x16_t m, int n);
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2]);
-
-void gift_64_vec_sliced_init(void);
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2]);
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2]);
-
-#endif
-\end{lstlisting}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.c}]
-#include "gift_vec_sliced.h"
-
-#include <stddef.h>
-#include <stdio.h>
-#include <string.h>
-
-static uint64_t perm_u64[16] = {
-        0x1410050115110400UL, 0x1410050115110400UL,
-        0x0501151104001410UL, 0x0501151104001410UL,
-        0x1511040014100501UL, 0x1511040014100501UL,
-        0x0400141005011511UL, 0x0400141005011511UL,
-        0x1612070317130602UL, 0x1612070317130602UL,
-        0x0703171306021612UL, 0x0703171306021612UL,
-        0x1713060216120703UL, 0x1713060216120703UL,
-        0x0602161207031713UL, 0x0602161207031713UL
-};
-
-static uint64_t perm_inv_u64[16] = {
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL
-};
-
-static uint8x16x4_t perm[2];
-static uint8x16x4_t perm_inv[2];
-
-static uint8x16_t pack_mask_0;
-static uint8x16_t pack_mask_1;
-static uint8x16_t pack_mask_2;
-
-static const int round_constant[] = {
-        // rounds 0-15
-        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
-        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
-        // rounds 16-31
-        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
-        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
-        // rounds 32-47
-        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
-        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
-};
-
-uint8x16_t shl(uint8x16_t v, int n)
-{
-        // TODO check out vst
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l1 = l1 << n | (l0 >> (64 - n));
-        l0 <<= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
-}
-
-uint8x16_t shr(uint8x16_t v, int n)
-{
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l0 = l0 >> n | (((l1 << (64 - n)) >> (64 - n)) << (64 - n));
-        l1 >>= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
-}
-
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
-{
-
-        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
-        *b = veorq_u8(*b, t);
-        *a = veorq_u8(*a, shl(t, n));
-}
-
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
-{
-        // take care not to shift mask bits out of the register
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[0].val[3], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[2], &m[1].val[3], pack_mask_0, 1);
-
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[0].val[3], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[1], &m[1].val[3], pack_mask_1, 2);
-
-        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[1].val[0], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[1].val[2], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[1].val[1], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[3], &m[1].val[3], pack_mask_2, 4);
-}
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
-{
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[3] = t;
-
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-        t            = veorq_u8(cs[1].val[0], vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(cs[1].val[3], cs[1].val[2]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[3] = t;
-}
-
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16_t t = cs[0].val[3];
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[3] = veorq_u8(cs[0].val[0], cs[0].val[2]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(t, vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-
-        t            = cs[1].val[3];
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[3] = veorq_u8(cs[1].val[0], cs[1].val[2]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(t, vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-}
-
-void gift_64_vec_sliced_permute(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm[1].val[3]);
-}
-
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm_inv[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm_inv[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm_inv[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm_inv[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm_inv[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm_inv[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm_inv[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm_inv[1].val[3]);
-}
-
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2])
-{
-        // TODO shift-and-or
-
-        uint64_t key_state[] = {key[0], key[1]};
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                int v = (key_state[0] >> 0 ) & 0xffff;
-                int u = (key_state[0] >> 16) & 0xffff;
-
-                // add round key (RK=U||V)
-                uint64_t round_key[8] = { 0x0UL };
-                for (size_t i = 0; i < 8; i++) {
-                        int key_bit_v   = (v >> (2 * i + 0)) & 0x1;
-                        int key_bit_u   = (u >> (2 * i + 0)) & 0x1;
-                        round_key[0] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[1] ^= (uint64_t)key_bit_u << (i * 8);
-
-                        key_bit_v   = (v >> (2 * i + 1)) & 0x1;
-                        key_bit_u   = (u >> (2 * i + 1)) & 0x1;
-                        round_key[4] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[5] ^= (uint64_t)key_bit_u << (i * 8);
-                }
-
-                // add single bit
-                round_key[7] ^= 1UL << (7 * 8);
-
-                // add round constants
-                round_key[3] ^= ((round_constant[round] >> 0) & 0x1) << (0 * 8);
-                round_key[7] ^= ((round_constant[round] >> 1) & 0x1) << (0 * 8);
-                round_key[3] ^= ((round_constant[round] >> 2) & 0x1) << (1 * 8);
-                round_key[7] ^= ((round_constant[round] >> 3) & 0x1) << (1 * 8);
-                round_key[3] ^= ((round_constant[round] >> 4) & 0x1) << (2 * 8);
-                round_key[7] ^= ((round_constant[round] >> 5) & 0x1) << (2 * 8);
-
-                // extend bits to bytes
-                for (size_t i = 0; i < 8; i++) {
-                        round_key[i] |= round_key[i] << 1;
-                        round_key[i] |= round_key[i] << 2;
-                        round_key[i] |= round_key[i] << 4;
-                }
-
-                // load into vector registers (upper and lower part are same)
-                round_keys[round][0].val[0] = vdupq_n_u64(round_key[0]);
-                round_keys[round][0].val[1] = vdupq_n_u64(round_key[1]);
-                round_keys[round][0].val[2] = vdupq_n_u64(round_key[2]);
-                round_keys[round][0].val[3] = vdupq_n_u64(round_key[3]);
-                round_keys[round][1].val[0] = vdupq_n_u64(round_key[4]);
-                round_keys[round][1].val[1] = vdupq_n_u64(round_key[5]);
-                round_keys[round][1].val[2] = vdupq_n_u64(round_key[6]);
-                round_keys[round][1].val[3] = vdupq_n_u64(round_key[7]);
-
-                // update key state
-                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
-                int k1 = (key_state[0] >> 16) & 0xffffUL;
-                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
-                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
-                key_state[0] >>= 32;
-                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
-                key_state[1] >>= 32;
-                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
-        }
-}
-
-void gift_64_vec_sliced_init(void)
-{
-        // permutations
-        perm[0] = vld1q_u8_x4((uint8_t*)&perm_u64[0]);
-        perm[1] = vld1q_u8_x4((uint8_t*)&perm_u64[8]);
-
-        // inverse permutations
-        perm_inv[0] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[0]);
-        perm_inv[1] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[8]);
-
-        // packing masks
-        pack_mask_0 = vdupq_n_u8(0x55);
-        pack_mask_1 = vdupq_n_u8(0x33);
-        pack_mask_2 = vdupq_n_u8(0x0f);
-}
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&m[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&m[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                gift_64_vec_sliced_subcells(s);
-                gift_64_vec_sliced_permute(s);
-
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&c[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&c[8], s[1]);
-}
-
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&c[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&c[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-
-                gift_64_vec_sliced_permute_inv(s);
-                gift_64_vec_sliced_subcells_inv(s);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&m[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&m[8], s[1]);
-}
-\end{lstlisting}
-
-\chapter*{Acknowledgements}
-
-I want to thank ...
-
-\printbibliography
-
-\end{document}

### 2023-03-23 Paper/paper.tex
move chapters into own files

--- 2023-03-18 Paper/paper.tex
@@
 
-\chapter*{Abstract}
-
-Lorem ipsum dolor \cite{gift:2017} sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
-nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
-fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-\chapter*{Declaration}
-
-I hereby declare that ...
+\include{chapters/abstract}
+\include{chapters/declaration}
 
@@
 
-\chapter{Introduction}
-\section{Notation}
-
-\section{Block ciphers}
-
-Securing communication channels between different parties has been a long-term
-subject of study for cryptographers and engineers which is essential to our
-modern world to cope with ever-increasing amounts of devices producing and
-sharing data. The main way to facilitate high-throughput, confidential
-communications nowadays is through the use of symmetric cryptography in which
-two parties share a common secret, called a key, which allows them to encrypt,
-share and subsequently decrypt messages to achieve confidentiality against
-third parties. Ciphers can be divided into two categories; block ciphers, which
-always encrypt fixed-sized messages called blocks, and stream ciphers, which
-continuously provide encryption for an arbitrarily long, constant stream of
-data.
-
-A block cipher can be defined as a bijection between the input block (the
-message) and the output block (the ciphertext). For any block cipher with block
-size $n$, we denote the key-dependent encryption and decryption functions as
-$E_K,D_K:\F{n}\rightarrow \F{n}$. The simplest way to
-characterize this bijection is through a lookup table which yields the highest
-possible performance as each block can be encrypted by one simple lookup
-depending on the key and the message. This is not practical though due to most
-ciphers working with block and key sizes $n,|K|\geq 64$. For a block cipher
-with $n=64,|K|=128$, a space of $2^{64}2^{128}64=2^{198}$ is necessary.
-Considering modern consumer hard disks being able to store data in the order of
-$2^{40}$, it is easy to see that a lookup table is wholly impractical. We
-therefore describe block ciphers algorithmically which opens up possibilities
-for different tradeoffs and security concerns.
-
-\subsection{GIFT}
-
-\texttt{GIFT}\cite{gift:2017}, first presented in the \textit{CHES 2017}
-cryptographic hardware and embedded systems conference, is a lightweight block
-cipher based on a previous design called \texttt|PRESENT|, developed in 2007. Its
-goal is to offer maximum security while being extremely light on resources.
-Modern battery-powered devices like RFID tags or low-latency operations like
-on-the-fly disc encryption present strong hardware and power constraints. GIFT
-aims to be a simple, low-energy cipher suited for these kinds of applications.
-
-\texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
-and \texttt{GIFT-128} working with 128-bit blocks. In both cases, the key is 128
-bits long. The design is a very simple, round-based substitution-permutation
-network (SPN). One round consists in a sequential application of the confusion
-layer by means of 4-bit S-boxes and subsequent diffusion through bit
-permutation. After the bit permutation, a round key is added to the cipher
-state and the single round is complete. \texttt{GIFT-64} uses 28 rounds while
-\texttt{GIFT-128} uses 40 rounds.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/GIFT-64.pdf}
-    \caption{Two rounds of GIFT-64}
-\end{figure}
-
-\subsubsection{Substitution layer}
-
-The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
-16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \texttt{GIFT-128}. The S-box
-$S:\F{4}\rightarrow \F{4}$ is defined as follows:
-
-\[
-    \begin{array}{l|cccccccccccccccc}
-        x & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & a & b & c & d & e & f \\
-        \hline
-        S(x) & 1 & a & 4 & c & 6 & f & 3 & 9 & 2 & d & b & 7 & 5 & 0 & 8 & e
-    \end{array}
-\]
-
-\subsubsection{Permutation layer}
-
-The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
-i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64} and
-\texttt{GIFT-128} can be expressed by:
-
-\begin{align*}
-    P_{64}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+16\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-    P_{128}(i)&=4\left\lfloor\frac{i}{16}\right\rfloor+32\left(\left(3\left\lfloor\frac{i\bmod 16}{4}\right\rfloor+(i\bmod 4)\right)\bmod 4\right)+(i\bmod 4) \\
-\end{align*}
-
-\subsubsection{Round key addition}
-
-The last step of each round consists in XORing a round key $R_i$ to the cipher
-state. The new cipher state $s_{i+1}$ after each full round is therefore given
-by
-
-\[
-    s_{i+1}=P(S(s_i))\oplus R_i
-\]
-
-\subsubsection{Round key extraction and key schedule}
-
-Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
-$K=k7||k6||\dots||k0$ denote the $128$-bit key state.
-
-\paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
-state. $u_i$ and $v_i$ are XORed to $r_{4i+1}$ and $r_{4i}$ of the round key
-$R$ respectively.
-
-\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k4||k1||k_0$ from
-the key state. $u_i$ and $v_i$ are XORed to $r_{4i+2}$ and $b_{4i+1}$ of the
-round key $R$ respectively.
-
-In both cases, we additionally XOR a round constant $C=c_5c_4c_3c_2c_1c_0$ to
-bit positions $n-1,23,19,15,11,7,3$. The round constants are generated using a
-6-bit affine linear-feedback shift register and have the following values:\\
-
-\begin{tabular}{r|l}
-    \textbf{Rounds} & \textbf{Constants} \\
-    \hline
-    \textbf{1 - 16} &  \small\texttt{01,03,07,0F,1F,3E,3D,3B,37,2F,1E,3C,39,33,27,0E} \\
-    \textbf{17 - 32} & \small\texttt{1D,3A,35,2B,16,2C,18,30,21,02,05,0B,17,2E,1C,38} \\
-    \textbf{33 - 48} & \small\texttt{31,23,06,0D,1B,36,2D,1A,34,29,12,24,08,11,22,04}
-\end{tabular}\\
-
-The key state is then updated by setting $k_1\leftarrow k_1\ggg 2$,
-$k_0\leftarrow k_0\ggg 12$ and rotating the new state $32$ bits to the right:
-
-\[
-    k_7||k_6||\dots||k_1||k_0\leftarrow k_1\ggg 2||k_0\ggg 12||k_7||k_6||\dots||k_3||k_2
-\]
-
-\subsection{Camellia}
-
-\section{The ARMv8 platform}
-
-With small devices, embedded processors and ASICs becoming ever more ubiquitous
-and essential in areas like medicine or automotive design, the need for ...
-
-\chapter{Implementation strategies}
-
-Due to the structural differences of SPN- and Feistel network-based ciphers, we
-shall analyze these two separately.
-
-\section{Strategies for SPN}
-
-Three implementation strategies for substitution-permutation networks are
-introduced by \cite{implx86:2014}:
-
-\begin{itemize}
-    \item Table-based implementations
-    \item \texttt{vperm} implementations
-    \item Bitslice implementations
-\end{itemize}
-
-\subsection{Table-based}
-
-Table-driven programming is a simple way to increase performance of operations
-by tabulating the results, therefore requiring only a single memory access to
-acquire the result. This approach is obviously limited to manageable table
-sizes, so while tabulating a function like the AES S-box
-$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space,
-tabulating the \texttt{GIFT} permutation layer
-$P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
-$2^{70}$ space, which is totally unfeasible.
-
-A common approach is to tabulate the output of each S-box, including the
-diffusion layer, and then XORing the results together. Let $n$ denote the
-internal cipher state size and $s$ the size of a single S-box in bits. For each
-S-box $S_i,i\in\{0,\dots,\frac{n}{s}\}$, we can construct a mapping
-$T_i:\F{s}\rightarrow \F{n}$ representing substitution with subsequent
-permutation of that single S-box. The cipher state before round key addition is
-then given by $\bigoplus_{i=0}^{\frac{n}{s}-1}{T_i(m_i)}$ for each $s$-bit
-message chunk $m_i$. This approach requires space of
-$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for \texttt{GIFT-64},
-results in a manageable size of $\frac{64^2 2^4}{4}=2^{14}$ bits which equals
-$16$ KiB.
-
-\subsubsection{Constructing the tables}
-
-For \texttt{GIFT-64}, table construction is relatively straightforward and can
-be done as follows:
-
-\begin{lstlisting}[caption={Table construction algorithm}, escapechar=|]
-    tables <- [][]
-    for sbox_index from 0 to 15 do
-        for sbox_input from 0 to 15 do|\label{lst:tablesbox}|
-            output <- sbox(sbox_input)
-            output <- permute(output << (4 * sbox_index))
-            tables[sbox_index][sbox_input] <- output
-\end{lstlisting}
-
-Implementing this algorithm gives us the following table representing the first
-and second S-box.
-
-\[
-    \begin{array}{l|l|l|c}
-        x & T_0(x) & T_1(x) & \dots \\
-        \hline
-        0x0 & 0x1               & 0x1000000000000   & \dots \\
-        0x1 & 0x8000000020000   & 0x800000002       & \dots \\
-        0x2 & 0x400000000       & 0x40000           & \dots \\
-        0x3 & 0x8000400000000   & 0x800040000       & \dots \\
-        0x4 & 0x400020000       & 0x40002           & \dots \\
-        0x5 & 0x8000400020001   & 0x1000800040002   & \dots \\
-        0x6 & 0x20001           & 0x1000000000002   & \dots \\
-        0x7 & 0x8000000000001   & 0x1000800000000   & \dots \\
-        0x8 & 0x20000           & 0x2               & \dots \\
-        0x9 & 0x8000400000001   & 0x1000800040000   & \dots \\
-        0xa & 0x8000000020001   & 0x1000800000002   & \dots \\
-        0xb & 0x400020001       & 0x1000000040002   & \dots \\
-        0xc & 0x400000001       & 0x1000000040000   & \dots \\
-        0xd & 0x0               & 0x0               & \dots \\
-        0xe & 0x8000000000000   & 0x800000000       & \dots \\
-        0xf & 0x8000400020000   & 0x800040002       & \dots
-    \end{array}
-\]
-
-The tables for \texttt{GIFT-128} can be generated in a similar way by looping
-through all 32 S-boxes instead of 16 on line \ref{lst:tablesbox}.
-
-\subsection{Using \texttt{vperm}}
-
-Nowadays, most instructions set architectures support single-instruction,
-multiple-data processing. The idea of such an SIMD system is to work on
-multiple data stored in vectors at once to speed up calculations. For A64, two
-types of vector processing are available:
-
-\begin{enumerate}
-    \item Advanced SIMD, known as NEON
-    \item Scalable Vector Extension (SVE)
-\end{enumerate}
-
-We will take a look at NEON as this is the type of vector processing supported
-by the Cortex-A73 processor.
-
-\subsubsection{ARM Neon}
-
-The register file of the NEON unit is made up of 32 quad-word (128-bit)
-registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
-registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
-sized lanes on which the vector instructions operate. Valid ways to interpret
-for example the register \texttt{V0} are:
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/V_register.pdf}
-    \caption{Divisions of the V0 register}
-\end{figure}
-
-NEON instructions interpret their operands' layouts (i.e. lane count and width)
-through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
-adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
-together and storing the result in \texttt{V0} can be done as follows:
-
-\begin{center}
-    \texttt{ADD V0.8H, V1.8H, V2.8H}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/vector_add.pdf}
-    \caption{Addition of two vector registers}
-\end{figure}
-
-The plenitude of different processing instructions allow flexible ways to
-further speed up algorithms having reached their optimizational limit on
-non-SIMD platforms. \texttt{vperm}, a general term standing for \textit{vector
-permute}, is a common instruction on SIMD machines. Called \texttt{TBL} on
-NEON, it is used for parallel table lookups and arbitrary permutations. It takes
-two inputs to perform a lanewise lookup:
-
-\begin{enumerate}
-    \item A register with lookup values
-    \item Two or more registers containing data
-\end{enumerate}
-
-\subsubsection{S-box lookup}
-
-This instruction can be used to implement S-box lookup of all 16 S-boxes in a
-single instruction. We do this by packing our 64-bit cipher state
-$s=s_{15}||s_{14}||\dots||s_0$ into a vector register $V_0$. Because we can
-only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane which
-neatly fits into the 128-bit registers. We then put the S-box itself into
-register $V_1$ which will be used as the data register for the table lookup.
-
-The confusion layer can now be performed through one \texttt{TBL} instruction:
-
-\begin{center}
-    \texttt{TBL V0.16B, V1.16B, V0.16B}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/tbl_example.pdf}
-    \caption{Performing the S-Box lookup in parallel}
-\end{figure}
-
-\subsection{Bitslicing}
-
-Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
-achieve a more efficient representation to operate on. The structure of
-\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
-cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
-i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
-individual S-boxes. This is equivalent to transposing the bit matrix.
-
-\[
-    S=\begin{bmatrix}
-        S_0\\
-        S_1\\
-        S_2\\
-        S_3
-    \end{bmatrix}
-    =\begin{bmatrix}
-        b_{60}b_{56}b_{52}\dots b_0\\
-        b_{61}b_{57}b_{53}\dots b_1\\
-        b_{62}b_{58}b_{54}\dots b_2\\
-        b_{63}b_{59}b_{55}\dots b_3
-    \end{bmatrix}
-\]
-
-\subsubsection{Parallel S-Boxes}
-
-This representation offers multiple advantages. We first note that computation
-of the S-box can be executed in parallel, similar to the \texttt{vperm}
-technique above. This can be done by finding an algorithmic way to apply the
-S-box which has already been proposed by the original \texttt{GIFT} authors:
-
-\begin{align*}
-    S_1&\leftarrow S_1\oplus (S_0\land S_2) \\
-    t&\leftarrow S_0\oplus (S_1\land S_3) \\
-    S_2&\leftarrow S_2\oplus (t\lor S_1) \\
-    S_0&\leftarrow S_3\oplus S_2 \\
-    S_1&\leftarrow S_1\oplus S_0 \\
-    S_0&\leftarrow \lnot S_0 \\
-    S_2&\leftarrow S_2\oplus (t\land S_1) \\
-    S_3&\leftarrow t
-\end{align*}
-
-This is very efficient as it only requires six XOR-, three AND and one OR
-operation.
-
-An important property of the permutation is the fact that bits always stay in
-their slice. This means we can decompose the permutation $P$ into four
-permutations $P_i,i\in\{0,1,2,3\}$ and apply these permutations
-separately to each slice. One possible way to implement a permutation $P_i$ in
-software is to mask off all bits individually, shift them to their correct
-position and OR them together:
-
-\[
-    P_i(S_i)=\bigvee_{k=0}^{15}{(S_i\land m_i) \ll s_i}
-\]
-
-This approach requires $47$ operations, meaning all four permutations require
-over $150$ operations which would present a major bottleneck to the round
-function. We can improve on this by working on multiple message blocks at once
-and using the aforementioned \texttt{vperm} instruction to implement the bit
-shuffling. We then need only four \texttt{TBL} instructions for the complete
-diffusion layer.
-
-\subsubsection{Using \texttt{vperm} for slice permutation}
-
-We cannot use the \texttt{TBL} instruction directly as we need to shuffle
-individual bits, but the smallest data we can operate on are bytes. We
-therefore encrypt $8n$ messages at once which allows us to create bytewise
-groupings. These messages are put into $4m$ registers with register $R_{4i}$
-containing $S_0$, register $R_{4i+1}$ containing $S_1$ and so forth. With block
-size $BS$ and register size $RS$, the following must hold:
-
-\[
-    8n\cdot BS=4m\cdot RS
-\]
-
-In the case of \texttt{GIFT-64} with $BS=64$ and ARM NEON with $RS=128$, we get
-
-\[
-    8n\cdot 64=4m\cdot 128\Leftrightarrow n=m
-\]
-
-$n=m=1$ would be a valid choice which yields eight messages divided into four
-registers. We choose $n=m=2$ so we can directly utilize the algorithm for bit
-packing presented by the original GIFT authors, although it is simple to adapt
-this algorithm to only four registers and eight messages by adjusting the
-\texttt{SWAPMOVE} shift and mask values.
-
-\subsubsection{Packing the data into bitslice format}
-
-Let $a,b,\dots,p$ be sixteen messages of length $64$ with subscripts denoting
-individual bits. We first put these messages into eight SIMD registers
-$V_0,V_1,\dots,V_7$:
-
-\begin{alignat*}{3}
-    V_0&=b||a\qquad V_4&&=j||&&i \\
-    V_1&=d||c\qquad V_5&&=l||&&k \\
-    V_2&=f||e\qquad V_6&&=n||&&m \\
-    V_3&=h||g\qquad V_7&&=p||&&o
-\end{alignat*}
-
-We then use the \texttt{SWAPMOVE} technique to bring the data into bitslice
-format. This operation operates on two registers $A,B$ using mask $M$ and shift
-value $N$. It swaps bits in $A$ masked by $(M\ll N)$ with bits in $B$ masked by
-$M$ in using only three XOR-, one AND- and two shift operations.
-
-\begin{align*}
-    &\text{SWAPMOVE}(A,B,M,N): \\
-    &\qquad T=((A\gg N)\oplus B)\land M \\
-    &\qquad B=B\oplus T \\
-    &\qquad A=A\oplus (T\ll N) \\
-\end{align*}
-
-One caveat of this approach is the fact that NEON registers cannot be shifted
-in their entirety due to the fact bits are not able to cross lanes. This leads
-to the problem of being able to shift at most two lanes of 64 bits at once. We
-thus need to implement the \texttt{shr(V,n)} and \texttt{shl(V,n)} operations
-on our own. This can be done by first extracting the 64-bit lanes $a,b$ out of
-$V=b||a$, shifting the lanes individually and finally shifting and ORing the
-crossing bits back into the other lane.
-
-\begin{alignat*}{2}
-    &\text{shl}(V,n): \\
-    &\qquad a,b&&=V[0],V[1] \\
-    &\qquad c&&=(a\gg (64-n)) \\
-    &\qquad a&&=(a\ll n) \\
-    &\qquad b&&=(b\ll n)\lor c \\
-    &\qquad V[0],V[1]&&=a,b
-\end{alignat*}
-
-The following operations group all $i$th bits of the messages $a,c,\dots,o$
-into bytes and puts these into the lower half of the registers $V_{i\bmod 8}$.
-The same is done for messages $b,d,\dots,p$, only differing in that the bytes
-are put into the upper half of the registers.
-
-\begin{align*}
-    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_4,V_5,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_6,V_7,0x5555\dots 55,1) \\
-    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_4,V_6,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_5,V_7,0x3333\dots 33,2) \\
-    &\text{SWAPMOVE}(V_0,V_4,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_1,V_5,0x0f0f\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4) \\
-\end{align*}
-
-With $Ax=o_xm_xk_xj_xg_xe_xc_xa_x$ and $Bx=p_xn_xl_xi_xh_xf_xd_xb_x$ denoting
-byte groups, our data now has the following permutation-friendly format:
-
-\[
-    \begin{array}{c|llllllll}
-        n & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
-        \hline
-        V_0 & A56 & A48 & A40 & A32 & A24 & A16 & A8 & A0 \\
-        V_1 & A57 & A49 & A41 & A33 & A25 & A17 & A9 & A1 \\
-        V_2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
-        V_3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
-        V_4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
-        V_5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
-        V_6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
-        V_7 & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
-    \end{array}
-\]
-
-\[
-    \begin{array}{c|llllllll}
-        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 \\
-        \hline
-        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8 & B0 \\
-        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9 & B1 \\
-        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 \\
-        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 \\
-        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 \\
-        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 \\
-        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 \\
-        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7
-    \end{array}
-\]
-
-We can now create permutation tables using the specification of the individual
-slice permutations $P_i$:
-
-\[
-    \begin{array}{c|llllllllllllllll}
-        j & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\
-        \hline
-        P_0(j) & 0 & 12 & 8 & 4 & 1 & 13 & 9 & 5 & 2 & 14 & 10 & 6 & 3 & 15 & 11 & 7 \\
-        P_1(j) & 4 & 0 & 12 & 8 & 5 & 1 & 13 & 9 & 6 & 2 & 14 & 10 & 7 & 3 & 15 & 11 \\
-        P_2(j) & 8 & 4 & 0 & 12 & 9 & 5 & 1 & 13 & 10 & 6 & 2 & 14 & 11 & 7 & 3 & 15 \\
-        P_3(j) & 12 & 8 & 4 & 0 & 13 & 9 & 5 & 1 & 14 & 10 & 6 & 2 & 15 & 11 & 7 & 3
-    \end{array}
-\]
-
-One thing to take note of is the original permutation values only showing where
-a given byte should land, not which byte belongs to a certain position - i.e.
-for $P_0$, byte 1 should land in position 12, but the byte belonging to
-position 1 is byte 4. Because \texttt{vperm} works in the latter way, we have
-to do some trivial rearrangements. In addition, because data for each slice is
-split between two registers $(V_0,V_4),(V_1,V_5),\dots$, the \texttt{vperm}
-operation needs to use these two registers as a data source. Using two data
-sources is supported by the \texttt{TBL} instruction by means of an additional
-parameter.
-
-Assuming the correct permutation values are put into registers
-$V_9,V_{10},\dots,V_{16}$, this now allows us to compute the permutation layer for
-all 16 blocks in only eight permutation instructions plus four additional vector
-copy instructions used for saving the original values of $V_0,V_1,V_2,V_3$
-before the permutation is applied.
-
-\begin{alignat*}{2}
-    &\texttt{MOV V17, V0}\qquad &&\texttt{MOV V18, V1} \\
-    &\texttt{MOV V19, V2}\qquad &&\texttt{MOV V20, V3} \\
-    &\texttt{TBL V0, \{V0, V4\}, V9}\qquad &&\texttt{TBL V1, \{V1, V5\}, V10} \\
-    &\texttt{TBL V2, \{V2, V6\}, V9}\qquad &&\texttt{TBL V3, \{V3, V7\}, V10} \\
-    &\texttt{TBL V4, \{V17, V4\}, V9}\qquad &&\texttt{TBL V5, \{V18, V5\}, V10} \\
-    &\texttt{TBL V6, \{V19, V6\}, V9}\qquad &&\texttt{TBL V7, \{V20, V7\}, V10}
-\end{alignat*}
-
-\subsubsection{Round key function}
-
-In contrast to packing and unpacking of data which is only done once in the
-beginning and end, a round key is derived for every round, so the round key
-derivation function needs to be as fast as possible. A simple but naive
-approach for one round would be to generate a single round key, copy it 15
-times and pack the resulting registers similar to how we proceed with the
-messages. Due to the cost of packing the messages, this is prohibitivly
-expensive. Because we know where each byte group ends up after packing, we can
-directly set the round key bits at the correct position. Extending these bits
-to bytes can then be done simply by repeatedly shifting and ORing the registers
-together.
-
-\subsection{Strategies for Camellia}
-
-\chapter{Implementation}
-
-We will provide and discuss implementations for the presented strategies in the
-C programming language. Although directly writing Assembler code could result
-in a small performance benefit, this generally increases the work necessary by
-an order of magnitude for only limited results. This is why we use NEON
-intrinsics.
-
-\subsection{NEON intrinsics}
-
-The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
-definitions including vector data types and C functions for working with these
-vectors. These functions are known as NEON intrinsics and give the programmer a
-high-level interface to most NEON instructions. Major advantages of this
-approach include the ease of development as the compiler takes over register
-allocation and load/store operations as well as performance benefits through
-compiler optimizations.
-
-Standard vector data types have the format \texttt{uintnxm\_t} with lane width
-$n$ in bits and and lane count $m$. Array types of the format
-\texttt{uintnxmxc\_t}, $c\in\{1,2,3,4\}$ are also defined which are used in
-operations requiring multiple parameters like \texttt{TBL} or pairwise
-load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
-after the operation name to indicate operation on a 128-bit register and end
-with the lane data format. Multiplying eight pairs of 16-bit numbers \texttt{a,b} for
-example can be done via the following:
-
-\begin{center}
-    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
-\end{center}
-
-The compiler allocates vector registers for \texttt{a}, \texttt{b} and
-\texttt{result} and assembles the intrinsic to \texttt{MUL Vr.8H, Va.8H,
-Vb.8H}. Necessary load and stores for the result and parameters are also
-handled automatically.
-
-\chapter{Evaluation}
-
-\chapter{Appendix}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.h}]
-#ifndef GIFT_VEC_SLICED_H
-#define GIFT_VEC_SLICED_H
-
-#include <stdint.h>
-#include <arm_neon.h>
-
-#define ROUNDS_GIFT_64 28
-
-// expose for benchmarking
-uint8x16_t shl(uint8x16_t v, int n);
-uint8x16_t shr(uint8x16_t v, int n);
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
-                                 uint8x16_t m, int n);
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2]);
-
-void gift_64_vec_sliced_init(void);
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2]);
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2]);
-
-#endif
-\end{lstlisting}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.c}]
-#include "gift_vec_sliced.h"
-
-#include <stddef.h>
-#include <stdio.h>
-#include <string.h>
-
-static uint64_t perm_u64[16] = {
-        0x1410050115110400UL, 0x1410050115110400UL,
-        0x0501151104001410UL, 0x0501151104001410UL,
-        0x1511040014100501UL, 0x1511040014100501UL,
-        0x0400141005011511UL, 0x0400141005011511UL,
-        0x1612070317130602UL, 0x1612070317130602UL,
-        0x0703171306021612UL, 0x0703171306021612UL,
-        0x1713060216120703UL, 0x1713060216120703UL,
-        0x0602161207031713UL, 0x0602161207031713UL
-};
-
-static uint64_t perm_inv_u64[16] = {
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL
-};
-
-static uint8x16x4_t perm[2];
-static uint8x16x4_t perm_inv[2];
-
-static uint8x16_t pack_mask_0;
-static uint8x16_t pack_mask_1;
-static uint8x16_t pack_mask_2;
-
-static const int round_constant[] = {
-        // rounds 0-15
-        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
-        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
-        // rounds 16-31
-        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
-        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
-        // rounds 32-47
-        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
-        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
-};
-
-uint8x16_t shl(uint8x16_t v, int n)
-{
-        // TODO check out vst
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l1 = l1 << n | (l0 >> (64 - n));
-        l0 <<= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
-}
-
-uint8x16_t shr(uint8x16_t v, int n)
-{
-        uint64_t l0 = vgetq_lane_u64(v, 0);
-        uint64_t l1 = vgetq_lane_u64(v, 1);
-        l0 = l0 >> n | (((l1 << (64 - n)) >> (64 - n)) << (64 - n));
-        l1 >>= n;
-        uint8x16_t ret;
-        ret = vsetq_lane_u64(l0, ret, 0);
-        ret = vsetq_lane_u64(l1, ret, 1);
-        return ret;
-}
-
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
-{
-
-        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
-        *b = veorq_u8(*b, t);
-        *a = veorq_u8(*a, shl(t, n));
-}
-
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
-{
-        // take care not to shift mask bits out of the register
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[0].val[3], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[2], &m[1].val[3], pack_mask_0, 1);
-
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[0].val[3], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[1], &m[1].val[3], pack_mask_1, 2);
-
-        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[1].val[0], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[1].val[2], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[1].val[1], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[3], &m[1].val[3], pack_mask_2, 4);
-}
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
-{
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[3] = t;
-
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-        t            = veorq_u8(cs[1].val[0], vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(cs[1].val[3], cs[1].val[2]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[3] = t;
-}
-
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16_t t = cs[0].val[3];
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[3] = veorq_u8(cs[0].val[0], cs[0].val[2]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(t, vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-
-        t            = cs[1].val[3];
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[3] = veorq_u8(cs[1].val[0], cs[1].val[2]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(t, vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-}
-
-void gift_64_vec_sliced_permute(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm[1].val[3]);
-}
-
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm_inv[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm_inv[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm_inv[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm_inv[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm_inv[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm_inv[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm_inv[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm_inv[1].val[3]);
-}
-
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2])
-{
-        // TODO shift-and-or
-
-        uint64_t key_state[] = {key[0], key[1]};
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                int v = (key_state[0] >> 0 ) & 0xffff;
-                int u = (key_state[0] >> 16) & 0xffff;
-
-                // add round key (RK=U||V)
-                uint64_t round_key[8] = { 0x0UL };
-                for (size_t i = 0; i < 8; i++) {
-                        int key_bit_v   = (v >> (2 * i + 0)) & 0x1;
-                        int key_bit_u   = (u >> (2 * i + 0)) & 0x1;
-                        round_key[0] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[1] ^= (uint64_t)key_bit_u << (i * 8);
-
-                        key_bit_v   = (v >> (2 * i + 1)) & 0x1;
-                        key_bit_u   = (u >> (2 * i + 1)) & 0x1;
-                        round_key[4] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[5] ^= (uint64_t)key_bit_u << (i * 8);
-                }
-
-                // add single bit
-                round_key[7] ^= 1UL << (7 * 8);
-
-                // add round constants
-                round_key[3] ^= ((round_constant[round] >> 0) & 0x1) << (0 * 8);
-                round_key[7] ^= ((round_constant[round] >> 1) & 0x1) << (0 * 8);
-                round_key[3] ^= ((round_constant[round] >> 2) & 0x1) << (1 * 8);
-                round_key[7] ^= ((round_constant[round] >> 3) & 0x1) << (1 * 8);
-                round_key[3] ^= ((round_constant[round] >> 4) & 0x1) << (2 * 8);
-                round_key[7] ^= ((round_constant[round] >> 5) & 0x1) << (2 * 8);
-
-                // extend bits to bytes
-                for (size_t i = 0; i < 8; i++) {
-                        round_key[i] |= round_key[i] << 1;
-                        round_key[i] |= round_key[i] << 2;
-                        round_key[i] |= round_key[i] << 4;
-                }
-
-                // load into vector registers (upper and lower part are same)
-                round_keys[round][0].val[0] = vdupq_n_u64(round_key[0]);
-                round_keys[round][0].val[1] = vdupq_n_u64(round_key[1]);
-                round_keys[round][0].val[2] = vdupq_n_u64(round_key[2]);
-                round_keys[round][0].val[3] = vdupq_n_u64(round_key[3]);
-                round_keys[round][1].val[0] = vdupq_n_u64(round_key[4]);
-                round_keys[round][1].val[1] = vdupq_n_u64(round_key[5]);
-                round_keys[round][1].val[2] = vdupq_n_u64(round_key[6]);
-                round_keys[round][1].val[3] = vdupq_n_u64(round_key[7]);
-
-                // update key state
-                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
-                int k1 = (key_state[0] >> 16) & 0xffffUL;
-                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
-                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
-                key_state[0] >>= 32;
-                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
-                key_state[1] >>= 32;
-                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
-        }
-}
-
-void gift_64_vec_sliced_init(void)
-{
-        // permutations
-        perm[0] = vld1q_u8_x4((uint8_t*)&perm_u64[0]);
-        perm[1] = vld1q_u8_x4((uint8_t*)&perm_u64[8]);
-
-        // inverse permutations
-        perm_inv[0] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[0]);
-        perm_inv[1] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[8]);
-
-        // packing masks
-        pack_mask_0 = vdupq_n_u8(0x55);
-        pack_mask_1 = vdupq_n_u8(0x33);
-        pack_mask_2 = vdupq_n_u8(0x0f);
-}
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&m[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&m[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                gift_64_vec_sliced_subcells(s);
-                gift_64_vec_sliced_permute(s);
-
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&c[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&c[8], s[1]);
-}
-
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&c[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&c[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-
-                gift_64_vec_sliced_permute_inv(s);
-                gift_64_vec_sliced_subcells_inv(s);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&m[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&m[8], s[1]);
-}
-\end{lstlisting}
-
-\chapter*{Acknowledgements}
-
-I want to thank ...
+\include{chapters/introduction}
+\include{chapters/implementation-strategies}
+\include{chapters/implementation}
+\include{chapters/evaluation}
+\include{chapters/acknowledgements}
+\include{chapters/appendix}
 

### 2023-03-23 impl/vector/gift_vec_sbox.c

@@ -4,6 +4,14 @@
 #include <arm_neon.h>
 #include <string.h>
 
+static uint64_t sbox_vec_u64[2] = {
+        0x09030f060c040a01UL, 0x0e080005070b0d02UL
+};
+
+static uint64_t sbox_vec_inv_u64[2] = {
+0x0b040c020608000dUL, 0x050f09030a01070eUL
+};
+
 static uint8x16_t sbox_vec;
 static uint8x16_t sbox_vec_inv;
 
@@ -166,12 +174,10 @@ void gift_64_vec_sbox_generate_round_keys(uint8x16_t round_keys[ROUNDS_GIFT_64],
 void gift_64_vec_sbox_init(void)
 {
         // construct sbox_vec
-        sbox_vec = vsetq_lane_u64(0x09030f060c040a01UL, sbox_vec, 0);
-        sbox_vec = vsetq_lane_u64(0x0e080005070b0d02UL, sbox_vec, 1);
+        sbox_vec = vld1q_u64(sbox_vec_u64);
 
         // construct sbox_vec_inv
-        sbox_vec_inv = vsetq_lane_u64(0x0b040c020608000dUL, sbox_vec_inv, 0);
-        sbox_vec_inv = vsetq_lane_u64(0x050f09030a01070eUL, sbox_vec_inv, 1);
+        sbox_vec_inv = vld1q_u64(sbox_vec_inv_u64);
 }
 
 uint64_t gift_64_vec_sbox_encrypt(const uint64_t m, const uint64_t key[2])

### 2023-03-23 Paper/chapters/implementation-strategies.tex
table-based implementation

@@
 
-\begin{lstlisting}[caption={Table construction algorithm}, escapechar=|]
+\begin{lstlisting}[caption={Table construction algorithm}, label={lst:tables}, escapechar=|]
     tables <- [][]
@@
 
-\subsection{Strategies for Camellia}
+\section{Strategies for Camellia}

### 2023-03-23 Paper/chapters/implementation.tex
table-based implementation

@@
 in a small performance benefit, this generally increases the work necessary by
-an order of magnitude for only limited results. This is why we use NEON
-intrinsics.
+an order of magnitude for only limited results. Instruction-level optimization
+and in particular register allocation is left to the compiler.
+
+\section{Table-based}
+
+We generate a 16 by 16 array of \texttt{uint64\_t} as explained in \ref{lst:tablesbox}.
+The round function then consists of extracting each nibble, looking up the
+corresponding table value and ORing this to the new cipher state.
+
+\begin{lstlisting}[caption={Table-based round function}, language=c]
+static const uint64_t table[16][16] = {
+        { 0x0000000000000001UL, 0x0008000000020000UL,
+          0x0000000400000000UL, 0x0008000400000000UL,
+          0x0000000400020000UL, 0x0008000400020001UL,
+          0x0000000000020001UL, 0x0008000000000001UL,
+          0x0000000000020000UL, 0x0008000400000001UL,
+          0x0008000000020001UL, 0x0000000400020001UL,
+          0x0000000400000001UL, 0x0000000000000000UL,
+          0x0008000000000000UL, 0x0008000400020000UL },
+        // ...
+        { 0x0000000010000000UL, 0x0000200000008000UL,
+          0x4000000000000000UL, 0x4000000000008000UL,
+          0x4000200000000000UL, 0x4000200010008000UL,
+          0x0000200010000000UL, 0x0000000010008000UL,
+          0x0000200000000000UL, 0x4000000010008000UL,
+          0x0000200010008000UL, 0x4000200010000000UL,
+          0x4000000010000000UL, 0x0000000000000000UL,
+          0x0000000000008000UL, 0x4000200000008000UL }
+};
+
+uint64_t gift_64_table_subperm(const uint64_t cipher_state)
+{
+        uint64_t new_cipher_state = 0;
+
+        for (size_t i = 0; i < 16; i++) {
+                int nibble = (cipher_state >> (i * 4)) & 0xf;
+                new_cipher_state |= table[i][nibble];
+        }
+
+        return new_cipher_state;
+}
+\end{lstlisting}
+
+The round function is similar for all strategies: for each round, we extract 16
+bits from the key state which are then XORed to the round key, together with
+the precomputed round constants and the single bit. Key state updating is done
+using masks and shifts only which gets optimized by the compiler through
+bit-field extraction instructions which can extract, shift and insert a bit
+field at once. Intermediate values are also optimized away.
+
+\begin{lstlisting}[caption={Round key generation function}, language=c]
+void gift_64_table_generate_round_keys(uint64_t round_keys[restrict ROUNDS_GIFT_64],
+                                       const uint64_t key[restrict 2])
+{
+        uint64_t key_state[] = {key[0], key[1]};
+        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
+                int v = (key_state[0] >> 0 ) & 0xffff;
+                int u = (key_state[0] >> 16) & 0xffff;
+
+                // add round key (RK=U||V)
+                round_keys[round] = 0UL;
+                for (size_t i = 0; i < 16; i++) {
+                        int key_bit_v   = (v >> i)  & 0x1;
+                        int key_bit_u   = (u >> i)  & 0x1;
+                        round_keys[round] ^= (uint64_t)key_bit_v << (i * 4 + 0);
+                        round_keys[round] ^= (uint64_t)key_bit_u << (i * 4 + 1);
+                }
+
+                // add single bit
+                round_keys[round] ^= 1UL << 63;
+
+                // add round constants
+                round_keys[round] ^= ((round_constant[round] >> 0) & 0x1) << 3;
+                round_keys[round] ^= ((round_constant[round] >> 1) & 0x1) << 7;
+                round_keys[round] ^= ((round_constant[round] >> 2) & 0x1) << 11;
+                round_keys[round] ^= ((round_constant[round] >> 3) & 0x1) << 15;
+                round_keys[round] ^= ((round_constant[round] >> 4) & 0x1) << 19;
+                round_keys[round] ^= ((round_constant[round] >> 5) & 0x1) << 23;
+
+                // update key state
+                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
+                int k1 = (key_state[0] >> 16) & 0xffffUL;
+                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
+                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
+                key_state[0] >>= 32;
+                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
+                key_state[1] >>= 32;
+                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
+        }
+}
+\end{lstlisting}
+
+While this approach does not rely on ARM-specific instructions, the next two
+strategies do. This is why we use NEON intrinsics.
+
+\section{Using \texttt{vperm}}
 
@@
 \texttt{b} and \texttt{result} and assembles the intrinsic to \texttt{MUL
-Vr.8H, Va.8H, Vb.8H}. Necessary load and stores for the result and parameters
+Vr.8H, Va.8H, Vb.8H}. Necessary loads and stores for the result and parameters
 are also handled automatically. Of special interest to us are the following
@@
 
-The implementations closely follow the aforementioned strategies;
-instruction-level optimization is left to the compiler.
+\section{Bitslicing}

### 2023-03-23 impl/benchmark.c

@@ -37,7 +37,6 @@ void benchmark_gift_64(void)
         uint64_t round_keys[ROUNDS_GIFT_64];
 
         uint64_t t0, t1;
-        // TODO take into account repeated execution due to cache effects
         MEASURE(gift_64_generate_round_keys(round_keys, key), t0, t1);
         for (int i = 0; i < 5; i++) {
                 MEASURE(gift_64_subcells(m), t0, t1);
@@ -60,7 +59,6 @@ void benchmark_gift_128(void)
         uint8_t round_keys[ROUNDS_GIFT_128][32];
 
         uint64_t t0, t1;
-        // TODO take into account repeated execution due to cache effects
         MEASURE(gift_128_generate_round_keys(round_keys, key), t0, t1);
         for (int i = 0; i < 5; i++) {
                 MEASURE(gift_128_subcells(m), t0, t1);
@@ -69,6 +67,24 @@ void benchmark_gift_128(void)
                 MEASURE(gift_128_permute(m), t0, t1);
         }
         MEASURE(gift_128_encrypt(c, m, key), t0, t1);
+}
+
+void benchmark_gift_64_table(void)
+{
+        uint64_t key[2];
+        rand_bytes((uint8_t*)key, sizeof(key));
+
+        uint64_t m;
+        rand_bytes((uint8_t*)&m, sizeof(m));
+
+        uint64_t round_keys[ROUNDS_GIFT_64];
+
+        uint64_t t0, t1;
+        while (1) {
+                rand_bytes((uint8_t*)key, sizeof(key));
+                MEASURE(gift_64_table_generate_round_keys(round_keys, key), t0, t1);
+                /* MEASURE(gift_64_table_encrypt(m, key), t0, t1); */
+        }
 }
 
 void benchmark_gift_64_sliced(void)
@@ -83,7 +99,6 @@ void benchmark_gift_64_sliced(void)
         uint8_t round_keys[ROUNDS_GIFT_128][32];
 
         uint64_t t0, t1;
-        // TODO take into account repeated execution due to cache effects
         MEASURE(gift_64_sliced_generate_round_keys(round_keys, key), t0, t1);
         MEASURE(gift_64_sliced_subcells(m), t0, t1);
         MEASURE(gift_64_sliced_permute(m), t0, t1);
@@ -105,7 +120,6 @@ void benchmark_gift_64_vec_sbox(void)
[…]

### 2023-03-23 impl/table/gift_table.c

@@ -31,7 +31,8 @@ static const uint64_t tables[16][16] = {
         { 0x0000000010000000UL, 0x0000200000008000UL, 0x4000000000000000UL, 0x4000000000008000UL, 0x4000200000000000UL, 0x4000200010008000UL, 0x0000200010000000UL, 0x0000000010008000UL, 0x0000200000000000UL, 0x4000000010008000UL, 0x0000200010008000UL, 0x4000200010000000UL, 0x4000000010000000UL, 0x0000000000000000UL, 0x0000000000008000UL, 0x4000200000008000UL }
 };
 
-void gift_64_table_generate_round_keys(uint64_t round_keys[ROUNDS_GIFT_64], const uint64_t key[2])
+void gift_64_table_generate_round_keys(uint64_t round_keys[restrict ROUNDS_GIFT_64],
+                                       const uint64_t key[restrict 2])
 {
         uint64_t key_state[] = {key[0], key[1]};
         for (int round = 0; round < ROUNDS_GIFT_64; round++) {
@@ -82,7 +83,7 @@ uint64_t gift_64_table_subperm(const uint64_t cipher_state)
         return new_cipher_state;
 }
 
-uint64_t gift_64_table_encrypt(const uint64_t m, const uint64_t key[2])
+uint64_t gift_64_table_encrypt(const uint64_t m, const uint64_t key[restrict 2])
 {
         uint64_t c = m;
 

### 2023-03-23 impl/table/gift_table.h

@@ -5,12 +5,12 @@
 
 #define ROUNDS_GIFT_64 28
 
-void gift_64_table_generate_round_keys(uint64_t round_keys[ROUNDS_GIFT_64],
-                                 const uint64_t key[2]);
+void gift_64_table_generate_round_keys(uint64_t round_keys[restrict ROUNDS_GIFT_64],
+                                 const uint64_t key[restrict 2]);
 
 uint64_t gift_64_table_subperm(const uint64_t cipher_state);
 
 // can only encrypt using table technique!
-uint64_t gift_64_table_encrypt(const uint64_t m, const uint64_t key[2]);
+uint64_t gift_64_table_encrypt(const uint64_t m, const uint64_t key[restrict 2]);
 
 #endif

### 2023-03-23 impl/vector/gift_vec_sliced.c

@@ -260,11 +260,14 @@ void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROU
                         round_key[i] |= round_key[i] << 4;
                 }
 
-                round_keys[round][0].val[0] = vld1q_u64(&round_key[0]);
-                round_keys[round][0].val[1] = vld1q_u64(&round_key[2]);
+                round_keys[round][0].val[0] = vsetq_lane_u64(round_key[0], round_keys[round][0].val[0], 0);
+                round_keys[round][0].val[0] = vsetq_lane_u64(round_key[1], round_keys[round][0].val[0], 0);
+                round_keys[round][0].val[1] = vsetq_lane_u64(round_key[2], round_keys[round][0].val[1], 0);
+                round_keys[round][0].val[1] = vsetq_lane_u64(round_key[3], round_keys[round][0].val[1], 0);
                 round_keys[round][0].val[2] = vdupq_n_u8(0);
-                round_keys[round][0].val[3] = vld1q_u64(&round_key[4]);
-                round_keys[round][1] = round_keys[round][0];
+                round_keys[round][0].val[3] = vsetq_lane_u64(round_key[4], round_keys[round][0].val[3], 0);
+                round_keys[round][0].val[3] = vsetq_lane_u64(round_key[5], round_keys[round][0].val[3], 0);
+                round_keys[round][1]        = round_keys[round][0];
 
                 // update key state
                 int k0 = (key_state[0] >> 0 ) & 0xffffUL;

### 2023-03-28 impl/table/gift_table.c
xor instead of or

@@ -77,7 +77,7 @@ uint64_t gift_64_table_subperm(const uint64_t cipher_state)
 
         for (size_t i = 0; i < 16; i++) {
                 int nibble = (cipher_state >> (i * 4)) & 0xf;
-                new_cipher_state |= tables[i][nibble];
+                new_cipher_state ^= tables[i][nibble];
         }
 
         return new_cipher_state;

### 2023-04-02 Paper/chapters/appendix.tex
remove implementation.tex, fancyhdr

@@
-\chapter{Appendix}
+\appendix
+\chapter{C implementations}
+\label{app:cimpl}
 
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.h}]
-#ifndef GIFT_VEC_SLICED_H
-#define GIFT_VEC_SLICED_H
+\section{Implementations for SPN}
+\subsection{Table-based}
 
-#include <stdint.h>
-#include <arm_neon.h>
+\lstinputlisting[language=c, caption={\texttt{gift\_table.h}}]{../impl/table/gift_table.h}
+\lstinputlisting[language=c, caption={\texttt{gift\_table.c}}]{../impl/table/gift_table.c}
 
-#define ROUNDS_GIFT_64 28
+\subsection{Using \texttt{vperm}}
 
-// expose for benchmarking
-uint8x16_t shl(uint8x16_t v, int n);
-uint8x16_t shr(uint8x16_t v, int n);
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
-                                 uint8x16_t m, int n);
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);
+\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.h}}]{../impl/vector/gift_vec_sbox.h}
+\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.c}}]{../impl/vector/gift_vec_sbox.c}
 
-void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2]);
+\subsection{Bitslicing}
 
-void gift_64_vec_sliced_init(void);
+\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.h}}]{../impl/vector/gift_vec_sliced.h}
+\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.c}}]{../impl/vector/gift_vec_sliced.c}
 
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2]);
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2]);
-
-#endif
-\end{lstlisting}
-
-\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.c}]
-#include "gift_vec_sliced.h"
-
-#include <stddef.h>
-#include <stdio.h>
-#include <string.h>
-
-static uint64_t perm_u64[16] = {
-        0x1410050115110400UL, 0x1410050115110400UL,
-        0x0501151104001410UL, 0x0501151104001410UL,
-        0x1511040014100501UL, 0x1511040014100501UL,
-        0x0400141005011511UL, 0x0400141005011511UL,
-        0x1612070317130602UL, 0x1612070317130602UL,
-        0x0703171306021612UL, 0x0703171306021612UL,
-        0x1713060216120703UL, 0x1713060216120703UL,
-        0x0602161207031713UL, 0x0602161207031713UL
-};
-
-static uint64_t perm_inv_u64[16] = {
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1317030712160206UL, 0x1317030712160206UL,
-        0x1511050114100400UL, 0x1511050114100400UL,
-        0x1713070316120602UL, 0x1713070316120602UL,
-        0x1115010510140004UL, 0x1115010510140004UL
-};
-
-static uint8x16x4_t perm[2];
-static uint8x16x4_t perm_inv[2];
-
-static uint8x16_t pack_mask_0;
-static uint8x16_t pack_mask_1;
-static uint8x16_t pack_mask_2;
-
-static const int round_constant[] = {
-        // rounds 0-15
-        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
-        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
-        // rounds 16-31
-        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
-        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
-        // rounds 32-47
-        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
-        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
-};
-
-uint8x16_t shl(uint8x16_t v, int n)
-{
-        uint64_t l[2];
-        vst1q_u64(l, v);
-        l[1] = l[1] << n | (l[0] >> (64 - n));
-        l[1] <<= n;
-        return vreinterpretq_u8_u64(vld1q_u64(l));
-}
-
-uint8x16_t shr(uint8x16_t v, int n)
-{
-        uint64_t l[2];
-        vst1q_u64(l, v);
-        l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
-        l[1] >>= n;
-        return vreinterpretq_u8_u64(vld1q_u64(l));
-}
-
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
-{
-
-        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
-        *b = veorq_u8(*b, t);
-        *a = veorq_u8(*a, shl(t, n));
-}
-
-void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
-{
-        // take care not to shift mask bits out of the register
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[0].val[3], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[1], pack_mask_0, 1);
-        gift_64_vec_sliced_swapmove(&m[1].val[2], &m[1].val[3], pack_mask_0, 1);
-
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[0].val[3], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[2], pack_mask_1, 2);
-        gift_64_vec_sliced_swapmove(&m[1].val[1], &m[1].val[3], pack_mask_1, 2);
-
-        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
-        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[1].val[0], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[1].val[2], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[1].val[1], pack_mask_2, 4);
-        gift_64_vec_sliced_swapmove(&m[0].val[3], &m[1].val[3], pack_mask_2, 4);
-}
-
-void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
-{
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[3] = t;
-
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-        t            = veorq_u8(cs[1].val[0], vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(cs[1].val[3], cs[1].val[2]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[3] = t;
-}
-
-void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16_t t = cs[0].val[3];
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
-        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
-        cs[0].val[3] = veorq_u8(cs[0].val[0], cs[0].val[2]);
-        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
-        cs[0].val[0] = veorq_u8(t, vandq_u8(cs[0].val[1], cs[0].val[3]));
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-
-        t            = cs[1].val[3];
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
-        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
-        cs[1].val[3] = veorq_u8(cs[1].val[0], cs[1].val[2]);
-        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
-        cs[1].val[0] = veorq_u8(t, vandq_u8(cs[1].val[1], cs[1].val[3]));
-        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
-}
-
-void gift_64_vec_sliced_permute(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm[1].val[3]);
-}
-
-void gift_64_vec_sliced_permute_inv(uint8x16x4_t cs[restrict 2])
-{
-        uint8x16x2_t pairs[4] = {
-                { .val = { cs[0].val[0], cs[1].val[0] }},
-                { .val = { cs[0].val[1], cs[1].val[1] }},
-                { .val = { cs[0].val[2], cs[1].val[2] }},
-                { .val = { cs[0].val[3], cs[1].val[3] }},
-        };
-
-        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm_inv[0].val[0]);
-        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm_inv[0].val[1]);
-        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm_inv[0].val[2]);
-        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm_inv[0].val[3]);
-
-        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm_inv[1].val[0]);
-        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm_inv[1].val[1]);
-        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm_inv[1].val[2]);
-        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm_inv[1].val[3]);
-}
-
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
-                                            const uint64_t key[restrict 2])
-{
-        // TODO shift-and-or
-
-        uint64_t key_state[] = {key[0], key[1]};
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                int v = (key_state[0] >> 0 ) & 0xffff;
-                int u = (key_state[0] >> 16) & 0xffff;
-
-                // add round key (RK=U||V)
-                uint64_t round_key[8] = { 0x0UL };
-                for (size_t i = 0; i < 8; i++) {
-                        int key_bit_v   = (v >> (2 * i + 0)) & 0x1;
-                        int key_bit_u   = (u >> (2 * i + 0)) & 0x1;
-                        round_key[0] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[1] ^= (uint64_t)key_bit_u << (i * 8);
-
-                        key_bit_v   = (v >> (2 * i + 1)) & 0x1;
-                        key_bit_u   = (u >> (2 * i + 1)) & 0x1;
-                        round_key[4] ^= (uint64_t)key_bit_v << (i * 8);
-                        round_key[5] ^= (uint64_t)key_bit_u << (i * 8);
-                }
-
-                // add single bit
-                round_key[7] ^= 1UL << (7 * 8);
-
-                // add round constants
-                round_key[3] ^= ((round_constant[round] >> 0) & 0x1) << (0 * 8);
-                round_key[7] ^= ((round_constant[round] >> 1) & 0x1) << (0 * 8);
-                round_key[3] ^= ((round_constant[round] >> 2) & 0x1) << (1 * 8);
-                round_key[7] ^= ((round_constant[round] >> 3) & 0x1) << (1 * 8);
-                round_key[3] ^= ((round_constant[round] >> 4) & 0x1) << (2 * 8);
-                round_key[7] ^= ((round_constant[round] >> 5) & 0x1) << (2 * 8);
-
-                // extend bits to bytes
-                for (size_t i = 0; i < 8; i++) {
-                        round_key[i] |= round_key[i] << 1;
-                        round_key[i] |= round_key[i] << 2;
-                        round_key[i] |= round_key[i] << 4;
-                }
-
-                // load into vector registers (upper and lower part are same)
-                round_keys[round][0].val[0] = vdupq_n_u64(round_key[0]);
-                round_keys[round][0].val[1] = vdupq_n_u64(round_key[1]);
-                round_keys[round][0].val[2] = vdupq_n_u64(round_key[2]);
-                round_keys[round][0].val[3] = vdupq_n_u64(round_key[3]);
-                round_keys[round][1].val[0] = vdupq_n_u64(round_key[4]);
-                round_keys[round][1].val[1] = vdupq_n_u64(round_key[5]);
-                round_keys[round][1].val[2] = vdupq_n_u64(round_key[6]);
-                round_keys[round][1].val[3] = vdupq_n_u64(round_key[7]);
-
-                // update key state
-                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
-                int k1 = (key_state[0] >> 16) & 0xffffUL;
-                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
-                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
-                key_state[0] >>= 32;
-                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
-                key_state[1] >>= 32;
-                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
-        }
-}
-
-void gift_64_vec_sliced_init(void)
-{
-        // permutations
-        perm[0] = vld1q_u8_x4((uint8_t*)&perm_u64[0]);
-        perm[1] = vld1q_u8_x4((uint8_t*)&perm_u64[8]);
-
-        // inverse permutations
-        perm_inv[0] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[0]);
-        perm_inv[1] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[8]);
-
-        // packing masks
-        pack_mask_0 = vdupq_n_u8(0x55);
-        pack_mask_1 = vdupq_n_u8(0x33);
-        pack_mask_2 = vdupq_n_u8(0x0f);
-}
-
-void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
-                                const uint64_t m[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&m[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&m[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
-                gift_64_vec_sliced_subcells(s);
-                gift_64_vec_sliced_permute(s);
-
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&c[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&c[8], s[1]);
-}
-
-void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
-                                const uint64_t c[restrict 16],
-                                const uint64_t key[restrict 2])
-{
-        uint8x16x4_t s[2];
-        s[0] = vld1q_u8_x4((uint8_t*)&c[0]);
-        s[1] = vld1q_u8_x4((uint8_t*)&c[8]);
-        gift_64_vec_sliced_bits_pack(s);
-
-        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
-        gift_64_vec_sliced_generate_round_keys(round_keys, key);
-
-        for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
-                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
-                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
-                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
-                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
-                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
-                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
-                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
-                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
-
-                gift_64_vec_sliced_permute_inv(s);
-                gift_64_vec_sliced_subcells_inv(s);
-        }
-
-        gift_64_vec_sliced_bits_pack(s);
-        vst1q_u8_x4((uint8_t*)&m[0], s[0]);
-        vst1q_u8_x4((uint8_t*)&m[8], s[1]);
-}
-\end{lstlisting}
+\chapter{Lorem dolor}

### 2023-04-02 Paper/chapters/implementation-strategies.tex
remove implementation.tex, fancyhdr

@@
     \item A register with lookup values
-    \item Two or more registers containing data
+    \item One or more registers containing data
 \end{enumerate}
@@
 We can now create permutation tables using the specification of the individual
-slice permutations $P_i$ which are then applied to $V_i$ and $V_{i+4}$:
+slice permutations $P_i$ which are then applied to $V_i$ and $V_{i+4}$
+respectively:
 
@@
 
+\section{Implementation}
+
+Implementations in the C programming language for the presented strategies can
+be found in Appendix \ref{app:cimpl}. Although directly writing Assembler code could
+result in a small performance benefit, this generally increases the work
+necessary by an order of magnitude for only limited results. Instruction-level
+optimization and in particular register allocation is left to the compiler.
+
+\subsection{NEON intrinsics}
+
+The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
+definitions including vector data types and C functions for working with these
+vectors. These functions are known as NEON intrinsics \cite{neonintr:2022} and
+give the programmer a high-level interface to most NEON instructions. Major
+advantages of this approach include the ease of development as the compiler
+takes over register allocation and load/store operations as well as performance
+benefits through compiler optimizations.
+
+Standard vector data types have the format \texttt{uintnxm\_t} with lane width
+$n$ in bits and and lane count $m$. Array types of the format
+\texttt{uintnxmxc\_t}, $c\in\{2,3,4\}$ are also defined which are used in
+operations requiring multiple parameters like \texttt{TBL} or pairwise
+load/stores. Intrinsics include the operation name and lane data format as well
+as an optional \texttt{q} suffix to indicate operation on a 128-bit register.
+Multiplying eight pairs of 16-bit numbers \texttt{a,b} for example can be done
+via the following:
+
+\begin{center}
+    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
+\end{center}
+
+In this case, the compiler allocates vector registers for \texttt{a},
+\texttt{b} and \texttt{result} and assembles the intrinsic to \texttt{MUL
+Vr.8H, Va.8H, Vb.8H}. Necessary loads and stores for the result and parameters
+are also handled automatically. Of special interest to us are the following
+intrinsics, each existing in different variants with different lane widths and
+also array types: \\
+
+\begin{table}[h!]
+    \centering
+    \footnotesize
+    \begin{tabularx}{\textwidth}{ll|X}
+        Intrinsic && Description \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vreinterpretq\_u8\_u64(uint64x2\_t)} & Explicit casting \\
+        \hline
+        \texttt{uint64\_t} & \texttt{vgetq\_lane\_u64(void)} & Extract a single lane \\
+        \hline
+        \texttt{void} & \texttt{vsetq\_lane\_u64(uint64\_t)} & Insert a single lane \\
+        \hline
+        \texttt{uint64x2\_t} & \texttt{vdupq\_n\_u64(uint64\_t)} & Initialize all lanes to same value \\
+        \hline
+        \texttt{void} & \texttt{vst1q\_u64(uint64\_t*, uint64x2\_t)} & Store from register to memory \\
+        \hline
+        \texttt{uint64x2\_t} & \texttt{vld1q\_u64(uint64\_t*, uint64x2\_t)} & Load from memory to register \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{veorq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise XOR \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vandq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise AND \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vorrq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise OR \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vmvnq\_u8(uint8x16\_t)} & bitwise NOT \\
+        \hline
+        \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
+    \end{tabularx}
+\end{table}
+
 \section{Strategies for Camellia}

### 2023-04-02 Paper/chapters/implementation-strategies.tex
begin work on benchmark

@@
     &\qquad c&&=(a\gg (64-n)) \\
-    &\qquad a&&=(a\ll n) \\
-    &\qquad b&&=(b\ll n)\lor c \\
-    &\qquad V[0],V[1]&&=a,b
+    &\qquad V[0]&&=(a\ll n) \\
+    &\qquad V[1]&&=(b\ll n)\lor c
 \end{alignat*}
 
-The following operations group all $i$th bits of the messages $a,c,\dots,o$
+The following operations group all $i$-th bits of the messages $a,c,\dots,o$
 into bytes and puts these into the lower half of the registers $V_{i\bmod 8}$.
@@
 
+\scriptsize
 \[
-    \scriptsize
     \begin{array}{c|llllllll|llllllll}
@@
     \end{tabularx}
+    \caption{Common NEON intrinsics}
 \end{table}

### 2023-04-02 impl/benchmark.c

@@ -13,11 +13,24 @@
 #include <stdint.h>
 #include <arm_neon.h>
 
-#define MEASURE(code, t0, t1)\
+#define N 10000
+
+#define MEASURE(code)\
+        {\
+        uint64_t t0, t1;\
         asm volatile("mrs %[c], PMCCNTR_EL0" : [c] "=r"(t0));\
         code;\
         asm volatile("mrs %[c], PMCCNTR_EL0" : [c] "=r"(t1));\
-        printf(#code ": took %ld cycles\n", t1 - t0);
+        printf(#code ": took %ld cycles\n", t1 - t0);\
+        }
+
+#define TIME(code)({\
+        uint64_t t0, t1;\
+        asm volatile("mrs %[c], PMCCNTR_EL0" : [c] "=r"(t0));\
+        code;\
+        asm volatile("mrs %[c], PMCCNTR_EL0" : [c] "=r"(t1));\
+        t1 - t0;\
+        })
 
 void rand_bytes(uint8_t m[], size_t n)
 {
@@ -28,155 +41,153 @@ void rand_bytes(uint8_t m[], size_t n)
 
 void benchmark_gift_64(void)
 {
+        printf("benchmarking GIFT_64...\n");
+
         uint64_t key[2];
-        rand_bytes((uint8_t*)key, sizeof(key));
-
         uint64_t m;
-        rand_bytes((uint8_t*)&m, sizeof(m));
-
         uint64_t round_keys[ROUNDS_GIFT_64];
 
-        uint64_t t0, t1;
-        MEASURE(gift_64_generate_round_keys(round_keys, key), t0, t1);
-        for (int i = 0; i < 5; i++) {
-                MEASURE(gift_64_subcells(m), t0, t1);
+        uint64_t cycles[4] = { 0UL };
+        for (int i = 0; i < N; i++) {
+                cycles[0] += TIME(gift_64_generate_round_keys(round_keys, key));
+                cycles[1] += TIME(gift_64_permute(m));
+                cycles[2] += TIME(gift_64_subcells(m));
+                cycles[3] += TIME(gift_64_encrypt(m, key));
         }
-        for (int i = 0; i < 5; i++) {
-                MEASURE(gift_64_permute(m), t0, t1);
-        }
-        MEASURE(gift_64_encrypt(m, key), t0, t1);
+
+        printf("%f %f %f %f\n",
+               cycles[0] / (float)N,
+               cycles[1] / (float)N,
+               cycles[2] / (float)N,
+               cycles[3] / (float)N);
[…]

### 2023-04-02 impl/naive/gift_sliced.c

@@ -198,7 +198,7 @@ void gift_64_sliced_permute_inv(uint64_t s[8])
         dst[63] = src[7][1];
 }
 
-void gift_64_sliced_encrypt(uint8_t c[8][8], const uint8_t m[8][8], const uint64_t key[2])
+void gift_64_sliced_encrypt(uint64_t c[8], const uint64_t m[8], const uint64_t key[2])
 {
         // generate round keys
         uint64_t round_keys[ROUNDS_GIFT_SLICED_64][8];
@@ -206,7 +206,7 @@ void gift_64_sliced_encrypt(uint8_t c[8][8], const uint8_t m[8][8], const uint64
 
         // copy to state (eight 64-bit "registers") and pack message bits
         uint64_t state[8];
-        memcpy(state, m, 8 * 8 * sizeof(m[0][0]));
+        memcpy(state, m, 8 * sizeof(m[0]));
         bits_pack(state);
 
         // round loop
@@ -235,7 +235,7 @@ void gift_64_sliced_encrypt(uint8_t c[8][8], const uint8_t m[8][8], const uint64
         memcpy(c, state, 8 * sizeof(state[0]));
 }
 
-void gift_64_sliced_decrypt(uint8_t m[8][8], const uint8_t c[8][8], const uint64_t key[2])
+void gift_64_sliced_decrypt(uint64_t m[8], const uint64_t c[8], const uint64_t key[2])
 {
         // generate round keys
         uint64_t round_keys[ROUNDS_GIFT_SLICED_64][8];
@@ -243,7 +243,7 @@ void gift_64_sliced_decrypt(uint8_t m[8][8], const uint8_t c[8][8], const uint64
 
         // copy to state (eight 64-bit "registers") and pack message bits
         uint64_t state[8];
-        memcpy(state, c, 8 * 8 * sizeof(m[0][0]));
+        memcpy(state, c, 8 * sizeof(m[0]));
         bits_pack(state);
 
         // round loop

### 2023-04-02 impl/naive/gift_sliced.h

@@ -12,5 +12,5 @@ void gift_64_sliced_subcells_inv(uint64_t s[8]);
 void gift_64_sliced_permute(uint64_t s[8]);
 void gift_64_sliced_permute_inv(uint64_t s[8]);
 
-void gift_64_sliced_encrypt(uint8_t c[8][8], const uint8_t m[8][8], const uint64_t key[2]);
-void gift_64_sliced_decrypt(uint8_t m[8][8], const uint8_t c[8][8], const uint64_t key[2]);
+void gift_64_sliced_encrypt(uint64_t c[8], const uint64_t m[8], const uint64_t key[2]);
+void gift_64_sliced_decrypt(uint64_t m[8], const uint64_t c[8], const uint64_t key[2]);

### 2023-04-02 impl/table/gift_table.c

@@ -5,11 +5,14 @@
 
 static const int round_constant[] = {
         // rounds 0-15
-        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B, 0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
+        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
+        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
         // rounds 16-31
-        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30, 0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
+        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
+        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
         // rounds 32-47
-        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A, 0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
+        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
+        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
 };
 
 static const uint64_t tables[16][16] = {

### 2023-04-03 impl/benchmark.c
benchmark

@@ -50,16 +50,16 @@ void benchmark_gift_64(void)
         uint64_t cycles[4] = { 0UL };
         for (int i = 0; i < N; i++) {
                 cycles[0] += TIME(gift_64_generate_round_keys(round_keys, key));
-                cycles[1] += TIME(gift_64_permute(m));
-                cycles[2] += TIME(gift_64_subcells(m));
+                cycles[1] += TIME(gift_64_subcells(m));
+                cycles[2] += TIME(gift_64_permute(m));
                 cycles[3] += TIME(gift_64_encrypt(m, key));
         }
 
         printf("%f %f %f %f\n",
-               cycles[0] / (float)N,
-               cycles[1] / (float)N,
-               cycles[2] / (float)N,
-               cycles[3] / (float)N);
+               cycles[0] / (float)N / 8.0f,
+               cycles[1] / (float)N / 8.0f,
+               cycles[2] / (float)N / 8.0f,
+               cycles[3] / (float)N / 8.0f);
 }
 
 void benchmark_gift_128(void)
@@ -67,22 +67,22 @@ void benchmark_gift_128(void)
         printf("benchmarking GIFT_128...\n");
 
         uint64_t key[2];
-        uint8_t m[32], c[32];
-        uint8_t round_keys[ROUNDS_GIFT_64][32];
+        uint8_t m[16], c[16];
+        uint8_t round_keys[ROUNDS_GIFT_128][32];
 
         uint64_t cycles[4] = { 0UL };
         for (int i = 0; i < N; i++) {
                 cycles[0] += TIME(gift_128_generate_round_keys(round_keys, key));
-                cycles[1] += TIME(gift_128_permute(m));
-                cycles[2] += TIME(gift_128_subcells(m));
+                cycles[1] += TIME(gift_128_subcells(m));
+                cycles[2] += TIME(gift_128_permute(m));
                 cycles[3] += TIME(gift_128_encrypt(c, m, key));
         }
 
         printf("%f %f %f %f\n",
-               cycles[0] / (float)N,
-               cycles[1] / (float)N,
-               cycles[2] / (float)N,
-               cycles[3] / (float)N);
+               cycles[0] / (float)N / 16.0f,
+               cycles[1] / (float)N / 16.0f,
+               cycles[2] / (float)N / 16.0f,
+               cycles[3] / (float)N / 16.0f);
 }
 
 void benchmark_gift_64_table(void)
[…]

### 2023-04-03 impl/naive/gift.c

@@ -109,7 +109,8 @@ uint64_t gift_64_permute_inv(const uint64_t cipher_state)
         return new_cipher_state;
 }
 
-void gift_64_generate_round_keys(uint64_t round_keys[ROUNDS_GIFT_64], const uint64_t key[2])
+void gift_64_generate_round_keys(uint64_t round_keys[restrict ROUNDS_GIFT_64],
+                                 const uint64_t key[restrict 2])
 {
         uint64_t key_state[] = {key[0], key[1]};
         for (int round = 0; round < ROUNDS_GIFT_64; round++) {
@@ -148,21 +149,21 @@ void gift_64_generate_round_keys(uint64_t round_keys[ROUNDS_GIFT_64], const uint
         }
 }
 
-void gift_128_subcells(uint8_t cipher_state[32])
+void gift_128_subcells(uint8_t cipher_state[restrict 32])
 {
         for (size_t i = 0; i < 32; i++) {
                 cipher_state[i] = sbox[cipher_state[i]];
         }
 }
 
-void gift_128_subcells_inv(uint8_t cipher_state[32])
+void gift_128_subcells_inv(uint8_t cipher_state[restrict 32])
 {
         for (size_t i = 0; i < 32; i++) {
                 cipher_state[i] = sbox_inv[cipher_state[i]];
         }
 }
 
-void gift_128_permute(uint8_t cipher_state[32])
+void gift_128_permute(uint8_t cipher_state[restrict 32])
 {
         uint8_t old_cipher_state[32];
         memcpy(old_cipher_state, cipher_state, sizeof(old_cipher_state));
@@ -175,7 +176,7 @@ void gift_128_permute(uint8_t cipher_state[32])
         }
 }
 
-void gift_128_permute_inv(uint8_t cipher_state[32])
+void gift_128_permute_inv(uint8_t cipher_state[restrict 32])
 {
         uint8_t old_cipher_state[32];
         memcpy(old_cipher_state, cipher_state, sizeof(old_cipher_state));
@@ -188,7 +189,8 @@ void gift_128_permute_inv(uint8_t cipher_state[32])
         }
 }
 
-void gift_128_generate_round_keys(uint8_t round_keys[ROUNDS_GIFT_128][32], const uint64_t key[2])
+void gift_128_generate_round_keys(uint8_t round_keys[restrict ROUNDS_GIFT_128][32],
+                                  const uint64_t key[restrict 2])
 {
[…]

### 2023-04-03 impl/naive/gift.h

@@ -14,16 +14,20 @@ uint64_t gift_64_permute_inv(const uint64_t cipher_state);
 void gift_64_generate_round_keys(uint64_t round_keys[ROUNDS_GIFT_64],
                                  const uint64_t key[2]);
 
-void gift_128_subcells(uint8_t cipher_state[32]);
-void gift_128_subcells_inv(uint8_t cipher_state[32]);
-void gift_128_permute(uint8_t cipher_state[32]);
-void gift_128_permute_inv(uint8_t cipher_state[32]);
-void gift_128_generate_round_keys(uint8_t round_keys[ROUNDS_GIFT_128][32],
-                                         const uint64_t key[2]);
+void gift_128_subcells(uint8_t cipher_state[restrict 32]);
+void gift_128_subcells_inv(uint8_t cipher_state[restrict 32]);
+void gift_128_permute(uint8_t cipher_state[restrict 32]);
+void gift_128_permute_inv(uint8_t cipher_state[restrict 32]);
+void gift_128_generate_round_keys(uint8_t round_keys[restrict ROUNDS_GIFT_128][32],
+                                         const uint64_t key[restrict 2]);
 
-uint64_t gift_64_encrypt(const uint64_t m, const uint64_t key[2]);
-uint64_t gift_64_decrypt(const uint64_t c, const uint64_t key[2]);
-void gift_128_encrypt(uint8_t c[16], const uint8_t m[16], const uint64_t key[2]);
-void gift_128_decrypt(uint8_t m[16], const uint8_t c[16], const uint64_t key[2]);
+uint64_t gift_64_encrypt(const uint64_t m, const uint64_t key[restrict 2]);
+uint64_t gift_64_decrypt(const uint64_t c, const uint64_t key[restrict 2]);
+void gift_128_encrypt(uint8_t c[restrict 16],
+                      const uint8_t m[restrict 16],
+                      const uint64_t key[restrict 2]);
+void gift_128_decrypt(uint8_t m[restrict 16],
+                      const uint8_t c[restrict 16],
+                      const uint64_t key[restrict 2]);
 
 #endif

### 2023-04-03 impl/table/gift_table.h

@@ -5,7 +5,7 @@
 
 #define ROUNDS_GIFT_64 28
 
-void gift_64_table_generate_round_keys(uint64_t round_keys[restrict ROUNDS_GIFT_64],
+void gift_64_table_generate_round_keys(uint64_t rks[restrict ROUNDS_GIFT_64],
                                  const uint64_t key[restrict 2]);
 
 uint64_t gift_64_table_subperm(const uint64_t cipher_state);

### 2023-04-03 impl/vector/gift_vec_sbox.c

@@ -51,7 +51,7 @@ static const size_t perm_64_inv[] = {
         4, 9, 14, 3, 20, 25, 30, 19, 36, 41, 46, 35, 52, 57, 62, 51
 };
 
-static const int round_constant[] = {
+static const int round_const[] = {
         // rounds 0-15
         0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B, 0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
         // rounds 16-31
@@ -76,19 +76,19 @@ uint8x16_t gift_64_vec_sbox_permute(const uint8x16_t cipher_state)
         uint64_t new_cipher_state = 0UL;
 
         // S-box 0-7
-        uint64_t boxes = vgetq_lane_u64(cipher_state, 0);
+        uint64_t boxes[2];
+        vst1q_u64(boxes, cipher_state);
         for (size_t box = 0; box < 8; box++) {
                 for (size_t i = 0; i < 4; i++) {
-                        int bit = (boxes >> (box * 8 + i)) & 0x1;
+                        int bit = (boxes[0] >> (box * 8 + i)) & 0x1;
                         new_cipher_state |= (uint64_t)bit << perm_64[box * 4 + i];
                 }
         }
 
         // S-box 8-15
-        boxes = vgetq_lane_u64(cipher_state, 1);
         for (size_t box = 0; box < 8; box++) {
                 for (size_t i = 0; i < 4; i++) {
-                        int bit = (boxes >> (box * 8 + i)) & 0x1;
+                        int bit = (boxes[1] >> (box * 8 + i)) & 0x1;
                         new_cipher_state |= (uint64_t)bit << perm_64[(box + 8) * 4 + i];
                 }
         }
@@ -128,7 +128,7 @@ uint8x16_t gift_64_vec_sbox_permute_inv(const uint8x16_t cipher_state)
         return ret;
 }
 
-void gift_64_vec_sbox_generate_round_keys(uint8x16_t round_keys[ROUNDS_GIFT_64],
+void gift_64_vec_sbox_generate_round_keys(uint8x16_t rks[ROUNDS_GIFT_64],
                                           const uint64_t key[2])
 {
         uint64_t key_state[] = {key[0], key[1]};
@@ -149,15 +149,15 @@ void gift_64_vec_sbox_generate_round_keys(uint8x16_t round_keys[ROUNDS_GIFT_64],
[…]

### 2023-04-03 impl/vector/gift_vec_sbox.h

@@ -11,7 +11,7 @@ uint8x16_t gift_64_vec_sbox_subcells(const uint8x16_t cipher_state);
 uint8x16_t gift_64_vec_sbox_subcells_inv(const uint8x16_t cipher_state);
 uint8x16_t gift_64_vec_sbox_permute(const uint8x16_t cipher_state);
 uint8x16_t gift_64_vec_sbox_permute_inv(const uint8x16_t cipher_state);
-void       gift_64_vec_sbox_generate_round_keys(uint8x16_t round_keys[ROUNDS_GIFT_64],
+void       gift_64_vec_sbox_generate_round_keys(uint8x16_t rks[ROUNDS_GIFT_64],
                                                 const uint64_t key[2]);
 
 // construct tables

### 2023-04-03 impl/vector/gift_vec_sliced.c

@@ -38,7 +38,7 @@ static uint8x16_t pack_mask_0;
 static uint8x16_t pack_mask_1;
 static uint8x16_t pack_mask_2;
 
-static const int round_constant[] = {
+static const int round_const[] = {
         // rounds 0-15
         0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B, 0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
         // rounds 16-31
@@ -114,8 +114,6 @@ void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
 
 void gift_64_vec_sliced_bits_unpack(uint8x16x4_t m[restrict 2])
 {
-        // same plaintext slice bits into same register (so we only have to do
-        // what we are doing here once instead of every round)
         uint8x16x2_t pairs[4] = {
                 { .val = { m[0].val[0], m[1].val[0] }},
                 { .val = { m[0].val[1], m[1].val[1] }},
@@ -153,8 +151,10 @@ void gift_64_vec_sliced_bits_unpack(uint8x16x4_t m[restrict 2])
 
 void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
 {
-        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
-        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
+        cs[0].val[1] = veorq_u8(cs[0].val[1],
+                                vandq_u8(cs[0].val[0], cs[0].val[2]));
+        uint8x16_t t = veorq_u8(cs[0].val[0],
+                                vandq_u8(cs[0].val[1], cs[0].val[3]));
         cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
         cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
         cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
@@ -162,8 +162,10 @@ void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
         cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
[…]

### 2023-04-03 impl/vector/gift_vec_sliced.h

@@ -17,7 +17,7 @@ void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
 void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
 void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
 void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
-void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
+void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t rks[restrict ROUNDS_GIFT_64][2],
                                             const uint64_t key[restrict 2]);
 
 void gift_64_vec_sliced_init(void);

### 2023-04-04 Paper/chapters/implementation-strategies.tex
benchmark
bibliography no ul

--- 2023-03-23 Paper/chapters/implementation-strategies.tex
@@
 
-\section{Strategies for SPN}
+\section{Strategies for \texttt{GIFT}}
 
@@
 
-Nowadays, most instructions set architectures support single-instruction,
-multiple-data processing. The idea of such an SIMD system is to work on
-multiple data stored in vectors at once to speed up calculations. For A64, two
-types of vector processing are available:
-
-\begin{enumerate}
-    \item Advanced SIMD, known as NEON
-    \item Scalable Vector Extension (SVE)
-\end{enumerate}
-
-We will take a look at NEON as this is the type of vector processing supported
-by the Cortex-A73 processor.
-
-\subsubsection{ARM Neon}
-
-The register file of the NEON unit is made up of 32 quad-word (128-bit)
-registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
-registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
-sized lanes on which the vector instructions operate. Valid ways to interpret
-for example the register \texttt{V0} are:
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/V_register.pdf}
-    \caption{Divisions of the V0 register}
-\end{figure}
-
-NEON instructions interpret their operands' layouts (i.e. lane count and width)
-through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
-adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
-together and storing the result in \texttt{V0} can be done as follows:
-
-\begin{center}
-    \texttt{ADD V0.8H, V1.8H, V2.8H}
-\end{center}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/vector_add.pdf}
-    \caption{Addition of two vector registers}
-\end{figure}
-
-The plenitude of different processing instructions allow flexible ways to
-further speed up algorithms having reached their optimizational limit on
-non-SIMD platforms. \texttt{vperm}, a general term standing for \textit{vector
-permute}, is a common instruction on SIMD machines. Called \texttt{TBL} on
-NEON, it is used for parallel table lookups and arbitrary permutations. It takes
-two inputs to perform a lanewise lookup:
+The plenitude of different processing instructions introduced by
+NEON\ref{ss:neon} allow flexible ways to further speed up algorithms having
+reached their optimizational limit on non-SIMD platforms. \texttt{vperm}, a
+general term standing for \textit{vector permute}, is a common instruction on
+SIMD machines. Called \texttt{TBL} on NEON, it is used for parallel table
+lookups and arbitrary permutations. It takes two inputs to perform a lanewise
+lookup:
 
@@
 \]
+\normalsize
 

### 2023-04-04 Paper/chapters/implementation.tex
xor instead of or
bibliography no ul

--- 2023-03-23 Paper/chapters/implementation.tex
@@
 
-We will provide and discuss implementations for the presented strategies in the
-C programming language. Although directly writing Assembler code could result
-in a small performance benefit, this generally increases the work necessary by
-an order of magnitude for only limited results. This is why we use NEON
-intrinsics.
-
-\subsection{NEON intrinsics}
-
-The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
-definitions including vector data types and C functions for working with these
-vectors. These functions are known as NEON intrinsics and give the programmer a
-high-level interface to most NEON instructions. Major advantages of this
-approach include the ease of development as the compiler takes over register
-allocation and load/store operations as well as performance benefits through
-compiler optimizations.
-
-Standard vector data types have the format \texttt{uintnxm\_t} with lane width
-$n$ in bits and and lane count $m$. Array types of the format
-\texttt{uintnxmxc\_t}, $c\in\{2,3,4\}$ are also defined which are used in
-operations requiring multiple parameters like \texttt{TBL} or pairwise
-load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
-after the operation name to indicate operation on a 128-bit register, and end
-with the lane data format. Multiplying eight pairs of 16-bit numbers
-\texttt{a,b} for example can be done via the following:
-
-\begin{center}
-    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
-\end{center}
-
-In this case, the compiler allocates vector registers for \texttt{a},
-\texttt{b} and \texttt{result} and assembles the intrinsic to \texttt{MUL
-Vr.8H, Va.8H, Vb.8H}. Necessary load and stores for the result and parameters
-are also handled automatically. Of special interest to us are the following
-intrinsics, each existing in different variants with different lane widths and
-also array types: \\
-
-\begin{table}[h!]
-    \centering
-    \footnotesize
-    \begin{tabularx}{\textwidth}{ll|X}
-        Intrinsic && Description \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vreinterpretq\_u8\_u64(uint64x2\_t)} & Explicit casting \\
-        \hline
-        \texttt{uint64\_t} & \texttt{vgetq\_lane\_u64(void)} & Extract a single lane \\
-        \hline
-        \texttt{void} & \texttt{vsetq\_lane\_u64(uint64\_t)} & Insert a single lane \\
-        \hline
-        \texttt{uint64x2\_t} & \texttt{vdupq\_n\_u64(uint64\_t)} & Initialize all lanes to same value \\
-        \hline
-        \texttt{void} & \texttt{vst1q\_u64(uint64\_t*, uint64x2\_t)} & Store from register to memory \\
-        \hline
-        \texttt{uint64x2\_t} & \texttt{vld1q\_u64(uint64\_t*, uint64x2\_t)} & Load from memory to register \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{veorq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise XOR \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vandq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise AND \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vorrq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise OR \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vmvnq\_u8(uint8x16\_t)} & bitwise NOT \\
-        \hline
-        \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
-    \end{tabularx}
-\end{table}
-
-The implementations closely follow the aforementioned strategies;
-instruction-level optimization is left to the compiler.
+Implementations in the C programming language for the presented strategies can
+be found in Appendix \ref{app:cimpl}. Although directly writing Assembler code could
+result in a small performance benefit, this generally increases the work
+necessary by an order of magnitude for only limited results. Instruction-level
+optimization and in particular register allocation is left to the compiler.

### 2023-04-04 Paper/chapters/introduction.tex
bibliography no ul

@@
 Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
-$K=k7||k6||\dots||k0$ denote the $128$-bit key state.
+$K=k_7||k_6||\dots||k_0$ denote the $128$-bit key state.
 
 \paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
-state. $u_i$ and $v_i$ are XORed to $r_{4i+1}$ and $r_{4i}$ of the round key
-$R$ respectively.
+state. These are then added to the 64-bit round key: $R_{4i+1}\leftarrow
+u_i,R_{4i}\leftarrow v_i$.
 
-\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k4||k1||k_0$ from
-the key state. $u_i$ and $v_i$ are XORed to $r_{4i+2}$ and $b_{4i+1}$ of the
-round key $R$ respectively.
+\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k_4||k_1||k_0$ from
+the key state. These are then added to the 128-bit round key: $R_{4i+2}\leftarrow
+u_i,R_{4i+1}\leftarrow v_i$.
 
@@
 
-With small devices, embedded processors and ASICs becoming ever more ubiquitous
-and essential in areas like medicine or automotive design, the need for ...
+With small devices and embedded processors becoming ever more ubiquitous and
+essential in areas like consumer electronics or industrial and IoT
+applications, the need for low-power, high-performance microprocessors has
+increased steadily. With more than 250 billion chips shipped, semiconductors
+designed by ARM power 95\% of mobile devices and have found a great many
+applications due to their high performance and low power consumption. The
+ODROID-N2+\cite{odroidn2} development board we are using is based on the big.LITTLE
+architecture and is powered by a quad-core ARM Cortex-A73 processor and a
+weaker dual-core ARM Cortex-A53 for power efficiency. Both these processors are
+part of the eight generation of ARM designs known as ARMv8\cite{armv8:2013}.
+
+ARMv8 defines three architecture profiles for different use cases as well as
+dynamic execution states with corresponding instruction sets. This work will
+focus on the A profile running in the AArch64 state utilizing the A64
+instruction set with NEON and crypto extensions.
+
+\begin{table}[h!]
+    \centering
+    \begin{tabularx}{\textwidth}{lX}
+        \toprule
+        Profile & Description \\
+        \midrule
+        Application (A) & Traditional use with virtual memory and privilege level support \\
+        Real-time (R) & Real-time, low-latency, deterministic embedded systems \\
+        Microcontroller (M) & Very low-power, fast-interrupt embedded systems \\
+        \bottomrule
+    \end{tabularx}
+    \caption{ARMv8 profiles}
+\end{table}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabularx}{\textwidth}{llX}
+        \toprule
+        Execution state & Usage & Instruction sets \\
+        \midrule
+        AArch32 & 32-bit compatibility & A32/T32 \\
+        AArch64 & 64-bit & A64 \\
+        \bottomrule
+    \end{tabularx}
+    \caption{ARMv8 execution states}
+\end{table}
+
+\subsection{General architecture}
+
+ARMv8 is a RISC architecture employing simple data processing instructions
+operating only on registers as well as dedicated load/store instructions to
+transfer data from register to memory and back. This enables faster execution
+of individual instructions, a simplier pipeline design, predictable instruction
+timings and fewer addressing modes.
+
+The A64 instruction set defines 31 64-bit general-purpose registers
+\texttt{X0-X30} which can also be accessed as 32-bit registers \texttt{W0-W30}.
+Values are loaded from and stored to memory using \texttt{LDR}/\texttt{STR}.
+Data processing instructions generally use explicit output registers instead of
+overwriting the first input register.
+
+\begin{table}[h!]
+    \centering
+    \small
+    \begin{tabularx}{\textwidth}{llX}
+        \toprule
+        Addressing mode & Example & Description \\
+        \midrule
+        Base register & \texttt{LDR W0, [X1]} & \texttt{W0 = *(X1);} \\
+        Offset & \texttt{LDR W0, [X1, \#12]} & \texttt{W0 = *(X1 + 12);} \\
+        Pre-indexing & \texttt{LDR W0, [X1, \#12]!} & \texttt{X1 += 12; W0 = *(X1)} \\
+        Post-indexing & \texttt{LDR W0, [X1], \#12} & \texttt{W0 = *(X1); X1 += 12} \\
+        \bottomrule
+    \end{tabularx}
+    \caption{AArch64 addressing modes}
+\end{table}
+
+\subsection{NEON}
+\label{ss:neon}
+
+ARMv8 supports single-instruction, multiple-data ARMv8 supports
+single-instruction, multiple-data (SIMD) processing. These systems allow the
+programmer to store multiple pieces of data in a vector and work on them in
+parallel to speed up calculations. The A64 instruction set defines two possible
+SIMD implementations.
+
+\begin{enumerate}
+    \item Advanced SIMD, known as NEON
+    \item Scalable Vector Extension (SVE)
+\end{enumerate}
+
+We will take a look at NEON as this is the type of vector processing supported
+by the Cortex-A73 processor.
+
+The register file of the NEON unit is made up of 32 quad-word (128-bit)
+registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
+registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
+sized lanes on which the vector instructions operate. Valid ways to interpret
+for example the register \texttt{V0} are:
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/V_register.pdf}
+    \caption{Divisions of the V0 register}
+\end{figure}
+
+NEON instructions interpret their operands' layouts (i.e. lane count and width)
+through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
+adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
+together and storing the result in \texttt{V0} can be done as follows:
+
+\begin{center}
+    \texttt{ADD V0.8H, V1.8H, V2.8H}
+\end{center}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/vector_add.pdf}
+    \caption{Addition of two vector registers}
+\end{figure}
+
+\subsection{NEON Intrinsics}
+
+The header file \texttt{<arm\_neon.h>} provides ARM-specific data and function
+definitions including vector data types and C functions for working with these
+vectors. These functions are known as NEON intrinsics \cite{neonintr:2022} and
+give the programmer a high-level interface to most NEON instructions. Major
+advantages of this approach include the ease of development as the compiler
+takes over register allocation and load/store operations as well as performance
+benefits through compiler optimizations.
+
+Standard vector data types have the format \texttt{uintnxm\_t} with lane width
+$n$ in bits and and lane count $m$. Array types of the format
+\texttt{uintnxmxc\_t}, $c\in\{2,3,4\}$ are also defined which are used in
+operations requiring multiple parameters like \texttt{TBL} or pairwise
+load/stores. Intrinsics include the operation name and lane data format as well
+as an optional \texttt{q} suffix to indicate operation on a 128-bit register.
+Multiplying eight pairs of 16-bit numbers \texttt{a,b} for example can be done
+via the following:
+
+\begin{center}
+    \texttt{uint16x8\_t result = vmulq\_u16(a, b);}
+\end{center}
+
+In this case, the compiler allocates vector registers for \texttt{a},
+\texttt{b} and \texttt{result} and assembles the intrinsic to \texttt{MUL
+Vr.8H, Va.8H, Vb.8H}. Necessary loads and stores for the result and parameters
+are also handled automatically. Of special interest to us are the following
+intrinsics, each existing in different variants with different lane widths and
+also array types: \\
+
+\begin{table}[h!]
+    \centering
+    \footnotesize
+    \begin{tabularx}{\textwidth}{llX}
+        \toprule
+        Intrinsic && Description \\
+        \midrule
+        \texttt{uint8x16\_t} & \texttt{vreinterpretq\_u8\_u64(uint64x2\_t)} & Explicit casting \\
+        \midrule
+        \texttt{uint64\_t} & \texttt{vgetq\_lane\_u64(void)} & Extract a single lane \\
+        \midrule
+        \texttt{void} & \texttt{vsetq\_lane\_u64(uint64\_t)} & Insert a single lane \\
+        \midrule
+        \texttt{uint64x2\_t} & \texttt{vdupq\_n\_u64(uint64\_t)} & Initialize all lanes to same value \\
+        \midrule
+        \texttt{void} & \texttt{vst1q\_u64(uint64\_t*, uint64x2\_t)} & Store from register to memory \\
+        \midrule
+        \texttt{uint64x2\_t} & \texttt{vld1q\_u64(uint64\_t*, uint64x2\_t)} & Load from memory to register \\
+        \midrule
+        \texttt{uint8x16\_t} & \texttt{veorq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise XOR \\
+        \midrule
+        \texttt{uint8x16\_t} & \texttt{vandq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise AND \\
+        \midrule
+        \texttt{uint8x16\_t} & \texttt{vorrq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise OR \\
+        \midrule
+        \texttt{uint8x16\_t} & \texttt{vmvnq\_u8(uint8x16\_t)} & bitwise NOT \\
+        \midrule
+        \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
+        \bottomrule
+    \end{tabularx}
+    \caption{Common NEON intrinsics}
+\end{table}

### 2023-04-04 impl/benchmark.c
throughput data

@@ -9,11 +9,13 @@
 #include <stdio.h>
 #include <stdlib.h>
 #include <time.h>
+#include <sys/time.h> // wall time via gettimeofday()
 
 #include <stdint.h>
 #include <arm_neon.h>
 
-#define N 10000
+#define NL 10000 // latency measurement count
+#define NT 1000000 // throughput measurement count
 
 #define MEASURE(code)\
         {\
@@ -32,14 +34,19 @@
         t1 - t0;\
         })
 
-void rand_bytes(uint8_t m[], size_t n)
+static void rand_bytes(uint8_t m[], size_t n)
 {
         for (size_t i = 0; i < n; i++) {
                 m[i] = rand();
         }
 }
 
-void benchmark_gift_64(void)
+static double elapsed_seconds(const struct timeval *st, const struct timeval *et)
+{
+        return (et->tv_sec - st->tv_sec) + (et->tv_usec - st->tv_usec) / 1000000.0;
+}
+
+static void benchmark_gift_64(void)
 {
         printf("benchmarking GIFT_64...\n");
 
@@ -48,21 +55,31 @@ void benchmark_gift_64(void)
         uint64_t round_keys[ROUNDS_GIFT_64];
 
         uint64_t cycles[4] = { 0UL };
-        for (int i = 0; i < N; i++) {
+        for (int i = 0; i < NL; i++) {
                 cycles[0] += TIME(gift_64_generate_round_keys(round_keys, key));
                 cycles[1] += TIME(gift_64_subcells(m));
                 cycles[2] += TIME(gift_64_permute(m));
-                cycles[3] += TIME(gift_64_encrypt(m, key));
+                cycles[3] += TIME(gift_64_encrypt(m, round_keys));
         }
 
         printf("%f %f %f %f\n",
-               cycles[0] / (float)N / 8.0f,
-               cycles[1] / (float)N / 8.0f,
-               cycles[2] / (float)N / 8.0f,
-               cycles[3] / (float)N / 8.0f);
+               cycles[0] / (float)NL / 8.0f,
+               cycles[1] / (float)NL / 8.0f,
+               cycles[2] / (float)NL / 8.0f,
+               cycles[3] / (float)NL / 8.0f);
+
+        struct timeval st, et;
+        gettimeofday(&st, NULL);
+        for (int i = 0; i < NT; i++) {
+                gift_64_encrypt(m, round_keys);
+        }
+        gettimeofday(&et, NULL);
[…]

### 2023-04-04 impl/naive/gift.c

@@ -230,13 +230,9 @@ void gift_128_generate_round_keys(uint8_t round_keys[restrict ROUNDS_GIFT_128][3
 }
 
 uint64_t gift_64_encrypt(const uint64_t m,
-                         const uint64_t key[restrict 2])
+                         const uint64_t round_keys[restrict ROUNDS_GIFT_64])
 {
         uint64_t c = m;
-
-        // generate round keys
-        uint64_t round_keys[ROUNDS_GIFT_64];
-        gift_64_generate_round_keys(round_keys, key);
 
         // round loop
         for (int round = 0; round < ROUNDS_GIFT_64; round++) {
@@ -249,13 +245,9 @@ uint64_t gift_64_encrypt(const uint64_t m,
 }
 
 uint64_t gift_64_decrypt(const uint64_t c,
-                         const uint64_t key[restrict 2])
+                         const uint64_t round_keys[restrict ROUNDS_GIFT_64])
 {
         uint64_t m = c;
-
-        // generate round keys
-        uint64_t round_keys[ROUNDS_GIFT_64];
-        gift_64_generate_round_keys(round_keys, key);
 
         // round loop (in reverse)
         for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
@@ -270,16 +262,12 @@ uint64_t gift_64_decrypt(const uint64_t c,
 
 void gift_128_encrypt(uint8_t c_[restrict 16],
                       const uint8_t m[restrict 16],
-                      const uint64_t key[restrict 2])
+                      const uint8_t round_keys[restrict ROUNDS_GIFT_128][32])
 {
         uint8_t c[32]; // one byte per s-box
         for (size_t i = 0; i < 32; i++) {
                 c[i] = (m[i / 2] >> (i % 2) * 4) & 0xf;
         }
-
-        // generate round keys
-        uint8_t round_keys[ROUNDS_GIFT_128][32];
-        gift_128_generate_round_keys(round_keys, key);
 
         // round loop
         for (int round = 0; round < ROUNDS_GIFT_128; round++) {
@@ -298,16 +286,12 @@ void gift_128_encrypt(uint8_t c_[restrict 16],
 
 void gift_128_decrypt(uint8_t m_[restrict 16],
                       const uint8_t c[restrict 16],
-                      const uint64_t key[restrict 2])
+                      const uint8_t round_keys[restrict ROUNDS_GIFT_128][32])
 {
[…]

### 2023-04-04 impl/table/gift_table.c

@@ -86,13 +86,10 @@ uint64_t gift_64_table_subperm(const uint64_t cipher_state)
         return new_cipher_state;
 }
 
-uint64_t gift_64_table_encrypt(const uint64_t m, const uint64_t key[restrict 2])
+uint64_t gift_64_table_encrypt(const uint64_t m,
+                               const uint64_t rks[restrict ROUNDS_GIFT_64])
 {
         uint64_t c = m;
-
-        // generate round keys
-        uint64_t rks[ROUNDS_GIFT_64];
-        gift_64_table_generate_round_keys(rks, key);
 
         // round loop
         for (int round = 0; round < ROUNDS_GIFT_64; round++) {

### 2023-04-04 impl/table/gift_table.h

@@ -11,6 +11,7 @@ void gift_64_table_generate_round_keys(uint64_t rks[restrict ROUNDS_GIFT_64],
 uint64_t gift_64_table_subperm(const uint64_t cipher_state);
 
 // can only encrypt using table technique!
-uint64_t gift_64_table_encrypt(const uint64_t m, const uint64_t key[restrict 2]);
+uint64_t gift_64_table_encrypt(const uint64_t m,
+                               const uint64_t rks[restrict ROUNDS_GIFT_64]);
 
 #endif

### 2023-04-04 impl/vector/gift_vec_sbox.c

@@ -9,33 +9,62 @@ static uint64_t sbox_vec_u64[2] = {
 };
 
 static uint64_t sbox_vec_inv_u64[2] = {
-0x0b040c020608000dUL, 0x050f09030a01070eUL
+        0x0b040c020608000dUL, 0x050f09030a01070eUL
 };
 
 static uint8x16_t sbox_vec;
 static uint8x16_t sbox_vec_inv;
 
-#define U64_TO_V128(V,M)\
-        V = vsetq_lane_u64(\
-        (uint64_t)((M >> 4 * 0) & 0xf) << 8 * 0 |\
-        (uint64_t)((M >> 4 * 1) & 0xf) << 8 * 1 |\
-        (uint64_t)((M >> 4 * 2) & 0xf) << 8 * 2 |\
-        (uint64_t)((M >> 4 * 3) & 0xf) << 8 * 3 |\
-        (uint64_t)((M >> 4 * 4) & 0xf) << 8 * 4 |\
-        (uint64_t)((M >> 4 * 5) & 0xf) << 8 * 5 |\
-        (uint64_t)((M >> 4 * 6) & 0xf) << 8 * 6 |\
-        (uint64_t)((M >> 4 * 7) & 0xf) << 8 * 7,\
-        V, 0);\
-        V = vsetq_lane_u64(\
-        (uint64_t)((M >> 4 * 8)  & 0xf) << 8 * 0 |\
-        (uint64_t)((M >> 4 * 9)  & 0xf) << 8 * 1 |\
-        (uint64_t)((M >> 4 * 10) & 0xf) << 8 * 2 |\
-        (uint64_t)((M >> 4 * 11) & 0xf) << 8 * 3 |\
-        (uint64_t)((M >> 4 * 12) & 0xf) << 8 * 4 |\
-        (uint64_t)((M >> 4 * 13) & 0xf) << 8 * 5 |\
-        (uint64_t)((M >> 4 * 14) & 0xf) << 8 * 6 |\
-        (uint64_t)((M >> 4 * 15) & 0xf) << 8 * 7,\
-        V, 1);
+// split S-box bits into vector lanes
+uint8x16_t gift_64_vec_sbox_bits_pack(const uint64_t a)
+{
+        uint8x16_t v;
+        v = vsetq_lane_u64(
+        (uint64_t)((a >> 4 * 0) & 0xf) << 8 * 0 |
[…]

### 2023-04-04 impl/vector/gift_vec_sbox.h

@@ -7,17 +7,21 @@
 #define ROUNDS_GIFT_64 28
 
 // expose for benchmarking
+uint8x16_t gift_64_vec_sbox_bits_pack(const uint64_t a);
+uint64_t   gift_64_vec_sbox_bits_unpack(const uint8x16_t a);
 uint8x16_t gift_64_vec_sbox_subcells(const uint8x16_t cipher_state);
 uint8x16_t gift_64_vec_sbox_subcells_inv(const uint8x16_t cipher_state);
 uint8x16_t gift_64_vec_sbox_permute(const uint8x16_t cipher_state);
 uint8x16_t gift_64_vec_sbox_permute_inv(const uint8x16_t cipher_state);
 void       gift_64_vec_sbox_generate_round_keys(uint8x16_t rks[ROUNDS_GIFT_64],
-                                                const uint64_t key[2]);
+                                                const uint64_t key[restrict 2]);
 
 // construct tables
 void gift_64_vec_sbox_init(void);
 
-uint64_t gift_64_vec_sbox_encrypt(const uint64_t m, const uint64_t key[2]);
-uint64_t gift_64_vec_sbox_decrypt(const uint64_t c, const uint64_t key[2]);
+uint64_t gift_64_vec_sbox_encrypt(const uint64_t m,
+                                  const uint8x16_t rks[restrict ROUNDS_GIFT_64]);
+uint64_t gift_64_vec_sbox_decrypt(const uint64_t c,
+                                  const uint8x16_t rks[restrict ROUNDS_GIFT_64]);
 
 #endif

### 2023-04-04 impl/vector/gift_vec_sliced.c

@@ -47,28 +47,29 @@ static const int round_const[] = {
         0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A, 0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
 };
 
-uint8x16_t shl(uint8x16_t v, int n)
+uint8x16_t shl(const uint8x16_t v, const int n)
 {
         uint64_t l[2];
         vst1q_u64(l, v);
         l[1] = (l[1] << n) | (l[0] >> (64 - n));
         l[0] <<= n;
-        return vreinterpretq_u8_u64(vld1q_u64(l));
+        return vld1q_u64(l);
 }
 
-uint8x16_t shr(uint8x16_t v, int n)
+uint8x16_t shr(const uint8x16_t v, const int n)
 {
         uint64_t l[2];
         vst1q_u64(l, v);
         l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
         l[1] >>= n;
-        return vreinterpretq_u8_u64(vld1q_u64(l));
+        return vld1q_u64(l);
 }
 
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
+void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
+                                 const uint8x16_t m, const int n)
 {
 
-        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
+        const uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
         *b = veorq_u8(*b, t);
         *a = veorq_u8(*a, shl(t, n));
 }
@@ -94,7 +95,7 @@ void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
 
         // same plaintext slice bits into same register (so we only have to do
         // what we are doing here once instead of every round)
-        uint8x16x2_t pairs[4] = {
+        const uint8x16x2_t pairs[4] = {
                 { .val = { m[0].val[0], m[1].val[0] }},
                 { .val = { m[0].val[1], m[1].val[1] }},
                 { .val = { m[0].val[2], m[1].val[2] }},
[…]

### 2023-04-06 Paper/chapters/implementation.tex
implementation chapter

@@
 Implementations in the C programming language for the presented strategies can
-be found in Appendix \ref{app:cimpl}. Although directly writing Assembler code could
-result in a small performance benefit, this generally increases the work
+be found in Appendix \ref{app:cimpl}. Although directly writing Assembler code
+could result in a small performance benefit, this generally increases the work
 necessary by an order of magnitude for only limited results. Instruction-level
 optimization and in particular register allocation is left to the compiler.
+Relying on the compiler mandates a closer study of the generated, optimized
+assembler. All source files were compiled using clang version 15.0.7 and
+optimization level O3.
+
+\section{Pipelining}
+
+Understanding certain choices requires an understanding of the Cortex-A73
+instruction pipeline\cite{a72opt:2015}. Being a superscalar processor, it is
+able to execute more than one instruction per clock cycle by dispatching
+instructions to different execution units working in parallel.
+
+\begin{figure}{h!}
+    \centering
+    \includegraphics[width=0.7\textwidth]{Figures/a72pipeline.pdf}
+    \caption{High-level overview of the Cortex-A72 instruction pipeline}
+\end{figure}
+
+The processor might for example store a calculation result, load a necessary
+value from memory and execute two SIMD operations at once, all in the same
+clock cycle. Modern compilers take advantage of this fact by reordering
+instructions such that all pipeline execution units stay as busy as possible
+and do not stall while having to wait for new instructions to be dispatched. A
+more thorough analysis will be presented for the bitsliced strategy of
+\texttt{GIFT}, but all implementations are heavily reliant on function
+inlining, instruction reordering and loop unrolling.
+
+\section{GIFT}
+\subsection{Table-based}
+
+This is the simplest strategy to implement. Indeed, its biggest advantage lies
+in its portability to other platforms without relying on specific features or
+extensions. The cipher state is stored as a 64-bit word and one round consists
+in extracting the 4-bit S-boxes, looking up table values, collecting these in
+an accumulator and finally adding the round key.
+
+\lstinputlisting[language=c, firstline=77, lastline=87]{../impl/table/gift_table.c}
+\lstinputlisting[language=c, firstline=89, lastline=101]{../impl/table/gift_table.c}
+
+\subsection{Using \texttt{vperm}}
+
+Implementation of the substitution layer requires the use of a single vector
+intrinsic. This mandates the packing of data into a vector register which in
+turn is disadvantageous to the permutation layer as we need to extract single
+bits. Packing and unpacking is nothing more than filling 8-bit vector lanes
+with 4-bit S-boxes and vice versa.
+
+\lstinputlisting[language=c, firstline=92, lastline=95]{../impl/vector/gift_vec_sbox.c}
+\lstinputlisting[language=c, firstline=102, lastline=117]{../impl/vector/gift_vec_sbox.c}
+\lstinputlisting[language=c, firstline=197, lastline=212]{../impl/vector/gift_vec_sbox.c}
+
+\subsection{Bitslicing}
+
+We will examine the round function in closer detail and compare the source code
+with the generated assembly.
+\pagebreak
+
+\begin{lstlisting}[language=c, caption={Round function src}]
+for (int round = 0; round < ROUNDS_GIFT_64; round++) {
+    gift_64_vec_sliced_subcells(s);
+    gift_64_vec_sliced_permute(s);
+
+    // round key addition
+    s[0].val[0] = veorq_u8(s[0].val[0], rks[round][0].val[0]);
+    s[0].val[1] = veorq_u8(s[0].val[1], rks[round][0].val[1]);
+    s[0].val[2] = veorq_u8(s[0].val[2], rks[round][0].val[2]);
+    s[0].val[3] = veorq_u8(s[0].val[3], rks[round][0].val[3]);
+    s[1].val[0] = veorq_u8(s[1].val[0], rks[round][1].val[0]);
+    s[1].val[1] = veorq_u8(s[1].val[1], rks[round][1].val[1]);
+    s[1].val[2] = veorq_u8(s[1].val[2], rks[round][1].val[2]);
+    s[1].val[3] = veorq_u8(s[1].val[3], rks[round][1].val[3]);
+}
+\end{lstlisting}
+
+\begin{minipage}{.45\textwidth}
+    \begin{lstlisting}[language={[ARM]Assembler}, caption={Round function asm}, escapechar=|]
+and     v20.16b, v17.16b, v6.16b
+add     x9, x20, x8
+eor     v20.16b, v20.16b, v16.16b
+add     x8, x8, #0x80|\label{lst:cntrincr}|
+and     v21.16b, v20.16b, v19.16b
+cmp     x8, #0xe70|\label{lst:cntrchk}|
+eor     v21.16b, v21.16b, v6.16b
+orr     v6.16b, v6.16b, v16.16b
+eor     v6.16b, v6.16b, v17.16b
+eor     v16.16b, v19.16b, v6.16b
+eor     v17.16b, v16.16b, v20.16b
+and     v19.16b, v21.16b, v17.16b
+eor     v6.16b, v19.16b, v6.16b
+and     v19.16b, v7.16b, v4.16b
+eor     v19.16b, v19.16b, v5.16b
+and     v20.16b, v19.16b, v18.16b
+eor     v20.16b, v20.16b, v4.16b
+orr     v4.16b, v4.16b, v5.16b
+eor     v4.16b, v4.16b, v7.16b
+eor     v5.16b, v18.16b, v4.16b
+eor     v7.16b, v5.16b, v19.16b
+and     v18.16b, v20.16b, v7.16b
+eor     v4.16b, v18.16b, v4.16b
+    \end{lstlisting}
+\end{minipage}\hfill
+\begin{minipage}{.45\textwidth}
+    \vspace*{7mm}
+    \begin{lstlisting}[language={[ARM]Assembler}, firstnumber=24, escapechar=|]
+tbl     v18.16b, {v6.16b}, v2.16b
+tbl     v19.16b, {v21.16b}, v3.16b
+tbl     v21.16b, {v4.16b}, v2.16b
+ldp     q6, q4, [x9, #-112]|\label{lst:floatload}|
+mvn     v16.16b, v16.16b
+tbl     v16.16b, {v16.16b}, v0.16b
+tbl     v17.16b, {v17.16b}, v1.16b
+mvn     v5.16b, v5.16b
+tbl     v5.16b, {v5.16b}, v0.16b
+ldp     q22, q23, [x9, #-80]
+eor     v6.16b, v6.16b, v16.16b
+eor     v16.16b, v4.16b, v17.16b
+tbl     v7.16b, {v7.16b}, v1.16b
+eor     v17.16b, v22.16b, v18.16b
+tbl     v20.16b, {v20.16b}, v3.16b
+ldp     q4, q18, [x9, #-48]
+eor     v19.16b, v23.16b, v19.16b
+eor     v4.16b, v4.16b, v5.16b
+ldp     q22, q23, [x9, #-16]
+eor     v5.16b, v18.16b, v7.16b
+eor     v7.16b, v22.16b, v21.16b
+eor     v18.16b, v23.16b, v20.16b
+b.ne    197e0
+    \end{lstlisting}
+\end{minipage}
+
+Without needing to dive too deep into details, it is obvious to see the
+compiler has inlined the two function calls to \texttt{subcells} and
+\texttt{permute} with lines 1 to 23 originating from the \texttt{subcells} and
+lines 24-46 from the \texttt{permute} and round key addition functions. It has
+chosen not to unroll the loop, but has moved the loop counter increment as well
+as the condition check in between the \texttt{subcells} instructions to line
+\ref{lst:cntrincr} and \ref{lst:cntrchk}. In addition, the loop counter serves
+a second purpose as an offset register and is therefore incremented by 0x80=128
+instead of just 1.
+
+Permutation (\texttt{tbl}) and round key addition (\texttt{eor}) instructions
+are interleaved. The compiler recognizes data dependencies and can therefore
+proceed with round key addition directly after a slice has been permutated
+without needing to wait for all permutations to finish. This is only logical
+considering the inner workings of the instruction pipeline: by interleaving
+NEON with regular logic and load instructions, the execution units are filled
+more evenly and pipeline stalls are prevented which speeds up computation.
+
+Round keys are loaded from memory a few instructions before they are needed;
+assuming all round keys are stored in the L1 cache, loading a
+floating-point/vector register takes 5 cycles. After the load has been
+dispatched to the load execution unit in line \ref{lst:floatload}, the
+processor happily continues chugging away at the instruction stream by issuing
+\texttt{tbl} and \texttt{mvn} $\mu$-ops to other execution units.
+
+These kinds of optimizations are pervasive when programming using higher level
+languages like C and modern-day compilers more often than not outperform even
+carefully handcrafted assembly.

### 2023-04-06 impl/test.c

@@ -44,7 +44,10 @@ void test_gift_64(void)
         uint64_t m = 0x4dcfd3bdd61810f0UL;
         uint64_t c_expected = 0xb11d30b8d39763e1UL;
         uint64_t c;
-        c = gift_64_encrypt(m, key);
+
+        uint64_t rks[ROUNDS_GIFT_64];
+        gift_64_generate_round_keys(rks, key);
+        c = gift_64_encrypt(m, rks);
         ASSERT_EQUALS(c, c_expected);
         return;
 
@@ -55,8 +58,9 @@ void test_gift_64(void)
                 m_rand((uint8_t*)&m, 8);
                 memset(&m, 0, sizeof(m));
 
-                c = gift_64_encrypt(m, key);
-                uint64_t m_actual = gift_64_decrypt(c, key);
+                gift_64_generate_round_keys(rks, key);
+                c = gift_64_encrypt(m, rks);
+                uint64_t m_actual = gift_64_decrypt(c, rks);
                 ASSERT_EQUALS(m, m_actual);
         }
 }
@@ -75,7 +79,10 @@ void test_gift_128(void)
                 0x06, 0x7b, 0xc0, 0x97, 0xa6, 0x38, 0x1f, 0xe8
         };
         uint8_t c[16];
-        gift_128_encrypt(c, m, key);
+
+        uint8_t rks[ROUNDS_GIFT_128][32];
+        gift_128_generate_round_keys(rks, key);
+        gift_128_encrypt(c, m, rks);
         ASSERT_TRUE(memcmp(c, c_expected, 16) == 0);
 
         // test encrypt-decrypt
@@ -84,9 +91,10 @@ void test_gift_128(void)
                 key_rand(key);
                 m_rand(m, 16);
 
-                gift_128_encrypt(c, m, key);
+                gift_128_generate_round_keys(rks, key);
+                gift_128_encrypt(c, m, rks);
                 uint8_t m_actual[16];
-                gift_128_decrypt(m_actual, c, key);
+                gift_128_decrypt(m_actual, c, rks);
                 ASSERT_TRUE(memcmp(m, m_actual, 8) == 0);
         }
 }
@@ -96,23 +104,17 @@ void test_gift_64_sliced(void)
         // test encrypt to known value (8 times the same)
[…]

### 2023-04-06 impl/vector/gift_vec_sliced.c

@@ -60,7 +60,7 @@ uint8x16_t shr(const uint8x16_t v, const int n)
 {
         uint64_t l[2];
         vst1q_u64(l, v);
-        l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
+        l[0] = l[0] >> n | (l[1] << (64 - n));
         l[1] >>= n;
         return vld1q_u64(l);
 }

### 2023-04-07 Paper/chapters/implementation.tex
start work on camellia

@@
 are interleaved. The compiler recognizes data dependencies and can therefore
-proceed with round key addition directly after a slice has been permutated
+proceed with round key addition immediately after a slice has been permuted
 without needing to wait for all permutations to finish. This is only logical
@@
 processor happily continues chugging away at the instruction stream by issuing
-\texttt{tbl} and \texttt{mvn} $\mu$-ops to other execution units.
+\texttt{tbl} and \texttt{mvn} $\mu$ops to other execution units.
 

### 2023-04-07 Paper/chapters/introduction.tex
throughput data
start work on camellia

@@
 designed by ARM power 95\% of mobile devices and have found a great many
-applications due to their high performance and low power consumption. The
-ODROID-N2+\cite{odroidn2} development board we are using is based on the big.LITTLE
-architecture and is powered by a quad-core ARM Cortex-A73 processor and a
-weaker dual-core ARM Cortex-A53 for power efficiency. Both these processors are
-part of the eight generation of ARM designs known as ARMv8\cite{armv8:2013}.
+applications due to their high performance and low power
+consumption\cite{armcompany}. The ODROID-N2+\cite{odroidn2} development board
+we are using is based on the big.LITTLE architecture and is powered by a
+quad-core ARM Cortex-A73 processor and a weaker dual-core ARM Cortex-A53 for
+power efficiency. Both these processors are part of the eight generation of ARM
+designs known as ARMv8\cite{armv8:2013}.
 
@@
         Offset & \texttt{LDR W0, [X1, \#12]} & \texttt{W0 = *(X1 + 12);} \\
-        Pre-indexing & \texttt{LDR W0, [X1, \#12]!} & \texttt{X1 += 12; W0 = *(X1)} \\
-        Post-indexing & \texttt{LDR W0, [X1], \#12} & \texttt{W0 = *(X1); X1 += 12} \\
+        Pre-indexing & \texttt{LDR W0, [X1, \#12]!} & \texttt{X1 += 12; W0 = *(X1);} \\
+        Post-indexing & \texttt{LDR W0, [X1], \#12} & \texttt{W0 = *(X1); X1 += 12;} \\
         \bottomrule
     \end{tabularx}
-    \caption{AArch64 addressing modes}
+    \caption{A64 addressing modes}
 \end{table}
@@
 
-ARMv8 supports single-instruction, multiple-data ARMv8 supports
-single-instruction, multiple-data (SIMD) processing. These systems allow the
-programmer to store multiple pieces of data in a vector and work on them in
-parallel to speed up calculations. The A64 instruction set defines two possible
-SIMD implementations.
+ARMv8 supports single-instruction, multiple-data (SIMD) processing. These systems
+allow the programmer to store multiple pieces of data in a vector and work on
+them in parallel to speed up calculations. The A64 instruction set defines two
+possible SIMD implementations:
 
@@
 The register file of the NEON unit is made up of 32 quad-word (128-bit)
-registers \texttt{V[0-31]}, each extending the standard 64-bit floating-point
-registers \mbox{\texttt{D[0-31]}}. These registers are divided into equally
-sized lanes on which the vector instructions operate. Valid ways to interpret
-for example the register \texttt{V0} are:
+registers \texttt{V0-V31}, each extending the standard 64-bit floating-point
+registers \mbox{\texttt{D0-31}}. These registers are divided into equally sized
+lanes on which vector instructions operate. Figure \ref{fig:regdivs} shows
+valid ways to interpret the register \texttt{V0}.
 
@@
     \caption{Divisions of the V0 register}
+    \label{fig:regdivs}
 \end{figure}
@@
 NEON instructions interpret their operands' layouts (i.e. lane count and width)
-through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. For example,
-adding eight 16-bit halfwords from register \texttt{V1} and \texttt{V2}
-together and storing the result in \texttt{V0} can be done as follows:
+through the use of suffixes such as \texttt{.4S} or \texttt{.8H}. Adding eight
+16-bit halfwords stored in \texttt{V1} and \texttt{V2} can be done as follows:
 
@@
         \toprule
-        Intrinsic && Description \\
-        \midrule
-        \texttt{uint8x16\_t} & \texttt{vreinterpretq\_u8\_u64(uint64x2\_t)} & Explicit casting \\
+        Intrinsic && Summary \\
         \midrule

### 2023-04-07 impl/camellia/naive.c

@@ -1,0 +1,145 @@
+#include "naive.h"
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+static const uint64_t keysched_const[] = {
+        0xa09e667f3bcc908bUL,
+        0xb67ae8584caa73b2UL,
+        0xc6ef372fe94f82beUL,
+        0x54ff53a5f1d36f1cUL,
+        0x10e527fade682d1dUL,
+        0xb05688c2b3e6c1fdUL
+};
+
+// rotates for values larger than 64 are done by swapping array elements
+#define rol128(_b, _a, _n){\
+        uint64_t _t = _a >> (64 - _n);\
+        _a = (_a << _n) | (_b >> (64 - _n));\
+        _b = (_b << _n) | _t;\
+}
+
+#define rol32expr(_a, _n)\
+        ((_a << _n) | (_a >> (32 - _n)))
+
+int main(int argc, char *argv[])
+{
+
+}
+
+uint64_t camellia_naive_S(uint64_t X, uint64_t k)
+{
+
+}
+
+uint64_t camellia_naive_P(uint64_t X, uint64_t k)
+{
+
+}
+
+uint64_t camellia_naive_F(uint64_t X, uint64_t k)
+{
+
+}
+
+uint64_t camellia_naive_FL(uint64_t X, uint64_t kl)
+{
+        uint32_t XL = (X >> 32);
+        uint32_t XR = X & 0xffffffff;
+
+        uint32_t klL = (kl >> 32);
+        uint32_t klR = kl & 0xffffffff;
+
+        uint32_t YR = rol32expr(XL & klL, 1) ^ XR;
+        uint32_t YL = (YR | klR) ^ XL;
+
+        return (uint64_t)YR | ((uint64_t)YL << 32);
+}
+
+uint64_t camellia_naive_FL_inv(uint64_t Y, uint64_t kl)
+{
+        uint32_t YL = (Y >> 32);
+        uint32_t YR = Y & 0xffffffff;
+
+        uint32_t klL = (kl >> 32);
+        uint32_t klR = kl & 0xffffffff;
+
+        uint32_t XL = (YR | klR) ^ YL;
+        uint32_t XR = rol32expr(XL & klL, 1) ^ YR;
+
+        return (uint64_t)XR | ((uint64_t)XL << 32);
+}
+
+void camellia_naive_feistel_round(uint64_t state[2], uint64_t kr)
+{
+        uint64_t Lr = state[0] ^ camellia_naive_F(state[1], kr);
[…]

### 2023-04-07 impl/camellia/naive.h

@@ -1,0 +1,26 @@
+#ifndef NAIVE_H
+#define NAIVE_H
+
+#include <stdint.h>
+
+// 128-bit Camellia
+
+struct camellia_keytable {
+        uint64_t kw[4]; // whitening
+        uint64_t ku[18]; // used in F function
+        uint64_t kl[4]; // used in FL layer
+};
+
+uint64_t camellia_naive_S(uint64_t X, uint64_t k);
+uint64_t camellia_naive_P(uint64_t X, uint64_t k);
+uint64_t camellia_naive_F(uint64_t X, uint64_t k);
+uint64_t camellia_naive_FL(uint64_t X, uint64_t kl);
+uint64_t camellia_naive_FL_inv(uint64_t X, uint64_t kl);
+void camellia_naive_feistel_round(uint64_t state[2], uint64_t kr);
+void camellia_naive_generate_round_keys(const uint64_t key[restrict 2],
+                                        struct camellia_keytable *restrict rks);
+
+void camellia_naive_encrypt(uint64_t m[restrict 2],
+                            uint64_t key[restrict 2]);
+
+#endif

### 2023-04-10 Paper/chapters/implementation.tex
camellia introduction

@@
 These kinds of optimizations are pervasive when programming using higher level
-languages like C and modern-day compilers more often than not outperform even
-carefully handcrafted assembly.
+languages like C and modern-day compilers more often than not outperform
+handcrafted assembly.

### 2023-04-10 Paper/chapters/introduction.tex
camellia introduction

@@
 \texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
-and \texttt{GIFT-128} working with 128-bit blocks. In both cases, the key is 128
+and GIFT-128 working with 128-bit blocks. In both cases, the key is 128
 bits long. The design is a very simple, round-based substitution-permutation
@@
 permutation. After the bit permutation, a round key is added to the cipher
-state and the single round is complete. \texttt{GIFT-64} uses 28 rounds while
-\texttt{GIFT-128} uses 40 rounds.
+state and the single round is complete. GIFT-64 uses 28 rounds while
+GIFT-128 uses 40 rounds.
 
@@
 The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
-16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \texttt{GIFT-128}. The S-box
+16 S-boxes for GIFT-64 and 32 S-boxes for GIFT-128. The S-box
 $S:\F{4}\rightarrow \F{4}$ is defined as follows:
@@
 
-The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
-i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64} and
-\texttt{GIFT-128} can be expressed by:
+The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)}$.
+The different permutations for GIFT-64 and GIFT-128 can be expressed by:
 
@@
 
-The last step of each round consists in XORing a round key $R_i$ to the cipher
-state. The new cipher state $s_{i+1}$ after each full round is therefore given
+The last step of each round consists in XORing a round key $R_{i}$ to the cipher
+state. The new cipher state $x_{i+1}$ after each full round is therefore given
 by
@@
 \[
-    s_{i+1}=P(S(s_i))\oplus R_i
+    x_{i+1}=P(S(x_i))\oplus R_i
 \]
@@
 
-Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
-$K=k_7||k_6||\dots||k_0$ denote the $128$-bit key state.
+Round key extraction differs for GIFT-64 and GIFT-128. Let
+$K_{(128)}=k_7||k_6||\dots||k_0$ denote the 128-bit key state.
 
-\paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
-state. These are then added to the 64-bit round key: $R_{4i+1}\leftarrow
-u_i,R_{4i}\leftarrow v_i$.
+\paragraph{GIFT-64}. We extract two words $U_{(16)}||V_{(16)}=k_1||k_0$ from the key
+state. These are then added to round key $R_{(64)}$: $R_{4i+1}\leftarrow
+U_i,R_{4i}\leftarrow V_i$.
 
-\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k_4||k_1||k_0$ from
-the key state. These are then added to the 128-bit round key: $R_{4i+2}\leftarrow
-u_i,R_{4i+1}\leftarrow v_i$.
+\paragraph{GIFT-128}. We extract two words
+$U_{(32)}||V_{(32)}=k_5||k_4||k_1||k_0$ from the key state. These are then
+added to round key $R_{(128)}$: $R_{4i+2}\leftarrow U_i,R_{4i+1}\leftarrow
+V_i$.
 
-In both cases, we additionally XOR a round constant $C=c_5c_4c_3c_2c_1c_0$ to
-bit positions $n-1,23,19,15,11,7,3$. The round constants are generated using a
-6-bit affine linear-feedback shift register and have the following values:\\
+In both cases, we additionally XOR a round constant $C_{(6)}$ to bit positions
+$n-1,23,19,15,11,7,3$. The round constants are generated using a 6-bit affine
+linear-feedback shift register and have the following values:\\
 
@@
 
-The key state is then updated by setting $k_1\leftarrow k_1\ggg 2$,
-$k_0\leftarrow k_0\ggg 12$ and rotating the new state $32$ bits to the right:
+The key state is then updated by individually rotating $k_1$ and $k_0$ and
+rotating the new state $32$ bits to the right:
 
@@
 \subsection{Camellia}
+
+Camellia\cite{camellia:2001} is a block cipher jointly developed by NTT and
+Mitsubishi Electric Corporation and first published in 2001. Following AES
+specifications, it is able to encrypt 128-bit blocks using either 128-, 196- or
+256-bit keys and claims to possess similar performance and security levels as
+the AES finalists.
+
+\subsubsection{Encryption overview}
+
+The encryption process has an 18-round Feistel structure for 128-bit keys and a
+24-round Feistel structure for 192/256-bit key and employs key whitening to
+increase security. First, subkeys $kw_{t(64)}(t=0,1,2,3)$,
+$k_{u(64)}(u=1,2,\dots,(17|23)$ and $kl_{v(64)}(v=0,1,2,3)$ are generated from
+the master key. Then, pre-whitening keys are applied to the left and right half
+of the plaintext $m_{(128)}=L_{(64)}||R_{(64)}$:
+
+\[
+    L\leftarrow L\oplus kw_0, R\leftarrow R\oplus kw_1
+\]
+
+The next steps differ for 128-bit and 192/256-bit keys:
+
+\begin{table}[h!]
+    \centering
+    \caption{Camellia encryption}
+    \begin{tabular}{lll}
+        \toprule
+        Round $r$ & 128-bit keys & 192/256-bit keys \\
+        \midrule
+        0-4 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
+        \midrule
+        5 & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_0,kl_1)\end{aligned}$} & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_0,kl_1)\end{aligned}$} \\
+        \midrule
+        6-10 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
+        \midrule
+        11 & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_2,kl_3)\end{aligned}$} & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_2,kl_3)\end{aligned}$} \\
+        \midrule
+        12-16 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
+        \midrule
+        17 & & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_4,kl_5)\end{aligned}$} \\
+        \midrule
+        18-23 & & $FE(L,R,k_r)$ \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+Finally, $R$ and $L$ are concatenated and XORed with the post-whitening keys to
+obtain the cipher text $c_{(128)}$:
+
+\[
+    c=(R\oplus kw_2)||(L\oplus kw_3)
+\]
+
+\subsubsection{Components}
+
+$FE$ and $FLA$ are macros for applying one round of the Feistel structure and
+$FL$ functions and are defined as follows:
 

### 2023-04-10 Paper/paper.tex
remove implementation.tex, fancyhdr
begin work on benchmark
benchmark
bibliography no ul
implementation chapter
camellia introduction

@@
 
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{ulem} % underline
+\usepackage{amsmath, amsfonts, amssymb, amsthm}
+\usepackage[normalem]{ulem} % underline (don't redefine emph to underline!)
 \usepackage{float} % H float specifier
-\usepackage{hyperref} % hyperlinks
 \usepackage{tabularx} % tabular with column specifier X to break lines
-% \usepackage{hyphenat}
-\usepackage{bookmark}
+% \usepackage{hyphenat} % disable/enable hyphenation
+\usepackage{bookmark} % this loads hyperref
+\usepackage{multirow}
+\usepackage{booktabs} % (top|mid|bottom)rule and other table stuff
+\usepackage{fancyhdr} % fancy headers
 \usepackage{systeme} % systems of equation
@@
 \usepackage{parcolumns} % multiple column parallel typesetting
-\usepackage[english]{babel} % comment this for Angelsächsisch
+\usepackage[english]{babel} % uncomment this for Angelsächsisch
 \usepackage{graphicx} % including graphics
@@
 \usepackage{listings} % source code listings
-\usepackage{biblatex} % better citation
-% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry
+\usepackage{caption} % customize captions in floats (rotate, sideways, ...)
+\usepackage{lstlangarm} % custom ARM Assembler keyword list
+\usepackage[backend=biber, sorting=none]{biblatex} % better citation
+\usepackage[a4paper]{geometry} % customize page geometry
+
+\pagestyle{fancy}
+\fancyhead[L]{\rightmark}
+\fancyhead[R]{\thepage}
+\fancyfoot{}
+
+% non-italic bibliography titles
+\DeclareFieldFormat{title}{``#1''}
+\DeclareFieldFormat{citetitle}{``#1''}
 

### 2023-04-11 Paper/chapters/introduction.tex
rearrange table captions
more intro

@@
 increase security. First, subkeys $kw_{t(64)}(t=0,1,2,3)$,
-$k_{u(64)}(u=1,2,\dots,(17|23)$ and $kl_{v(64)}(v=0,1,2,3)$ are generated from
-the master key. Then, pre-whitening keys are applied to the left and right half
-of the plaintext $m_{(128)}=L_{(64)}||R_{(64)}$:
+$k_{u(64)}(u=0,1,\dots,(17|23)$ and $kl_{v(64)}(v=0,1,2,3)$ are generated from
+the master key. Then, pre-whitening keys are applied to the plaintext
+$m_{(128)}=L_{(64)}||R_{(64)}$:
 
 \[
-    L\leftarrow L\oplus kw_0, R\leftarrow R\oplus kw_1
+    (L||R)\leftarrow (L||R)\oplus (kw_0||kw_1)
 \]
 
-The next steps differ for 128-bit and 192/256-bit keys:
+The next steps differ for 128-bit and 192/256-bit keys in the number of rounds:
 
@@
         \midrule
-        0-4 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
+        0-4 & $(L||R)\leftarrow FE(L,R,k_r)$ & $(L||R)\leftarrow FE(L,R,k_r)$ \\
         \midrule
-        5 & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_0,kl_1)\end{aligned}$} & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_0,kl_1)\end{aligned}$} \\
+        5 & {$\!\begin{aligned}&(L||R)\leftarrow FE(L,R)\\&(L||R)\leftarrow FLL(L,R,kl_0,kl_1)\end{aligned}$} & {$\!\begin{aligned}&(L||R)\leftarrow FE(L,R)\\&(L||R)\leftarrow FLL(L,R,kl_0,kl_1)\end{aligned}$} \\
         \midrule
-        6-10 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
+        6-10 & $(L||R)\leftarrow FE(L,R,k_r)$ & $(L||R)\leftarrow FE(L,R,k_r)$ \\
         \midrule
-        11 & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_2,kl_3)\end{aligned}$} & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_2,kl_3)\end{aligned}$} \\
+        11 & {$\!\begin{aligned}&(L||R)\leftarrow FE(L,R)\\&(L||R)\leftarrow FLL(L,R,kl_2,kl_3)\end{aligned}$} & {$\!\begin{aligned}&(L||R)\leftarrow FE(L,R)\\&(L||R)\leftarrow FLL(L,R,kl_2,kl_3)\end{aligned}$} \\
         \midrule
-        12-16 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
+        12-16 & $(L||R)\leftarrow FE(L,R,k_r)$ & $(L||R)\leftarrow FE(L,R,k_r)$ \\
         \midrule
-        17 & & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_4,kl_5)\end{aligned}$} \\
+        17 & & {$\!\begin{aligned}&(L||R)\leftarrow FE(L,R)\\&(L||R)\leftarrow FLL(L,R,kl_4,kl_5)\end{aligned}$} \\
         \midrule
-        18-23 & & $FE(L,R,k_r)$ \\
+        18-23 & & $(L||R)\leftarrow FE(L,R,k_r)$ \\
         \bottomrule
@@
 \[
-    c=(R\oplus kw_2)||(L\oplus kw_3)
+    c=(R||L)\oplus (kw_2)||kw_3)
 \]
@@
 
-$FE$ and $FLA$ are macros for applying one round of the Feistel structure and
-$FL$ functions and are defined as follows:
+We will give an overview of the main functional components of Camellia.
+
+\paragraph{Feistel round function $FE$}
+
+\begin{align*}
+    FE:(\F{64})^3 &\rightarrow (\F{64})^2 \\
+    (L_{(64)},R_{(64)},k_{(64)})&\mapsto (R\oplus F(L,k), L)
+\end{align*}
+
+\paragraph{SP-function $F$}
+
+\begin{align*}
+    F:(\F{64})^2 &\rightarrow \F{64} \\
+    (X_{(64)},k_{(64)})&\mapsto P(S(X\oplus k))
+\end{align*}
+
+\paragraph{Substitution function $F$}
+
+\begin{align*}
+    S:\F{64} &\rightarrow \F{64} \\
+    l_{0(8)||l_{1(8)||l_{2(8)||l_{3(8)||l_{4(8)||l_{5(8)||l_{6(8)||l_{7(8)||&\mapsto P(S(X\oplus k))
+\end{align*}
+
+\paragraph{$FL$ layer function $FLL$}
+
+\begin{align*}
+    FLL:(\F{64})^4 &\rightarrow (\F{64})^2 \\
+    (L_{(64)},R_{(64)},k_{0(64)},k_{1(64)})&\mapsto (FL(L,k_0),FL^{-1}(R,k_1))
+\end{align*}
 
@@
     \centering
+    \caption{ARMv8 profiles}
     \begin{tabularx}{\textwidth}{lX}
@@
     \end{tabularx}
-    \caption{ARMv8 profiles}
 \end{table}
@@
     \centering
+    \caption{ARMv8 execution states}
     \begin{tabularx}{\textwidth}{llX}
@@
     \end{tabularx}
-    \caption{ARMv8 execution states}
 \end{table}
@@
     \small
+    \caption{A64 addressing modes}
     \begin{tabularx}{\textwidth}{llX}
@@
     \end{tabularx}
-    \caption{A64 addressing modes}
 \end{table}
@@
     \footnotesize
+    \caption{Common NEON intrinsics}
     \begin{tabularx}{\textwidth}{llX}
@@
     \end{tabularx}
-    \caption{Common NEON intrinsics}
 \end{table}

### 2023-04-14 Paper/chapters/implementation-strategies.tex
more absorption

@@
 \section{Strategies for Camellia}
+
+\subsection{Platform-independent techniques}
+
+The original paper proposes various platform-independent ways to implement
+Camellia efficiently. Only some of these apply to the ARMv8 architecture since
+features like the inline barrel shifter and bitfield manipulation instructions
+generally offer better performance.
+
+\begin{description}
+    \item[XOR cancellation property in key schedule:]
+        \begin{align*}
+            K_A&\leftarrow K_L\oplus K_R \\
+            K_A&\leftarrow FE(FE(K_A, \Sigma_0=0xa09e667f3bcc908b), \Sigma_1=0xb67ae8584caa73b2) \\
+            K_A&\leftarrow K_A\oplus K_L \\
+        \end{align*}
+
+        causes cancellations, allowing us to eliminate some operations by
+        replacing it with the following:
+
+        \begin{align*}
+            K_{A_L}&\leftarrow F(K_{L_R}\oplus F(K_{L_L}, \Sigma_0)) \\
+            K_{A_R}&\leftarrow F(K_{L_L}\oplus \Sigma_1)
+        \end{align*}
+
+    \item[Absorption of whitening keys:] Key additions $kw_1,kw_3$ can be
+        absorped into other subkeys:
+
+\end{description}

### 2023-04-14 Paper/chapters/introduction.tex
camellia intro FL
more absorption

@@
 
-\subsubsection{Encryption overview}
+\subsubsection{Encryption}
 
@@
 
+\subsubsection{Key schedule}
+
+The master key $K$ is split into two parts $K=K_{L(128)}||K_{R(128)}$ with
+$K_R=0$ for 128-bit keys. Then, two variables $K_{A(128)},K_{B(128)}$ are
+generated by repeated application of the round function with key constants
+$\Sigma_i,i=(0,1,\dots,5)$:
+
+\begin{align*}
+    K_A&\leftarrow K_L\oplus K_R \\
+    K_A&\leftarrow FE(FE(K_A, \Sigma_0=0xa09e667f3bcc908b), \Sigma_1=0xb67ae8584caa73b2) \\
+    K_A&\leftarrow K_A\oplus K_L \\
+    K_A&\leftarrow FE(FE(K_A, \Sigma_2=0xc6ef372fe94f82be), \Sigma_3=0x54ff53a5f1d36f1c) \\
+    K_B&\leftarrow FE(FE(K_A, \Sigma_4=0x10e527fade682d1d), \Sigma_5=0xb05688c2b3e6c1fd)
+\end{align*}
+
+Subkeys are then created by rotating $K_L,K_R,K_A,K_B$:
+
+\begin{table}[h!]
+    \centering
+    \scriptsize
+    \caption{Subkeys for 128-bit key}
+    \begin{tabular}{lll|lll}
+        \toprule
+        Usage & Subkey & Value & Usage & Subkey & Value \\
+        \midrule
+        \multirow{2}{*}{Prewhitening} & $kw_{0(64)}$ & $(K_L{\lll_0})_{L(64)}$ & $F(\text{Round 9})$ & $k_{9(64)}$ & $(K_L{\lll_{60}})_{R(64)}$ \\
+                                      & $kw_{1(64)}$ & $(K_L{\lll_0})_{R(64)}$ & $F(\text{Round 10})$ & $k_{10(64)}$ & $(K_A{\lll_{60}})_{L(64)}$ \\
+        $F(\text{Round 0})$ & $k_{0(64)}$ & $(K_A{\lll_0})_{L(64)}$            & $F(\text{Round 11})$ & $k_{11(64)}$ & $(K_A{\lll_{60}})_{R(64)}$ \\
+        $F(\text{Round 1})$ & $k_{1(64)}$ & $(K_A{\lll_0})_{R(64)}$            & $FL$        & $kl_{2(64)}$ & $(K_L{\lll_{77}})_{L(64)}$ \\
+        $F(\text{Round 2})$ & $k_{2(64)}$ & $(K_L{\lll_{15}})_{L(64)}$         & $FL^{-1}$   & $kl_{3(64)}$ & $(K_L{\lll_{77}})_{R(64)}$ \\
+        $F(\text{Round 3})$ & $k_{3(64)}$ & $(K_L{\lll_{15}})_{R(64)}$         & $F(\text{Round 12})$ & $k_{12(64)}$ & $(K_L{\lll_{94}})_{L(64)}$ \\
+        $F(\text{Round 4})$ & $k_{4(64)}$ & $(K_A{\lll_{15}})_{L(64)}$         & $F(\text{Round 13})$ & $k_{13(64)}$ & $(K_L{\lll_{94}})_{R(64)}$ \\
+        $F(\text{Round 5})$ & $k_{5(64)}$ & $(K_A{\lll_{15}})_{R(64)}$         & $F(\text{Round 14})$ & $k_{14(64)}$ & $(K_A{\lll_{94}})_{L(64)}$ \\
+        $FL$        & $kl_{0(64)}$ & $(K_A{\lll_{30}})_{L(64)}$                & $F(\text{Round 15})$ & $k_{15(64)}$ & $(K_L{\lll_{94}})_{R(64)}$ \\
+        $FL^{-1}$   & $kl_{1(64)}$ & $(K_A{\lll_{30}})_{R(64)}$                & $F(\text{Round 16})$ & $k_{16(64)}$ & $(K_A{\lll_{111}})_{L(64)}$ \\
+        $F(\text{Round 6})$ & $k_{6(64)}$ & $(K_L{\lll_{45}})_{L(64)}$         & $F(\text{Round 17})$ & $k_{17(64)}$ & $(K_A{\lll_{111}})_{R(64)}$ \\
+        $F(\text{Round 7})$ & $k_{7(64)}$ & $(K_L{\lll_{45}})_{R(64)}$         & \multirow{2}{*}{Postwhitening} & $kw_{2(64)}$ & $(K_A{\lll_{111}})_{L(64)}$ \\
+        $F(\text{Round 8})$ & $k_{8(64)}$ & $(K_A{\lll_{45}})_{L(64)}$         &                                & $kw_{3(64)}$ & $(K_A{\lll_{111}})_{R(64)}$ \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+Subkey generation for 192/256-bit keys are generated in a similar way.
+
 \subsubsection{Components}
@@
 
-\paragraph{Feistel round function $FE$}
+\begin{description}
+\item[Feistel round function $FE$:]
 
@@
 
-\paragraph{SP-function $F$}
+\item[SP-function $F$:]
 
@@
 
-\paragraph{Substitution function $F$}
+\item[Substitution function $S$:]
 
@@
     S:\F{64} &\rightarrow \F{64} \\
-    l_{0(8)||l_{1(8)||l_{2(8)||l_{3(8)||l_{4(8)||l_{5(8)||l_{6(8)||l_{7(8)||&\mapsto P(S(X\oplus k))
+    \begin{aligned}
+        l_{0(8)}||l_{1(8)}||l_{2(8)}||l_{3(8)}&||\\
+        l_{4(8)}||l_{5(8)}||l_{6(8)}||l_{7(8)}&
+    \end{aligned}&\mapsto
+    \begin{aligned}
+        s_0(l_0)||s_1(l_1)||s_2(l_2)||s_3(l_3)&||\\
+        s_1(l_4)||s_2(l_5)||s_3(l_6)||s_0(l_7)&
+    \end{aligned},
 \end{align*}
 
-\paragraph{$FL$ layer function $FLL$}
+with 8-bit S-boxes $s_0,s_1,s_2,s_3:\F{8}\rightarrow \F{8}$.
+
+\item[Permutation function $P$:]
+
+\begin{align*}
+    P:\F{64} &\rightarrow \F{64} \\
+    \begin{pmatrix}z_7\\z_6\\z_5\\z_4\\z_3\\z_2\\z_1\\z_0\end{pmatrix}&\mapsto
+    \begin{pmatrix}
+        0 & 1 & 1 & 1 & 1 & 0 & 0 & 1 \\
+        1 & 0 & 1 & 1 & 1 & 1 & 0 & 0 \\
+        1 & 1 & 0 & 1 & 0 & 1 & 1 & 0 \\
+        1 & 1 & 1 & 0 & 0 & 0 & 1 & 1 \\
+        0 & 1 & 1 & 1 & 1 & 1 & 1 & 0 \\
+        1 & 0 & 1 & 1 & 0 & 1 & 1 & 1 \\
+        1 & 1 & 0 & 1 & 1 & 0 & 1 & 1 \\
+        1 & 1 & 1 & 0 & 1 & 1 & 0 & 1
+    \end{pmatrix}\begin{pmatrix}z_7\\z_6\\z_5\\z_4\\z_3\\z_2\\z_1\\z_0\end{pmatrix}
+\end{align*}
+
+\item[$FL$ layer function $FLL$:]
 
@@
     FLL:(\F{64})^4 &\rightarrow (\F{64})^2 \\
-    (L_{(64)},R_{(64)},k_{0(64)},k_{1(64)})&\mapsto (FL(L,k_0),FL^{-1}(R,k_1))
+    (X_{L(64)},X_{R(64)},k_{0(64)},k_{1(64)})&\mapsto (FL(X_L,k_0),FL^{-1}(X_R,k_1))
 \end{align*}
+
+\item[$FL$:]
+
+\begin{align*}
+    FL:(\F{64})^2 &\rightarrow \F{64} \\
+    (X_{L(32)}||X_{R(32)},k_{L(32)}||k_{R(32)})&\mapsto (Y_{L(32)}||Y_{R(32)}),
+\end{align*}
+
+where
+
+\begin{align*}
+    Y_{R(32)}&=((X_L\cap k_{L})\lll_1)\oplus X_{R} \\
+    Y_{L(32)}&=(Y_R\cup k_{R})\oplus X_{L}
+\end{align*}
+
+\item[$FL^{-1}$:]
+
+\begin{align*}
+    FL^{-1}:(\F{64})^2 &\rightarrow \F{64} \\
+    (Y_{L(32)}||Y_{R(32)},k_{L(32)}||k_{R(32)})&\mapsto (X_{L(32)}||X_{R(32)}),
+\end{align*}
+
+where
+
+\begin{align*}
+    X_{L(32)}&=(Y_R\cup k_{R})\oplus Y_{L} \\
+    X_{R(32)}&=((X_L\cap k_{L})\lll_1)\oplus Y_{R}
+\end{align*}
+\end{description}
 

### 2023-04-14 impl/benchmark.c

@@ -6,6 +6,7 @@
 #include "vector/gift_vec_sbox.h"
 #include "vector/gift_vec_sliced.h"
 #include "camellia/naive.h"
+#include "camellia/spec_opt.h"
 
 #include <stdio.h>
 #include <stdlib.h>
@@ -275,6 +276,58 @@ static void benchmark_camellia_naive(void)
         printf("throughput: %f MiB/s\n", megs / seconds);
 }
 
+static void benchmark_camellia_spec_opt(void)
+{
+        printf("Benchmaring CAMELLIA_SPEC_OPT 128-bit...\n");
+
+        uint64_t key[2];
+        uint64_t m[2], c[2];
+        struct camellia_keys_128 rks;
+
+        uint64_t cycles[2] = { 0UL };
+        for (int i = 0; i < NL; i++) {
+                cycles[0] += TIME(camellia_spec_opt_generate_round_keys_128(key, &rks));
+                cycles[1] += TIME(camellia_spec_opt_encrypt_128(c, m, &rks));
+        }
+
+        printf("%f %f\n",
+               cycles[0] / (float)NL / 16.0f,
+               cycles[1] / (float)NL / 16.0f);
+
+        struct timeval st, et;
+        gettimeofday(&st, NULL);
+        for (int i = 0; i < NT; i++) {
+                camellia_spec_opt_encrypt_128(c, m, &rks);
+        }
+        gettimeofday(&et, NULL);
+        double seconds = elapsed_seconds(&st, &et);
+        double megs = NT * sizeof(m) / (1024 * 1024);
+        printf("throughput: %f MiB/s\n", megs / seconds);
+
+        printf("Benchmaring CAMELLIA_SPEC_OPT 256-bit...\n");
+        uint64_t key_256[4];
+        struct camellia_keys_256 rks_256;
+
+        memset(cycles, 0, sizeof(cycles));
+        for (int i = 0; i < NL; i++) {
+                cycles[0] += TIME(camellia_spec_opt_generate_round_keys_256(key_256, &rks_256));
+                cycles[1] += TIME(camellia_spec_opt_encrypt_256(c, m, &rks_256));
+        }
+
+        printf("%f %f\n",
+               cycles[0] / (float)NL / 16.0f,
+               cycles[1] / (float)NL / 16.0f);
+
+        gettimeofday(&st, NULL);
+        for (int i = 0; i < NT; i++) {
[…]

### 2023-04-14 impl/camellia/camellia_keys.h

@@ -1,0 +1,19 @@
+#ifndef CAMELLIA_KEYS
+#define CAMELLIA_KEYS
+
+#include <stdint.h>
+
+struct camellia_keys_128 {
+        uint64_t kw[4]; // whitening
+        uint64_t ku[18]; // used in F function
+        int64_t kl[4]; // used in FL layer
+};
+
+struct camellia_keys_256 {
+        uint64_t kw[4]; // whitening
+        uint64_t ku[24]; // used in F function
+        uint64_t kl[6]; // used in FL layer
+};
+
+
+#endif

### 2023-04-14 impl/camellia/spec_opt.c

@@ -1,4 +1,4 @@
-#include "naive.h"
+#include "spec_opt.h"
 
 #include <stdio.h>
 #include <stdlib.h>
@@ -169,9 +169,9 @@ static const uint8_t s4[256] =
 }
 
 #define expr_rol32_1(_a)\
-        (((_a) << 1) | ((_a) >> 31))
+        (((_a) << 1) + ((_a) >> 31))
 
-uint64_t camellia_naive_S(uint64_t X)
+uint64_t camellia_spec_opt_S(uint64_t X)
 {
         uint64_t result = 0UL;
         result |= (uint64_t)s1[(X >> 56) & 0xff] << 56;
@@ -186,7 +186,7 @@ uint64_t camellia_naive_S(uint64_t X)
         return result;
 }
 
-uint64_t camellia_naive_P(uint64_t X)
+uint64_t camellia_spec_opt_P(uint64_t X)
 {
         uint8_t z1 = X >> 56;
         uint8_t z2 = X >> 48;
@@ -210,12 +210,12 @@ uint64_t camellia_naive_P(uint64_t X)
         return result;
 }
 
-uint64_t camellia_naive_F(uint64_t X, const uint64_t k)
+uint64_t camellia_spec_opt_F(uint64_t X, const uint64_t k)
 {
-        return camellia_naive_P(camellia_naive_S(X ^ k));
+        return camellia_spec_opt_P(camellia_spec_opt_S(X ^ k));
 }
 
-uint64_t camellia_naive_FL(uint64_t X, const uint64_t kl)
+uint64_t camellia_spec_opt_FL(uint64_t X, const uint64_t kl)
 {
         const uint32_t XL = (X >> 32);
         const uint32_t XR = (X >> 0);
@@ -229,7 +229,7 @@ uint64_t camellia_naive_FL(uint64_t X, const uint64_t kl)
         return ((uint64_t)YL << 32) | (uint64_t)YR;
 }
 
-uint64_t camellia_naive_FL_inv(uint64_t Y, const uint64_t kl)
+uint64_t camellia_spec_opt_FL_inv(uint64_t Y, const uint64_t kl)
 {
         const uint32_t YL = (Y >> 32);
         const uint32_t YR = (Y >> 0);
@@ -243,184 +243,259 @@ uint64_t camellia_naive_FL_inv(uint64_t Y, const uint64_t kl)
         return ((uint64_t)XL << 32) | (uint64_t)XR;
 }
 
[…]

### 2023-04-14 impl/camellia/spec_opt.h

@@ -1,0 +1,37 @@
+#ifndef SPEC_OPT_H
+#define SPEC_OPT_H
+
+// 128-bit and 256-bit Camellia
+
+#include "camellia_keys.h"
+#include <stdint.h>
+
+uint64_t camellia_spec_opt_S(uint64_t X);
+uint64_t camellia_spec_opt_P(uint64_t X);
+uint64_t camellia_spec_opt_F(uint64_t X, const uint64_t k);
+uint64_t camellia_spec_opt_FL(uint64_t X, const uint64_t kl);
+uint64_t camellia_spec_opt_FL_inv(uint64_t X, const uint64_t kl);
+void camellia_spec_opt_feistel_round(uint64_t state[2], const uint64_t kr);
+void camellia_spec_opt_feistel_round_inv(uint64_t state[2], const uint64_t kr);
+void camellia_spec_opt_generate_round_keys_128(const uint64_t key[restrict 2],
+                                            struct camellia_keys_128 *restrict rks);
+void camellia_spec_opt_generate_round_keys_256(const uint64_t key[restrict 4],
+                                            struct camellia_keys_256 *restrict rks);
+
+void camellia_spec_opt_encrypt_128(uint64_t c[restrict 2],
+                            const uint64_t m[restrict 2],
+                            struct camellia_keys_128 *restrict rks);
+
+void camellia_spec_opt_decrypt_128(uint64_t m[restrict 2],
+                            const uint64_t c[restrict 2],
+                            struct camellia_keys_128 *restrict rks);
+
+void camellia_spec_opt_encrypt_256(uint64_t c[restrict 2],
+                            const uint64_t m[restrict 2],
+                            struct camellia_keys_256 *restrict rks);
+
+void camellia_spec_opt_decrypt_256(uint64_t m[restrict 2],
+                            const uint64_t c[restrict 2],
+                            struct camellia_keys_256 *restrict rks);
+
+#endif

### 2023-04-14 impl/test.c

@@ -5,6 +5,7 @@
 #include "vector/gift_vec_sliced.h"
 
 #include "camellia/naive.h"
+#include "camellia/spec_opt.h"
 
 #include <stdio.h>
 #include <stdlib.h>
@@ -279,8 +280,8 @@ void test_camellia_naive(void)
         for (int i = 0; i < 256; i++) {
                 m_rand((uint8_t*)m, sizeof(m));
                 key_rand(key);
-                ASSERT_EQUALS(camellia_naive_FL(camellia_naive_FL_inv(m[0], key[0]), key[0]), m[0]);
-                ASSERT_EQUALS(camellia_naive_FL_inv(camellia_naive_FL(m[1], key[1]), key[1]), m[1]);
+                ASSERT_EQUALS(camellia_naive_FL(camellia_spec_opt_FL_inv(m[0], key[0]), key[0]), m[0]);
+                ASSERT_EQUALS(camellia_naive_FL_inv(camellia_spec_opt_FL(m[1], key[1]), key[1]), m[1]);
 
                 uint64_t mc[2];
                 memcpy(mc, m, sizeof(mc));
@@ -347,6 +348,85 @@ void test_camellia_naive(void)
         }
 }
 
+void test_camellia_spec_opt(void)
+{
+        uint64_t m[2], c[2];
+        uint64_t key[2];
+        struct camellia_keys_128 rks_128;
+        struct camellia_keys_256 rks_256;
+
+        printf("testing CAMELLIA_SPEC_OPT FL-FL_inv and feistel-feistel_inv...\n");
+        for (int i = 0; i < 256; i++) {
+                m_rand((uint8_t*)m, sizeof(m));
+                key_rand(key);
+                ASSERT_EQUALS(camellia_spec_opt_FL(camellia_spec_opt_FL_inv(m[0], key[0]), key[0]), m[0]);
+                ASSERT_EQUALS(camellia_spec_opt_FL_inv(camellia_spec_opt_FL(m[1], key[1]), key[1]), m[1]);
+
+                uint64_t mc[2];
+                memcpy(mc, m, sizeof(mc));
+                camellia_spec_opt_feistel_round(mc, key[0]);
+                camellia_spec_opt_feistel_round_inv(mc, key[0]);
+                ASSERT_TRUE(memcmp(mc, m, sizeof(mc)) == 0);
+                camellia_spec_opt_feistel_round_inv(mc, key[1]);
+                camellia_spec_opt_feistel_round(mc, key[1]);
[…]

### 2023-04-16 Paper/chapters/implementation-strategies.tex
ps tabulation paper

@@
             K_A&\leftarrow K_L\oplus K_R \\
-            K_A&\leftarrow FE(FE(K_A, \Sigma_0=0xa09e667f3bcc908b), \Sigma_1=0xb67ae8584caa73b2) \\
+            K_A&\leftarrow FE(FE(K_A, \Sigma_0=\mathrm{0xa09e667f3bcc908b}), \Sigma_1=\mathrm{0xb67ae8584caa73b2}) \\
             K_A&\leftarrow K_A\oplus K_L \\
@@
 
-    \item[Absorption of whitening keys:] Key additions $kw_1,kw_3$ can be
-        absorped into other subkeys:
+    \item[Absorption of whitening keys:] Whitening keys $kw_1,kw_3$ can be
+        absorbed into other subkeys to save two XOR operations
+
+    \item[Efficiently computing $P\circ S$:] This technique applies to 64-bit processors.
+        By preparing tables
+
+        \begin{alignat*}{9}
+            & SP_0(y_{0(8)})=(&s_0(y_0),&& s_0(y_0),&& s_0(y_0),&& 0,&& s_0(y_0),&& 0,&& 0,&& s_0(y_0)&) \\
+            & SP_1(y_{1(8)})=(&0,&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& 0,&& 0&) \\
+            & SP_2(y_{2(8)})=(&s_2(y_2),&& 0,&& s_2(y_2),&& s_2(y_2),&& 0,&& s_2(y_2),&& s_2(y_2),&& 0&) \\
+            & SP_3(y_{3(8)})=(&s_3(y_3),&& s_3(y_3),&& 0,&& s_3(y_3),&& 0,&& 0,&& s_3(y_3),&& s_3(y_3)&) \\
+            & SP_4(y_{4(8)})=(&0,&& s_2(y_4),&& s_2(y_4),&& s_2(y_4),&& 0,&& s_2(y_4),&& s_2(y_4),&& s_2(y_4)&) \\
+            & SP_5(y_{5(8)})=(&s_3(y_5),&& 0,&& s_3(y_5),&& s_3(y_5),&& s_3(y_5),&& 0,&& s_3(y_5),&& s_3(y_5)&) \\
+            & SP_6(y_{6(8)})=(&s_4(y_6),&& s_4(y_6),&& 0,&& s_4(y_6),&& s_4(y_6),&& s_4(y_6),&& 0,&& s_4(y_6)&) \\
+            & SP_7(y_{7(8)})=(&s_1(y_7),&& s_1(y_7),&& s_1(y_7),&& 0,&& s_1(y_7),&& s_1(y_7),&& s_1(y_7),&& 0&)
+        \end{alignat*}
+
+        , we can compute $(P\circ S)(X_{(64)})$ with only 8 table lookups and 7
+        XORs in the following way:
+
+        \[
+            (z_1', z_2', z_3', z_4', z_5', z_6', z_7', z_8')\leftarrow\bigoplus_{i=0}^7 SP_i(y_i)
+        \]
 

### 2023-04-16 Paper/chapters/introduction.tex
ps tabulation paper

@@
     K_A&\leftarrow K_L\oplus K_R \\
-    K_A&\leftarrow FE(FE(K_A, \Sigma_0=0xa09e667f3bcc908b), \Sigma_1=0xb67ae8584caa73b2) \\
+    K_A&\leftarrow FE(FE(K_A, \Sigma_0\mathrm{0xa09e667f3bcc908b}), \Sigma_1\mathrm{0xb67ae8584caa73b2}) \\
     K_A&\leftarrow K_A\oplus K_L \\
-    K_A&\leftarrow FE(FE(K_A, \Sigma_2=0xc6ef372fe94f82be), \Sigma_3=0x54ff53a5f1d36f1c) \\
-    K_B&\leftarrow FE(FE(K_A, \Sigma_4=0x10e527fade682d1d), \Sigma_5=0xb05688c2b3e6c1fd)
+    K_A&\leftarrow FE(FE(K_A, \Sigma_2\mathrm{0xc6ef372fe94f82be}), \Sigma_3\mathrm{0x54ff53a5f1d36f1c}) \\
+    K_B&\leftarrow FE(FE(K_A, \Sigma_4\mathrm{0x10e527fade682d1d}), \Sigma_5\mathrm{0xb05688c2b3e6c1fd})
 \end{align*}
@@
     \scriptsize
-    \caption{Subkeys for 128-bit key}
+    \caption{Subkey creation for 128-bit keys}
     \begin{tabular}{lll|lll}
@@
 
-Subkey generation for 192/256-bit keys are generated in a similar way.
+Subkeys for 192/256-bit keys are generated in a similar way.
 

### 2023-05-08 Paper/chapters/implementation-strategies.tex
camellia galois affine isomorphism

@@
 
-Bitslicing refers to the technique of splitting up $n$ bits into $m$ slices to
-achieve a more efficient representation to operate on. The structure of
+Bitsliced implementation techniques were first introduced to improve the performance
+of DES in 1997 and work by viewing a processor with $n$-bit registers as a
+machine capable of executing $n$ bitwise operations at
+once\cite{desslicing:1997}. Bitslicing offers a performance advantage by
+splitting up $n$ bits into $m$ slices to achieve a more efficient
+representation which can exploit this bitwise parallelism. The structure of
 \texttt{GIFT} naturally offers possibilities for bitslicing. We split the
@@
 of the S-box can be executed in parallel, similar to the \texttt{vperm}
-technique above. This can be done by finding an algorithmic way to apply the
-S-box which has already been proposed by the original \texttt{GIFT} authors:
+technique above. This can be done by finding a bitwise instruction sequence to
+apply the S-box which has already been proposed by the original \texttt{GIFT}
+authors:
 
@@
     \item[XOR cancellation property in key schedule:]
+        While deriving $K_A$ in the key schedule, the instruction sequence
+
         \begin{align*}
             K_A&\leftarrow K_L\oplus K_R \\
-            K_A&\leftarrow FE(FE(K_A, \Sigma_0=\mathrm{0xa09e667f3bcc908b}), \Sigma_1=\mathrm{0xb67ae8584caa73b2}) \\
+            K_A&\leftarrow FE(FE(K_A, \Sigma_0), \Sigma_1) \\
             K_A&\leftarrow K_A\oplus K_L \\
@@
 
-    \item[Efficiently computing $P\circ S$:] This technique applies to 64-bit processors.
+    \item[Efficiently computing $(P\circ S)$:] This technique applies to 64-bit processors.
         By preparing tables
@@
 
-        , we can compute $(P\circ S)(X_{(64)})$ with only 8 table lookups and 7
+        , we can compute $(P\circ S)(X_{(64)})$ using only 8 table lookups and 7
         XORs in the following way:
@@
 \end{description}
+
+\subsection{Byteslicing}
+
+A byte-sliced implementation strategy for ARMv8 can be derived from already
+existing x86-optimized implementations utilizing the AES-NI advanced encryption
+standard instruction set \cite{bcfastimplx86:2013}. NEON itself possesses
+cryptographic extensions for finite field arithmetic and AES as well as SHA
+calculations. These can be used to produce an accelerated Camellia implementation
+due to the algebraic similarity of the AES- and Camellia S-boxes.
+
+\subsubsection{Hardware-accelerated Camellia S-box}

### 2023-05-08 impl/benchmark.c

@@ -230,7 +230,7 @@ static void benchmark_camellia_naive(void)
 
         uint64_t key[2];
         uint64_t m[2], c[2];
-        struct camellia_keys_128 rks;
+        struct camellia_rks_128 rks;
 
         uint64_t cycles[2] = { 0UL };
         for (int i = 0; i < NL; i++) {
@@ -254,7 +254,7 @@ static void benchmark_camellia_naive(void)
 
         printf("Benchmaring CAMELLIA_NAIVE 256-bit...\n");
         uint64_t key_256[4];
-        struct camellia_keys_256 rks_256;
+        struct camellia_rks_256 rks_256;
 
         memset(cycles, 0, sizeof(cycles));
         for (int i = 0; i < NL; i++) {
@@ -282,7 +282,7 @@ static void benchmark_camellia_spec_opt(void)
 
         uint64_t key[2];
         uint64_t m[2], c[2];
-        struct camellia_keys_128 rks;
+        struct camellia_rks_128 rks;
 
         uint64_t cycles[2] = { 0UL };
         for (int i = 0; i < NL; i++) {

### 2023-05-08 impl/camellia/bytesliced.c

@@ -1,0 +1,151 @@
+#include "bytesliced.h"
+#include "bytesliced.h"
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+static const uint64_t keysched_const[] = {
+        0xa09e667f3bcc908bUL,
+        0xb67ae8584caa73b2UL,
+        0xc6ef372fe94f82beUL,
+        0x54ff53a5f1d36f1cUL,
+        0x10e527fade682d1dUL,
+        0xb05688c2b3e6c1fdUL
+};
+
+static uint64_t rol1_tbl_u64[] = {
+        0x060504030201001fUL, 0x0e0d0c0b0a090807UL
+};
+
+static uint64_t rol8_tbl_u64[] = {
+        0x1f1e1d1c1b1a1918UL, 0x0706050403020100UL
+};
+
+static uint64_t shiftrows_inv_u64[2] = {
+        0x0b0e0104070a0d00UL, 0x0306090c0f020508UL
+};
+
+static uint64_t prefilter_0_u64[][2] = {
+        { 0x862b832eed40e845UL, 0x88258d20e34ee64bUL },
+        { 0x2a7bdb8aa0f15100UL, 0x2372d283a9f85809UL }
+};
+
+static uint64_t prefilter_3_u64[][2] = {
+        { 0x25204e4b2b2e4045UL, 0x74711f1a7a7f1114UL },
+        { 0x7283f8097b8af100UL, 0xdf2e55a4d6275cadUL }
+};
+
+static uint64_t postfilter_0_u64[][2] = {
+        { 0x31c1c2323fcfcc3cUL, 0xd12122d2df2f2cdcUL },
+        { 0xa8512ed77f86f900UL, 0x0cf58a73db225da4UL }
+};
+
+static uint64_t postfilter_1_u64[][2] = {
+        { 0x628385647e9f9978UL, 0xa34244a5bf5e58b9UL },
+        { 0x51a25caffe0df300UL, 0x18eb15e6b744ba49UL }
+};
+
+static uint64_t postfilter_2_u64[][2] = {
+        { 0x98e061199fe7661eUL, 0xe8901169ef97166eUL },
+        { 0x54a817ebbf43fc00UL, 0x06fa45b9ed11ae52UL }
+};
+
+static uint8x16_t rol1_tbl;
+static uint8x16_t rol8_tbl;
+static uint8x16_t lower_4_bits_mask;
+static uint8x16_t shiftrows_inv;
+
[…]

### 2023-05-08 impl/camellia/bytesliced.h

@@ -1,0 +1,44 @@
+#ifndef BYTESLICED_H
+#define BYTESLICED_H
+
+// 128-bit bitsliced camellia
+
+#include <stdint.h>
+#include <arm_neon.h>
+
+struct camellia_rks_sliced_128 {
+        uint8x16x4_t kw[2][4]; // whitening
+        uint8x16x4_t ku[2][18]; // used in F function
+        uint8x16x4_t kl[2][4]; // used in FL layer
+};
+
+void camellia_sliced_F(uint8x16x4_t X[restrict 2],
+                       const uint8x16x4_t k[restrict 2]);
+void camellia_sliced_FL(uint8x16x4_t X[restrict 2],
+                        const uint8x16x4_t kl[restrict 2]);
+void camellia_sliced_FL_inv(uint8x16x4_t X[restrict 2],
+                            const uint8x16x4_t kl[restrict 2]);
+void camellia_sliced_feistel_round(uint8x16x4_t state[restrict 2],
+                                   const uint8x16x4_t kr[restrict 2]);
+void camellia_sliced_feistel_round_inv(uint8x16x4_t state[restrict 2],
+                                       const uint8x16x4_t kr[restrict 2]);
+void camellia_sliced_generate_round_keys_128(const uint64_t key[restrict 2],
+                                             struct camellia_rks_sliced_128 *restrict rks);
+
+void camellia_sliced_pack(uint8x16x4_t packed[restrict 2],
+                          const uint64_t x[restrict 2][8]);
+
+void camellia_sliced_unpack(uint64_t x[restrict 2][8],
+                            const uint8x16x4_t packed[restrict 2]);
+
+void camellia_sliced_init(void);
+
+void camellia_sliced_encrypt_128(uint64_t c[restrict 2][8],
+                                 const uint64_t m[restrict 2][8],
+                                 struct camellia_rks_sliced_128 *restrict rks);
+
+void camellia_sliced_decrypt_128(uint64_t m[restrict 2][8],
+                                 const uint64_t c[restrict 2][8],
+                                 struct camellia_rks_sliced_128 *restrict rks);
+
+#endif

### 2023-05-08 impl/camellia/camellia_keys.h

@@ -3,17 +3,16 @@
 
 #include <stdint.h>
 
-struct camellia_keys_128 {
+struct camellia_rks_128 {
         uint64_t kw[4]; // whitening
         uint64_t ku[18]; // used in F function
-        int64_t kl[4]; // used in FL layer
+        uint64_t kl[4]; // used in FL layer
 };
 
-struct camellia_keys_256 {
+struct camellia_rks_256 {
         uint64_t kw[4]; // whitening
         uint64_t ku[24]; // used in F function
         uint64_t kl[6]; // used in FL layer
 };
-
 
 #endif

### 2023-05-08 impl/camellia/naive.c

@@ -258,7 +258,7 @@ void camellia_naive_feistel_round_inv(uint64_t state[2], const uint64_t kr)
 }
 
 void camellia_naive_generate_round_keys_128(const uint64_t key[restrict 2],
-                                        struct camellia_keys_128 *restrict rks)
+                                        struct camellia_rks_128 *restrict rks)
 {
         uint64_t KL[2], KA[2];
         memcpy(KL, key, sizeof(KL));
@@ -315,7 +315,7 @@ void camellia_naive_generate_round_keys_128(const uint64_t key[restrict 2],
 }
 
 void camellia_naive_generate_round_keys_256(const uint64_t key[restrict 4],
-                                            struct camellia_keys_256 *restrict rks)
+                                            struct camellia_rks_256 *restrict rks)
 {
         uint64_t KL[2], KR[2], KA[2], KB[2];
         memcpy(KL, &key[0], sizeof(KL));
@@ -389,7 +389,7 @@ void camellia_naive_generate_round_keys_256(const uint64_t key[restrict 4],
 
 void camellia_naive_encrypt_128(uint64_t c[restrict 2],
                             const uint64_t m[restrict 2],
-                            struct camellia_keys_128 *restrict rks)
+                            struct camellia_rks_128 *restrict rks)
 {
         memcpy(c, m, sizeof(c[0]) * 2);
 
@@ -422,7 +422,7 @@ void camellia_naive_encrypt_128(uint64_t c[restrict 2],
 
 void camellia_naive_decrypt_128(uint64_t m[restrict 2],
                             const uint64_t c[restrict 2],
-                            struct camellia_keys_128 *restrict rks)
+                            struct camellia_rks_128 *restrict rks)
 {
         memcpy(m, c, sizeof(m[0]) * 2);
 
@@ -456,7 +456,7 @@ void camellia_naive_decrypt_128(uint64_t m[restrict 2],
 
 void camellia_naive_encrypt_256(uint64_t c[restrict 2],
                             const uint64_t m[restrict 2],
-                            struct camellia_keys_256 *restrict rks)
+                            struct camellia_rks_256 *restrict rks)
 {
         memcpy(c, m, sizeof(c[0]) * 2);
 
@@ -496,7 +496,7 @@ void camellia_naive_encrypt_256(uint64_t c[restrict 2],
 
 void camellia_naive_decrypt_256(uint64_t m[restrict 2],
                             const uint64_t c[restrict 2],
[…]

### 2023-05-08 impl/camellia/naive.h

@@ -14,24 +14,24 @@ uint64_t camellia_naive_FL_inv(uint64_t X, const uint64_t kl);
 void camellia_naive_feistel_round(uint64_t state[2], const uint64_t kr);
 void camellia_naive_feistel_round_inv(uint64_t state[2], const uint64_t kr);
 void camellia_naive_generate_round_keys_128(const uint64_t key[restrict 2],
-                                            struct camellia_keys_128 *restrict rks);
+                                            struct camellia_rks_128 *restrict rks);
 void camellia_naive_generate_round_keys_256(const uint64_t key[restrict 4],
-                                            struct camellia_keys_256 *restrict rks);
+                                            struct camellia_rks_256 *restrict rks);
 
 void camellia_naive_encrypt_128(uint64_t c[restrict 2],
                             const uint64_t m[restrict 2],
-                            struct camellia_keys_128 *restrict rks);
+                            struct camellia_rks_128 *restrict rks);
 
 void camellia_naive_decrypt_128(uint64_t m[restrict 2],
                             const uint64_t c[restrict 2],
-                            struct camellia_keys_128 *restrict rks);
+                            struct camellia_rks_128 *restrict rks);
 
 void camellia_naive_encrypt_256(uint64_t c[restrict 2],
                             const uint64_t m[restrict 2],
-                            struct camellia_keys_256 *restrict rks);
+                            struct camellia_rks_256 *restrict rks);
 
 void camellia_naive_decrypt_256(uint64_t m[restrict 2],
                             const uint64_t c[restrict 2],
-                            struct camellia_keys_256 *restrict rks);
+                            struct camellia_rks_256 *restrict rks);
 
 #endif

### 2023-05-08 impl/camellia/spec_opt.c

@@ -88,7 +88,7 @@ void camellia_spec_opt_feistel_round_inv(uint64_t state[2], const uint64_t kr)
 }
 
 void camellia_spec_opt_generate_round_keys_128(const uint64_t key[restrict 2],
-                                        struct camellia_keys_128 *restrict rks)
+                                        struct camellia_rks_128 *restrict rks)
 {
         uint64_t KL[2], KA[2];
 
@@ -106,7 +106,7 @@ void camellia_spec_opt_generate_round_keys_128(const uint64_t key[restrict 2],
         camellia_spec_opt_feistel_round(KA, keysched_const[2]);
         camellia_spec_opt_feistel_round(KA, keysched_const[3]);
 
-        struct camellia_keys_128 keys;
+        struct camellia_rks_128 keys;
 
         // KL-dependent subkeys
         keys.kw[0] = KL[0]; keys.kw[1] = KL[1];
@@ -145,7 +145,7 @@ void camellia_spec_opt_generate_round_keys_128(const uint64_t key[restrict 2],
 
 void camellia_spec_opt_encrypt_128(uint64_t c[restrict 2],
                             const uint64_t m[restrict 2],
-                            struct camellia_keys_128 *restrict rks)
+                            struct camellia_rks_128 *restrict rks)
 {
         memcpy(c, m, sizeof(c[0]) * 2);
 
@@ -181,7 +181,7 @@ void camellia_spec_opt_encrypt_128(uint64_t c[restrict 2],
 
 void camellia_spec_opt_decrypt_128(uint64_t m[restrict 2],
                             const uint64_t c[restrict 2],
-                            struct camellia_keys_128 *restrict rks)
+                            struct camellia_rks_128 *restrict rks)
 {
         memcpy(m, c, sizeof(m[0]) * 2);
 

### 2023-05-08 impl/test.c

@@ -6,6 +6,7 @@
 
 #include "camellia/naive.h"
 #include "camellia/spec_opt.h"
+#include "camellia/bytesliced.h"
 
 #include <stdio.h>
 #include <stdlib.h>
@@ -273,8 +274,8 @@ void test_camellia_naive(void)
 {
         uint64_t m[2], c[2];
         uint64_t key[2];
-        struct camellia_keys_128 rks_128;
-        struct camellia_keys_256 rks_256;
+        struct camellia_rks_128 rks_128;
+        struct camellia_rks_256 rks_256;
 
         printf("testing CAMELLIA_NAIVE FL-FL_inv and feistel-feistel_inv...\n");
         for (int i = 0; i < 256; i++) {
@@ -352,7 +353,7 @@ void test_camellia_spec_opt(void)
 {
         uint64_t m[2], c[2];
         uint64_t key[2];
-        struct camellia_keys_128 rks_128;
+        struct camellia_rks_128 rks_128;
 
         printf("testing CAMELLIA_SPEC_OPT FL-FL_inv and feistel-feistel_inv...\n");
         for (int i = 0; i < 256; i++) {
@@ -397,15 +398,47 @@ void test_camellia_spec_opt(void)
         }
 }
 
+#include <arm_neon.h>
+
+void test_camellia_sliced(void)
+{
+        camellia_sliced_init();
+
+        /* uint64_t test[2] = { */
+        /*         0x8796a5b4c3d2e1f0UL, 0x0f1e2d3c4b5a6978UL */
+        /* }; */
+
+        uint64_t test[2] = {
+                0x0706050403020100UL, 0x0f0e0d0c0b0a0908UL
+        };
+
+        uint8x16_t a = vld1q_u8((uint8_t*)&test[0]);
+        uint8x16x4_t data[2];
+        uint8x16x4_t key[2];
+        for (size_t i = 0; i < 2; i++) {
+                for (size_t j = 0; j < 4; j++) {
+                        data[i].val[j] = a;
+                        key[i].val[j] = vdupq_n_u8(0x0);
+                }
+        }
+
+
+        camellia_sliced_F(data, key);
+        printf("%lx %lx\n",
+               vgetq_lane_u64(data[0].val[0], 1),
+               vgetq_lane_u64(data[0].val[0], 0));
+}
+
 int main(int argc, char *argv[])
 {
         srand(time(NULL));
-        test_gift_64();
-        test_gift_128();
-        test_gift_64_sliced();
[…]

### 2023-05-08 impl/vector/gift_vec_sliced.c

@@ -4,17 +4,17 @@
 #include <stdio.h>
 #include <string.h>
 
-static uint64_t pack_shf_u64[4] = {
+static uint64_t pack_shf_u64[] = {
         0x1303120211011000UL, 0x1707160615051404UL, // S0/S1/S2/S3
         0x1b0b1a0a19091808UL, 0x1f0f1e0e1d0d1c0cUL, // S4/S5/S6/S7
 };
 
-static uint64_t pack_shf_inv_u64[4] = {
+static uint64_t pack_shf_inv_u64[] = {
         0x0e0c0a0806040200UL, 0x1e1c1a1816141210UL, // S0/S1/S2/S3
         0x0f0d0b0907050301UL, 0x1f1d1b1917151311UL, // S4/S5/S6/S7
 };
 
-static uint64_t perm_u64[8] = {
+static uint64_t perm_u64[] = {
         0x0f0b07030c080400UL, 0x0d0905010e0a0602UL, // S0/S4
         0x0c0804000d090501UL, 0x0e0a06020f0b0703UL, // S1/S5
         0x0d0905010e0a0602UL, 0x0f0b07030c080400UL, // S2/S6
@@ -22,7 +22,7 @@ static uint64_t perm_u64[8] = {
 };
 
 
-static uint64_t perm_inv_u64[8] = {
+static uint64_t perm_inv_u64[] = {
         0x05090d0104080c00UL, 0x070b0f03060a0e02UL, // S0/S4
         0x090d0105080c0004UL, 0x0b0f03070a0e0206UL, // S1/S5
         0x0d0105090c000408UL, 0x0f03070b0e02060aUL, // S2/S6

### 2023-05-08 scripts/filters.py

@@ -1,0 +1,15 @@
+#!/usr/bin/env python3
+
+import numpy as np
+
+psi_prev = np.array([
+    (0, 1, 0, 0, 1, 0, 1, 0),
+    (0, 1, 1, 0, 0, 1, 1, 0),
+    (0, 0, 1, 0, 0, 1, 1, 1),
+    (0, 0, 1, 0, 1, 0, 0, 0),
+    (1, 1, 0, 0, 1, 0, 1, 0),
+    (1, 1, 0, 1, 1, 0, 0, 1),
+    (1, 1, 0, 1, 1, 0, 1, 0),
+    (1, 1, 0, 1, 1, 1, 1, 1)])
+
+print(psi_prev @ psi_prev % 2)

### 2023-05-11 Paper/chapters/implementation-strategies.tex
bytesliced implementation packing

@@
 
+% TODO change shr/shl implementation to only use vector registers
 One caveat of this approach is the fact that NEON registers cannot be shifted
@@
 
-A byte-sliced implementation strategy for ARMv8 can be derived from already
-existing x86-optimized implementations utilizing the AES-NI advanced encryption
-standard instruction set \cite{bcfastimplx86:2013}. NEON itself possesses
-cryptographic extensions for finite field arithmetic and AES as well as SHA
-calculations. These can be used to produce an accelerated Camellia implementation
-due to the algebraic similarity of the AES- and Camellia S-boxes.
+\subsubsection{Packing}
+
+We encrypt 16 plaintext blocks in parallel so we can fill 16 vector registers
+with register $i$ containing all $i$-th bytes. This lends itself well to a NEON
+implementation since S-boxes as well as the FL layer, Feistel round and
+packing/unpacking functions can be implemented efficiently.
 
 \subsubsection{Hardware-accelerated Camellia S-box}
+
+A byte-sliced implementation strategy for the S-box on ARMv8 can be derived
+from already existing x86-optimized implementations utilizing the AES-NI
+advanced encryption standard instruction set \cite{bcfastimplx86:2013}. NEON
+itself possesses cryptographic extensions for finite field arithmetic and AES
+as well as SHA calculations. These can be used to produce an accelerated
+Camellia implementation due to the algebraic similarity of the AES- and
+Camellia S-boxes.
+
+The AES S-box works by multiplicatively inverting $x\in\F{8}$ over GF$(2^8)$ and then
+applying an affine transformation $A$:
+
+\[
+    s(x):\F{8}\rightarrow\F{8}, x\mapsto A(x^{-1}_{\text{GF}(2^8)})
+\]
+
+The Camellia S-box $s_0$ is defined as follows:
+
+\[
+    s_0(x):\F{8}\rightarrow\F{8}, x\mapsto h(g(f(x\oplus 0xc5)))\oplus 0x6e
+\]
+
+with affine transformations $h,f$ and the multiplicative inversion function $g$
+in the composite field GF$((2^4)^2)$. Because Galois fields of equal size are
+isomorphic, there are affine isomorphisms $\delta,\delta^{-1}$ between
+GF($2^8$) and GF($(2^4)^2$) respectively\cite{satohmorioka:2003}:
+
+\begin{alignat*}{2}
+    \delta&:\text{GF}(2^8)&&\rightarrow\text{GF}((2^4)^2) \\
+    \delta^{-1}&:\text{GF}((2^4)^2)&&\rightarrow\text{GF}(2^8)
+\end{alignat*}
+
+NEON provides the \texttt{AESE} instruction for one round of AES encryption
+which works on a 128-bit vector register. By applying the inverse of ShiftRow
+to x, we can apply the AES S-box to 16 bytes at once.
+
+\[
+    \texttt{AESE}(x)=\text{SubBytes}(\text{ShiftRow(x)})\Leftrightarrow \texttt{AESE}(\text{ShiftRow}^{-1}(x))=\text{SubBytes}(x)
+\]
+
+We can then reverse the affine transformation $A$ due to bijectivity and
+extract the multiplicative inverse of all 16 vector bytes in GF$(2^8)$.
+
+\[
+    x^{-1}_{\text{GF}(2^8)}=A^{-1}(A(x^{-1}_{\text{GF}(2^8)}))=A^{-1}(\text{SubBytes}(x))=A^{-1}(\texttt{AESE}(\text{ShiftRow}^{-1}(x)))
+\]
+
+Using the affine isomorphism, we transform this inverse into the inverse in
+GF$((2^4)^2)$ which is equal to $g(x)$:
+
+\[
+    g(x)=x^{-1}_{\text{GF}((2^4)^2)}=\delta(x^{-1}_{\text{GF}(2^8)})=\delta(A^{-1}(\texttt{AESE}(\text{ShiftRow}^{-1}(x))))
+\]
+
+We now redefine $h,f$ as $H,F$ such that they include addition of constants
+$0xc5$ and $0x6e$. We can now use the inverse in conjunction with $H$ and $F$
+and an additional input transformation to calculate the Camellia S-box:
+
+\begin{align*}
+    s_0(x)&=h(g(f(x\oplus 0xc5)))\oplus 0x6e \\
+          &=H(g(F(x))) \\
+          &=H(\delta(A^{-1}(\texttt{AESE}(\text{ShiftRow}^{-1}(\delta^{-1}(F(x)))))))
+\end{align*}
+
+Because of different bit endianness of the matrices representing
+$\delta,\delta^{-1}$, we define an additional bit-swapping function:
+
+\[
+    S=S^{-1}:\begin{pmatrix}x_0\\x_1\\x_2\\x_3\\x_4\\x_5\\x_6\\x_7\end{pmatrix}\mapsto
+    \begin{pmatrix}
+        0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
+        0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
+        0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
+        0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
+        0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
+        0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
+        0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+        1 & 0 & 0 & 0 & 0 & 0 & 0 & 0
+    \end{pmatrix}
+    \begin{pmatrix}x_0\\x_1\\x_2\\x_3\\x_4\\x_5\\x_6\\x_7\end{pmatrix}
+\]
+
+The final equation serving as the basis for our implementation then becomes
+
+\[
+    s_0(x)=S^{-1}(H(\delta(S(A^{-1}(\texttt{AESE}(\text{ShiftRow}^{-1}(S^{-1}(\delta^{-1}(F(S(x))))))))))
+\]
+
+While appearing to be computationally too intensive for any performance gains,
+we shall notice that all of the transformations are affine and can therefore be
+combined into a single operation by simple matrix multiplication. We combine
+them into a pre-filter function $\theta_0$ and a post-filter function $\psi_0$:
+
+\begin{align*}
+    \theta_0(x)&=S^{-1}(\delta^{-1}(F(S(x)))) \\
+    \psi_0(x)&=S^{-1}(H(\delta(S(A^{-1}(x)))))
+\end{align*}
+
+This simplifies our final equation:
+
+\[
+    s_0(x)=\psi_0(\texttt{AESE}(\text{ShiftRow}^{-1}(\theta_0(x))))
+\]
+
+The other S-boxes $s_1,s_2,s_3$ are defined in terms of $s_0$ and rotations. We
+can include these rotations by constructing additional S-box-specific filters
+$\theta_3,\psi_1,\psi_2$:
+
+\begin{alignat*}{2}
+    s_1(x)&=s_0(x)\lll_1&&=\psi_1(\texttt{AESE}(\text{ShiftRow}^{-1}(\theta_0(x)))) \\
+    s_2(x)&=s_0(x)\ggg_1&&=\psi_2(\texttt{AESE}(\text{ShiftRow}^{-1}(\theta_0(x)))) \\
+    s_3(x)&=s_0(x\lll_1)&&=\psi_0(\texttt{AESE}(\text{ShiftRow}^{-1}(\theta_3(x))))
+\end{alignat*}
+
+\subsubsection{Fast matrix multiplication}
+
+Efficient application of the filter functions is essential to performance and
+can be implemented in parallel by use of the \texttt{TBL} instruction. First
+note that matrix-vector multiplication $\F{8\times 8}\cdot\F{8}\rightarrow
+\F{8}$ can be decomposed into two multiplications $\F{8\times
+4}\cdot\F{4}\rightarrow \F{8}$ with subsequent addition of the results:
+
+\[
+    \begin{pmatrix}
+        1 & 1 & 1 & 0 & 1 & 1 & 0 & 1 \\
+        0 & 0 & 1 & 1 & 0 & 0 & 1 & 0 \\
+        1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 \\
+        1 & 0 & 1 & 1 & 0 & 0 & 1 & 1 \\
+        0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 \\
+        1 & 0 & 1 & 0 & 0 & 1 & 0 & 0 \\
+        0 & 0 & 1 & 0 & 1 & 1 & 0 & 0 \\
+        1 & 0 & 0 & 0 & 0 & 1 & 1 & 0 \\
+    \end{pmatrix}\cdot
+    \begin{pmatrix}x_0\\x_1\\x_2\\x_3\\x_4\\x_5\\x_6\\x_7\end{pmatrix}=
+    \begin{pmatrix}
+        1 & 1 & 0 & 1 \\
+        0 & 0 & 1 & 0 \\
+        0 & 0 & 0 & 0 \\
+        0 & 0 & 1 & 1 \\
+        1 & 1 & 0 & 0 \\
+        0 & 1 & 0 & 0 \\
+        1 & 1 & 0 & 0 \\
+        0 & 1 & 1 & 0 \\
+    \end{pmatrix}\cdot
+    \begin{pmatrix}x_0\\x_1\\x_2\\x_3\end{pmatrix}\oplus
+    \begin{pmatrix}
+        1 & 1 & 1 & 0 \\
+        0 & 0 & 1 & 1 \\
+        1 & 1 & 0 & 1 \\
+        1 & 0 & 1 & 1 \\
+        0 & 0 & 0 & 0 \\
+        1 & 0 & 1 & 0 \\
+        0 & 0 & 1 & 0 \\
+        1 & 0 & 0 & 0 \\
+    \end{pmatrix}\cdot
+    \begin{pmatrix}x_4\\x_5\\x_6\\x_7\end{pmatrix}
+\]
+
+We can then tabulate the decomposed multiplications since each input now
+consists of only four bits which allows us to fit the 16 possible 8-bit results
+into a single vector register. A single matrix multiplication is then executed
+for 16 bytes in parallel by use of two \texttt{TBL} and one XOR instruction with
+an additional AND and SHR operation for masking off the lower/upper 4 bits.

### 2023-05-11 impl/benchmark.c

@@ -232,15 +232,27 @@ static void benchmark_camellia_naive(void)
         uint64_t m[2], c[2];
         struct camellia_rks_128 rks;
 
-        uint64_t cycles[2] = { 0UL };
+        uint64_t cycles[8] = { 0UL };
         for (int i = 0; i < NL; i++) {
                 cycles[0] += TIME(camellia_naive_generate_round_keys_128(key, &rks));
-                cycles[1] += TIME(camellia_naive_encrypt_128(c, m, &rks));
+                cycles[1] += TIME(camellia_naive_S(m[0]));
+                cycles[2] += TIME(camellia_naive_P(m[1]));
+                cycles[3] += TIME(camellia_naive_F(m[0], key[0]));
+                cycles[4] += TIME(camellia_naive_FL(m[0], key[0]));
+                cycles[5] += TIME(camellia_naive_FL_inv(m[1], key[1]));
+                cycles[6] += TIME(camellia_naive_feistel_round(m, key[1]));
+                cycles[7] += TIME(camellia_naive_encrypt_128(c, m, &rks));
         }
 
-        printf("%f %f\n",
+        printf("%f %f %f %f %f %f %f %f\n",
                cycles[0] / (float)NL / 16.0f,
-               cycles[1] / (float)NL / 16.0f);
+               cycles[1] / (float)NL / 16.0f,
+               cycles[2] / (float)NL / 16.0f,
+               cycles[3] / (float)NL / 16.0f,
+               cycles[4] / (float)NL / 16.0f,
+               cycles[5] / (float)NL / 16.0f,
+               cycles[6] / (float)NL / 16.0f,
+               cycles[7] / (float)NL / 16.0f);
 
         struct timeval st, et;
         gettimeofday(&st, NULL);
@@ -284,15 +296,23 @@ static void benchmark_camellia_spec_opt(void)
         uint64_t m[2], c[2];
         struct camellia_rks_128 rks;
 
-        uint64_t cycles[2] = { 0UL };
+        uint64_t cycles[6] = { 0UL };
         for (int i = 0; i < NL; i++) {
                 cycles[0] += TIME(camellia_spec_opt_generate_round_keys_128(key, &rks));
-                cycles[1] += TIME(camellia_spec_opt_encrypt_128(c, m, &rks));
+                cycles[1] += TIME(camellia_spec_opt_F(m[0], key[0]));
[…]

### 2023-05-11 impl/camellia/bytesliced.c

@@ -14,54 +14,17 @@ static const uint64_t keysched_const[] = {
         0xb05688c2b3e6c1fdUL
 };
 
-static uint64_t rol1_tbl_u64[] = {
-        0x060504030201001fUL, 0x0e0d0c0b0a090807UL
-};
 
-static uint64_t rol8_tbl_u64[] = {
-        0x1f1e1d1c1b1a1918UL, 0x0706050403020100UL
-};
+static uint8x16x4_t pack_group;
+static uint8x16x4_t pack_single;
 
-static uint64_t shiftrows_inv_u64[2] = {
-        0x0b0e0104070a0d00UL, 0x0306090c0f020508UL
-};
-
-static uint64_t prefilter_0_u64[][2] = {
-        { 0x862b832eed40e845UL, 0x88258d20e34ee64bUL },
-        { 0x2a7bdb8aa0f15100UL, 0x2372d283a9f85809UL }
-};
-
-static uint64_t prefilter_3_u64[][2] = {
-        { 0x25204e4b2b2e4045UL, 0x74711f1a7a7f1114UL },
-        { 0x7283f8097b8af100UL, 0xdf2e55a4d6275cadUL }
-};
-
-static uint64_t postfilter_0_u64[][2] = {
-        { 0x31c1c2323fcfcc3cUL, 0xd12122d2df2f2cdcUL },
-        { 0xa8512ed77f86f900UL, 0x0cf58a73db225da4UL }
-};
-
-static uint64_t postfilter_1_u64[][2] = {
-        { 0x628385647e9f9978UL, 0xa34244a5bf5e58b9UL },
-        { 0x51a25caffe0df300UL, 0x18eb15e6b744ba49UL }
-};
-
-static uint64_t postfilter_2_u64[][2] = {
-        { 0x98e061199fe7661eUL, 0xe8901169ef97166eUL },
-        { 0x54a817ebbf43fc00UL, 0x06fa45b9ed11ae52UL }
-};
-
-static uint8x16_t rol1_tbl;
-static uint8x16_t rol8_tbl;
 static uint8x16_t lower_4_bits_mask;
 static uint8x16_t shiftrows_inv;
 
 // store two for low ([0]) and high ([1]) bits of matrix multiplication
 static uint8x16x2_t prefilter_0;
-static uint8x16_t   prefilter_0_xor;
 static uint8x16x2_t prefilter_3;  // s3(x) = s0(x <<< 1)
[…]

### 2023-05-11 impl/camellia/bytesliced.h

@@ -1,44 +1,44 @@
 #ifndef BYTESLICED_H
 #define BYTESLICED_H
 
-// 128-bit bitsliced camellia
+// 128-bit bitsliced camellia with 16 blocks encrypted in parallel
 
 #include <stdint.h>
 #include <arm_neon.h>
 
 struct camellia_rks_sliced_128 {
-        uint8x16x4_t kw[2][4]; // whitening
-        uint8x16x4_t ku[2][18]; // used in F function
-        uint8x16x4_t kl[2][4]; // used in FL layer
+        uint8x16x4_t kw[4][2]; // whitening
+        uint8x16x4_t ku[18][2]; // used in F function
+        uint8x16x4_t kl[4][2]; // used in FL layer
 };
 
 void camellia_sliced_F(uint8x16x4_t X[restrict 2],
                        const uint8x16x4_t k[restrict 2]);
 void camellia_sliced_FL(uint8x16x4_t X[restrict 2],
                         const uint8x16x4_t kl[restrict 2]);
-void camellia_sliced_FL_inv(uint8x16x4_t X[restrict 2],
+void camellia_sliced_FL_inv(uint8x16x4_t Y[restrict 2],
                             const uint8x16x4_t kl[restrict 2]);
-void camellia_sliced_feistel_round(uint8x16x4_t state[restrict 2],
+void camellia_sliced_feistel_round(uint8x16x4_t state[restrict 4],
                                    const uint8x16x4_t kr[restrict 2]);
-void camellia_sliced_feistel_round_inv(uint8x16x4_t state[restrict 2],
+void camellia_sliced_feistel_round_inv(uint8x16x4_t state[restrict 4],
                                        const uint8x16x4_t kr[restrict 2]);
 void camellia_sliced_generate_round_keys_128(const uint64_t key[restrict 2],
                                              struct camellia_rks_sliced_128 *restrict rks);
 
-void camellia_sliced_pack(uint8x16x4_t packed[restrict 2],
-                          const uint64_t x[restrict 2][8]);
+void camellia_sliced_pack(uint8x16x4_t packed[restrict 4],
+                          const uint64_t x[restrict 16][2]);
 
-void camellia_sliced_unpack(uint64_t x[restrict 2][8],
-                            const uint8x16x4_t packed[restrict 2]);
+void camellia_sliced_unpack(uint64_t x[restrict 16][2],
+                            const uint8x16x4_t packed[restrict 4]);
 
 void camellia_sliced_init(void);
 
[…]

### 2023-05-11 impl/camellia/naive.c

@@ -168,8 +168,10 @@ static const uint8_t s4[256] =
         (_a) = ((_a) << (_n)) | ((_t) >> (64 - (_n)));\
 }
 
-#define expr_rol32_1(_a)\
-        (((_a) << 1) | ((_a) >> 31))
+static uint32_t rol32_1(uint32_t a)
+{
+        return (a << 1) | (a >> 31);
+}
 
 uint64_t camellia_naive_S(uint64_t X)
 {
@@ -223,7 +225,7 @@ uint64_t camellia_naive_FL(uint64_t X, const uint64_t kl)
         const uint32_t klL = (kl >> 32);
         const uint32_t klR = (kl >> 0);
 
-        const uint32_t YR = expr_rol32_1(XL & klL) ^ XR;
+        const uint32_t YR = rol32_1(XL & klL) ^ XR;
         const uint32_t YL = (YR | klR) ^ XL;
 
         return ((uint64_t)YL << 32) | (uint64_t)YR;
@@ -238,7 +240,7 @@ uint64_t camellia_naive_FL_inv(uint64_t Y, const uint64_t kl)
         const uint32_t klR = (kl >> 0);
 
         const uint32_t XL = (YR | klR) ^ YL;
-        const uint32_t XR = expr_rol32_1(XL & klL) ^ YR;
+        const uint32_t XR = rol32_1(XL & klL) ^ YR;
 
         return ((uint64_t)XL << 32) | (uint64_t)XR;
 }

### 2023-05-11 impl/test.c

@@ -398,19 +398,29 @@ void test_camellia_spec_opt(void)
         }
 }
 
-#include <arm_neon.h>
-
 void test_camellia_sliced(void)
 {
         camellia_sliced_init();
 
-        /* uint64_t test[2] = { */
-        /*         0x8796a5b4c3d2e1f0UL, 0x0f1e2d3c4b5a6978UL */
-        /* }; */
+        uint64_t test[16][2];
+        /* for (size_t i = 0; i < 16; i++) { */
+        /*         test[i][0] = 0x8796a5b4c3d2e1f0UL; */
+        /*         test[i][1] = 0x0f1e2d3c4b5a6978UL; */
+        /* } */
 
-        uint64_t test[2] = {
-                0x0706050403020100UL, 0x0f0e0d0c0b0a0908UL
-        };
+        for (size_t i = 0; i < 16; i++) {
+                test[i][0] = 0x0706050403020100UL;
+                test[i][1] = 0x0f0e0d0c0b0a0908UL;
+        }
+
+        uint8x16x4_t packed[4];
+        camellia_sliced_pack(packed, test);
+        for (size_t i = 0; i < 4; i++) {
+                for (size_t j = 0; j < 4; j++) {
+                        printf("%lx %lx\n", vgetq_lane_u64(packed[i].val[j], 1),
+                              vgetq_lane_u64(packed[i].val[j], 0));
+                }
+        }
 
         uint8x16_t a = vld1q_u8((uint8_t*)&test[0]);
         uint8x16x4_t data[2];
@@ -423,10 +433,10 @@ void test_camellia_sliced(void)
         }
 
 
-        camellia_sliced_F(data, key);
-        printf("%lx %lx\n",
-               vgetq_lane_u64(data[0].val[0], 1),
-               vgetq_lane_u64(data[0].val[0], 0));
+        /* camellia_sliced_F(data, key); */
+        /* printf("%lx %lx\n", */
+        /*        vgetq_lane_u64(data[0].val[0], 1), */
+        /*        vgetq_lane_u64(data[0].val[0], 0)); */
 }
 
 int main(int argc, char *argv[])

### 2023-05-16 Paper/chapters/implementation-strategies.tex
camellia_128_sliced working

@@
 thus need to implement the \texttt{shr(V,n)} and \texttt{shl(V,n)} operations
-on our own. This can be done by first extracting the 64-bit lanes $a,b$ out of
-$V=b||a$, shifting the lanes individually and finally shifting and ORing the
-crossing bits back into the other lane.
+on our own. This can be done by extracting and shifting the overflow $ov$ out
+of $V=b||a$, shifting the lanes individually and finally ORing the overflow
+bits to the corresponding vector element.
 
@@
     &\text{shl}(V,n): \\
-    &\qquad a,b&&=V[0],V[1] \\
-    &\qquad c&&=(a\gg (64-n)) \\
-    &\qquad V[0]&&=(a\ll n) \\
-    &\qquad V[1]&&=(b\ll n)\lor c
+    &\qquad ov&&=V[0]\land(\lnot(\texttt{0xffffffffffffffff}\gg_n))\gg_{63-n} \\
+    &\qquad V[0]&&=V[0]\gg_n \\
+    &\qquad V[1]&&=(V[1]\gg_n)\lor ov
 \end{alignat*}
@@
     &\text{SWAPMOVE}(V_0,V_4,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_1,V_5,0x0f0f\dots 0f,4) \\
-    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4) \\
+    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4)
 \end{align*}
@@
 
-\subsubsection{Packing}
+Because Camellia is a byte-oriented block cipher, we can pursue a similar
+strategy as for GIFT: find a convenient data packing format and a way to apply
+necessary operations in parallel.
 
-We encrypt 16 plaintext blocks in parallel so we can fill 16 vector registers
-with register $i$ containing all $i$-th bytes. This lends itself well to a NEON
-implementation since S-boxes as well as the FL layer, Feistel round and
-packing/unpacking functions can be implemented efficiently.
+We choose to encrypt 16 plaintext blocks at once so we can fill 16 vector
+registers with register $V_i$ containing all $i$-th bytes. This lends itself well
+to a NEON implementation since S-boxes as well as the FL layer, Feistel round
+and packing/unpacking functions can be implemented efficiently.
+
+\subsubsection{Packing and unpacking}
+
+Packing and unpacking of data can be done efficiently by use of 32 \texttt{TBL}
+instructions and is summarized in Figure \ref{fig:byteslicing}. After
+packing, every $i$-th register will contain all $i$-th bytes of the 16 input
+blocks. Unpacking is done in a similar way with different permutation tables.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/byteslicing.pdf}
+    \caption{Camellia byte-sliced packing function with 32 \texttt{TBL} operations.
+    Curly bracets represent a 4-element vector array being used as source registers.}
+    \label{fig:byteslicing}
+\end{figure}
 
@@
 in the composite field GF$((2^4)^2)$. Because Galois fields of equal size are
-isomorphic, there are affine isomorphisms $\delta,\delta^{-1}$ between
+isomorphic, there exist affine isomorphisms $\delta,\delta^{-1}$ between
 GF($2^8$) and GF($(2^4)^2$) respectively\cite{satohmorioka:2003}:
@@
 an additional AND and SHR operation for masking off the lower/upper 4 bits.
+
+\subsubsection{Permutation}
+
+Since all bytes of the same position are collected in the same register,
+applying the permutation can be realized in 16 XOR operations as described in
+the Camellia specification with the difference that we choose to not swap the
+lower and higher 4 bytes, but rather compensate for this by choosing the
+correct bytes when XORing with the other half later in the Feistel round.
+
+% TODO add graphic here
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/camelliaround.pdf}
+    \caption{Camellia byte-sliced Feistel round function}
+\end{figure}
+
+\subsubsection{FL layer}
+
+The two functions $FL,FL^{-1}$ can be implemented in a straightforward manner
+once 1-bit left rotation has been defined for the bytesliced representation.
+This can be facilitated by extracting the highest bit of each byte, shifting it
+all the way to the right and storing it as overflow $o_i$. $o_i$ is then added
+to the left-shifted value of byte $(i + 4) (\bmod 4)$ for the final result.
+
+% TODO add graphic here

### 2023-05-16 impl/benchmark.c

@@ -7,6 +7,7 @@
 #include "vector/gift_vec_sliced.h"
 #include "camellia/naive.h"
 #include "camellia/spec_opt.h"
+#include "camellia/bytesliced.h"
 
 #include <stdio.h>
 #include <stdlib.h>
@@ -66,10 +67,10 @@ static void benchmark_gift_64(void)
         }
 
         printf("%f %f %f %f\n",
-               cycles[0] / (float)NL / 8.0f,
-               cycles[1] / (float)NL / 8.0f,
-               cycles[2] / (float)NL / 8.0f,
-               cycles[3] / (float)NL / 8.0f);
+               cycles[0] / ((float)NL * 8.0f),
+               cycles[1] / ((float)NL * 8.0f),
+               cycles[2] / ((float)NL * 8.0f),
+               cycles[3] / ((float)NL * 8.0f));
 
         struct timeval st, et;
         gettimeofday(&st, NULL);
@@ -99,10 +100,10 @@ static void benchmark_gift_128(void)
         }
 
         printf("%f %f %f %f\n",
-               cycles[0] / (float)NL / 16.0f,
-               cycles[1] / (float)NL / 16.0f,
-               cycles[2] / (float)NL / 16.0f,
-               cycles[3] / (float)NL / 16.0f);
+               cycles[0] / ((float)NL * 16.0f),
+               cycles[1] / ((float)NL * 16.0f),
+               cycles[2] / ((float)NL * 16.0f),
+               cycles[3] / ((float)NL * 16.0f));
 
         struct timeval st, et;
         gettimeofday(&st, NULL);
@@ -131,9 +132,9 @@ static void benchmark_gift_64_table(void)
         }
 
         printf("%f %f %f\n",
-               cycles[0] / (float)NL / 8.0f,
-               cycles[1] / (float)NL / 8.0f,
-               cycles[2] / (float)NL / 8.0f);
+               cycles[0] / ((float)NL * 8.0f),
+               cycles[1] / ((float)NL * 8.0f),
+               cycles[2] / ((float)NL * 8.0f));
 
         struct timeval st, et;
         gettimeofday(&st, NULL);
@@ -167,12 +168,12 @@ static void benchmark_gift_64_vec_sbox(void)
         }
 
         printf("%f %f %f %f %f %f\n",
-               cycles[0] / (float)NL / 8.0f,
-               cycles[1] / (float)NL / 8.0f,
-               cycles[2] / (float)NL / 8.0f,
[…]

### 2023-05-16 impl/camellia/bytesliced.h

@@ -3,14 +3,12 @@
 
 // 128-bit bitsliced camellia with 16 blocks encrypted in parallel
 
+#include "camellia_keys.h"
+
 #include <stdint.h>
 #include <arm_neon.h>
 
-struct camellia_rks_sliced_128 {
-        uint8x16x4_t kw[4][2]; // whitening
-        uint8x16x4_t ku[18][2]; // used in F function
-        uint8x16x4_t kl[4][2]; // used in FL layer
-};
+void rol32_1(uint8x16x4_t *a);
 
 void camellia_sliced_F(uint8x16x4_t X[restrict 2],
                        const uint8x16x4_t k[restrict 2]);
@@ -22,8 +20,8 @@ void camellia_sliced_feistel_round(uint8x16x4_t state[restrict 4],
                                    const uint8x16x4_t kr[restrict 2]);
 void camellia_sliced_feistel_round_inv(uint8x16x4_t state[restrict 4],
                                        const uint8x16x4_t kr[restrict 2]);
-void camellia_sliced_generate_round_keys_128(const uint64_t key[restrict 2],
-                                             struct camellia_rks_sliced_128 *restrict rks);
+void camellia_sliced_generate_round_keys_128(struct camellia_rks_sliced_128 *restrict rks,
+                                             const uint64_t key[2]);
 
 void camellia_sliced_pack(uint8x16x4_t packed[restrict 4],
                           const uint64_t x[restrict 16][2]);

### 2023-05-16 impl/camellia/camellia_keys.h

@@ -2,6 +2,7 @@
 #define CAMELLIA_KEYS
 
 #include <stdint.h>
+#include <arm_neon.h>
 
 struct camellia_rks_128 {
         uint64_t kw[4]; // whitening
@@ -13,6 +14,12 @@ struct camellia_rks_256 {
         uint64_t kw[4]; // whitening
         uint64_t ku[24]; // used in F function
         uint64_t kl[6]; // used in FL layer
+};
+
+struct camellia_rks_sliced_128 {
+        uint8x16x4_t kw[4][2]; // whitening
+        uint8x16x4_t ku[18][2]; // used in F function
+        uint8x16x4_t kl[4][2]; // used in FL layer
 };
 
 #endif

### 2023-05-16 impl/camellia/naive.c

@@ -259,8 +259,8 @@ void camellia_naive_feistel_round_inv(uint64_t state[2], const uint64_t kr)
         state[1] = Rr1;
 }
 
-void camellia_naive_generate_round_keys_128(const uint64_t key[restrict 2],
-                                        struct camellia_rks_128 *restrict rks)
+void camellia_naive_generate_round_keys_128(struct camellia_rks_128 *restrict rks,
+                                            const uint64_t key[restrict 2])
 {
         uint64_t KL[2], KA[2];
         memcpy(KL, key, sizeof(KL));
@@ -316,8 +316,8 @@ void camellia_naive_generate_round_keys_128(const uint64_t key[restrict 2],
                 rks->kw[2] = KA[0]; rks->kw[3] = KA[1];
 }
 
-void camellia_naive_generate_round_keys_256(const uint64_t key[restrict 4],
-                                            struct camellia_rks_256 *restrict rks)
+void camellia_naive_generate_round_keys_256(struct camellia_rks_256 *restrict rks,
+                                            const uint64_t key[restrict 4])
 {
         uint64_t KL[2], KR[2], KA[2], KB[2];
         memcpy(KL, &key[0], sizeof(KL));
@@ -390,8 +390,8 @@ void camellia_naive_generate_round_keys_256(const uint64_t key[restrict 4],
 }
 
 void camellia_naive_encrypt_128(uint64_t c[restrict 2],
-                            const uint64_t m[restrict 2],
-                            struct camellia_rks_128 *restrict rks)
+                                const uint64_t m[restrict 2],
+                                struct camellia_rks_128 *restrict rks)
 {
         memcpy(c, m, sizeof(c[0]) * 2);
 
@@ -411,6 +411,7 @@ void camellia_naive_encrypt_128(uint64_t c[restrict 2],
         c[0] = camellia_naive_FL(c[0], rks->kl[2]);
         c[1] = camellia_naive_FL_inv(c[1], rks->kl[3]);
 
+
         for (int i = 0; i < 6; i++) {
                 camellia_naive_feistel_round(c, rks->ku[i + 12]);
         }
@@ -423,8 +424,8 @@ void camellia_naive_encrypt_128(uint64_t c[restrict 2],
 }
 
 void camellia_naive_decrypt_128(uint64_t m[restrict 2],
-                            const uint64_t c[restrict 2],
-                            struct camellia_rks_128 *restrict rks)
+                                const uint64_t c[restrict 2],
[…]

### 2023-05-16 impl/camellia/naive.h

@@ -13,25 +13,26 @@ uint64_t camellia_naive_FL(uint64_t X, const uint64_t kl);
 uint64_t camellia_naive_FL_inv(uint64_t X, const uint64_t kl);
 void camellia_naive_feistel_round(uint64_t state[2], const uint64_t kr);
 void camellia_naive_feistel_round_inv(uint64_t state[2], const uint64_t kr);
-void camellia_naive_generate_round_keys_128(const uint64_t key[restrict 2],
-                                            struct camellia_rks_128 *restrict rks);
-void camellia_naive_generate_round_keys_256(const uint64_t key[restrict 4],
-                                            struct camellia_rks_256 *restrict rks);
+void camellia_naive_generate_round_keys_128(struct camellia_rks_128 *restrict rks,
+                                            const uint64_t key[restrict 2]);
+
+void camellia_naive_generate_round_keys_256(struct camellia_rks_256 *restrict rks,
+                                            const uint64_t key[restrict 4]);
 
 void camellia_naive_encrypt_128(uint64_t c[restrict 2],
-                            const uint64_t m[restrict 2],
-                            struct camellia_rks_128 *restrict rks);
+                                const uint64_t m[restrict 2],
+                                struct camellia_rks_128 *restrict rks);
 
 void camellia_naive_decrypt_128(uint64_t m[restrict 2],
-                            const uint64_t c[restrict 2],
-                            struct camellia_rks_128 *restrict rks);
+                                const uint64_t c[restrict 2],
+                                struct camellia_rks_128 *restrict rks);
 
 void camellia_naive_encrypt_256(uint64_t c[restrict 2],
-                            const uint64_t m[restrict 2],
-                            struct camellia_rks_256 *restrict rks);
+                                const uint64_t m[restrict 2],
+                                struct camellia_rks_256 *restrict rks);
 
 void camellia_naive_decrypt_256(uint64_t m[restrict 2],
-                            const uint64_t c[restrict 2],
-                            struct camellia_rks_256 *restrict rks);
+                                const uint64_t c[restrict 2],
+                                struct camellia_rks_256 *restrict rks);
 
 #endif

### 2023-05-16 impl/camellia/spec_opt.c

@@ -87,8 +87,8 @@ void camellia_spec_opt_feistel_round_inv(uint64_t state[2], const uint64_t kr)
         state[1] = Rr;
 }
 
-void camellia_spec_opt_generate_round_keys_128(const uint64_t key[restrict 2],
-                                        struct camellia_rks_128 *restrict rks)
+void camellia_spec_opt_generate_round_keys_128(struct camellia_rks_128 *restrict rks,
+                                               const uint64_t key[restrict 2])
 {
         uint64_t KL[2], KA[2];
 
@@ -98,8 +98,6 @@ void camellia_spec_opt_generate_round_keys_128(const uint64_t key[restrict 2],
         // compute KA
         memcpy(KA, KL, sizeof(KA));
 
-        /* camellia_spec_opt_feistel_round(KA, keysched_const[0]); */
-        /* camellia_spec_opt_feistel_round(KA, keysched_const[1]); */
         KA[0] = camellia_spec_opt_F(KL[1] ^ camellia_spec_opt_F(KL[0], keysched_const[0]),
                                     keysched_const[1]);
         KA[1] = camellia_spec_opt_F(KL[0], keysched_const[0]);
@@ -152,21 +150,21 @@ void camellia_spec_opt_encrypt_128(uint64_t c[restrict 2],
         c[0] ^= rks->kw[0];
         c[1] ^= rks->kw[1];
 
-        for (int i = 0; i < 6; i++) {
+        for (size_t i = 0; i < 6; i++) {
                 camellia_spec_opt_feistel_round(c, rks->ku[i + 0]);
         }
 
         c[0] = camellia_spec_opt_FL(c[0], rks->kl[0]);
         c[1] = camellia_spec_opt_FL_inv(c[1], rks->kl[1]);
 
-        for (int i = 0; i < 6; i++) {
+        for (size_t i = 0; i < 6; i++) {
                 camellia_spec_opt_feistel_round(c, rks->ku[i + 6]);
         }
 
         c[0] = camellia_spec_opt_FL(c[0], rks->kl[2]);
         c[1] = camellia_spec_opt_FL_inv(c[1], rks->kl[3]);
 
-        for (int i = 0; i < 6; i++) {
+        for (size_t i = 0; i < 6; i++) {
                 camellia_spec_opt_feistel_round(c, rks->ku[i + 12]);
         }
 
[…]

### 2023-05-16 impl/camellia/spec_opt.h

@@ -11,8 +11,8 @@ uint64_t camellia_spec_opt_FL(uint64_t X, const uint64_t kl);
 uint64_t camellia_spec_opt_FL_inv(uint64_t X, const uint64_t kl);
 void camellia_spec_opt_feistel_round(uint64_t state[2], const uint64_t kr);
 void camellia_spec_opt_feistel_round_inv(uint64_t state[2], const uint64_t kr);
-void camellia_spec_opt_generate_round_keys_128(const uint64_t key[restrict 2],
-                                            struct camellia_rks_128 *restrict rks);
+void camellia_spec_opt_generate_round_keys_128(struct camellia_rks_128 *restrict rks,
+                                            const uint64_t key[restrict 2]);
 
 void camellia_spec_opt_encrypt_128(uint64_t c[restrict 2],
                             const uint64_t m[restrict 2],

### 2023-05-16 impl/test.c

@@ -1,3 +1,5 @@
+#pragma clang optimize off
+
 #include "naive/gift.h"
 #include "naive/gift_sliced.h"
 #include "table/gift_table.h"
@@ -15,7 +17,7 @@
 #include <string.h>
 
 // no testing framework necessary for this small project
-#define ASSERT_EQUALS(x,y)\
+#define ASSERT_EQUALS(x, y)\
         if (x != y) {\
                 fprintf(stderr, "ASSERT_EQUALS failed on line %d: %lx != %lx\n", __LINE__, x, y);\
                 exit(-1);\
@@ -301,7 +303,7 @@ void test_camellia_naive(void)
                 0x6767313854966973, 0x0857065648eabe43
         };
 
-        camellia_naive_generate_round_keys_128(key, &rks_128);
+        camellia_naive_generate_round_keys_128(&rks_128, key);
         camellia_naive_encrypt_128(c, m, &rks_128);
         ASSERT_EQUALS(c[0], c_expected[0]);
         ASSERT_EQUALS(c[1], c_expected[1]);
@@ -310,7 +312,7 @@ void test_camellia_naive(void)
         for (int i = 0; i < 100; i++) {
                 m_rand((uint8_t*)m, sizeof(m));
                 key_rand(key);
-                camellia_naive_generate_round_keys_128(key, &rks_128);
+                camellia_naive_generate_round_keys_128(&rks_128, key);
 
                 uint64_t m_decr[2];
                 camellia_naive_encrypt_128(c, m, &rks_128);
@@ -330,7 +332,7 @@ void test_camellia_naive(void)
         c_expected[0] = 0x9acc237dff16d76cUL;
         c_expected[1] = 0x20ef7c919e3a7509UL;
 
-        camellia_naive_generate_round_keys_256(key_256, &rks_256);
+        camellia_naive_generate_round_keys_256(&rks_256, key_256);
         camellia_naive_encrypt_256(c, m, &rks_256);
         ASSERT_EQUALS(c[0], c_expected[0]);
         ASSERT_EQUALS(c[1], c_expected[1]);
@@ -339,7 +341,7 @@ void test_camellia_naive(void)
         for (int i = 0; i < 100; i++) {
                 m_rand((uint8_t*)m, sizeof(m));
                 m_rand((uint8_t*)key_256, sizeof(key_256));
-                camellia_naive_generate_round_keys_256(key, &rks_256);
[…]

### 2023-05-16 impl/vector/gift_vec_sliced.c

@@ -47,51 +47,56 @@ static const int round_const[] = {
         0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A, 0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
 };
 
-uint8x16_t shl(const uint8x16_t v, const int n)
-{
-        uint64_t l[2];
-        vst1q_u64(l, v);
-        l[1] = (l[1] << n) | (l[0] >> (64 - n));
-        l[0] <<= n;
-        return vld1q_u64(l);
+/*
+uint8x16_t shl(const uint8x16_t v, const int n) */
+#define shl(_a, v, n)                                                       \
+{                                                                           \
+        uint64x2_t _overflow = vshrq_n_u64(v, 64 - n);                        \
+        _overflow = vextq_u64(vdupq_n_u64(0x0), _overflow, 1);                \
+        _a = vorrq_u8(v, _overflow);                                         \
 }
 
-uint8x16_t shr(const uint8x16_t v, const int n)
-{
-        uint64_t l[2];
-        vst1q_u64(l, v);
-        l[0] = l[0] >> n | (l[1] << (64 - n));
-        l[1] >>= n;
-        return vld1q_u64(l);
+/*
+uint8x16_t shr(const uint8x16_t v, const int n) */
+#define shr(_a, v, n)                                                       \
+{                                                                           \
+        uint64x2_t _overflow = vshrq_n_u64(vshlq_n_u64(v, 64 - n), 64 - n);    \
+        _overflow = vextq_u64(_overflow, vdupq_n_u64(0x0), 1);                \
+        _a = vorrq_u8(v, _overflow);                                         \
 }
 
+/* implemented as a macro so we can use vshlq_n_u8 with variable n
 void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
-                                 const uint8x16_t m, const int n)
-{
-
-        const uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
-        *b = veorq_u8(*b, t);
-        *a = veorq_u8(*a, shl(t, n));
[…]

### 2023-05-16 scripts/camellia_perm.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import numpy as np
+
+z1 = np.array([1, 0, 1, 1, 0, 1, 1, 1])
+z2 = np.array([1, 1, 0, 1, 1, 0, 1, 1]) ^ z1
+z3 = np.array([1, 1, 1, 0, 1, 1, 0, 1]) ^ z2
+z4 = np.array([0, 1, 1, 1, 1, 1, 1, 0]) ^ z3
+z5 = np.array([1, 1, 0, 0, 0, 1, 1, 1]) ^ z4
+z6 = np.array([0, 1, 1, 0, 1, 0, 1, 1]) ^ z5
+z7 = np.array([0, 0, 1, 1, 1, 1, 0, 1]) ^ z6
+z8 = np.array([1, 0, 0, 1, 1, 1, 1, 0]) ^ z7
+
+print(z1)
+print(z2)
+print(z3)
+print(z4)
+print(z5)
+print(z6)
+print(z7)
+print(z8)

### 2023-05-16 scripts/camellia_sliced_perm.py

@@ -1,0 +1,39 @@
+#!/usr/bin/env python3
+
+import numpy as np
+
+first = np.array([
+    (0, 1, 1, 1),
+    (1, 0, 1, 1),
+    (1, 1, 0, 1),
+    (1, 1, 1, 0),
+    (0, 1, 1, 1),
+    (1, 0, 1, 1),
+    (1, 1, 0, 1),
+    (1, 1, 1, 0)
+])
+
+second = np.array([
+    (1, 0, 0, 1),
+    (1, 1, 0, 0),
+    (0, 1, 1, 0),
+    (0, 0, 1, 1),
+    (1, 1, 1, 0),
+    (0, 1, 1, 1),
+    (1, 0, 1, 1),
+    (1, 1, 0, 1)
+])
+
+for i in range(15, -1, -1):
+    vec = np.array([(i >> 0) & 0x1, (i >> 1) & 0x1, (i >> 2) & 0x1, (i >> 3) & 0x1]).reshape(4, 1)
+    result = ((first @ vec) % 2) * np.array([1, 2, 4, 8, 16, 32, 64, 128]).reshape(8, 1)
+    print(hex(result.sum())[2:], end=' ')
+
+print('')
+
+for i in range(15, -1, -1):
+    vec = np.array([(i >> 0) & 0x1, (i >> 1) & 0x1, (i >> 2) & 0x1, (i >> 3) & 0x1]).reshape(4, 1)
+    result = ((second @ vec) % 2) * np.array([1, 2, 4, 8, 16, 32, 64, 128]).reshape(8, 1)
+    print(hex(result.sum())[2:], end=' ')
+
+print('')

### 2023-05-17 Paper/chapters/implementation.tex
camellia galois affine isomorphism
add macros to paper

@@
 
+Lots of operations require extraction of a certain number of consecutive bits,
+usually referred to as a bitfield. Indices for table lookups are generally
+attained by right-shifting a larger value stored in a register, then applying
+an AND operation to get the lowest $n$ bits and finally writing the result into
+the beginning of another register. Due to its usefulness, this operation is
+implemented as \texttt{UBFX} for an unsigned bitfield extract and can be used
+to implement S-box extraction for subperm lookups which would otherwise take
+two or three instructions.
+
 \subsection{Using \texttt{vperm}}
@@
 \subsection{Bitslicing}
+\subsubsection{Shifts for swapmove}
 
+Implementing shift functions for 128-bit NEON registers by extracting the
+overflow and adding it back in later can be realized using 5 instructions. It
+would be most useful for the function to take a shift amount parameter, but
+most NEON intrinsics encode their parameters into the final machine code such
+that parameters need to be compile-time constants. We will therefore implement
+\texttt{shl}, \texttt{shr} and \texttt{swapmove} as C macros utilizing the
+\texttt{vextq\_u64} intrinsic to swap the two 64-bit vector elements:
+
+\begin{lstlisting}{language=c, caption={Shift and swapmove}}
+
+/*
+uint8x16_t shl(const uint8x16_t v, const int n) */
+#define shl(_a, v, n)                                                       \
+{                                                                           \
+        uint64x2_t _overflow = vshrq_n_u64(v, 64 - n);                      \
+        _overflow = vextq_u64(vdupq_n_u64(0x0), _overflow, 1);              \
+        _a = vorrq_u8(vshlq_n_u64(v, n), _overflow);                        \
+}
+
+/*
+uint8x16_t shr(const uint8x16_t v, const int n) */
+#define shr(_a, v, n)                                                       \
+{                                                                           \
+        uint64x2_t _overflow = vshlq_n_u64(v, 64 - n);                      \
+        _overflow = vextq_u64(_overflow, vdupq_n_u64(0x0), 1);              \
+        _a = vorrq_u8(vshrq_n_u64(v, n), _overflow);                        \
+}
+
+/* implemented as a macro so we can use vshlq_n_u8 with variable n
+void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
+                                 const uint8x16_t m, const int n) */
+#define gift_64_vec_sliced_swapmove(a, b, m, n)                             \
+{                                                                           \
+        uint8x16_t _a;                                                      \
+        shr(_a, a, n);                                                      \
+        const uint8x16_t _t = vandq_u8(veorq_u8(_a, b), m);                 \
+        b = veorq_u8(b, _t);                                                \
+        shl(_a, _t, n);                                                     \
+        a = veorq_u8(a, _a);                                                \
+}
+\end{lstlisting}
+
+\subsubsection{Round function}
 We will examine the round function in closer detail and compare the source code
 with the generated assembly.
-\pagebreak
 
-\begin{lstlisting}[language=c, caption={Round function src}]
+\begin{lstlisting}[language=c, caption={Round function}]
 for (int round = 0; round < ROUNDS_GIFT_64; round++) {
@@
 
-Without needing to dive too deep into details, it is obvious to see the
-compiler has inlined the two function calls to \texttt{subcells} and
-\texttt{permute} with lines 1 to 23 originating from the \texttt{subcells} and
-lines 24-46 from the \texttt{permute} and round key addition functions. It has
-chosen not to unroll the loop, but has moved the loop counter increment as well
-as the condition check in between the \texttt{subcells} instructions to line
-\ref{lst:cntrincr} and \ref{lst:cntrchk}. In addition, the loop counter serves
-a second purpose as an offset register and is therefore incremented by 0x80=128
-instead of just 1.
+It is obvious to see the compiler has inlined the two function calls to
+\texttt{subcells} and \texttt{permute} with lines 1 to 23 originating from the
+\texttt{subcells} and lines 24-46 from the \texttt{permute} and round key
+addition functions. It has chosen not to unroll the loop, but has moved the
+loop counter increment as well as the condition check in between the
+\texttt{subcells} instructions to line \ref{lst:cntrincr} and
+\ref{lst:cntrchk}. In addition, the loop counter serves a second purpose as an
+offset register and is therefore incremented by 0x80=128 instead of just 1.
 

### 2023-05-17 impl/benchmark.c

@@ -375,10 +375,10 @@ int main(int argc, char *argv[])
         /* benchmark_gift_128(); */
         /* benchmark_gift_64_table(); */
         /* benchmark_gift_64_vec_sbox(); */
-        /* benchmark_gift_64_vec_sliced(); */
-        benchmark_camellia_naive();
-        benchmark_camellia_spec_opt();
-        benchmark_camellia_sliced();
+        benchmark_gift_64_vec_sliced();
+        /* benchmark_camellia_naive(); */
+        /* benchmark_camellia_spec_opt(); */
+        /* benchmark_camellia_sliced(); */
 }
 
 #pragma clang optimize on

### 2023-05-18 Paper/chapters/appendix.tex
rename files, add implementation sample code

@@
 
-\lstinputlisting[language=c, caption={\texttt{gift\_table.h}}]{../impl/table/gift_table.h}
-\lstinputlisting[language=c, caption={\texttt{gift\_table.c}}]{../impl/table/gift_table.c}
+\lstinputlisting[language=c, caption={\texttt{gift\_table.h}}]{../impl/gift/table.h}
+\lstinputlisting[language=c, caption={\texttt{gift\_table.c}}]{../impl/gift/table.c}
 
@@
 
-\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.h}}]{../impl/vector/gift_vec_sbox.h}
-\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.c}}]{../impl/vector/gift_vec_sbox.c}
+\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.h}}]{../impl/gift/vec_sbox.h}
+\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.c}}]{../impl/gift/vec_sbox.c}
 
@@
 
-\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.h}}]{../impl/vector/gift_vec_sliced.h}
-\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.c}}]{../impl/vector/gift_vec_sliced.c}
+\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.h}}]{../impl/gift/vec_sliced.h}
+\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.c}}]{../impl/gift/vec_sliced.c}
 

### 2023-05-18 Paper/chapters/implementation-strategies.tex
add macros to paper
rename files, add implementation sample code

@@
 on our own. This can be done by extracting and shifting the overflow $ov$ out
-of $V=b||a$, shifting the lanes individually and finally ORing the overflow
-bits to the corresponding vector element.
+of $V=V[1]||V[0]$, shifting the lanes individually and finally ORing the
+overflow bits to the corresponding vector element.
 
@@
     &\text{shl}(V,n): \\
-    &\qquad ov&&=V[0]\land(\lnot(\texttt{0xffffffffffffffff}\gg_n))\gg_{63-n} \\
-    &\qquad V[0]&&=V[0]\gg_n \\
-    &\qquad V[1]&&=(V[1]\gg_n)\lor ov
+    &\qquad ov&&=V[0]\gg_{64-n} \\
+    &\qquad V[0]&&=V[0]\ll \\
+    &\qquad V[1]&&=(V[1]\ll)\lor ov
 \end{alignat*}
@@
 The following operations group all $i$-th bits of the messages $a,c,\dots,o$
-into bytes and puts these into the lower half of the registers $V_{i\bmod 8}$.
+into bytes and put these into the lower half of the registers $V_{i\bmod 8}$.
 The same is done for messages $b,d,\dots,p$, only differing in that the bytes
@@
 
-We choose to encrypt 16 plaintext blocks at once so we can fill 16 vector
-registers with register $V_i$ containing all $i$-th bytes. This lends itself well
-to a NEON implementation since S-boxes as well as the FL layer, Feistel round
-and packing/unpacking functions can be implemented efficiently.
+We choose a bytesliced representation by encrypting 16 plaintext blocks at once
+so we can fill 16 vector registers with register $V_i$ containing all $i$-th
+bytes. This lends itself well to a NEON implementation since S-boxes as well as
+the FL layer, Feistel round and packing/unpacking functions can be implemented
+efficiently.
 
@@
     \includegraphics[width=\textwidth]{Figures/byteslicing.pdf}
-    \caption{Camellia byte-sliced packing function with 32 \texttt{TBL} operations.
-    Curly bracets represent a 4-element vector array being used as source registers.}
+    \caption{Camellia bytesliced packing function with 32 \texttt{TBL} operations.
+    Curly brackets represent a 4-element vector array being used as source registers.}
     \label{fig:byteslicing}
@@
 
-A byte-sliced implementation strategy for the S-box on ARMv8 can be derived
+A bytesliced implementation strategy for the S-box on ARMv8 can be derived
 from already existing x86-optimized implementations utilizing the AES-NI
@@
 
-% TODO add graphic here
-
 \begin{figure}[h!]
@@
     \includegraphics[width=\textwidth]{Figures/camelliaround.pdf}
-    \caption{Camellia byte-sliced Feistel round function}
+    \caption{Camellia bytesliced Feistel round function}
 \end{figure}
@@
 This can be facilitated by extracting the highest bit of each byte, shifting it
-all the way to the right and storing it as overflow $o_i$. $o_i$ is then added
+all the way to the right and storing it as overflow $ov_i$. $ov_i$ is then added
 to the left-shifted value of byte $(i + 4) (\bmod 4)$ for the final result.
 
-% TODO add graphic here
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{Figures/camellia_bytesliced_rol.pdf}
+    \caption{Camellia bytesliced 1-bit left rotate}
+\end{figure}

### 2023-05-18 Paper/chapters/implementation.tex
rename files, add implementation sample code

@@
 
-\lstinputlisting[language=c, firstline=77, lastline=87]{../impl/table/gift_table.c}
-\lstinputlisting[language=c, firstline=89, lastline=101]{../impl/table/gift_table.c}
+\lstinputlisting[language=c, caption={GIFT-64 table subperm}, firstline=77, lastline=87]{../impl/gift/table.c}
+\lstinputlisting[language=c, caption={GIFT-64 table encrypt}, firstline=89, lastline=101]{../impl/gift/table.c}
 
@@
 to implement S-box extraction for subperm lookups which would otherwise take
-two or three instructions.
+two or three instructions. Interestingly, the AArch64 instruction set makes
+heavy use of instruction aliasing. The logical shift left instruction
+\texttt{lsl} for example is an alias of \texttt{UBFX} which itself is an alias
+for \texttt{UBFM}.
+
+Another keyword aiding in optimization is \texttt{restrict} which can be used
+for pointer and array function parameters; the programmer can add this keyword
+to parameters to tell the compiler they are never aliased by any other pointers
+which allows the compiler to rearrange instructions and eliminate loads.
 
@@
 
-\lstinputlisting[language=c, firstline=92, lastline=95]{../impl/vector/gift_vec_sbox.c}
-\lstinputlisting[language=c, firstline=102, lastline=117]{../impl/vector/gift_vec_sbox.c}
-\lstinputlisting[language=c, firstline=197, lastline=212]{../impl/vector/gift_vec_sbox.c}
+\lstinputlisting[language=c, caption={\texttt{vperm} S-box}, firstline=95, lastline=98]{../impl/gift/vec_sbox.c}
+\lstinputlisting[language=c, caption={\texttt{vperm} permutation}, firstline=105, lastline=120]{../impl/gift/vec_sbox.c}
+\lstinputlisting[language=c, caption={\texttt{vperm} encrypt function}, firstline=200, lastline=215]{../impl/gift/vec_sbox.c}
 
 \subsection{Bitslicing}
+
+\subsubsection{A note on data storage}
+
+NEON only provides 32 vector registers. While this is more than the 16 YMM
+registers offered by Intel's AVX-256 vector extension, it is not enough to
+accompany all 28 round keys for GIFT-64 plus the 16 registers representing
+cipher state at once. Because loading and storing single registers is
+inefficient, data is represented using the vector array type
+\texttt{uint8x16x4\_t}. Loads and stores are then assembled such that the whole
+array is loaded or stored in a single instruction. Loading a single vector from
+memory for example has a latency of 5 cycles while loading four vectors into a
+vector array can be done in 8 cycles which results in an amortized cost of 2
+cycles per vector. Vectors are grouped into vector arrays as often as possible
+to reduce the number of necessary loads and stores.
+
 \subsubsection{Shifts for swapmove}
@@
 
-\begin{lstlisting}{language=c, caption={Shift and swapmove}}
-
-/*
-uint8x16_t shl(const uint8x16_t v, const int n) */
-#define shl(_a, v, n)                                                       \
-{                                                                           \
-        uint64x2_t _overflow = vshrq_n_u64(v, 64 - n);                      \
-        _overflow = vextq_u64(vdupq_n_u64(0x0), _overflow, 1);              \
-        _a = vorrq_u8(vshlq_n_u64(v, n), _overflow);                        \
-}
-
-/*
-uint8x16_t shr(const uint8x16_t v, const int n) */
-#define shr(_a, v, n)                                                       \
-{                                                                           \
-        uint64x2_t _overflow = vshlq_n_u64(v, 64 - n);                      \
-        _overflow = vextq_u64(_overflow, vdupq_n_u64(0x0), 1);              \
-        _a = vorrq_u8(vshrq_n_u64(v, n), _overflow);                        \
-}
-
-/* implemented as a macro so we can use vshlq_n_u8 with variable n
-void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
-                                 const uint8x16_t m, const int n) */
-#define gift_64_vec_sliced_swapmove(a, b, m, n)                             \
-{                                                                           \
-        uint8x16_t _a;                                                      \
-        shr(_a, a, n);                                                      \
-        const uint8x16_t _t = vandq_u8(veorq_u8(_a, b), m);                 \
-        b = veorq_u8(b, _t);                                                \
-        shl(_a, _t, n);                                                     \
-        a = veorq_u8(a, _a);                                                \
-}
-\end{lstlisting}
+\lstinputlisting[language=c, caption={Bitsliced GIFT swapmove and shift macros}, firstline=50, lastline=79]{../impl/gift/vec_sliced.c}
 
@@
 with the generated assembly.
+
+\lstinputlisting[language=c, caption={Bitsliced GIFT S-box}, firstline=157, lastline=180]{../impl/gift/vec_sliced.c}
+\lstinputlisting[language=c, caption={Bitsliced GIFT permutation}, firstline=205, lastline=216]{../impl/gift/vec_sliced.c}
 
@@
 dispatched to the load execution unit in line \ref{lst:floatload}, the
-processor happily continues chugging away at the instruction stream by issuing
-\texttt{tbl} and \texttt{mvn} $\mu$ops to other execution units.
+processor continues processing the instruction stream by issuing \texttt{tbl}
+and \texttt{mvn} $\mu$ops to other execution units.
 
@@
 handcrafted assembly.
+
+\section{Camellia}
+\subsection{Optimized non-SIMD implementation}
+
+An optimized non-SIMD implementation based on the platform-independent
+techniques presented in the original paper is relatively easy to achieve since
+all functional components are already well defined. One thing to take note of
+however is the fact that the specification is based on a big endian
+representation while ARMv8 is a little endian machine. The problem of
+endianness manifests itself whenever a memory-register interaction takes place,
+i.e. for loads and stores. Arithmetic on a register is unaffected since a
+register is always treated as one large number with no conception of addresses.
+
+We will store input data and arrays in Camellia byte-order such that a lower
+array element actually represents a higher numerical value, i.e. the numerical
+value of
+
+\begin{center}
+    \texttt{\{ 0x0123456789abcdefUL, 0xfedcba9876543210UL \}}
+\end{center}
+
+is equal to
+
+\begin{center}
+    \texttt{0x0123456789abcdeffedcba9876543210}
+\end{center}
+
+which is equal to the byte string
+
+\begin{center}
+    \texttt{01 23 45 67 89 ab cd ef fe dc ba 98 76 54 32 10}
+\end{center}
+
+This allows us to change the endianness of the implementation while keeping
+input in the original Camellia byte order. A change in implementation
+endianness manifest itself for example in table lookups where, according to
+Camellia specification, the first byte is expected to be the most significant
+byte:
+
+\lstinputlisting[language=c, caption={Camellia optimized SP function with reverse byte order}, firstline=29, lastline=45]{../impl/camellia/spec_opt.c}
+
+\subsection{Bytesliced implementation}
+
+\subsubsection{Key generation}
+
+A possible approach to storing keys is storing them in bytesliced format so
+they can later be loaded from memory and used directly since we only have 32
+vector registers in total and half of them are already occupied by the current
+cipher state. This can be implemented by first generating 64 bit round keys and
+then filling register $V_i$ with the $i$-th byte by means of
+\texttt{vdupq\_n\_u8} for each key.
+
+\lstinputlisting[language=c, caption={Bytesliced Camellia round key generation}, firstline=237, lastline=266]{../impl/camellia/bytesliced.c}
+
+Another way to implement key generation is not to store the keys bytesliced,
+but rather to store the 64 bit values and create necessary vector registers on
+the fly. This saves 3120 bytes of memory and could show some usefulness for
+memory-constrained environments, but benchmarking shows a performance drop of
+about $3.6\%$. This is due to the fact that \texttt{vdupq\_n\_u8} has a latency
+of 8 cycles and a throughput of only 1, making it a great deal slower than
+simple vector array loads.
+
+\subsubsection{F-function}
+
+The s-function implements matrix multiplication with a given pre- and
+postfilter using the aforementioned strategy.
+
+\lstinputlisting[language=c, caption={Bytesliced Camellia S-box}, firstline=40, lastline=65]{../impl/camellia/bytesliced.c}
+
+Just as before we need to invert the byte order since the first vector $V_0$
+contains all least significant bytes, but the Camellia specification places the
+most significant bytes at the beginning.
+
+\lstinputlisting[language=c, caption={Bytesliced Camellia F-function}, firstline=67, lastline=109]{../impl/camellia/bytesliced.c}

### 2023-05-18 impl/benchmark.c

@@ -1,10 +1,11 @@
 #pragma clang optimize off
 
-#include "naive/gift.h"
-#include "naive/gift_sliced.h"
-#include "table/gift_table.h"
-#include "vector/gift_vec_sbox.h"
-#include "vector/gift_vec_sliced.h"
+#include "gift/naive.h"
+#include "gift/sliced.h"
+#include "gift/table.h"
+#include "gift/vec_sbox.h"
+#include "gift/vec_sliced.h"
+
 #include "camellia/naive.h"
 #include "camellia/spec_opt.h"
 #include "camellia/bytesliced.h"
@@ -375,10 +376,10 @@ int main(int argc, char *argv[])
         /* benchmark_gift_128(); */
         /* benchmark_gift_64_table(); */
         /* benchmark_gift_64_vec_sbox(); */
-        benchmark_gift_64_vec_sliced();
+        /* benchmark_gift_64_vec_sliced(); */
         /* benchmark_camellia_naive(); */
         /* benchmark_camellia_spec_opt(); */
-        /* benchmark_camellia_sliced(); */
+        benchmark_camellia_sliced();
 }
 
 #pragma clang optimize on

### 2023-05-18 impl/camellia/bytesliced.c

@@ -1,10 +1,10 @@
-#include "bytesliced.h"
-
-#include "spec_opt.h" // need the spec_opt key schedule
-
-#include <stdio.h>
-#include <stdlib.h>
+#include <arm_neon.h>
+#include <stdint.h>
+#include <stddef.h>
 #include <string.h>
+
+#include "bytesliced.h"
+#include "spec_opt.h" // need the spec_opt key schedule
 
 static uint8x16x4_t pack_group;
 static uint8x16x4_t pack_group_inv;

### 2023-05-18 impl/camellia/bytesliced.h

@@ -1,12 +1,12 @@
-#ifndef BYTESLICED_H
-#define BYTESLICED_H
+#pragma once
 
 // 128-bit bitsliced camellia with 16 blocks encrypted in parallel
 
-#include "camellia_keys.h"
 
 #include <stdint.h>
 #include <arm_neon.h>
+
+#include "camellia_keys.h"
 
 void rol32_1(uint8x16x4_t *a);
 
@@ -38,5 +38,3 @@ void camellia_sliced_encrypt_128(uint64_t c[restrict 16][2],
 void camellia_sliced_decrypt_128(uint64_t m[restrict 16][2],
                                  const uint64_t c[restrict 16][2],
                                  struct camellia_rks_sliced_128 *restrict rks);
-
-#endif

### 2023-05-18 impl/camellia/camellia_keys.h

@@ -1,5 +1,4 @@
-#ifndef CAMELLIA_KEYS
-#define CAMELLIA_KEYS
+#pragma once
 
 #include <stdint.h>
 #include <arm_neon.h>
@@ -21,5 +20,3 @@ struct camellia_rks_sliced_128 {
         uint8x16x4_t ku[18][2]; // used in F function
         uint8x16x4_t kl[4][2]; // used in FL layer
 };
-
-#endif

### 2023-05-18 impl/camellia/spec_opt.c

@@ -1,9 +1,9 @@
+#include <stdint.h>
+#include <stddef.h>
+#include <string.h>
+
 #include "spec_opt.h"
 #include "spec_opt_table.h"
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
 
 static const uint64_t keysched_const[] = {
         0xa09e667f3bcc908bUL,
@@ -21,11 +21,10 @@ static const uint64_t keysched_const[] = {
         (_a) = ((_a) << (_n)) | ((_t) >> (64 - (_n)));\
 }
 
-#define expr_rol32_1(_a)\
-        (((_a) << 1) | ((_a) >> 31))
-
-#define expr_rol32_8(_a)\
-        (((_a) << 8) | ((_a) >> 24))
+static uint32_t rol32_1(uint32_t a)
+{
+        return (a << 1) | (a >> 31);
+}
 
 uint64_t camellia_spec_opt_F(uint64_t X, const uint64_t k)
 {
@@ -53,7 +52,7 @@ uint64_t camellia_spec_opt_FL(uint64_t X, const uint64_t kl)
         const uint32_t klL = (kl >> 32);
         const uint32_t klR = (kl >> 0);
 
-        const uint32_t YR = expr_rol32_1(XL & klL) ^ XR;
+        const uint32_t YR = rol32_1(XL & klL) ^ XR;
         const uint32_t YL = (YR | klR) ^ XL;
 
         return ((uint64_t)YL << 32) | (uint64_t)YR;
@@ -68,7 +67,7 @@ uint64_t camellia_spec_opt_FL_inv(uint64_t Y, const uint64_t kl)
         const uint32_t klR = (kl >> 0);
 
         const uint32_t XL = (YR | klR) ^ YL;
-        const uint32_t XR = expr_rol32_1(XL & klL) ^ YR;
+        const uint32_t XR = rol32_1(XL & klL) ^ YR;
 
         return ((uint64_t)XL << 32) | (uint64_t)XR;
 }
@@ -143,7 +142,7 @@ void camellia_spec_opt_generate_round_keys_128(struct camellia_rks_128 *restrict
 
 void camellia_spec_opt_encrypt_128(uint64_t c[restrict 2],
                             const uint64_t m[restrict 2],
-                            struct camellia_rks_128 *restrict rks)
+                            const struct camellia_rks_128 *restrict rks)
 {
         memcpy(c, m, sizeof(c[0]) * 2);
 
[…]

### 2023-05-18 impl/camellia/spec_opt.h

@@ -1,10 +1,10 @@
-#ifndef SPEC_OPT_H
-#define SPEC_OPT_H
+#pragma once
 
 // 128-bit and 256-bit Camellia
 
-#include "camellia_keys.h"
 #include <stdint.h>
+
+#include "camellia_keys.h"
 
 uint64_t camellia_spec_opt_F(uint64_t X, const uint64_t k);
 uint64_t camellia_spec_opt_FL(uint64_t X, const uint64_t kl);
@@ -16,10 +16,8 @@ void camellia_spec_opt_generate_round_keys_128(struct camellia_rks_128 *restrict
 
 void camellia_spec_opt_encrypt_128(uint64_t c[restrict 2],
                             const uint64_t m[restrict 2],
-                            struct camellia_rks_128 *restrict rks);
+                            const struct camellia_rks_128 *restrict rks);
 
 void camellia_spec_opt_decrypt_128(uint64_t m[restrict 2],
                             const uint64_t c[restrict 2],
-                            struct camellia_rks_128 *restrict rks);
-
-#endif
+                            const struct camellia_rks_128 *restrict rks);

### 2023-05-18 impl/camellia/spec_opt_table.h

@@ -1,5 +1,4 @@
-#ifndef SPEC_OPT_TABLE_H
-#define SPEC_OPT_TABLE_H
+#pragma once
 
 #include <stdint.h>
 
@@ -11,5 +10,3 @@ extern const uint64_t SP4[256];
 extern const uint64_t SP5[256];
 extern const uint64_t SP6[256];
 extern const uint64_t SP7[256];
-
-#endif

### 2023-05-18 impl/gift/naive.h

@@ -1,5 +1,4 @@
-#ifndef GIFT_H
-#define GIFT_H
+#pragma once
 
 #include <stdint.h>
 
@@ -31,5 +30,3 @@ void gift_128_encrypt(uint8_t c[restrict 16],
 void gift_128_decrypt(uint8_t m[restrict 16],
                       const uint8_t c[restrict 16],
                       const uint8_t round_keys[restrict ROUNDS_GIFT_128][32]);
-
-#endif

### 2023-05-18 impl/gift/sliced.c

@@ -1,7 +1,7 @@
-#include "gift_sliced.h"
+#include <stdint.h>
+#include <stddef.h>
 
-#include <stdio.h>
-#include <string.h>
+#include "sliced.h"
 
 static const int round_constant[] = {
         // rounds 0-15

### 2023-05-18 impl/gift/sliced.h

@@ -1,3 +1,5 @@
+#pragma once
+
 #include <stdint.h>
 
 #define ROUNDS_GIFT_SLICED_64 28

### 2023-05-18 impl/gift/table.c

@@ -1,7 +1,7 @@
-#include "gift_table.h"
+#include <stdint.h>
+#include <stddef.h>
 
-#include <stdlib.h>
-#include <string.h>
+#include "table.h"
 
 static const int round_const[] = {
         // rounds 0-15

### 2023-05-18 impl/gift/table.h

@@ -1,5 +1,4 @@
-#ifndef GIFT_TABLE_H
-#define GIFT_TABLE_H
+#pragma once
 
 #include <stdint.h>
 
@@ -13,5 +12,3 @@ uint64_t gift_64_table_subperm(const uint64_t cipher_state);
 // can only encrypt using table technique!
 uint64_t gift_64_table_encrypt(const uint64_t m,
                                const uint64_t rks[restrict ROUNDS_GIFT_64]);
-
-#endif

### 2023-05-18 impl/gift/vec_sbox.c

@@ -1,8 +1,8 @@
-#include "gift_vec_sbox.h"
-
-#include <stdint.h>
 #include <arm_neon.h>
-#include <string.h>
+#include <stdint.h>
+#include <stddef.h>
+
+#include "vec_sbox.h"
 
 static uint64_t sbox_vec_u64[2] = {
         0x09030f060c040a01UL, 0x0e080005070b0d02UL
@@ -46,22 +46,25 @@ uint8x16_t gift_64_vec_sbox_bits_pack(const uint64_t a)
 uint64_t gift_64_vec_sbox_bits_unpack(const uint8x16_t v)
 {
         uint64_t a = 0UL;
-        a |= (uint64_t)vgetq_lane_u8(v, 0) << 4 * 0;
-        a |= (uint64_t)vgetq_lane_u8(v, 1) << 4 * 1;
-        a |= (uint64_t)vgetq_lane_u8(v, 2) << 4 * 2;
-        a |= (uint64_t)vgetq_lane_u8(v, 3) << 4 * 3;
-        a |= (uint64_t)vgetq_lane_u8(v, 4) << 4 * 4;
-        a |= (uint64_t)vgetq_lane_u8(v, 5) << 4 * 5;
-        a |= (uint64_t)vgetq_lane_u8(v, 6) << 4 * 6;
-        a |= (uint64_t)vgetq_lane_u8(v, 7) << 4 * 7;
-        a |= (uint64_t)vgetq_lane_u8(v, 8) << 4 * 8;
-        a |= (uint64_t)vgetq_lane_u8(v, 9) << 4 * 9;
-        a |= (uint64_t)vgetq_lane_u8(v, 10) << 4 * 10;
-        a |= (uint64_t)vgetq_lane_u8(v, 11) << 4 * 11;
-        a |= (uint64_t)vgetq_lane_u8(v, 12) << 4 * 12;
-        a |= (uint64_t)vgetq_lane_u8(v, 13) << 4 * 13;
-        a |= (uint64_t)vgetq_lane_u8(v, 14) << 4 * 14;
-        a |= (uint64_t)vgetq_lane_u8(v, 15) << 4 * 15;
+        uint64_t lane = vgetq_lane_u64(v, 0);
+        a = (uint64_t)((lane >> 8 * 0) & 0xf) << 4 * 0 |
+            (uint64_t)((lane >> 8 * 1) & 0xf) << 4 * 1 |
[…]

### 2023-05-18 impl/gift/vec_sbox.h

@@ -1,8 +1,7 @@
-#ifndef GIFT_VEC_SBOX_H
-#define GIFT_VEC_SBOX_H
+#pragma once
 
-#include <stdint.h>
 #include <arm_neon.h>
+#include <stdint.h>
 
 #define ROUNDS_GIFT_64 28
 
@@ -23,5 +22,3 @@ uint64_t gift_64_vec_sbox_encrypt(const uint64_t m,
                                   const uint8x16_t rks[restrict ROUNDS_GIFT_64]);
 uint64_t gift_64_vec_sbox_decrypt(const uint64_t c,
                                   const uint8x16_t rks[restrict ROUNDS_GIFT_64]);
-
-#endif

### 2023-05-18 impl/gift/vec_sliced.c

@@ -1,8 +1,8 @@
-#include "gift_vec_sliced.h"
-
+#include <arm_neon.h>
+#include <stdint.h>
 #include <stddef.h>
-#include <stdio.h>
-#include <string.h>
+
+#include "vec_sliced.h"
 
 static uint64_t pack_shf_u64[] = {
         0x1303120211011000UL, 0x1707160615051404UL, // S0/S1/S2/S3

### 2023-05-18 impl/gift/vec_sliced.h

@@ -1,5 +1,4 @@
-#ifndef GIFT_VEC_SLICED_H
-#define GIFT_VEC_SLICED_H
+#pragma once
 
 #include <stdint.h>
 #include <arm_neon.h>
@@ -29,5 +28,3 @@ void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
 void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
                                 const uint64_t c[restrict 16],
                                 const uint8x16x4_t rks[restrict ROUNDS_GIFT_64][2]);
-
-#endif

### 2023-05-18 Paper/chapters/appendix.tex
benchmark plot

@@
 \appendix
-\chapter{C implementations}
+\chapter{Detailed benchmarking results}
+\label{app:benchmarks}
+
+\section{GIFT}
+
+\begin{table}[!htbp]
+    \centering
+    \caption{Benchmarks for GIFT}
+    \scriptsize
+    \begin{tabular}{llcccc}
+        \toprule
+        & & \multicolumn{2}{c}{Cortex-A53} & \multicolumn{2}{c}{Cortex-A73} \\
+        \cmidrule(lr){3-4}\cmidrule(lr){5-6}
+        Strategy & Component & $lat$ (c/B) & $thr$ (MiB/s) & $lat$ (c/B) & $thr$ (MiB/s) \\
+        \midrule
+        \multirow{2}{*}{Naive GIFT-64} & \texttt{round\_keys} & 223.54 & \multirow{2}{*}{1.22} & 190.34 & \multirow{2}{*}{2.33} \\
+                                                & \texttt{encrypt} & 1367.66 & &  830.49 & \\
+        \cmidrule(lr){2-2}
+                                                & \texttt{ subcells} & 4.64 & & 3.13 & \\
+                                                & \texttt{  permute} & 36.68 & & 21.09 & \\
+        \midrule
+        \multirow{2}{*}{Naive GIFT-128} & \texttt{round\_keys} & 182.91 & \multirow{2}{*}{0.51} & 167.16 & \multirow{2}{*}{0.79} \\
+                                                 & \texttt{encrypt} & 3532.52 & & 2615.89 & \\
+        \cmidrule(lr){2-2}
+                                                 & \texttt{ subcells} & 6.91 & & 2.73 & \\
+                                                 & \texttt{ permute} & 82.68 & & 61.27 & \\
+        \midrule
+        \multirow{2}{*}{Table-driven} & \texttt{round\_keys} & 223.81 & \multirow{2}{*}{13.59} & 190.11 & \multirow{2}{*}{16.09} \\
+                                      & \texttt{encrypt} & 122.11 & & 119.62 & \\
+        \cmidrule(lr){2-2}
+                                      & \texttt{ subperm} & 5.15 & & 4.47 & \\
+        \midrule
+        \multirow{2}{*}{\texttt{vperm} S-box} & \texttt{round\_keys} & 308.12 & \multirow{2}{*}{1.10} & 270.69 & \multirow{2}{*}{1.58} \\
+                                              & \texttt{encrypt} & 1514.28 & & 1218.18 & \\
+        \cmidrule(lr){2-2}
+                                              & \texttt{ subcells} & 1.63 & & 1.13 & \\
+                                              & \texttt{ permute} & 53.18 & & 42.40 & \\
+                                              & \texttt{ pack} & 7.51 & & 1.24 & \\
+                                              & \texttt{ unpack} & 7.39 & & 4.67 & \\
+        \midrule
+        \multirow{2}{*}{Bitsliced} & \texttt{round\_keys} & 19.93 & \multirow{2}{*}{108.80} & 16.12 & \multirow{2}{*}{150.50} \\
+                                   & \texttt{encrypt} & 16.69 & & 13.98 & \\
+        \cmidrule(lr){2-2}
+                                   & \texttt{ subcells} & 0.41 & & 0.06 & \\
+                                   & \texttt{ permute} & 0.39 & & 0.18 & \\
+                                   & \texttt{ pack} & 1.81 & & 1.34 & \\
+                                   & \texttt{ unpack} & 1.68 & & 1.31 & \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\section{Camellia}
+
+\begin{table}[!htpb]
+    \centering
+    \caption{Benchmarks for Camellia}
+    \scriptsize
+    \begin{tabular}{llcccc}
+        \toprule
+        & & \multicolumn{2}{c}{Cortex-A53} & \multicolumn{2}{c}{Cortex-A73} \\
+        \cmidrule(lr){3-4}\cmidrule(lr){5-6}
+        Strategy & Component & $lat$ (c/B) & $thr$ (MiB/s) & $lat$ (c/B) & $thr$ (MiB/s) \\
+        \midrule
+        \multirow{2}{*}{Naive Camellia-128} & \texttt{round\_keys} & 15.26 & \multirow{2}{*}{33.74} & 11.08 & \multirow{2}{*}{46.71} \\
+                                                & \texttt{encrypt} & 53.44 & &  43.51 & \\
+        \cmidrule(lr){2-2}
+                                                & \texttt{ feistel\_round} & 4.02 & & 2.66 & \\
+                                                & \texttt{ S} & 2.01 & & 0.94 & \\
+                                                & \texttt{ F} & 3.13 & & 2.25 & \\
+                                                & \texttt{ P} & 2.03 & & 1.63 & \\
+                                                & \texttt{ FL} & 1.06 & & 0.72 & \\
+        \midrule
+        \multirow{2}{*}{Naive Camellia-256} & \texttt{round\_keys} & 22.01 & \multirow{2}{*}{25.50} & 16.99 & \multirow{2}{*}{35.16} \\
+                                                 & \texttt{encrypt} & 70.55 & & 58.27 & \\
+        \midrule
+        \multirow{2}{*}{Optimized Camellia-128} & \texttt{round\_keys} & 8.18 & \multirow{2}{*}{77.07} & 5.93 & \multirow{2}{*}{107.31} \\
+                                                & \texttt{encrypt} & 23.63 & & 19.83 & \\
+        \cmidrule(lr){2-2}
+                                                & \texttt{ feistel\_round} & 2.32 & & 1.56 & \\
+                                                & \texttt{ F} & 2.21 & & 1.43 & \\
+                                                & \texttt{ FL} & 1.06 & & 0.69 & \\
+        \midrule
+        \multirow{2}{*}{Bytesliced Camellia-128} & \texttt{round\_keys} & 2.59 & \multirow{2}{*}{106.20} & 2.19 & \multirow{2}{*}{138.42} \\
+                                                & \texttt{encrypt} & 17.14 & & 15.19 & \\
+        \cmidrule(lr){2-2}
+                                                & \texttt{ feistel\_round} & 1.00 & & 0.69 & \\
+                                                & \texttt{ F} & 0.79 & & 0.58 & \\
+                                                & \texttt{ FL} & 0.36 & & 0.12 & \\
+                                                & \texttt{ pack} & 1.04 & & 0.96 & \\
+                                                & \texttt{ unpack} & 0.91 & & 0.83 & \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\chapter{C source code}
 \label{app:cimpl}

### 2023-05-18 Paper/chapters/implementation.tex
benchmark plot

@@
 
-Just as before we need to invert the byte order since the first vector $V_0$
-contains all least significant bytes, but the Camellia specification places the
-most significant bytes at the beginning.
+Just as before we need to invert the byte order since the first vector
+\texttt{X[0].val[0]} contains all least significant bytes, but the Camellia
+specification places the most significant bytes at the beginning.
 

### 2023-05-18 impl/benchmark.c

@@ -372,14 +372,14 @@ static void benchmark_camellia_sliced(void)
 int main(int argc, char *argv[])
 {
         srand(time(NULL));
-        /* benchmark_gift_64(); */
+        benchmark_gift_64();
         /* benchmark_gift_128(); */
         /* benchmark_gift_64_table(); */
-        /* benchmark_gift_64_vec_sbox(); */
+        benchmark_gift_64_vec_sbox();
         /* benchmark_gift_64_vec_sliced(); */
         /* benchmark_camellia_naive(); */
         /* benchmark_camellia_spec_opt(); */
-        benchmark_camellia_sliced();
+        /* benchmark_camellia_sliced(); */
 }
 
 #pragma clang optimize on

### 2023-05-18 impl/gift/naive.c

@@ -90,7 +90,7 @@ uint64_t gift_64_permute(const uint64_t cipher_state)
         uint64_t new_cipher_state = 0UL;
 
         for (size_t i = 0; i < 64; i++) {
-                int bit = (cipher_state >> i) & 0x1;
+                const int bit = (cipher_state >> i) & 0x1;
                 new_cipher_state |= (uint64_t)bit << perm_64[i];
         }
 

### 2023-05-19 Paper/chapters/appendix.tex
finish draft

@@
 
-\section{Implementations for SPN}
+\section{GIFT}
 \subsection{Table-based}
 
-\lstinputlisting[language=c, caption={\texttt{gift\_table.h}}]{../impl/gift/table.h}
-\lstinputlisting[language=c, caption={\texttt{gift\_table.c}}]{../impl/gift/table.c}
+\lstinputlisting[language=c, caption={\texttt{gift/table.h}}]{../impl/gift/table.h}
+\lstinputlisting[language=c, caption={\texttt{gift/table.c}}]{../impl/gift/table.c}
 
@@
 
-\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.h}}]{../impl/gift/vec_sbox.h}
-\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.c}}]{../impl/gift/vec_sbox.c}
+\lstinputlisting[language=c, caption={\texttt{gift/vec\_sbox.h}}]{../impl/gift/vec_sbox.h}
+\lstinputlisting[language=c, caption={\texttt{gift/vec\_sbox.c}}]{../impl/gift/vec_sbox.c}
 
@@
 
-\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.h}}]{../impl/gift/vec_sliced.h}
-\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.c}}]{../impl/gift/vec_sliced.c}
+\lstinputlisting[language=c, caption={\texttt{gift/vec\_sliced.h}}]{../impl/gift/vec_sliced.h}
+\lstinputlisting[language=c, caption={\texttt{gift/vec\_sliced.c}}]{../impl/gift/vec_sliced.c}
 
-\chapter{Lorem dolor}
+\section{Camellia}
+
+\subsection{Optimized non-SIMD}
+
+\lstinputlisting[language=c, caption={\texttt{camellia/spec\_opt.h}}]{../impl/camellia/spec_opt.h}
+\lstinputlisting[language=c, caption={\texttt{camellia/spec\_opt.c}}]{../impl/camellia/spec_opt.c}
+
+\subsection{Byteslicing}
+
+\lstinputlisting[language=c, caption={\texttt{camellia/bytesliced.h}}]{../impl/camellia/bytesliced.h}
+\lstinputlisting[language=c, caption={\texttt{camellia/bytesliced.c}}]{../impl/camellia/bytesliced.c}

### 2023-05-19 Paper/chapters/implementation-strategies.tex
finish draft

@@
 
-Bitsliced implementation techniques were first introduced to improve the performance
-of DES in 1997 and work by viewing a processor with $n$-bit registers as a
-machine capable of executing $n$ bitwise operations at
-once\cite{desslicing:1997}. Bitslicing offers a performance advantage by
+Bitsliced implementation techniques were first introduced to improve the
+performance of DES in 1997 and work by viewing a processor with $n$-bit
+registers as a machine capable of executing $n$ bitwise operations at
+once \cite{desslicing:1997}. Bitslicing offers a performance advantage by
 splitting up $n$ bits into $m$ slices to achieve a more efficient
@@
 this algorithm to only four registers and eight messages by adjusting the
-\texttt{SWAPMOVE} shift and mask values.
+\texttt{SWAPMOVE} shift and mask values. Encrypting multiple blocks in parallel
+is a realistic use case since usually a lot of blocks are encrypted under the
+same key.
 
@@
 
-\subsection{Platform-independent techniques}
+\subsection{Optimized non-SIMD}
 
@@
 isomorphic, there exist affine isomorphisms $\delta,\delta^{-1}$ between
-GF($2^8$) and GF($(2^4)^2$) respectively\cite{satohmorioka:2003}:
+GF($2^8$) and GF($(2^4)^2$) respectively \cite{satohmorioka:2003}:
 
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{Figures/camellia_bytesliced_rol.pdf}
+    \includegraphics[width=0.8\textwidth]{Figures/camellia_bytesliced_rol.pdf}
     \caption{Camellia bytesliced 1-bit left rotate}

### 2023-05-19 Paper/chapters/introduction.tex
rename files, add implementation sample code
finish draft

@@
 
-\texttt{GIFT}\cite{gift:2017}, first presented in the \textit{CHES 2017}
+\texttt{GIFT} \cite{gift:2017}, first presented in the \textit{CHES 2017}
 cryptographic hardware and embedded systems conference, is a lightweight block
@@
 
-Camellia\cite{camellia:2001} is a block cipher jointly developed by NTT and
+Camellia \cite{camellia:2001} is a block cipher jointly developed by NTT and
 Mitsubishi Electric Corporation and first published in 2001. Following AES
@@
 designed by ARM power 95\% of mobile devices and have found a great many
-applications due to their high performance and low power
-consumption\cite{armcompany}. The ODROID-N2+\cite{odroidn2} development board
-we are using is based on the big.LITTLE architecture and is powered by a
-quad-core ARM Cortex-A73 processor and a weaker dual-core ARM Cortex-A53 for
-power efficiency. Both these processors are part of the eight generation of ARM
-designs known as ARMv8\cite{armv8:2013}.
+applications due to their high performance and low power consumption
+\cite{armcompany}. The ODROID-N2+ \cite{odroidn2} development board we are
+using for testing is powered by an Amlogic S922X processor with four ARM
+Cortex-A73 and two weaker ARM Cortex-A53 cores based on the big.LITTLE
+architecture combining slower, more power efficient cores with faster, more
+power hungry ones. Both these cores are part of the eight generation of ARM
+designs known as ARMv8 \cite{armv8:2013} released in 2011 and succeeding the
+older 32-bit ARMv7.
 
@@
         \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
+        \midrule
+        \texttt{uint64x2\_t} & \texttt{vextq\_u64(uint64x2\_t, uint64x2\_t, int)} & extract from pair of vectors \\
         \bottomrule

### 2023-05-20 Paper/chapters/abstract.tex
bibliography no ul
add abstract

@@
 
-Lorem ipsum dolor \cite{gift:2017} sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
-nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
-fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+Processor designs licensed by ARM are ubiquitous in many areas due to their low
+power consumption and heat generation. Being used in nearly all modern
+smartphones and tablets as well as in many desktop computers and IoT devices,
+the need for securing communications and confidentiality through encryption
+arises. Designing cryptographic algorithms for such low-power environments is
+an active area of research and various designs have been proposed. We examine
+in detail the two block ciphers GIFT and Camellia. For that, we first give an
+introduction of the designs as well as the ARMv8 platform we develop for. We
+then present multiple implementation strategies which leverage ARMv8 SIMD
+processing facilities and realize these in the C programming language using
+NEON intrinsics. In the end, we present performance benchmarks and a brief
+comparison to conclude this work. We find bit- and bytesliced implementations
+employing vector instructions to be the fastest.

### 2023-05-20 Paper/chapters/introduction.tex
add abstract

@@
 \section{Notation}
+
+\begin{center}
+    \begin{tabular}{rl}
+        $\ll_n;\gg_n$ & left; right shift by $n$ bits \\
+        $\lll_n;\ggg_n$ & left; right rotate by $n$ bits \\
+        $\land$ & AND \\
+        $\lor$ & OR \\
+        $\oplus$ & XOR \\
+        $A_{(32)}$ & $A$ is 32 bits long \\
+        $||$ & concatenation
+    \end{tabular}
+\end{center}
 
@@
     \centering
-    \footnotesize
+    \scriptsize
     \caption{Common NEON intrinsics}
@@
         Intrinsic && Summary \\
-        \midrule
-        \texttt{uint64\_t} & \texttt{vgetq\_lane\_u64(void)} & Extract a single lane \\
-        \midrule
-        \texttt{void} & \texttt{vsetq\_lane\_u64(uint64\_t)} & Insert a single lane \\
         \midrule

### 2023-05-30 Paper/chapters/implementation-strategies.tex
finalize

@@
 
-\section{Strategies for \texttt{GIFT}}
+\section{Strategies for GIFT}
 
@@
 sizes, so while tabulating a function like the AES S-box
-$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space,
-tabulating the \texttt{GIFT} permutation layer
-$P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
+$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space, tabulating the
+GIFT permutation layer $P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
 $2^{70}$ space, which is totally unfeasible.
@@
 message chunk $m_i$. This approach requires space of
-$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for \texttt{GIFT-64},
+$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for GIFT-64,
 results in a manageable size of $\frac{64^2 2^4}{4}=2^{14}$ bits which equals
@@
 
-For \texttt{GIFT-64}, table construction is relatively straightforward and can
+For GIFT-64, table construction is relatively straightforward and can
 be done as follows:
@@
 
-The tables for \texttt{GIFT-128} can be generated in a similar way by looping
+The tables for GIFT-128 can be generated in a similar way by looping
 through all 32 S-boxes instead of 16 on line \ref{lst:tablesbox}.
@@
 representation which can exploit this bitwise parallelism. The structure of
-\texttt{GIFT} naturally offers possibilities for bitslicing. We split the
+GIFT naturally offers possibilities for bitslicing. We split the
 cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
@@
 technique above. This can be done by finding a bitwise instruction sequence to
-apply the S-box which has already been proposed by the original \texttt{GIFT}
-authors:
+apply the S-box which has already been proposed by the original GIFT authors:
 
@@
 
-In the case of \texttt{GIFT-64} with $BS=64$ and ARM NEON with $RS=128$, we get
+In the case of GIFT-64 with $BS=64$ and ARM NEON with $RS=128$, we get
 
@@
 The original paper proposes various platform-independent ways to implement
-Camellia efficiently. Only some of these apply to the ARMv8 architecture since
-features like the inline barrel shifter and bitfield manipulation instructions
-generally offer better performance.
+Camellia efficiently \cite{camellia:2001}. Only some of these apply to the
+ARMv8 architecture since features like the inline barrel shifter and bitfield
+manipulation instructions generally offer better performance.
 
@@
             & SP_7(y_{7(8)})=(&s_1(y_7),&& s_1(y_7),&& s_1(y_7),&& 0,&& s_1(y_7),&& s_1(y_7),&& s_1(y_7),&& 0&)
-        \end{alignat*}
+        \end{alignat*},
 
-        , we can compute $(P\circ S)(X_{(64)})$ using only 8 table lookups and 7
+        we can compute $(P\circ S)(X_{(64)})$ using only 8 table lookups and 7
         XORs in the following way:

### 2023-05-30 Paper/chapters/implementation.tex
finish draft
finalize

@@
 Implementations in the C programming language for the presented strategies can
-be found in Appendix \ref{app:cimpl}. Although directly writing Assembler code
+be found in Appendix \ref{app:cimpl}. Although directly writing assembler code
 could result in a small performance benefit, this generally increases the work
@@
 Relying on the compiler mandates a closer study of the generated, optimized
-assembler. All source files were compiled using clang version 15.0.7 and
-optimization level O3.
+assembler code. All source files were compiled using clang version 15.0.7 at
+optimization level -O3.
 
@@
 
-Understanding certain choices requires an understanding of the Cortex-A73
-instruction pipeline\cite{a72opt:2015}. Being a superscalar processor, it is
-able to execute more than one instruction per clock cycle by dispatching
-instructions to different execution units working in parallel.
+Understanding certain implementation choices requires an understanding of the
+Cortex-A73 instruction pipeline \cite{a72opt:2015}. Being a superscalar
+processor, it is able to execute more than one instruction per clock cycle by
+dispatching instructions to different execution units working in parallel.
 
-\begin{figure}{h!}
+\begin{figure}[h!]
     \centering
@@
     \caption{High-level overview of the Cortex-A72 instruction pipeline}
+    \label{fig:pipelining}
 \end{figure}
@@
 to implement S-box extraction for subperm lookups which would otherwise take
-two or three instructions. Interestingly, the AArch64 instruction set makes
+two or three instructions. Interestingly, the A64 instruction set makes
 heavy use of instruction aliasing. The logical shift left instruction
@@
 We will examine the round function in closer detail and compare the source code
-with the generated assembly.
+with the generated assembler code.
 
@@
 \begin{minipage}{.45\textwidth}
-    \begin{lstlisting}[language={[ARM]Assembler}, caption={Round function asm}, escapechar=|]
+    \begin{lstlisting}[language={[ARM]Assembler}, caption={\small Round function assembly}, escapechar=|]
 and     v20.16b, v17.16b, v6.16b
@@
 languages like C and modern-day compilers more often than not outperform
-handcrafted assembly.
+handcrafted assembler code.
 
@@
 An optimized non-SIMD implementation based on the platform-independent
-techniques presented in the original paper is relatively easy to achieve since
-all functional components are already well defined. One thing to take note of
-however is the fact that the specification is based on a big endian
-representation while ARMv8 is a little endian machine. The problem of
-endianness manifests itself whenever a memory-register interaction takes place,
-i.e. for loads and stores. Arithmetic on a register is unaffected since a
-register is always treated as one large number with no conception of addresses.
+techniques presented in the original paper \cite{camellia:2001} is relatively
+easy to achieve since all functional components are already well defined. One
+thing to take note of however is the fact that the specification is based on a
+big endian representation while ARMv8 is a little endian machine. The problem
+of endianness manifests itself whenever a memory-register interaction takes
+place, i.e. for loads and stores. Arithmetic on a register is unaffected since
+a register is always treated as one large number with no conception of
+addresses.
 

### 2023-05-30 Paper/chapters/introduction.tex
finalize

--- 2023-05-19 Paper/chapters/introduction.tex
@@
 \chapter{Introduction}
-\section{Notation}
+
+\section{Motivation}
+
+With small devices and embedded processors becoming ever more ubiquitous and
+essential in areas like consumer electronics or industrial and IoT
+applications, the need for low-power, high-performance microprocessors has
+increased steadily. With more than 250 billion chips shipped, semiconductors
+designed by ARM power 95\% of mobile devices and have found a great many
+applications due to their high performance and low power consumption
+\cite{armcompany}. Following the proliferation of these systems, the need for
+new and efficient cryptographic algorithms to encrypt and decrypt data arises.
+
+Two promising ciphers are GIFT and Camellia. Both of these have not yet been
+implemented for 64-bit ARMv8 - implementations only exist for 32-bit in the case
+of GIFT \cite{fixslicing:2020} and Intel AVX-2 and AES-NI in the case of
+Camellia \cite{bcfastimplx86:2013}. Our contribution is to fill this gap and
+produce efficient implementations of GIFT and Camellia for 64-bit ARMv8.
 
@@
 \[
-    k_7||k_6||\dots||k_1||k_0\leftarrow k_1\ggg 2||k_0\ggg 12||k_7||k_6||\dots||k_3||k_2
+    k_7||k_6||\dots||k_1||k_0\leftarrow k_1\ggg_2||k_0\ggg_{12}||k_7||k_6||\dots||k_3||k_2
 \]
@@
 
-With small devices and embedded processors becoming ever more ubiquitous and
-essential in areas like consumer electronics or industrial and IoT
-applications, the need for low-power, high-performance microprocessors has
-increased steadily. With more than 250 billion chips shipped, semiconductors
-designed by ARM power 95\% of mobile devices and have found a great many
-applications due to their high performance and low power consumption
-\cite{armcompany}. The ODROID-N2+ \cite{odroidn2} development board we are
-using for testing is powered by an Amlogic S922X processor with four ARM
-Cortex-A73 and two weaker ARM Cortex-A53 cores based on the big.LITTLE
-architecture combining slower, more power efficient cores with faster, more
-power hungry ones. Both these cores are part of the eight generation of ARM
-designs known as ARMv8 \cite{armv8:2013} released in 2011 and succeeding the
-older 32-bit ARMv7.
+The ODROID-N2+ \cite{odroidn2} development board we are using for testing is
+powered by an Amlogic S922X processor with four ARM Cortex-A73 and two weaker
+ARM Cortex-A53 cores based on the big.LITTLE architecture combining slower,
+more power efficient cores with faster, more power hungry ones. Both these
+cores are part of the eight generation of ARM designs known as ARMv8
+\cite{armv8:2013} released in 2011 and succeeding the older 32-bit ARMv7.
 
@@
 dynamic execution states with corresponding instruction sets. This work will
-focus on the A profile running in the AArch64 state utilizing the A64
+focus on the A profile running in the AArch64 state utilizing the 64-bit A64
 instruction set with NEON and crypto extensions.
